### 2018-01-30 written_docs/draft.Rmd
Oh goody, start some writing!

---
title             : "Race-based disparities in allocation of discipline are associated with county-level rates of bias"
shorttitle        : "Discipline Disparities"

author: 
  - name          : "Travis Riddle"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : ""
    email         : "triddle@princeton.edu"
  - name          : "Stacey Sinclair"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Princeton University"
  - id            : "2"
    institution   : "Princeton University"

author_note: >

  Enter author note here.

abstract: >
  Enter abstract here.

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---

```{r include = FALSE}
library("papaja")
```
# Introduction

# Methods
## Data Sources
```{r summary-data}
library(dplyr)
library(forcats)
library(haven)
schools_dat <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')

schools_dat %>%
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_ENR_HI_M:TOT_ENR_F) %>%
  gather(group, total_number, SCH_ENR_HI_M:TOT_ENR_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>%
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='NR_', 
                          biracial='TR_', white='WH_')) %>%
  group_by(COMBOKEY, group) %>%
  mutate(total_number = sum(total_number)) %>%
  ungroup() %>%
  select(-prefix, -gender) %>%
  distinct() -> enrollment

enrollment %>% group_by(group) %>% summarise(sum(total_number)) -> enrollment_totals
```
We used three distinct data sources for the work described here. The academic disciplinary data is part of the Civil Rights Data Collection (CRDC) through the US Department of Education. The dataset we used comes from the 2013-2014 academic year and has data on "all public local and educational agencies and schools, including long-term secure juvenile justice facilities, charter schools, alternative schools, and schools serving students with disabilities." In total, the CRDC data represents `r dim(schools_dat)[1]` institutions enrolling approximately 50 million students, of which approximately 25 million are white and 7.8 million are Black.

We obtained county-level demographic information for use as covariates and state-level demographic information for use in post-stratification from the US Census Bureau. For both county-level and state-level demographics, we use 5-year estimatesfor the period ending in 2014 from the American Community Survey, which surveys around 295k households per month.

```{r iat-data}
df <- rbind(read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2013.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2012.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2010.sav'))

df2 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2011.sav')
df3 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2009.sav')
df4 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2008.sav')
df5 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2007.sav')
df6 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2006.sav')
df7 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2005.sav')
df8 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2004.sav')
df9 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2002-2003.sav')
df10 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2014.sav')

"%ni%" <- Negate("%in%")

df %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age)  -> subdat

df2 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all,  
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df3 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df4 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df5 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df6 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df7 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df8 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df9 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(raceomb = ethnic) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df10 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

subdat %>%
  mutate(explicit_bias=tblack_0to10) %>%
  #filter(!is.na(twhite_0to10)) %>%
  #mutate(explicit_bias = twhite_0to10 - tblack_0to10) %>%
  mutate(age_bin = cut(age, breaks=c(14, 24, 34, 54, 75, 120))) %>%
  mutate(race = fct_recode(as.character(raceomb), 
                           'Black'='5', 'White'='6')) %>%
  filter(!is.na(age_bin)) %>%
  filter(race=='White') %>%
  filter(STATE %ni% 
           c('AA', 'AE', 'AP', 'AS', 'FM', 'GU', 'MH', 'MP', 'PR', 'VI')) %>%
  mutate(county_id = paste(STATE, CountyNo, sep='-')) -> individual_data
```

We used measurements of implicit and explicit bias available from data collected through Project Implicit [@xu2014psychology]. From this data, we used only respondents who had geographic information that would allow us to place them in a United States county, identified as White, and visited the site before 2015. This consisted of approximately 1.1 million total respondents.

## Measures

```{r summary_counts}
schools_dat %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
         SCH_DISCWODIS_ISS_HI_M:TOT_DISCWODIS_ISS_F, 
         SCH_DISCWDIS_ISS_IDEA_HI_M:TOT_DISCWDIS_ISS_IDEA_F) %>% 
  gather(group, number, SCH_DISCWODIS_ISS_HI_M:TOT_DISCWDIS_ISS_IDEA_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>% 
  separate(group, into=c('prefix', 'group'), -4) %>%
  separate(prefix, into=c('prefix', 'disability'), -3) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='SS_', 
                          total='EA_', biracial='TR_', white='WH_')) -> tempout_susp
%>%
  mutate(disability=fct_recode(disability, not_disabled='_I', disabled='ID',
                               not_disabled='S_', disabled='A_')) %>%
  select(-prefix) -> susp_inschool

susp_inschool %>%
  filter(group=='black'|group=='white') %>%
  group_by(COMBOKEY, LEA_STATE, LEA_NAME, LEAID,
           CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
  summarise(number = sum(number, na.rm=T)) %>%
  left_join(enrollment) %>%
  mutate(proportion = number/total_number) -> tempout

tempout$number[which(tempout$number<0)] <- NA

tempout %>% 
  ungroup() %>%
  filter(is.finite(proportion)) %>%
  filter(total_number>number) %>%
  mutate(metric='inschool_susp') %>%
  mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
  distinct() -> subdat

###############
schools_dat %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:SCH_PSDISC_SINGOOS_IDEA_F) %>% 
  filter(SCH_GRADE_PS=='YES') %>%
  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:SCH_PSDISC_SINGOOS_IDEA_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>% 
  separate(group, into=c('prefix', 'group'), -4) %>%
  separate(prefix, into=c('prefix', 'disability'), -3) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='OS_', 
                          biracial='TR_', white='WH_')) -> tempout

%>%
  mutate(disability=fct_recode(disability, not_disabled='_I', disabled='ID',
                             not_disabled='GO', disabled='A_')) %>%
  select(-prefix) -> ps_susp

df_school %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
         SCH_GRADE_PS, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>% 
  filter(SCH_GRADE_PS=='YES') %>%
  gather(group, number_2, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>%
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='OS_', 
                          biracial='TR_', white='WH_')) %>%
  select(COMBOKEY, group, gender, number_2) %>%
  right_join(ps_susp) %>%
  filter(number>=0 & number_2 >= 0) %>%
  mutate(number = number+number_2) -> ps_susp

ps_susp %>%
  left_join(enrollment) %>%
  group_by(COMBOKEY, LEA_STATE, LEA_NAME, 
           CCD_LATCOD, CCD_LONCOD,  SCH_NAME, group) %>%
  summarise(number = sum(number),
            total_number = sum(total_number)) %>%
  mutate(proportion = number/total_number) -> full_dat

full_dat %>% 
  filter(group=='black'|group=='white') %>%
  filter(is.finite(proportion)) %>%
  filter(number>=0) %>%
  filter(total_number>number) %>%
  mutate(group=droplevels(group)) %>%
  ungroup() %>%
  mutate(metric='preschool_susp') %>%
  mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
  rbind(subdat) -> subdat

df_school %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
         SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>% 
  gather(group, number, SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>% 
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='XP_', 
                          biracial='TR_', white='WH_')) %>%
  select(-prefix) -> ps_exp

ps_exp %>%
  left_join(enrollment) %>%
  group_by(COMBOKEY, LEA_STATE, LEA_NAME, 
           CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
  summarise(number = sum(number),
            total_number = sum(total_number)) %>%
  mutate(proportion = number/total_number) -> full_dat

full_dat %>% 
  filter(group=='black'|group=='white') %>%
  filter(is.finite(proportion)) %>%
  filter(number>=0) %>%
  filter(total_number>number) %>%
  ungroup() %>%
  mutate(metric='preschool_expulsion') %>%
  mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
  rbind(subdat) -> subdat

```
The primary outcome in this analysis is a count of the number of students by race (black and white) who received one of several types of disciplinary action. Originally, and as specified in our preanalysis plan, our analyses focused on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.  However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes. In particular, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others. Additionally, the number of preschool students who are expelled or suspended is vanishingly smallseveral of these outcomes are not used for discipline, but rather as forms of treatment for students who pose threats to themselves or others (seclusion, physical restraint, mechanical restraint), and the number of preschool expulsions and suspensions are vanishingly small (10 expulsions, )

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

### 2018-01-30 iat_analysis.R
rewrite IAT files with all bias scores the same direction

@@ -150,8 +150,7 @@ df10 %>%
 
 subdat %>%
   mutate(explicit_bias=tblack_0to10) %>%
-  #filter(!is.na(twhite_0to10)) %>%
-  #mutate(explicit_bias = twhite_0to10 - tblack_0to10) %>%
+  mutate(explicit_bias_diff = twhite_0to10 - tblack_0to10) %>%
   mutate(age_bin = cut(age, breaks=c(14, 24, 34, 54, 75, 120))) %>%
   mutate(race = fct_recode(as.character(raceomb), 
                            'Black'='5', 'White'='6')) %>%
@@ -219,16 +218,30 @@ individual_data %>%
   left_join(covs, by='state') %>%
   left_join(covs_ed, by='state') %>%
   left_join(covs_income, by='state') %>%
-  mutate_at(vars(white_prop:poverty), scale) -> individual_data
+  mutate_at(vars(white_prop:poverty), scale) -> individual_data_base
 
 individual.model.bias <- lmer(D_biep.White_Good_all ~ white_prop + black_prop + 
-                                b.w.ratio + col_grads + unemp + income + poverty +
-                                (1|age_bin) + (1|county_id) + (1|state_abb), 
-                         data=individual_data)
+                                b.w.ratio + col_grads + unemp + income + 
+                                poverty + (1|age_bin) + (1|county_id) + 
+                                (1|state_abb), data=individual_data_base)
 individual.model.explicit <- lmer(explicit_bias ~ white_prop + black_prop + 
-                                  b.w.ratio + col_grads + unemp + income + poverty +
-                                  (1|age_bin) + (1|county_id) + (1|state_abb), 
-                                data=individual_data)
+                                    b.w.ratio + col_grads + unemp + income + 
+                                    poverty + (1|age_bin) + (1|county_id) + 
+                                    (1|state_abb), data=individual_data_base)
+individual_data %>%
+  left_join(df_states) %>%
+  left_join(df_acs_counts[,1:4], by=c('county_id', 'age_bin')) %>%
+  left_join(covs, by='state') %>%
+  left_join(covs_ed, by='state') %>%
+  left_join(covs_income, by='state') %>%
+  filter(!is.na(explicit_bias_diff)) %>%
+  mutate_at(vars(white_prop:poverty), scale) -> individual_data_diff
+
[…]

### 2018-01-30 aggregate_data.R
reorganize to write all estimates to same files

@@ -15,7 +15,7 @@ district_content <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CR
 df_district <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_LEA.csv')
 school_content <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH_content.csv')
 df_school <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')
-county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means_explicit_diff.csv', 
+county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv', 
                          colClasses = 'character')
 
 # Get enrollment figures
@@ -307,8 +307,8 @@ schools_loc$county_name[which(schools_loc$county_id=='VA-510')] <- 'Alexandria c
 county_means %>%
   select(-X) %>%
   right_join(schools_loc) %>%
-  select(county_id, county_name, state_abb, 
-         bias, warmth, weighted_bias, weighted_warmth, 
+  select(county_id, county_name, state_abb, bias, explicit, explicit_diff, 
+         weighted_bias, weighted_explicit, weighted_explicit_diff, 
          COMBOKEY, group, number, total_number, metric, LEAID) -> tempout
 
 ################
@@ -348,9 +348,9 @@ error_second <- c('0100002', '0500394', '0500390', '0409734', '0400144',
                   '5000005', '5308700', '5500035', '5600015', '5680251')
 
 tempout %>%
-  filter(COMBOKEY %ni% exclude$COMBOKEY) %>%
-  filter(LEAID %ni% error_elem) %>%
-  filter(LEAID %ni% error_second) -> tempout
+  mutate(exclude = COMBOKEY %in% exclude$COMBOKEY | 
+           LEAID %in% error_elem |
+           LEAID %in% error_second) -> tempout
 
 ################
 # append covariates #
@@ -421,27 +421,11 @@ county_teacher_estimates <- read.csv('/Users/travis/Documents/gits/educational_d
 
 schools_loc %>%
   left_join(county_teacher_estimates) %>%
-  filter(!is.na(county_bias)) %>%
+  filter(!is.na(teacher_bias)) %>%
   left_join(states) %>%
   left_join(covs) %>%
[…]

### 2018-01-30 consensusMCMC_model_prep.R
delete some superfluous code

@@ -1,10 +1,7 @@
 library(dplyr)
 set.seed(42)
 dfs <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/full_model_data.csv')
-dfss <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/selected_model_data_ucla_excl.csv')
-dfss_diff <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/selected_model_data_ucla_excl_exp_diff.csv')
 dft <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/teacher_model_data.csv')
-dft_diff <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/teacher_model_data_expdiff.csv')
 
 ######## write consensus MCMC files
 dfs %>% 
@@ -23,20 +20,13 @@ teacher_counties %>%
 
 dfs %>% 
   left_join(county_assignments) -> county_groups #append the grouping information
-dfss %>%
-  left_join(county_assignments) -> subcounty_groups
-dfss_diff %>%
-  left_join(county_assignments) -> subcounty_groups_diff
 dft %>%
   left_join(teacher_county_assignments) -> teacher_groups
-dft_diff %>%
-  left_join(teacher_county_assignments) -> teacher_groups_diff
 
 write.csv(county_groups, file='/Users/travis/Documents/gits/educational_disparities/cluster/data/county_grouping.csv', row.names = F)
-write.csv(subcounty_groups, file='data/subcounty_grouping_ucla_excl.csv', row.names=F)
-write.csv(subcounty_groups_diff, file='data/subcounty_grouping_diff.csv', row.names=F)
 write.csv(teacher_groups, file='data/county_teacher_grouping.csv', row.names=F)
-write.csv(teacher_groups_diff, file='/Users/travis/Documents/gits/educational_disparities/cluster/data/county_teacher_grouping_expdiff.csv', row.names=F)
+
+#### note that the script writing below is from when the model dataframes were written as separate files
 
 ######################## write raw files #############
 

### 2018-01-30 written_docs/draft.Rmd
first draft of methods

@@
 
-bibliography      : ["r-references.bib"]
+bibliography      : ["ref.bib"]
 
-figsintext        : no
+figsintext        : yes
 figurelist        : no
@@
 class             : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_word
 ---
@@
 library("papaja")
+knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message=FALSE, warning=FALSE, results='hide')
 ```
@@
 library(haven)
+library(tidyr)
+county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv', 
+                         colClasses = 'character')
 schools_dat <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')
@@
 We obtained county-level demographic information for use as covariates and state-level demographic information for use in post-stratification from the US Census Bureau.
-For both county-level and state-level demographics, we use 5-year estimatesfor the period ending in 2014 from the American Community Survey, which surveys around 295k households per month.
+For both county-level and state-level demographics, we use 5-year estimates for the period ending in 2014 from the American Community Survey, which surveys around 295k households per month.
 
@@
   mutate(county_id = paste(STATE, CountyNo, sep='-')) -> individual_data
+
 ```
@@
 From this data, we used only respondents who had geographic information that would allow us to place them in a United States county, identified as White, and visited the site before 2015.
-This consisted of approximately 1.1 million total respondents.
+This consisted of approximately 1.1 million total respondents from `r dim(county_means)[1]`.
+
+```{r teachers-summary}
+
+educators <- c('25-2000', '25-3000')
+
+df %>%
+  filter(CountyNo!='' &
+           !is.na(D_biep.White_Good_all) &
+           !is.na(tblack_0to10) &
+           !is.na(age)) %>%
+  select(CountyNo, STATE, D_biep.White_Good_all, 
+         twhite_0to10, tblack_0to10, raceomb, age, occupation)  -> subdat
+
+df2 %>%
+  filter(CountyNo!='' &
+           !is.na(D_biep.White_Good_all) &
+           !is.na(tblack_0to10) &
+           !is.na(age)) %>%
+  select(CountyNo, STATE, D_biep.White_Good_all, 
+         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
+  rbind(subdat) -> subdat
+
+df3 %>%
+  filter(CountyNo!='' &
+           !is.na(D_biep.White_Good_all) &
+           !is.na(tblack_0to10) &
+           !is.na(age)) %>%
+  select(CountyNo, STATE, D_biep.White_Good_all, 
+         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
+  rbind(subdat) -> subdat
+
+df4 %>%
+  filter(CountyNo!='' &
+           !is.na(D_biep.White_Good_all) &
+           !is.na(tblack_0to10) &
+           !is.na(age)) %>%
+  select(CountyNo, STATE, D_biep.White_Good_all, 
+         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
+  rbind(subdat) -> subdat
+
+df5 %>%
+  filter(CountyNo!='' &
+           !is.na(D_biep.White_Good_all) &
+           !is.na(tblack_0to10) &
+           !is.na(age)) %>%
+  select(CountyNo, STATE, D_biep.White_Good_all, 
+         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
+  rbind(subdat) -> subdat
+
+df6 %>%
+  filter(CountyNo!='' &
+           !is.na(D_biep.White_Good_all) &
+           !is.na(tblack_0to10) &
+           !is.na(age)) %>%
+  select(CountyNo, STATE, D_biep.White_Good_all, 
+         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
+  rbind(subdat) -> subdat
+
+df10 %>%
+  filter(CountyNo!='' &
+           !is.na(D_biep.White_Good_all) &
+           !is.na(tblack_0to10) &
+           !is.na(age)) %>%
+  select(CountyNo, STATE, D_biep.White_Good_all, 
+         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
+  rbind(subdat) -> subdat
+
+subdat %>%
+  filter(raceomb==6) %>%
+  mutate(explicit_diff=twhite_0to10 - tblack_0to10,
+         explicit = tblack_0to10*-1) %>%
+  filter(STATE %ni% 
+           c('AA', 'AE', 'AP', 'AS', 'FM', 'GU', 'MH', 'MP', 'PR', 'VI')) %>%
+  filter(occupation %in% educators) %>%
+  mutate(county_id = paste(STATE, CountyNo, sep='-')) -> individual_data_teachers
+
+teacher_estimates <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_teacher_means.csv')
+```
+
+A subset of years in the Project Implicit data also collected occupational information from respondents.
+As identified in our pre-analysis plan, we took advantage of the presence of primary and secondary educators in these data to test whether any associations between bias and race-based differences in the rates of disciplinary action were stronger among these respondents.
+Filtering for only white individuals who identified as primary, secondary, special education, and other teachers and instructors (occupation codes 25-2000 and 25-3000) reduced the dataset to `r dim(individual_data_teachers)[1]` respondents.
+In order to assure that our estimates were reasonably stable, we limited analysis to only counties that had at least 50 respondents.
+As such, our teacher analysis is limited to just `r dim(teacher_estimates)[1]` counties.
+Additionally, because we do not know of any state-level demographic estimates for teachers, we are unable to perform post-stratification for these data.
 
@@
 ```{r summary_counts}
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
-         SCH_DISCWODIS_ISS_HI_M:TOT_DISCWODIS_ISS_F, 
-         SCH_DISCWDIS_ISS_IDEA_HI_M:TOT_DISCWDIS_ISS_IDEA_F) %>% 
-  gather(group, number, SCH_DISCWODIS_ISS_HI_M:TOT_DISCWDIS_ISS_IDEA_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>% 
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  separate(prefix, into=c('prefix', 'disability'), -3) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='SS_', 
-                          total='EA_', biracial='TR_', white='WH_')) -> tempout_susp
-%>%
-  mutate(disability=fct_recode(disability, not_disabled='_I', disabled='ID',
-                               not_disabled='S_', disabled='A_')) %>%
-  select(-prefix) -> susp_inschool
-
-susp_inschool %>%
-  filter(group=='black'|group=='white') %>%
-  group_by(COMBOKEY, LEA_STATE, LEA_NAME, LEAID,
-           CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
-  summarise(number = sum(number, na.rm=T)) %>%
-  left_join(enrollment) %>%
-  mutate(proportion = number/total_number) -> tempout
-
-tempout$number[which(tempout$number<0)] <- NA
-
-tempout %>% 
-  ungroup() %>%
-  filter(is.finite(proportion)) %>%
-  filter(total_number>number) %>%
-  mutate(metric='inschool_susp') %>%
-  mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
-  distinct() -> subdat
 
@@
 schools_dat %>%
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
-         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:SCH_PSDISC_SINGOOS_IDEA_F) %>% 
-  filter(SCH_GRADE_PS=='YES') %>%
-  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:SCH_PSDISC_SINGOOS_IDEA_F) %>%
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
+  gather(group, total_number, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
   separate(group, into=c('group', 'gender'), -2) %>%
   separate(group, into=c('prefix', 'group'), -4) %>%
-  separate(prefix, into=c('prefix', 'disability'), -3) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='NR_', 
+                          biracial='TR_', white='WH_')) %>%
+  filter(total_number>=0) %>%
+  group_by(COMBOKEY, group) %>%
+  mutate(total_number = sum(total_number)) %>%
+  ungroup() %>%
+  select(-prefix, -gender) %>%
+  distinct() -> ps_enrollment
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
+         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>% 
+  filter(SCH_GRADE_PS=='YES') %>%
+  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>% 
+  separate(group, into=c('prefix', 'group'), -4) %>%
   mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                           hispanic='HI_', pac_isl='HP_', total='OS_', 
-                          biracial='TR_', white='WH_')) -> tempout
-
-%>%
-  mutate(disability=fct_recode(disability, not_disabled='_I', disabled='ID',
-                             not_disabled='GO', disabled='A_')) %>%
+                          biracial='TR_', white='WH_')) %>%
   select(-prefix) -> ps_susp
 
-df_school %>% 
+schools_dat %>% 
   select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
@@
 
-ps_susp %>%
-  left_join(enrollment) %>%
-  group_by(COMBOKEY, LEA_STATE, LEA_NAME, 
-           CCD_LATCOD, CCD_LONCOD,  SCH_NAME, group) %>%
-  summarise(number = sum(number),
-            total_number = sum(total_number)) %>%
-  mutate(proportion = number/total_number) -> full_dat
-
-full_dat %>% 
-  filter(group=='black'|group=='white') %>%
-  filter(is.finite(proportion)) %>%
-  filter(number>=0) %>%
-  filter(total_number>number) %>%
-  mutate(group=droplevels(group)) %>%
-  ungroup() %>%
-  mutate(metric='preschool_susp') %>%
-  mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
-  rbind(subdat) -> subdat
-
-df_school %>% 
+schools_dat %>% 
   select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
@@
                           biracial='TR_', white='WH_')) %>%
+  filter(number>=0) %>%
   select(-prefix) -> ps_exp
 
-ps_exp %>%
-  left_join(enrollment) %>%
-  group_by(COMBOKEY, LEA_STATE, LEA_NAME, 
-           CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
-  summarise(number = sum(number),
-            total_number = sum(total_number)) %>%
-  mutate(proportion = number/total_number) -> full_dat
-
-full_dat %>% 
-  filter(group=='black'|group=='white') %>%
-  filter(is.finite(proportion)) %>%
-  filter(number>=0) %>%
-  filter(total_number>number) %>%
-  ungroup() %>%
-  mutate(metric='preschool_expulsion') %>%
-  mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
-  rbind(subdat) -> subdat
+ps_susp %>% group_by(group) %>% summarise(n_students=sum(number)) -> susp_summ
+ps_exp %>% group_by(group) %>% summarise(n_students=sum(number)) -> exp_summ
+ps_enrollment %>% group_by(group) %>% summarise(n_students=sum(total_number)) -> enr_summ
 
@@
 In particular, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others.
-Additionally, the number of preschool students who are expelled or suspended is vanishingly smallseveral of these outcomes are not used for discipline, but rather as forms of treatment for students who pose threats to themselves or others (seclusion, physical restraint, mechanical restraint), and the number of preschool expulsions and suspensions are vanishingly small (10 expulsions, )
+Additionally, the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]`, total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely.
+We additionally discovered that counts of one expulsion category (expulsion under zero-tolerance policies) overlapped with counts in other categories, and so excluded this category.
+Finally, for expository reasons, we opted to combine the remaining two expulsion categories to yield one overall count of the number of students expelled.
+**Results with the uncombined counts largely mirror the results here.**
 
-## Material
+For both the post-stratification procedure and the actual statistical models used for inference, we used the same set of covariates.
+These covariates were at the state-level for post-stratification, and were at the county-level for the statistical models used for final estimation and inference.
+Specifically, our population based covariates were the total population count, the proportion of the population that is black, the proportion of the population that is white and the ratio of black-to-white persons.
+We also used socioeconomic covariates.
+We used the percent of individuals aged 16 or over who were in the labor force, but unemployed, the median household income, and the percentage of all families whose income is below the poverty line.
 
-## Procedure
+Finally, for the implicit and explicit bias measures, we relied on two primary variables from the Project Implicit data.
+Implicit bias was assessed via an Implicit Association Test (IAT).
+This test uses a speeded dual-categorization task in which individuals must quickly categorize black and white faces and "good" and "bad" words with key presses.
+The difference in how quickly and accurately participants are able to pair white faces with "good" words and black faces with "bad words" in comparison to the inverse is thought to reflect implicit associations between the two races and positive and negative affective reactions.
+This association is reflected in the IAT D-score, which we used as a measure of implicit bias.
+We preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e.
+*how warm or cold do you feel towards Blacks?
+0=very cold, 10=very warm).
+However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so we also report models using this metric of explicit bias.
 
 ## Data analysis
-We used `r cite_r("r-references.bib")` for all our analyses.
+Data analysis proceded in two steps.
+We first estimated county-level implict and bias.
+These estimates took two forms - either simple county-based averages, or post-stratified estimates.
+Post-stratification is a valuable procedure in obtaining accurate geographical population-based estimates because it allows a non-representative sample (e.g Project Implicit) to more closely resemble the true population, and it regularizes extreme observations with little data to support them (e.g.
+a county with only a handful of respondents with especially high or low scores) [@park2004bayesian, @gelman1997poststratification].
+Following past work [@leitner2016blacks], we identified age as one dimentions along which IAT respondents strongly differed from the general population.
+Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories in each county (15-24, 25-34, 35-54, 55-75, and 75+).
+We next fit multilevel models estimating bias (implicit and explicit biases seperately) as a function of our state-level covariates (the "fixed" effects), and allowed the estimates to vary by age bin, county, and state (the "random" effects).
+Next, we determined the population of whites in each county in these age groups using the American Community Survey's 5-year estimates ending in 2014.
+Finally, we used our estimated models to predict the expected response for each age bin, in each county.
+Our final county-level estimates is the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county.
+As a result of this procedure, we can be confident that our county-level estimates should more closely approximate what our estimates would look like if the Project Implicit data were truly representative along the age dimension in all counties.
+
+After obtaining these estimates, we use them as predictors in bayesian multilevel logistic regressions. Formally, the likelihood for a given observation is written as a binomial functions:
+
+$$
+{n\choose y}\pi^y(1-\pi)^{n-y}
+$$
+
+Where $y$ is the observed count of incidents (e.g.
+number of black or white students suspended), $n$ is the number of at-risk students (e.g.
+total number of black or white students), and $\pi = g^{-1}(\eta)$ is the probability of the incident ocurring.
+For this analysis, the linear predictor takes the form of a multilevel model with a set of effects that vary over county:
+
+$$
+\eta = \alpha + X\beta + \gamma_{county}
+$$
+
+Where $\alpha$ is an intercept that is constant across observations, $\beta$ represents a set of effects that are also constant across observations (i.e.
+fixed effects), and $\gamma_{county}$ represents intercepts and effects of ethnicity that vary across the counties (i.e.
+random efffects).
+
+In addition to the covariates described above, we also include effects of race, implicit bias, explicit bias, and the two-way interactions between implicit bias and race and explicit bias and race.
+We fit separate models for each of the outcomes, each of the two types of explicit biases (raw temperature toward blacks, and the difference in reported warmth between whites and blacks), and for teachers and general public.
+We also estimate models with raw and poststratified estimates.
+
+Because of the computational demands of fitting such a high-dimensional model to such a large dataset (the full model would consist of over 6k parameters to approximately 170k observations), we used a consensus monte carlo algorithm to obtain approximate posterior distributions for the parameters of interest [@scott2016bayes].
+The approximate posteriors derived from this algorithm have been shown to be nearly indistinguishable from the true posterior, a result we verified using a small subset of our own data.
+
+All numerical predictor variables were standardized at the appropriate level (county, state) before model estimation to help with estimation efficiency and interpretability.
+We set priors for the intercept and coefficients in the bayesian models to be weakly informative normal distributions centered on zero with a standard deviation of five.
+All other parameters were left to default values.
+Data analysis was done in R [@rcitation] version 3.3.2 running under OS X 10.11.6.
+Post-stratification was done with lme4, version 1.1.14 [@bates2015fitting].
+Final model fitting was done on the university cluster running Springdale Linux, release 6.9 using rstanarm, version 2.17.2 [@rstanarm2016].
+We used the implementation of the consensus monte carlo algorithm found in parallelMCMCcombine, version 1.0 [@miroshnikov2014parallel].
+Figures were made with ggplot2, version 2.2.1 [@wickham2009ggplot2], with data manipulation done using dplyr version 0.7.2 [@wickham2017dplyr] and tidyr, version 0.7.1 [@wickham2017tidyr].
+Full session info can be found on the OSF page (....)
 

### 2018-02-04 aggregate_data.R
start aggregating original metrics

@@ -36,6 +36,7 @@ df_school %>%
 # relevant metrics #
 ################ 
 
+# in-school suspension
 df_school %>% 
   select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
          SCH_DISCWODIS_ISS_HI_M:TOT_DISCWODIS_ISS_F, 
@@ -69,6 +70,7 @@ tempout %>%
   mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
   distinct() -> subdat
 
+#out of school suspension
 df_school %>% 
   select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
          SCH_DISCWODIS_SINGOOS_HI_M:TOT_DISCWODIS_SINGOOS_F,
@@ -122,6 +124,7 @@ tempout %>%
   distinct() %>%
   rbind(subdat) -> subdat
 
+#expulsion with education
 df_school %>% 
   select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
          SCH_DISCWODIS_EXPWE_HI_M:TOT_DISCWODIS_EXPWE_F,
@@ -137,6 +140,7 @@ df_school %>%
                                not_disabled='E_', disabled='A_')) %>%
   select(-prefix) -> exp_w_ed
 
+#expulsion without education
 df_school %>% 
   select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
          SCH_DISCWODIS_EXPWOE_HI_M:TOT_DISCWODIS_EXPWOE_F,
@@ -176,6 +180,7 @@ tempout %>%
   distinct() %>%
   rbind(subdat) -> subdat
 
+#law enforcement
 df_school %>% 
   select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
          SCH_DISCWODIS_REF_HI_M:TOT_DISCWODIS_REF_F,
@@ -210,6 +215,7 @@ tempout %>%
   distinct() %>%
   rbind(subdat) -> subdat
 
+#inschool arrest
 df_school %>% 
   select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
          SCH_DISCWODIS_ARR_HI_M:TOT_DISCWODIS_ARR_F,
@@ -429,3 +435,348 @@ schools_loc %>%
            LEAID %in% error_second) -> out
 
[…]

### 2018-02-04 consensusMCMC_model_prep.R
write files for nested model

@@ -13,7 +13,7 @@ dft %>%
   distinct() -> teacher_counties
 
 counties %>% 
-  mutate(grouping=sample(1:12, n(), replace=T)) -> county_assignments
+  mutate(grouping=sample(1:14, n(), replace=T)) -> county_assignments
 
 teacher_counties %>%
   mutate(grouping=sample(1:12, n(), replace=T)) -> teacher_county_assignments
@@ -49,9 +49,7 @@ string3 <- ") %>% filter(metric=='"
 string4 <- "') %>% select(county_id, COMBOKEY:metric) %>% left_join(scaled_county_level) -> mod.dat
 
 m <- stan_glmer(cbind(number, total_number-number) ~ group + bias + warmth + 
-group:bias + group:warmth + total_pop + unemp_rate + 
-med_income + poverty_rate + col_grads + white_prop + 
-black_prop + b.w.ratio + 
+group:bias + group:warmth + + 
 (group|county_id), data=mod.dat, prior = normal(0,5), 
 prior_intercept = normal(0,5), family=binomial, adapt_delta=.99)
 
@@ -266,7 +264,6 @@ for(i in rownames(table(dft$metric))){
   }
 }
 
-
 ######################## write weighted files w/3 way interaction #############
 
 #write files
@@ -307,3 +304,44 @@ for(i in rownames(table(dfs$metric))){
   }
 }
 
+
+######################## write weighted files for UCLA exclusions, explicit diff, nested intercept #############
+
+#write files
+string1 <- "library(dplyr)
+library(tidyr)
+library(forcats)
+library(ggplot2)
+library(rstanarm)
+
+options(mc.cores = parallel::detectCores())
+
+df <- read.csv('/home/triddle/educational_disparities/cluster/data/county_grouping.csv')
+
+df %>% filter(metric=='"
+
+string2 <- "') %>% filter(exclude==FALSE) %>% select(county_id, bias:weighted_explicit_diff, total_pop, unemp_rate:b.w.ratio) %>% distinct() %>% mutate_at(vars(-county_id), scale) -> scaled_county_level
+
+df %>% filter(grouping=="
+
+string3 <- ") %>% filter(metric=='"
+string4 <- "') %>% filter(exclude==FALSE) %>% select(county_id, COMBOKEY:metric) %>% left_join(scaled_county_level) -> mod.dat
+
[…]

### 2018-02-04 iat_analysis.R
write one output

@@ -361,4 +361,4 @@ individual_data %>%
   distinct() -> county_teacher_estimates
 
 write.csv(county_teacher_estimates, row.names = F,
-            file = '/Users/travis/Documents/gits/educational_disparities/output/county_teacher_means_expdiff.csv')
+            file = '/Users/travis/Documents/gits/educational_disparities/output/county_teacher_means.csv')

### 2018-02-04 written_docs/draft.Rmd
start with writing

@@
 ---
-title             : "Race-based disparities in allocation of discipline are associated with county-level rates of bias"
+title             : "Race-based disparities in allocation of academic disciplinary actions are associated with county-level rates of bias"
 shorttitle        : "Discipline Disparities"
@@
   - id            : "1"
-    institution   : "Princeton University"
-  - id            : "2"
     institution   : "Princeton University"
@@
 library("papaja")
-knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message=FALSE, warning=FALSE, results='hide')
+knitr::opts_chunk$set(echo = FALSE, cache=TRUE, cache.lazy = FALSE, 
+                      message=FALSE, warning=FALSE, results='hide')
 ```
@@
 schools_dat <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')
+model.data <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/full_model_data.csv')
 
-schools_dat %>%
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_ENR_HI_M:TOT_ENR_F) %>%
-  gather(group, total_number, SCH_ENR_HI_M:TOT_ENR_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>%
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='NR_', 
-                          biracial='TR_', white='WH_')) %>%
-  group_by(COMBOKEY, group) %>%
-  mutate(total_number = sum(total_number)) %>%
-  ungroup() %>%
-  select(-prefix, -gender) %>%
-  distinct() -> enrollment
+model.data %>%
+  filter(exclude==FALSE) %>%
+  select(COMBOKEY, group, total_number) %>%
+  distinct() %>%
+  mutate(total_schools = n_distinct(COMBOKEY)) %>%
+  group_by(group) %>%
+  summarise(ns = sum(total_number),
+            n_sch = n_distinct(COMBOKEY),
+            total_schools = max(total_schools)) -> enrollment_totals
 
-enrollment %>% group_by(group) %>% summarise(sum(total_number)) -> enrollment_totals
 ```
@@
 The academic disciplinary data is part of the Civil Rights Data Collection (CRDC) through the US Department of Education.
-The dataset we used comes from the 2013-2014 academic year and has data on "all public local and educational agencies and schools, including long-term secure juvenile justice facilities, charter schools, alternative schools, and schools serving students with disabilities." In total, the CRDC data represents `r dim(schools_dat)[1]` institutions enrolling approximately 50 million students, of which approximately 25 million are white and 7.8 million are Black.
+The dataset we used comes from the 2013-2014 academic year and has data on "all public local and educational agencies and schools, including long-term secure juvenile justice facilities, charter schools, alternative schools, and schools serving students with disabilities." In total, the CRDC data represents `r dim(schools_dat)[1]` institutions enrolling approximately 50 million students, of which approximately 25 million are white and 7.8 million are Black[^1].
+Previous work using these data have identified a number of districts whose data are in error, and have excluded juvenile justice facilities, as these institutions constitute dramatically different educational environments, where the meaning of disciplinary actions may be quite different [@losen2015we].
+After these exclusions are applied, the final sample used for modeling consists of `r enrollment_totals$total_schools[1]` institutions, enrolling `r round(sum(enrollment_totals$ns)/1000000, 1)` million black or white students, of which `r  round(enrollment_totals$ns[2]/1000000, 1)` million are white and `r  round(enrollment_totals$ns[1]/1000000, 1)` million are black.
+
+[^1]: We note that there are a number of differences between the analyses we registered and those presented in the main text.
+Our general conclusions are largely the same for both sets of analyses.
+We opted to report the modified analyses for reasons of clarity and to remain congruent with previous research on the same topics.
+The registered analyses can be found, in full, in the appendix.
 
@@
 From this data, we used only respondents who had geographic information that would allow us to place them in a United States county, identified as White, and visited the site before 2015.
-This consisted of approximately 1.1 million total respondents from `r dim(county_means)[1]`.
+This consisted of approximately 1.1 million total respondents from `r dim(county_means)[1]` counties.
 
@@
 
-```
-The primary outcome in this analysis is a count of the number of students by race (black and white) who received one of several types of disciplinary action.
-Originally, and as specified in our preanalysis plan, our analyses focused on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.
+#Originally, and as specified in our preanalysis plan, our analyses focused on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.
 However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes.
 In particular, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others.
-Additionally, the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]`, total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely.
+Additionally, the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]` total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely.
 We additionally discovered that counts of one expulsion category (expulsion under zero-tolerance policies) overlapped with counts in other categories, and so excluded this category.
@@
 
-For both the post-stratification procedure and the actual statistical models used for inference, we used the same set of covariates.
-These covariates were at the state-level for post-stratification, and were at the county-level for the statistical models used for final estimation and inference.
+ #We preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e.
+*how warm or cold do you feel towards Blacks?
+0=very cold, 10=very warm*).
+However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so we also report models using this metric of explicit bias.
+
+#s, each of the two types of explicit biases (raw temperature toward blacks, and the difference in reported warmth between whites and blacks), and for teachers and general public.
+We also estimate models with raw and poststratified estimates.
+
+```
+The primary outcome in this analysis is a count of the number of students by race (black and white) who received one of several types of disciplinary action.
+We report here rates of out-of-school suspension, in-school suspension, school-related arrests, law enforcement referrals, and total number of expulsions of any type.
+
+For both the post-stratification procedure (described below) and the actual statistical models used for inference, we used the same set of covariates.
+These covariates were at the state-level for post-stratification and were at the county-level for the statistical models used for final estimation and inference.
 Specifically, our population based covariates were the total population count, the proportion of the population that is black, the proportion of the population that is white and the ratio of black-to-white persons.
 We also used socioeconomic covariates.
-We used the percent of individuals aged 16 or over who were in the labor force, but unemployed, the median household income, and the percentage of all families whose income is below the poverty line.
+We used the percent of individuals aged 16 or over who were in the labor force but unemployed, the median household income, and the percentage of all families whose income is below the poverty line.
 
@@
 This test uses a speeded dual-categorization task in which individuals must quickly categorize black and white faces and "good" and "bad" words with key presses.
-The difference in how quickly and accurately participants are able to pair white faces with "good" words and black faces with "bad words" in comparison to the inverse is thought to reflect implicit associations between the two races and positive and negative affective reactions.
+The difference in how quickly and accurately participants are able to pair white faces with "good" words and black faces with "bad" words in comparison to the inverse is thought to reflect implicit associations between the two races and positive and negative affective reactions.
 This association is reflected in the IAT D-score, which we used as a measure of implicit bias.
-We preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e.
-*how warm or cold do you feel towards Blacks?
-0=very cold, 10=very warm).
-However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so we also report models using this metric of explicit bias.
+Our measure of explicit bias is the difference between reported warmth towards whites (i.e.
+*how warm or cold do you feel towards Whites?
+0=very cold, 10=very warm*) and reported warmth toward blacks.
 
@@
 Data analysis proceded in two steps.
-We first estimated county-level implict and bias.
-These estimates took two forms - either simple county-based averages, or post-stratified estimates.
+We first estimated county-level implict and explicit bias using multilevel regression and post-stratification.
 Post-stratification is a valuable procedure in obtaining accurate geographical population-based estimates because it allows a non-representative sample (e.g Project Implicit) to more closely resemble the true population, and it regularizes extreme observations with little data to support them (e.g.
-a county with only a handful of respondents with especially high or low scores) [@park2004bayesian, @gelman1997poststratification].
-Following past work [@leitner2016blacks], we identified age as one dimentions along which IAT respondents strongly differed from the general population.
-Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories in each county (15-24, 25-34, 35-54, 55-75, and 75+).
+a county with only a handful of respondents with especially high or low scores) [@park2004bayesian; @gelman1997poststratification].
+Following past work [@leitner2016blacks], we identified age as one dimension along which IAT respondents differed from the general population in ways that could bias our conclusions [@gonsalkorale2009aging].
+Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories (15-24, 25-34, 35-54, 55-75, and 75+).
 We next fit multilevel models estimating bias (implicit and explicit biases seperately) as a function of our state-level covariates (the "fixed" effects), and allowed the estimates to vary by age bin, county, and state (the "random" effects).
@@
 Finally, we used our estimated models to predict the expected response for each age bin, in each county.
-Our final county-level estimates is the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county.
+Our final county-level estimates are the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county.
 As a result of this procedure, we can be confident that our county-level estimates should more closely approximate what our estimates would look like if the Project Implicit data were truly representative along the age dimension in all counties.
 
-After obtaining these estimates, we use them as predictors in bayesian multilevel logistic regressions. Formally, the likelihood for a given observation is written as a binomial functions:
+After obtaining these estimates, we use them as predictors in bayesian multilevel logistic regressions. Formally, the likelihood for a given observation is written as a binomial function:
 
@@
 In addition to the covariates described above, we also include effects of race, implicit bias, explicit bias, and the two-way interactions between implicit bias and race and explicit bias and race.
-We fit separate models for each of the outcomes, each of the two types of explicit biases (raw temperature toward blacks, and the difference in reported warmth between whites and blacks), and for teachers and general public.
-We also estimate models with raw and poststratified estimates.
+We fit separate models for each of the outcomes.
 
-Because of the computational demands of fitting such a high-dimensional model to such a large dataset (the full model would consist of over 6k parameters to approximately 170k observations), we used a consensus monte carlo algorithm to obtain approximate posterior distributions for the parameters of interest [@scott2016bayes].
+Because of the computational demands of fitting such a high-dimensional model to such a large dataset (the full model for each metric would consist of over 6k parameters to approximately 170k observations), we used a consensus monte carlo algorithm to obtain approximate posterior distributions for the parameters of interest [@scott2016bayes].
 The approximate posteriors derived from this algorithm have been shown to be nearly indistinguishable from the true posterior, a result we verified using a small subset of our own data.
@@
 Figures were made with ggplot2, version 2.2.1 [@wickham2009ggplot2], with data manipulation done using dplyr version 0.7.2 [@wickham2017dplyr] and tidyr, version 0.7.1 [@wickham2017tidyr].
-Full session info can be found on the OSF page (....)
+A full report of session information can be found on the OSF page (....)
 

### 2018-02-04 written_docs/draft.Rmd
more writing

@@
 ## Measures
-
-```{r summary_counts}
-
-###############
-schools_dat %>%
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
-  gather(group, total_number, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>%
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='NR_', 
-                          biracial='TR_', white='WH_')) %>%
-  filter(total_number>=0) %>%
-  group_by(COMBOKEY, group) %>%
-  mutate(total_number = sum(total_number)) %>%
-  ungroup() %>%
-  select(-prefix, -gender) %>%
-  distinct() -> ps_enrollment
-
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
-         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>% 
-  filter(SCH_GRADE_PS=='YES') %>%
-  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>% 
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='OS_', 
-                          biracial='TR_', white='WH_')) %>%
-  select(-prefix) -> ps_susp
-
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
-         SCH_GRADE_PS, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>% 
-  filter(SCH_GRADE_PS=='YES') %>%
-  gather(group, number_2, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>%
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='OS_', 
-                          biracial='TR_', white='WH_')) %>%
-  select(COMBOKEY, group, gender, number_2) %>%
-  right_join(ps_susp) %>%
-  filter(number>=0 & number_2 >= 0) %>%
-  mutate(number = number+number_2) -> ps_susp
-
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
-         SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>% 
-  gather(group, number, SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>% 
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='XP_', 
-                          biracial='TR_', white='WH_')) %>%
-  filter(number>=0) %>%
-  select(-prefix) -> ps_exp
-
-ps_susp %>% group_by(group) %>% summarise(n_students=sum(number)) -> susp_summ
-ps_exp %>% group_by(group) %>% summarise(n_students=sum(number)) -> exp_summ
-ps_enrollment %>% group_by(group) %>% summarise(n_students=sum(total_number)) -> enr_summ
-
-#Originally, and as specified in our preanalysis plan, our analyses focused on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.
-However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes.
-In particular, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others.
-Additionally, the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]` total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely.
-We additionally discovered that counts of one expulsion category (expulsion under zero-tolerance policies) overlapped with counts in other categories, and so excluded this category.
-Finally, for expository reasons, we opted to combine the remaining two expulsion categories to yield one overall count of the number of students expelled.
-**Results with the uncombined counts largely mirror the results here.**
-
- #We preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e.
-*how warm or cold do you feel towards Blacks?
-0=very cold, 10=very warm*).
-However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so we also report models using this metric of explicit bias.
-
-#s, each of the two types of explicit biases (raw temperature toward blacks, and the difference in reported warmth between whites and blacks), and for teachers and general public.
-We also estimate models with raw and poststratified estimates.
-
-```
 The primary outcome in this analysis is a count of the number of students by race (black and white) who received one of several types of disciplinary action.
@@
 
+\newpage
+# Appendix
+## Preregistered analysis
+```{r summary_counts}
+
+###############
+schools_dat %>%
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
+  gather(group, total_number, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>%
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='NR_', 
+                          biracial='TR_', white='WH_')) %>%
+  filter(total_number>=0) %>%
+  group_by(COMBOKEY, group) %>%
+  mutate(total_number = sum(total_number)) %>%
+  ungroup() %>%
+  select(-prefix, -gender) %>%
+  distinct() -> ps_enrollment
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
+         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>% 
+  filter(SCH_GRADE_PS=='YES') %>%
+  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>% 
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='OS_', 
+                          biracial='TR_', white='WH_')) %>%
+  select(-prefix) -> ps_susp
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
+         SCH_GRADE_PS, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>% 
+  filter(SCH_GRADE_PS=='YES') %>%
+  gather(group, number_2, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>%
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='OS_', 
+                          biracial='TR_', white='WH_')) %>%
+  select(COMBOKEY, group, gender, number_2) %>%
+  right_join(ps_susp) %>%
+  filter(number>=0 & number_2 >= 0) %>%
+  mutate(number = number+number_2) -> ps_susp
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
+         SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>% 
+  gather(group, number, SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>% 
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='XP_', 
+                          biracial='TR_', white='WH_')) %>%
+  filter(number>=0) %>%
+  select(-prefix) -> ps_exp
+
+ps_susp %>% group_by(group) %>% summarise(n_students=sum(number)) -> susp_summ
+ps_exp %>% group_by(group) %>% summarise(n_students=sum(number)) -> exp_summ
+ps_enrollment %>% group_by(group) %>% summarise(n_students=sum(total_number)) -> enr_summ
+
+#s,
+
+```
+In our preanalysis plan, we specified our analyses to focus on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.
+However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes.
+In particular, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others.
+Additionally, the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]` total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely.
+We additionally discovered that counts of one expulsion category (expulsion under zero-tolerance policies) overlapped with counts in other categories, and so excluded this category.
+To remain consistent with previous studies, we opted to combine the remaining two expulsion categories to yield one overall count of the number of students expelled.
+
+We also preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e.
+*how warm or cold do you feel towards Blacks?
+0=very cold, 10=very warm*).
+However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so we in the main text, we report models using this metric of explicit bias.
+Additionally, we preregistered analyses with poststratified estimates (as presented in the main text) along with raw, county-based means.
+Finally, we had not known about the issues with juvenile justice facilities, or with the school districts with reporting errors.
+Here, we present the results of the preregistered analyses exactly.
+
+#Simple county means
+
+#Post stratified estimates
+
+#Teacher analyses
+
 \setlength{\parindent}{-0.5in}

### 2018-02-05 written_docs/draft.Rmd
cross val with disparities

@@
 0=very cold, 10=very warm*).
-However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so we in the main text, we report models using this metric of explicit bias.
+However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so in the main text, we report models using this metric of explicit bias.
 Additionally, we preregistered analyses with poststratified estimates (as presented in the main text) along with raw, county-based means.
@@
 #Simple county means
+```{r cross_val_county_means}
+"%ni%" <- Negate("%in%")
+df <- read.csv('/Users/travis/Documents/gits/educational_disparities/cluster/data/county_grouping.csv')
+f <- list.files('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_raw/corporal/')
+
+test_summary_dat <- data.frame(file=f, 
+                               r_2 = NA,
+                               r_2_lower=NA,
+                               r_2_upper=NA)
+j=1
+for (i in f){
+  print(i)
+  load(paste('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_raw/inschool_susp/', i, sep=''))
+  model <- m
+  testing_model <- sample(f[f %ni% i], 1)
+  load(paste('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_raw/inschool_susp/', testing_model, sep=''))
+  m$data$bias <- m$data$bias[,1]
+  m$data$warmth <- m$data$warmth[,1]
+  m$data$weighted_bias <- m$data$weighted_bias[,1]
+  m$data$weighted_warmth <- m$data$weighted_warmth[,1]
+  m$data$total_pop <- m$data$total_pop[,1]
+  m$data$unemp_rate <- m$data$unemp_rate[,1]
+  m$data$med_income <- m$data$med_income[,1]
+  m$data$poverty_rate <- m$data$poverty_rate[,1]
+  m$data$col_grads <- m$data$col_grads[,1]
+  m$data$white_prop <- m$data$white_prop[,1]
+  m$data$black_prop <- m$data$black_prop[,1]
+  m$data$b.w.ratio <- m$data$b.w.ratio[,1]
+  m$data %>%
+    filter(!is.na(bias)) %>%
+    filter(!is.na(warmth)) %>%
+    filter(!is.na(weighted_bias)) %>%
+    filter(!is.na(weighted_warmth)) %>%
+    filter(!is.na(total_pop)) -> testing_data
+  y_hat <- data.frame(t(posterior_predict(m, draws = 250, newdata = testing_data)))
+  y_hat$index <- seq(1, length(y_hat$X1))
+  testing_data <- cbind(testing_data, y_hat)
+  testing_data %>%
+    gather(sample, pred, X1:X250) %>%
+    group_by(sample) %>%
+    summarise(r = cor(number, pred)^2) %>%
+    summarise(r_2 = mean(r),
+              r_2_lower = quantile(r, .25),
+              r_2_upper = quantile(r, .75)) -> r2
+  test_summary_dat[j,2:4] <- r2
+  j = j+1
+}
+
+test_summary_datweighted <- data.frame(file=f, 
+                               r_2 = NA,
+                               r_2_lower=NA,
+                               r_2_upper=NA)
+j=1
+for (i in f){
+  print(i)
+  load(paste('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_weighted/older_run/inschool_susp/', i, sep=''))
+  model <- m
+  testing_model <- sample(f[f %ni% i], 1)
+  load(paste('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_weighted/older_run/inschool_susp/', testing_model, sep=''))
+  m$data$bias <- m$data$bias[,1]
+  m$data$warmth <- m$data$warmth[,1]
+  m$data$weighted_bias <- m$data$weighted_bias[,1]
+  m$data$weighted_warmth <- m$data$weighted_warmth[,1]
+  m$data$total_pop <- m$data$total_pop[,1]
+  m$data$unemp_rate <- m$data$unemp_rate[,1]
+  m$data$med_income <- m$data$med_income[,1]
+  m$data$poverty_rate <- m$data$poverty_rate[,1]
+  m$data$col_grads <- m$data$col_grads[,1]
+  m$data$white_prop <- m$data$white_prop[,1]
+  m$data$black_prop <- m$data$black_prop[,1]
+  m$data$b.w.ratio <- m$data$b.w.ratio[,1]
+  m$data %>%
+    filter(!is.na(bias)) %>%
+    filter(!is.na(warmth)) %>%
+    filter(!is.na(weighted_bias)) %>%
+    filter(!is.na(weighted_warmth)) %>%
+    filter(!is.na(total_pop)) -> testing_data
+  y_hat <- data.frame(t(posterior_predict(m, draws = 250, newdata = testing_data)))
+  y_hat$index <- seq(1, length(y_hat$X1))
+  testing_data <- cbind(testing_data, y_hat)
+  testing_data %>%
+    gather(sample, pred, X1:X250) %>%
+    group_by(sample) %>%
+    summarise(r = cor(number, pred)^2) %>%
+    summarise(r_2 = mean(r),
+              r_2_lower = quantile(r, .25),
+              r_2_upper = quantile(r, .75)) -> r2
+  test_summary_datweighted[j,2:4] <- r2
+  j = j+1
+}
+
+```
 

### 2018-02-26 written_docs/draft.Rmd
There's some content here.

--- 2018-02-04 written_docs/draft.Rmd
@@
 ---
-title             : "Race-based disparities in allocation of academic disciplinary actions are associated with county-level rates of bias"
-shorttitle        : "Discipline Disparities"
-
+title: Race-based disparities in allocation of academic disciplinary actions are associated
+  with county-level rates of bias
 author:
-  - name          : "Travis Riddle"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : ""
-    email         : "triddle@princeton.edu"
-  - name          : "Stacey Sinclair"
-    affiliation   : "1"
-
+- address: ''
+  affiliation: '1'
+  corresponding: yes
+  email: triddle@princeton.edu
+  name: Travis Riddle
+- affiliation: '1'
+  name: Stacey Sinclair
 affiliation:
-  - id            : "1"
-    institution   : "Princeton University"
-
-author_note: >
-
+- id: '1'
+  institution: Princeton University
+output:
+  papaja::apa6_word: default
+  papaja::apa6_pdf: default
+bibliography: ref.bib
+class: man
+figsintext: yes
+figurelist: no
+footnotelist: no
+keywords: keywords
+lang: english
+lineno: yes
+author_note: |
   Enter author note here.
-
-abstract: >
+shorttitle: Discipline Disparities
+tablelist: no
+abstract: |
   Enter abstract here.
-
-keywords          : "keywords"
-wordcount         : "X"
-
-bibliography      : ["ref.bib"]
-
-figsintext        : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-lineno            : yes
-
-lang              : "english"
-class             : "man"
-output            : papaja::apa6_word
+wordcount: X
 ---
@@
 library(tidyr)
+library(ggplot2)
 county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv', 
@@
 Previous work using these data have identified a number of districts whose data are in error, and have excluded juvenile justice facilities, as these institutions constitute dramatically different educational environments, where the meaning of disciplinary actions may be quite different [@losen2015we].
-After these exclusions are applied, the final sample used for modeling consists of `r enrollment_totals$total_schools[1]` institutions, enrolling `r round(sum(enrollment_totals$ns)/1000000, 1)` million black or white students, of which `r  round(enrollment_totals$ns[2]/1000000, 1)` million are white and `r  round(enrollment_totals$ns[1]/1000000, 1)` million are black.
+After these exclusions are applied, the final sample used for modeling consists of `r enrollment_totals$total_schools[1]` institutions, enrolling `r round(sum(enrollment_totals$ns)/1000000, 1)` million black or white students, of which `r  round(enrollment_totals$ns[2]/1000000)` million are white and `r  round(enrollment_totals$ns[1]/1000000)` million are black.
 
@@
            !is.na(age)) %>%
-  select(CountyNo, STATE, D_biep.White_Good_all, 
-         tblack_0to10, twhite_0to10, raceomb, age)  -> subdat
+  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
+         tblack_0to10, twhite_0to10, raceomb, age) %>%
+  filter(raceomb==5 | raceomb==6) -> subdat
 
@@
            !is.na(age)) %>%
-  select(CountyNo, STATE, D_biep.White_Good_all,  
+  dplyr::select(CountyNo, STATE, D_biep.White_Good_all,  
          tblack_0to10, twhite_0to10, raceomb, age) %>%
@@
            !is.na(age)) %>%
-  select(CountyNo, STATE, D_biep.White_Good_all, 
+  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
          tblack_0to10, twhite_0to10, raceomb, age) %>%
@@
   mutate(raceomb = ethnic) %>%
-  select(CountyNo, STATE, D_biep.White_Good_all, 
+  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
          tblack_0to10, twhite_0to10, raceomb, age) %>%
@@
          twhite_0to10=twhites_0to10) %>%
-  select(CountyNo, STATE, D_biep.White_Good_all, 
+  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
          tblack_0to10, twhite_0to10, raceomb, age) %>%
@@
 
+## Measures
+The primary outcome in this analysis is a count of the number of students by race (black and white) who received one of several types of disciplinary action.
+We report here rates of out-of-school suspension, in-school suspension, school-related arrests, law enforcement referrals, and total number of expulsions of any type.
+
+For both the post-stratification procedure (described below) and the actual statistical models used for inference, we used the same set of covariates.
+These covariates were at the state-level for post-stratification and were at the county-level for the statistical models used for final estimation and inference.
+Specifically, our population based covariates were the total population count, the proportion of the population that is black, the proportion of the population that is white and the ratio of black-to-white persons.
+We also used socioeconomic covariates.
+We used the percent of individuals aged 16 or over who were in the labor force but unemployed, the median household income, and the percentage of all families whose income is below the poverty line.
+
+Finally, for the implicit and explicit bias measures, we relied on two primary variables from the Project Implicit data.
+Implicit bias was assessed via an Implicit Association Test (IAT).
+This test uses a speeded dual-categorization task in which individuals must quickly categorize black and white faces and "good" and "bad" words with key presses.
+The difference in how quickly and accurately participants are able to pair white faces with "good" words and black faces with "bad" words in comparison to the inverse is thought to reflect implicit associations between the two races and positive and negative affective reactions.
+This association is indicated in the IAT D-score, which we used as a measure of implicit bias.
+Our measure of explicit bias is the difference between reported warmth towards whites (i.e.
+*how warm or cold do you feel towards Whites?
+0=very cold, 10=very warm*) and reported warmth toward blacks.
+
+## Data analysis
+Data analysis proceded in two steps.
+We first estimated county-level implict and explicit bias using multilevel regression and post-stratification.
+Post-stratification is a valuable procedure in obtaining accurate geographical population-based estimates because it allows a non-representative sample (e.g Project Implicit) to more closely resemble the true population, and it regularizes extreme observations with little data to support them (e.g.
+a county with only a handful of respondents with especially high or low scores) [@park2004bayesian; @gelman1997poststratification].
+Following past work [@leitner2016blacks], we identified age as one dimension along which IAT respondents differed from the general population in ways that could bias our conclusions [@gonsalkorale2009aging].
+Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories (15-24, 25-34, 35-54, 55-75, and 75+).
+We next fit multilevel models estimating bias (implicit and explicit biases seperately) as a function of our state-level covariates (the "fixed" effects), and allowed the estimates to vary by age bin, county, and state (the "random" effects).
+Next, we determined the population of whites in each county in these age groups using the American Community Survey's 5-year estimates ending in 2014.
+Finally, we used our estimated models to predict the expected response for each age bin, in each county.
+Our final county-level estimates are the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county.
+As a result of this procedure, we can be confident that our county-level estimates should more closely approximate what our estimates would look like if the Project Implicit data were truly representative along the age dimension in all counties.
+
+After obtaining these estimates, we use them as predictors in bayesian multilevel logistic regressions. Formally, the likelihood for a given observation is written as a binomial function:
+
+$$
+{n\choose y}\pi^y(1-\pi)^{n-y}
+$$
+
+Where $y$ is the observed count of incidents (e.g.
+number of black or white students suspended), $n$ is the number of at-risk students (e.g.
+total number of black or white students), and $\pi = g^{-1}(\eta)$ is the probability of the incident ocurring.
+For this analysis, the linear predictor takes the form of a multilevel model with a set of effects that vary over county:
+
+$$
+\eta = \alpha + X\beta + \gamma_{county}
+$$
+
+Where $\alpha$ is an intercept that is constant across observations, $\beta$ represents a set of effects that are also constant across observations (i.e.
+fixed effects), and $\gamma_{county}$ represents intercepts and effects of ethnicity that vary across the counties (i.e.
+random efffects).
+
+In addition to the covariates described above, we also include effects of race, implicit bias, explicit bias, and the two-way interactions between implicit bias and race and explicit bias and race.
+We fit separate models for each of the outcomes.
+
+Because of the computational demands of fitting such a high-dimensional model to such a large dataset (the full model for each metric would consist of over 6k parameters to approximately 170k observations), we used a consensus monte carlo algorithm to obtain approximate posterior distributions for the parameters of interest [@scott2016bayes].
+The approximate posteriors derived from this algorithm have been shown to be nearly indistinguishable from the true posterior, a result we verified using a small subset of our own data.
+
+All numerical predictor variables were standardized at the appropriate level (county, state) before model estimation to help with estimation efficiency and interpretability.
+We set priors for the intercept and coefficients in the bayesian models to be weakly informative normal distributions centered on zero with a standard deviation of five.
+All other parameters were left to default values.
+Data analysis was done in R [@rcitation] version 3.3.2 running under OS X 10.11.6.
+Post-stratification was done with lme4, version 1.1.14 [@bates2015fitting].
+Final model fitting was done on the university cluster running Springdale Linux, release 6.9 using rstanarm, version 2.17.2 [@rstanarm2016].
+We used the implementation of the consensus monte carlo algorithm found in parallelMCMCcombine, version 1.0 [@miroshnikov2014parallel].
+Figures were made with ggplot2, version 2.2.1 [@wickham2009ggplot2], with data manipulation done using dplyr version 0.7.2 [@wickham2017dplyr] and tidyr, version 0.7.1 [@wickham2017tidyr].
+A full report of session information can be found on the OSF page (....)
+
+# Results
+## Project Implicit Estimates
+```{r exp-diff}
+individual_data %>%
+  mutate(explicit_bias_diff = twhite_0to10-tblack_0to10) -> individual_data
+```
+We first report the results of estimating the implicit and explicit biases in from Project Implicit data.
+Overall, the individuals in project implicit show a pro-white bias in both implicit (*mean* = `r round(mean(individual_data$D_biep.White_Good_all), 2)`, *sd* = `r round(sd(individual_data$D_biep.White_Good_all), 2)`), and explicit measures (*mean* = `r round(mean(individual_data$explicit_bias_diff, na.rm=T), 2)` *sd* = `r round(sd(individual_data$explicit_bias_diff, na.rm=T), 2)`).
+When aggregated at the county level and adjusted with poststratification, the summary statistics across counties are similar in terms of their location, but as expected, the variability is much diminished ($mean_{implicit}$ = `r round(mean(as.numeric(county_means$weighted_bias), na.rm=T), 2)`, $sd_{implicit}$ = `r round(sd(as.numeric(county_means$weighted_bias), na.rm=T), 2)`; $mean_{explicit}$ = `r round(mean(as.numeric(county_means$weighted_explicit_diff), na.rm=T), 2)`, $sd_{explicit}$ = `r round(sd(as.numeric(county_means$weighted_explicit_diff), na.rm=T), 2)`, where on both scales 0 = no bias, and positive numbers indicate a pro-white bias).
+
+## Disciplinary action frequency
+
+```{r disc-count, results='asis'}
+model.data %>% 
+  filter(exclude==FALSE) %>%
+  group_by(group, metric) %>% 
+  summarise(students = sum(number)) %>%
+  mutate(metric = fct_recode(metric, 
+                             'expulsions' = 'expulsion_combined',
+                             'school arrests' = 'in_school_arrest',
+                             'in-school suspension' = 'inschool_susp',
+                             'law enforcement referral' = 'law_enforcement',
+                             'out-of-school suspension' = 'oos_susp')) -> counts_table
+
+apa_table(counts_table, caption = 'Count of students by race receiving each type of disciplinary action')
+```
+
+Table \@ref(tab:disc-count) shows the number of students of each race who were reported having received each of the actions under consideration.
+The counts range from a low of just `r counts_table$students[7]` white students arrested to a high of `r counts_table$students[5]` black students receiving an out-of-school suspension.
+Considering the vast differences in the overall number of black and white students, this simple count already illustrates that black students are disciplined at rates far higher than their white counterparts.
+
+## Associations across county
+Figure \@ref(fig:overall-associations) shows the estimate of primary interest for each of the models.
+The estimates displayed are the coefficients for the interaction between race and each of the two bias measurements.
+Given that African Americans are the baseline group, negative values for this coefficient indicate that as the bias in a county increases, the gap between the probability of a black student being disciplined and the probability of a white student being disciplined grows.
+
+```{r overall-associations, fig.cap="Association between each metric and county-level estimates of explicit and implicit bias.
+Negative values indicate that the rate of increase (or decrease) for blacks is faster (or slower) than for whites.
+Point is the mean of the posterior and error bars represent 95% bayesian uncertainty intervals."}
+#the p_neg for oos_susp & implicit bias is 100%, so when reporting the number in-text, i just replace it with 1 to prevent it from being rendered as NA in the document.
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/mw_uclaexcl_diff/'
+plot.dat <- data.frame(group = NA, weighted_bias = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(weighted_bias = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg=NA, metric=NA)
+
+# weighted bias w/exclusions & explicit difference
+metrics <- list.files(base_path)
+#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+  raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    df <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'weighted_bias')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate) %>%
+      filter(!is.na(county_id)) -> raw_rate
+    posterior_combo[,,j] <- t(df[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[13,]),
+             lower=quantile(p_cons[13,], .025),
+             upper=quantile(p_cons[13,], .975),
+             p_neg = prop.table(table(p_cons[13,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Expulsion' = 'expulsion_combined')
+plot.dat3_all <-plot.dat3
+plot.dat3_all$bias <- 'Implicit'
+
+#weighted warmth w/exclusions & explicit difference
+plot.dat <- data.frame(group = NA, weighted_warmth = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(weighted_warmth = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg = NA, metric=NA)
+
+for(i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, weighted_warmth=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    df <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'weighted_warmth')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate_warmth) %>%
+      filter(!is.na(county_id)) -> raw_rate_warmth
+    posterior_combo[,,j] <- t(df[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[14,]),
+             lower=quantile(p_cons[14,], .025),
+             upper=quantile(p_cons[14,], .975),
+             p_neg = prop.table(table(p_cons[14,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$bias <- 'Explicit'
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Expulsion' = 'expulsion_combined')
+plot.dat3_all <- rbind(plot.dat3_all, plot.dat3)
+ggplot(plot.dat3_all, aes(x=metric, y=est, group=bias, color=bias, shape=bias)) +
+  geom_point(position=position_dodge(width=.5), size=2) +
+  geom_errorbar(aes(ymin=lower, ymax=upper), 
+                width=.5, position=position_dodge(width=.5)) +
+  theme_classic() +
+  coord_flip() +
+  geom_hline(yintercept=0) +
+  ylab('Black-White Difference in log-odds slope') +
+  theme(axis.title.y = element_blank(),
+        legend.title = element_blank())
+```
+
+Several patterns are apparent from this figure.
+First, with the exceptions of implicit bias and expulsions and law enforcement referrals, all estimates are directionally consistent with higher levels of bias leading to larger differences between groups.
+Second, these effects are especially consistent for the two types of suspensions.
+The largest effect estimated is between implicit bias and out-of-school suspensions.
+The difference in the slope of the association between implicit bias and the log of the odds for out-of-school suspensions between white and black students is estimated to be `r round(plot.dat3_all$est[1], 2)`, with 95% of the posterior distribution between `r round(plot.dat3_all$lower[1], 2)` and `r round(plot.dat3_all$upper[1], 2)` and a proportion `r afex::round_ps(1)` of the posterior distribution consistent with a negative effect.
+
+Although not nearly as large of a difference, we have similar certainty with respect to the association between out-of-school suspensions and explicit bias.
+The difference in the slope of the association between explicit bias and the log of the odds for out-of-school suspensions between white and black students is estimated to be `r round(plot.dat3_all$est[6], 2)`, with 95% of the posterior distribution between `r round(plot.dat3_all$lower[6], 2)` and `r round(plot.dat3_all$upper[6], 2)` and a proportion `r afex::round_ps(plot.dat3_all$p_neg[6])` of the posterior distribution consistent with a negative effect.
+
+The estimated associations for in-school suspensions are smaller still, but are generally consistent with effects of the same direction.
+For implicit bias, the relevant parameter is estimated at `r round(plot.dat3_all$est[3], 2)` [`r round(plot.dat3_all$lower[3], 2)`, `r round(plot.dat3_all$upper[3], 2)`] $p_{neg}$ = `r round(plot.dat3_all$p_neg[3], 2)`, and for explicit bias, the effect is slightly larger, and a positive effect is essentially not credible, given the data and model `r round(plot.dat3_all$est[8], 2)` [`r round(plot.dat3_all$lower[8], 2)`, `r round(plot.dat3_all$upper[8], 2)`] $p_{neg}$ `r afex::round_ps(plot.dat3_all$p_neg[8])`.
+
+Other outcomes are estimated with less precision, or with patterns that are inconsistent between implicit and explicit bias.
+For instance, examining the associations for expulsions, the effect of explicit biases are generally in the expected direction (*est* = `r round(plot.dat3_all$est[10], 2)`, [`r round(plot.dat3_all$lower[10], 2)`, `r round(plot.dat3_all$upper[10], 2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[10])`), but the effect for implicit bias is estimated to be close to zero, with a enough uncertainty (*est* = `r round(plot.dat3_all$est[5], 2)`, [`r round(plot.dat3_all$lower[5], 2)`, `r round(plot.dat3_all$upper[5], 2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[5])`) to make it difficult to claim an effect of one direction or the other.
+The model indicates similar uncertainty with respect to school-related arrests and both estimates of bias ($est_{explicit}$ = `r round(plot.dat3_all$est[9], 2)`, [`r round(plot.dat3_all$lower[9], 2)`, `r round(plot.dat3_all$upper[9], 2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[9])`; $est_{implicit}$ = `r round(plot.dat3_all$est[4], 2)`, [`r round(plot.dat3_all$lower[4], 2)`, `r round(plot.dat3_all$upper[4], 2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[4])`).
+
+```{r detail-fig-exp, fig.cap="Association between bias and expulsions.
+Top: Association between bias and the estimated probability of expulsion.
+Line is the mean of the posterior.
+Bands indicate 95% uncertainty intervals; Bottom: Association between bias and the relative risk ratio for black students to white students.
+Points represent counties, whose size are scaled to the number of schools in that county."}
+library(gridExtra)
+i <- 'expulsion_combined'
+pth <- paste(base_path, i, '/', sep='')
+
+files <- list.files(pth)
+j <- 1
+posterior_combo <- array(0, dim=c(14,4000,length(files)))
+raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, 
+                       weighted_warmth=NA, group=NA, nincidents=NA, 
+                       nstudents=NA, rate=NA)
+for (k in files){
+  #print(i)
+  load(paste(pth,k,sep=''))
+  df <- as.matrix(m)
+  m$data %>% 
+    select(county_id, COMBOKEY) %>% 
+    distinct() %>% 
+    group_by(county_id) %>% 
+    summarise(nschools=n()) -> n_schools
+  m$data %>%
+    group_by(county_id, group) %>%
+    summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+    mutate(rate=nincidents/nstudents) -> rate
+  n_schools %>%
+    left_join(m$data[, c('county_id', 'weighted_bias', 'weighted_warmth')]) %>%
+    distinct() %>%
+    left_join(rate) %>%
+    rbind(raw_rate) %>%
+    filter(!is.na(county_id)) -> raw_rate
+  posterior_combo[,,j] <- t(df[,c(1:14)])
+  j <- j+1
+}
+p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+for (p in 1:14){
+  m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+}
+for (p in 1:14){
+  m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+}
+for (p in 1:14){
+  m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+}
+for (p in 1:14){
+  m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+}
+
+newdat <- expand.grid(group = rownames(table(m$data$group)),
+                      total_pop=0,
+                      unemp_rate=0,
+                      med_income=0,
+                      poverty_rate=0,
+                      col_grads=0,
+                      white_prop=0,
+                      black_prop=0,
+                      b.w.ratio=0,
+                      weighted_bias = seq(-2.5,2.5,.25),
+                      weighted_warmth = 0,
+                      number=0,
+                      total_number=1)
+y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)
+
+cbind(newdat, t(y_hat)) %>% 
+  mutate(index=row_number()) %>%
+  gather(sample, value, `1`:`4000`) -> posterior_distribution
+
+posterior_distribution %>%
+  select(group, weighted_bias, sample, value, index) %>%
+  group_by(group, weighted_bias) %>%
+  summarise(est = mean(value),
+            lower = quantile(value, .025),
+            upper = quantile(value, .975)) %>%
+  ungroup() %>%
+  mutate(bias = 'Implicit') -> plot.dat
+
+newdat <- expand.grid(group = rownames(table(m$data$group)),
+                      total_pop=0,
+                      unemp_rate=0,
+                      med_income=0,
+                      poverty_rate=0,
+                      col_grads=0,
+                      white_prop=0,
+                      black_prop=0,
+                      b.w.ratio=0,
+                      weighted_bias = 0,
+                      weighted_warmth = seq(-2.5,2.5,.25),
+                      number=0,
+                      total_number=1)
+y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)
+
+cbind(newdat, t(y_hat)) %>% 
+  mutate(index=row_number()) %>%
+  gather(sample, value, `1`:`4000`) -> posterior_distribution
+
+posterior_distribution %>%
+  select(group, weighted_warmth, sample, value, index) %>%
+  mutate(weighted_bias = weighted_warmth) %>%
+  group_by(group, weighted_bias) %>%
+  summarise(est = mean(value),
+            lower = quantile(value, .025),
+            upper = quantile(value, .975)) %>%
+  ungroup() %>%
+  mutate(bias = 'Explicit') %>%
+  rbind(plot.dat) -> plot.dat
+
+ggplot(plot.dat, aes(x=weighted_bias, y=est, group=group, linetype=group)) +
+  geom_line(aes(color=group)) +
+  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.15) +
+  theme_classic() +
+  ylab('Probability \n of Expulsion') +
+  xlab('Standardized Bias') + 
+  facet_wrap(~bias, ncol=2) +
+  xlim(-3.5, 3.5) +
+  theme(legend.title = element_blank(),
+        axis.title.x = element_blank()) -> p1
+
+posterior_distribution %>%
+  select(group, weighted_warmth, sample, value) %>%
+  spread(group, value) %>%
+  mutate(ratio = black/white) %>%
+  mutate(weighted_bias = weighted_warmth) %>%
+  group_by(weighted_bias) %>%
+  summarise(est = mean(ratio),
+            lower = quantile(ratio, .025),
+            upper = quantile(ratio, .975)) %>%
+  ungroup() %>%
+  mutate(bias = 'Explicit') -> plot.dat
+
+newdat <- expand.grid(group = rownames(table(m$data$group)),
+                      total_pop=0,
+                      unemp_rate=0,
+                      med_income=0,
+                      poverty_rate=0,
+                      col_grads=0,
+                      white_prop=0,
+                      black_prop=0,
+                      b.w.ratio=0,
+                      weighted_bias = seq(-2.5,2.5,.25),
+                      weighted_warmth = 0,
+                      number=0,
+                      total_number=1)
+y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)
+
+cbind(newdat, t(y_hat)) %>% 
+  mutate(index=row_number()) %>%
+  gather(sample, value, `1`:`4000`) -> posterior_distribution
+
+posterior_distribution %>%
+  select(group, weighted_bias, sample, value) %>%
+  spread(group, value) %>%
+  mutate(ratio = black/white) %>%
+  group_by(weighted_bias) %>%
+  summarise(est = mean(ratio),
+            lower = quantile(ratio, .025),
+            upper = quantile(ratio, .975)) %>%
+  ungroup() %>%
+  mutate(bias = 'Implicit') %>%
+  rbind(plot.dat) -> plot.dat
+
+raw_rate %>%
+  mutate(Implicit = weighted_bias,
+         Explicit = weighted_warmth) %>%
+  select(-weighted_bias, -weighted_warmth) %>%
+  gather(bias, value, Implicit:Explicit) %>%
+  select(county_id, nschools, bias, group, rate, value) %>%
+  spread(group, rate) %>%
+  mutate(odds_r = black/white) -> raw_or
+
+ggplot(plot.dat, aes(x=weighted_bias, y=est)) +
+  geom_point(data=raw_or, aes(x=value, y=odds_r, size=nschools), alpha=.075) +
+  geom_line() +
+  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.25) +
+  theme_classic() +
+  #scale_color_manual(values=c('grey', 'darkorange'), guide=F) + 
+  scale_alpha_manual(values=c(.2, 1), guide=F) +
+  ylim(0,9) +
+  xlim(-3.5, 3.5) + 
+  xlab('Standardized Bias') +
+  ylab('Relative \n Risk Ratio') +
+  geom_hline(yintercept=1) +
+  facet_wrap(~bias) -> p2
+
+p1$data %>% filter(weighted_bias %in% c(0, 1)) -> rates
+p2$data %>% filter(weighted_bias %in% c(0, 1)) -> risks
+
+grid.arrange(p1, p2)
+```
+
+To better illustrate the nature of these relationships, figure \@ref(fig:detail-fig-exp) shows the estimated probabilities of expulsion for black and white students as a function of bias, along with the relative risk ratio for black students.
+The relative risk ratio is the ratio of the probability that a black student will be expelled to the probability that a white student will get expelled.
+Values over 1 reflect higher levels of punishment for black students.
+As previously indicated in table \@ref(tab:disc-count), the top part of this figure makes plain the higher probability of expulsion for black children.
+In a county at the mean of the distribution of bias, approximately `r round(rates$est*100, 2)[1]`% of black students are expected to be expelled [`r round(rates$lower*100, 2)[1]`, `r round(rates$upper*100, 2)[1]`].
+The corresponding rate for white students is much lower, with about `r round(rates$est*100, 2)[3]`% expected to be expelled [`r round(rates$lower*100, 2)[3]`, `r round(rates$upper*100, 2)[3]`].
+Moving to a county one standard deviation above the mean of explicit bias has the effect of increasing the estimated percentage of black students expected to be expelled to `r round(rates$est*100, 2)[2]`% [`r round(rates$lower*100, 2)[2]`, `r round(rates$upper*100, 2)[2]`], while the percentage of white students expected to be expelled would decline to `r round(rates$est*100, 2)[4]`% [`r round(rates$lower*100, 2)[4]`, `r round(rates$upper*100, 2)[4]`].
+The same movement for implicit bias would slightly increase the expected expulsions for black students ( `r round(rates$est*100, 2)[6]`% [`r round(rates$lower*100, 2)[6]`, `r round(rates$upper*100, 2)[6]`), and increase the expected expulsions for white students a very small amount more to (`r round(rates$est*100, 2)[8]`% [`r round(rates$lower*100, 2)[8]`, `r round(rates$upper*100, 2)[8]`).
+In real terms, in a county at the mean of the distributions of bias, for every white student expelled, we should expect `r round(risks$est[1], 2)` [`r round(risks$lower[1], 2)`, `r round(risks$est[1], 2)`] black students to be expelled.
+If we move to a county one standard deviation above the mean of explicit bias, the ratio of black to white students expelled increases to `r round(risks$est[4], 2)` [`r round(risks$lower[4], 2)`, `r round(risks$est[4], 2)`], while the same movement for implicit bias slightly decreases the ratio to `r round(risks$est[2], 2)` [`r round(risks$lower[2], 2)`, `r round(risks$est[2], 2)`].
+
+```{r detail-fig-susp, fig.cap="Association between bias and out-of-school suspensions Top: Association between bias and the estimated probability of suspension Line is the mean of the posterior.
+Bands indicate 95% uncertainty intervals; Bottom: Association between bias and the relative risk ratio for black students to white students.
+Points represent counties, whose size are scaled to the number of schools in that county."}
+i <- 'oos_susp'
+pth <- paste(base_path, i, '/', sep='')
+
+files <- list.files(pth)
+j <- 1
+posterior_combo <- array(0, dim=c(14,4000,length(files)))
+raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, 
+                       weighted_warmth=NA, group=NA, nincidents=NA, 
+                       nstudents=NA, rate=NA)
+for (k in files){
+  #print(i)
+  load(paste(pth,k,sep=''))
+  df <- as.matrix(m)
+  m$data %>% 
+    select(county_id, COMBOKEY) %>% 
+    distinct() %>% 
+    group_by(county_id) %>% 
+    summarise(nschools=n()) -> n_schools
+  m$data %>%
+    group_by(county_id, group) %>%
+    summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+    mutate(rate=nincidents/nstudents) -> rate
+  n_schools %>%
+    left_join(m$data[, c('county_id', 'weighted_bias', 'weighted_warmth')]) %>%
+    distinct() %>%
+    left_join(rate) %>%
+    rbind(raw_rate) %>%
+    filter(!is.na(county_id)) -> raw_rate
+  posterior_combo[,,j] <- t(df[,c(1:14)])
+  j <- j+1
+}
+p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+for (p in 1:14){
+  m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+}
+for (p in 1:14){
+  m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+}
+for (p in 1:14){
+  m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+}
+for (p in 1:14){
+  m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+}
+
+newdat <- expand.grid(group = rownames(table(m$data$group)),
+                      total_pop=0,
+                      unemp_rate=0,
+                      med_income=0,
+                      poverty_rate=0,
+                      col_grads=0,
+                      white_prop=0,
+                      black_prop=0,
+                      b.w.ratio=0,
+                      weighted_bias = seq(-2.5,2.5,.25),
+                      weighted_warmth = 0,
+                      number=0,
+                      total_number=1)
+y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)
+
+cbind(newdat, t(y_hat)) %>% 
+  mutate(index=row_number()) %>%
+  gather(sample, value, `1`:`4000`) -> posterior_distribution
+
+posterior_distribution %>%
+  select(group, weighted_bias, sample, value, index) %>%
+  group_by(group, weighted_bias) %>%
+  summarise(est = mean(value),
+            lower = quantile(value, .025),
+            upper = quantile(value, .975)) %>%
+  ungroup() %>%
+  mutate(bias = 'Implicit') -> plot.dat
+
+newdat <- expand.grid(group = rownames(table(m$data$group)),
+                      total_pop=0,
+                      unemp_rate=0,
+                      med_income=0,
+                      poverty_rate=0,
+                      col_grads=0,
+                      white_prop=0,
+                      black_prop=0,
+                      b.w.ratio=0,
+                      weighted_bias = 0,
+                      weighted_warmth = seq(-2.5,2.5,.25),
+                      number=0,
+                      total_number=1)
+y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)
+
+cbind(newdat, t(y_hat)) %>% 
+  mutate(index=row_number()) %>%
+  gather(sample, value, `1`:`4000`) -> posterior_distribution
+
+posterior_distribution %>%
+  select(group, weighted_warmth, sample, value, index) %>%
+  mutate(weighted_bias = weighted_warmth) %>%
+  group_by(group, weighted_bias) %>%
+  summarise(est = mean(value),
+            lower = quantile(value, .025),
+            upper = quantile(value, .975)) %>%
+  ungroup() %>%
+  mutate(bias = 'Explicit') %>%
+  rbind(plot.dat) -> plot.dat
+
+ggplot(plot.dat, aes(x=weighted_bias, y=est, group=group, linetype=group)) +
+  geom_line(aes(color=group)) +
+  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.15) +
+  theme_classic() +
+  ylab('Probability of \n out-of-school suspension') +
+  xlab('Standardized Bias') + 
+  facet_wrap(~bias, ncol=2) +
+  xlim(-3.5, 3.5) +
+  theme(legend.title = element_blank(),
+        axis.title.x = element_blank()) -> p1
+
+posterior_distribution %>%
+  select(group, weighted_warmth, sample, value) %>%
+  spread(group, value) %>%
+  mutate(ratio = black/white) %>%
+  mutate(weighted_bias = weighted_warmth) %>%
+  group_by(weighted_bias) %>%
+  summarise(est = mean(ratio),
+            lower = quantile(ratio, .025),
+            upper = quantile(ratio, .975)) %>%
+  ungroup() %>%
+  mutate(bias = 'Explicit') -> plot.dat
+
+newdat <- expand.grid(group = rownames(table(m$data$group)),
+                      total_pop=0,
+                      unemp_rate=0,
+                      med_income=0,
+                      poverty_rate=0,
+                      col_grads=0,
+                      white_prop=0,
+                      black_prop=0,
+                      b.w.ratio=0,
+                      weighted_bias = seq(-2.5,2.5,.25),
+                      weighted_warmth = 0,
+                      number=0,
+                      total_number=1)
+y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)
+
+cbind(newdat, t(y_hat)) %>% 
+  mutate(index=row_number()) %>%
+  gather(sample, value, `1`:`4000`) -> posterior_distribution
+
+posterior_distribution %>%
+  select(group, weighted_bias, sample, value) %>%
+  spread(group, value) %>%
+  mutate(ratio = black/white) %>%
+  group_by(weighted_bias) %>%
+  summarise(est = mean(ratio),
+            lower = quantile(ratio, .025),
+            upper = quantile(ratio, .975)) %>%
+  ungroup() %>%
+  mutate(bias = 'Implicit') %>%
+  rbind(plot.dat) -> plot.dat
+
+raw_rate %>%
+  mutate(Implicit = weighted_bias,
+         Explicit = weighted_warmth) %>%
+  select(-weighted_bias, -weighted_warmth) %>%
+  gather(bias, value, Implicit:Explicit) %>%
+  select(county_id, nschools, bias, group, rate, value) %>%
+  spread(group, rate) %>%
+  mutate(odds_r = black/white) -> raw_or
+
+ggplot(plot.dat, aes(x=weighted_bias, y=est)) +
+  geom_point(data=raw_or, aes(x=value, y=odds_r, size=nschools), alpha=.075) +
+  geom_line() +
+  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.25) +
+  theme_classic() +
+  #scale_color_manual(values=c('grey', 'darkorange'), guide=F) + 
+  scale_alpha_manual(values=c(.2, 1), guide=F) +
+  ylim(0,9) +
+  xlim(-3.5, 3.5) + 
+  xlab('Standardized Bias') +
+  ylab('Relative \n Risk Ratio') +
+  geom_hline(yintercept=1) +
+  facet_wrap(~bias) -> p2
+
+p1$data %>% filter(weighted_bias %in% c(0, 1)) -> rates
+p2$data %>% filter(weighted_bias %in% c(0, 1)) -> risks
+
+grid.arrange(p1, p2)
+```
+
+Figure \@ref(fig:detail-fig-susp) shows similar patterns, but for out-of-school suspensions.
+In a county at the mean of the distribution of bias, approximately `r round(rates$est*100, 1)[1]`% of black students are expected to be suspended [`r round(rates$lower*100, 1)[1]`, `r round(rates$upper*100, 1)[1]`].
+The corresponding rate for white students is just over half that for black students, with about `r round(rates$est*100, 1)[3]`% expected to be suspended [`r round(rates$lower*100, 1)[3]`, `r round(rates$upper*100, 1)[3]`].
+Moving to a county one standard deviation above the mean of explicit bias has the effect of slightly decreasing the estimated percentage of black students expected to be suspended to `r round(rates$est*100, 1)[2]`% [`r round(rates$lower*100, 1)[2]`, `r round(rates$upper*100, 1)[2]`], while the percentage of white students expected to be suspended would decrease at a faster rate to `r round(rates$est*100, 1)[4]`% [`r round(rates$lower*100, 1)[4]`, `r round(rates$upper*100, 1)[4]`].
+The same movement for implicit bias would dramatically increase the expected suspensions for black students (`r round(rates$est*100, 1)[6]`% [`r round(rates$lower*100, 1)[6]`, `r round(rates$upper*100, 1)[6]`), and increase the expected suspensions for white students a much smaller amount to (`r round(rates$est*100, 1)[8]`% [`r round(rates$lower*100, 1)[8]`, `r round(rates$upper*100, 1)[8]`).
+In real terms, in a county at the mean of the distributions of bias, for every white student expelled, we should expect `r round(risks$est[1], 2)` [`r round(risks$lower[1], 2)`, `r round(risks$est[1], 2)`] black students to be expelled.
+If we move to a county one standard deviation above the mean of explicit bias, the ratio of black to white students expelled increases to `r round(risks$est[4], 2)` [`r round(risks$lower[4], 2)`, `r round(risks$est[4], 2)`], while the same movement for implicit bias slightly decreases the ratio to `r round(risks$est[2], 2)` [`r round(risks$lower[2], 2)`, `r round(risks$est[2], 2)`].`
+
+# Discussion
+
+\newpage
+
+# References
+```{r create_r-references}
+r_refs(file = "r-references.bib")
+```
+
+\newpage
+# Appendix
+## Preregistered analysis
+```{r summary_counts}
+
+###############
+schools_dat %>%
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
+  gather(group, total_number, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>%
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='NR_', 
+                          biracial='TR_', white='WH_')) %>%
+  filter(total_number>=0) %>%
+  group_by(COMBOKEY, group) %>%
+  mutate(total_number = sum(total_number)) %>%
+  ungroup() %>%
+  select(-prefix, -gender) %>%
+  distinct() -> ps_enrollment
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
+         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>% 
+  filter(SCH_GRADE_PS=='YES') %>%
+  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>% 
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='OS_', 
+                          biracial='TR_', white='WH_')) %>%
+  select(-prefix) -> ps_susp
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
+         SCH_GRADE_PS, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>% 
+  filter(SCH_GRADE_PS=='YES') %>%
+  gather(group, number_2, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>%
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='OS_', 
+                          biracial='TR_', white='WH_')) %>%
+  select(COMBOKEY, group, gender, number_2) %>%
+  right_join(ps_susp) %>%
+  filter(number>=0 & number_2 >= 0) %>%
+  mutate(number = number+number_2) -> ps_susp
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
+         SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>% 
+  gather(group, number, SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>% 
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='XP_', 
+                          biracial='TR_', white='WH_')) %>%
+  filter(number>=0) %>%
+  select(-prefix) -> ps_exp
+
+ps_susp %>% group_by(group) %>% summarise(n_students=sum(number)) -> susp_summ
+ps_exp %>% group_by(group) %>% summarise(n_students=sum(number)) -> exp_summ
+ps_enrollment %>% group_by(group) %>% summarise(n_students=sum(total_number)) -> enr_summ
+
+#s,
+
+```
+In our preanalysis plan, we specified our analyses to focus on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.
+However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes.
+In particular, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others.
+Additionally, the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]` total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely.
+We additionally discovered that counts of one expulsion category (expulsion under zero-tolerance policies) overlapped with counts in other categories, and so excluded this category.
+Finally, to remain consistent with previous studies, we opted to combine the remaining two expulsion categories to yield one overall count of the number of students expelled.
+
+We also preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e.
+*how warm or cold do you feel towards Blacks?
+0=very cold, 10=very warm*).
+However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so in the main text, we report models using this metric of explicit bias.
+Additionally, we preregistered analyses with poststratified estimates (as presented in the main text) along with raw, county-based means.
+We also had not known about the issues with juvenile justice facilities, or with the school districts with reporting errors, and so excluded these schools from analyses in the main text.
+Finally, we also registered an exploratory analysis using the responses of people who visited Project Implicit and identified themselves as teachers.
+Ultimately, we believe that there were too few individuals in the data to obtain reliable county-level estimates for this subgroup, and so this analysis is also limited to the appendix.
+Here, we present the results of the preregistered analyses exactly.
+Table \@ref(tab:all-results) presents the estimates (operationalized as the mean of the posterior), 95% uncertainty intervals, and the proportion of the posterior that is consistent with a negative effect for all of the outcomes for each of these models (*Model A*: raw county means estimates; *Model B*: post-stratified estimates using simple warmth score, *Model C*: teacher analyses).
+
+```{r all-results, results='asis'}
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_raw/'
+plot.dat <- data.frame(group = NA, weighted_bias = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(weighted_bias = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg=NA, metric=NA)
+
+metrics <- list.files(base_path)
+#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+  raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    df <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'weighted_bias')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate) %>%
+      filter(!is.na(county_id)) -> raw_rate
+    posterior_combo[,,j] <- t(df[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[13,]),
+             lower=quantile(p_cons[13,], .025),
+             upper=quantile(p_cons[13,], .975),
+             p_neg = prop.table(table(p_cons[13,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Corporal Punishment' = 'corporal',
+                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
+                               'Expulsion with Educational Services' = 'expulsion_w_ed',
+                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
+                               'Mechanical Restraint' = 'mechanical_restraint',
+                               'Physical Restraint' = 'physical_restraint',
+                               'Preschool Expulsion' = 'preschool_expulsion',
+                               'Preschool Suspension' = 'preschool_susp',
+                               'Seclusion' = 'seclusion')
+plot.dat3$model <- 'A'
+plot.dat3_all <-plot.dat3
+plot.dat3_all$bias <- 'Implicit'
+
+#weighted warmth w/exclusions & explicit difference
+plot.dat <- data.frame(group = NA, weighted_warmth = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(weighted_warmth = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg = NA, metric=NA)
+
+for(i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, weighted_warmth=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    df <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'weighted_warmth')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate_warmth) %>%
+      filter(!is.na(county_id)) -> raw_rate_warmth
+    posterior_combo[,,j] <- t(df[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[14,]),
+             lower=quantile(p_cons[14,], .025),
+             upper=quantile(p_cons[14,], .975),
+             p_neg = prop.table(table(p_cons[14,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$bias <- 'Explicit'
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Corporal Punishment' = 'corporal',
+                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
+                               'Expulsion with Educational Services' = 'expulsion_w_ed',
+                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
+                               'Mechanical Restraint' = 'mechanical_restraint',
+                               'Physical Restraint' = 'physical_restraint',
+                               'Preschool Expulsion' = 'preschool_expulsion',
+                               'Preschool Suspension' = 'preschool_susp',
+                               'Seclusion' = 'seclusion')
+plot.dat3$model <- 'A'
+table_dat <- rbind(plot.dat3_all, plot.dat3)
+
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_weighted/older_run/'
+plot.dat <- data.frame(group = NA, weighted_bias = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(weighted_bias = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg=NA, metric=NA)
+
+# weighted bias w/exclusions & explicit difference
+metrics <- list.files(base_path)
+
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+  raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    df <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'weighted_bias')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate) %>%
+      filter(!is.na(county_id)) -> raw_rate
+    posterior_combo[,,j] <- t(df[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[13,]),
+             lower=quantile(p_cons[13,], .025),
+             upper=quantile(p_cons[13,], .975),
+             p_neg = prop.table(table(p_cons[13,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Corporal Punishment' = 'corporal',
+                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
+                               'Expulsion with Educational Services' = 'expulsion_w_ed',
+                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
+                               'Mechanical Restraint' = 'mechanical_restraint',
+                               'Physical Restraint' = 'physical_restraint',
+                               'Preschool Expulsion' = 'preschool_expulsion',
+                               'Preschool Suspension' = 'preschool_susp',
+                               'Seclusion' = 'seclusion')
+plot.dat3$model <- 'B'
+plot.dat3_all <-plot.dat3
+plot.dat3_all$bias <- 'Implicit'
+
+#weighted warmth w/exclusions & explicit difference
+plot.dat <- data.frame(group = NA, weighted_warmth = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(weighted_warmth = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg = NA, metric=NA)
+
+for(i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, weighted_warmth=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    df <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'weighted_warmth')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate_warmth) %>%
+      filter(!is.na(county_id)) -> raw_rate_warmth
+    posterior_combo[,,j] <- t(df[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[14,]),
+             lower=quantile(p_cons[14,], .025),
+             upper=quantile(p_cons[14,], .975),
+             p_neg = prop.table(table(p_cons[14,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$bias <- 'Explicit'
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Corporal Punishment' = 'corporal',
+                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
+                               'Expulsion with Educational Services' = 'expulsion_w_ed',
+                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
+                               'Mechanical Restraint' = 'mechanical_restraint',
+                               'Physical Restraint' = 'physical_restraint',
+                               'Preschool Expulsion' = 'preschool_expulsion',
+                               'Preschool Suspension' = 'preschool_susp',
+                               'Seclusion' = 'seclusion')
+plot.dat3$model <- 'B'
+plot.dat3 %>%
+  rbind(plot.dat3_all) %>%
+  rbind(table_dat) -> table_dat
+
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/teacher_metrics/'
+plot.dat <- data.frame(group = NA, county_bias = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(county_bias = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg=NA, metric=NA)
+
+# weighted bias w/exclusions & explicit difference
+metrics <- list.files(base_path)
+
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+  raw_rate <- data.frame(county_id=NA, nschools=NA, county_bias=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    df <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'county_bias')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate) %>%
+      filter(!is.na(county_id)) -> raw_rate
+    posterior_combo[,,j] <- t(df[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[13,]),
+             lower=quantile(p_cons[13,], .025),
+             upper=quantile(p_cons[13,], .975),
+             p_neg = prop.table(table(p_cons[13,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Corporal Punishment' = 'corporal',
+                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
+                               'Expulsion with Educational Services' = 'expulsion_w_ed',
+                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
+                               'Mechanical Restraint' = 'mechanical_restraint',
+                               'Physical Restraint' = 'physical_restraint',
+                               'Preschool Expulsion' = 'preschool_expulsion',
+                               'Preschool Suspension' = 'preschool_susp',
+                               'Seclusion' = 'seclusion')
+plot.dat3$model <- 'C'
+plot.dat3_all <-plot.dat3
+plot.dat3_all$bias <- 'Implicit'
+
+plot.dat <- data.frame(group = NA, county_warmth = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(county_warmth = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg = NA, metric=NA)
+
+for(i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, county_warmth=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    df <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'county_warmth')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate_warmth) %>%
+      filter(!is.na(county_id)) -> raw_rate_warmth
+    posterior_combo[,,j] <- t(df[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[14,]),
+             lower=quantile(p_cons[14,], .025),
+             upper=quantile(p_cons[14,], .975),
+             p_neg = prop.table(table(p_cons[14,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$bias <- 'Explicit'
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Corporal Punishment' = 'corporal',
+                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
+                               'Expulsion with Educational Services' = 'expulsion_w_ed',
+                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
+                               'Mechanical Restraint' = 'mechanical_restraint',
+                               'Physical Restraint' = 'physical_restraint',
+                               'Preschool Expulsion' = 'preschool_expulsion',
+                               'Preschool Suspension' = 'preschool_susp',
+                               'Seclusion' = 'seclusion')
+plot.dat3$model <- 'C'
+plot.dat3 %>%
+  rbind(plot.dat3_all) %>%
+  rbind(table_dat) -> table_dat
+# ggplot(table_dat, aes(x=metric, y=est, group=model, color=model, shape=model)) +
+#   geom_point(position=position_dodge(width=.5), size=2) +
+#   geom_errorbar(aes(ymin=lower, ymax=upper), 
+#                 width=.5, position=position_dodge(width=.5)) +
+#   theme_classic() +
+#   coord_flip() +
+#   geom_hline(yintercept=0) +
+#   ylab('Black-White Difference in log-odds slope') +
+#   theme(axis.title.y = element_blank(),
+#         legend.title = element_blank(),
+#         axis.title.x = element_text(size=14),
+#         axis.text = element_text(size=12),
+#         legend.text = element_text(size=12),
+#         strip.text = element_text(size=12)) +
+#   facet_wrap(~bias)
+table_dat$p_neg[which(is.na(table_dat$p_neg))] <- 1
+table_dat %>%
+  mutate(UI = paste('[', round(lower, 2), ', ', round(upper, 2), ']', sep='')) %>%
+  select(est, UI, p_neg, metric, bias, model) %>%
+  mutate(val = paste(round(est, 2), ', ', UI, sep=''),
+         model = paste('Model ', model, sep='')) %>%
+  select(val, metric, bias, model) %>%
+  spread(model, val) %>%
+  arrange(bias, metric) -> tabdat
+
+papaja::apa_table(tabdat, caption = 'Effect estimates (operationalized as the mean of the posterior distribution for the difference in log-odds slope between black and white students) and 95% uncertainty intervals for all of the outcomes for each model (*Model A*: raw county means estimates; *Model B*: post-stratified estimates using simple warmth score, *Model C*: teacher analyses)', small = T)
+```
+
+#County-level estimates
+
+```{r raw-county-estimates, fig.cap='Difference in raw county means and post-stratifed estimates of implicit bias as a function of the number of observations in each county.
+The X axis is on the log scale.'}
+individual_data %>%
+  group_by(county_id) %>%
+  summarise(n_obs=n()) %>%
+  left_join(county_means) -> cm
+
+cm %>%
+  mutate(post_diff = as.numeric(bias) - as.numeric(weighted_bias)) %>%
+  ggplot(aes(x=n_obs, y=post_diff)) + 
+  geom_point() + 
+  scale_x_log10(breaks=c(1, 10, 100, 1000, 10000)) +
+  theme_classic() +
+  ylab('Raw Bias - Weighted Bias difference') +
+  xlab('Number of Observations')
+
+```
+
+Examining the models fit with simple county means, we find that the county-level estimates vary more than with the post-stratified estimates.
+This pattern is expected, as one desireable feature of post-stratification is that it regularizes extreme observations without a large amount of data to suppor them.
+This phenomena is illustrated in figure \@ref(fig:raw-county-estimates), which shows the distribution of differences between the raw county means and the post-stratified estimates of implicit bias as a function of the number of observations in each county.
+More specifically, while the mean of the county-level means does not meaningfully differ from the post-stratified estimates, the standard deviation is much larger ($M_{implicit}$ = `r round(mean(as.numeric(cm$bias), na.rm=T), 2)`, $SD_{implicit}$ = `r round(sd(as.numeric(cm$bias), na.rm=T), 2)`; $M_{explicit}$ = `r round(mean(as.numeric(cm$explicit), na.rm=T), 2)`, $SD_{explicit}$ = `r round(sd(as.numeric(cm$explicit), na.rm=T), 2)`)[^2]
+
+[^2]: This explicit estimate is based on the simple warmth towards blacks question, and not the difference between warmth towards whites and warmth twoards blacks.
+
+The county-level implicit bias estimates for the poststratified model are as reported in the main text.
+The estimates for explicit bias are slightly different, as they are not computed based on the difference in reporting warmth toward whites and warmth toward blacks.
+Specifically, the average amount of bias at the county-level is `r round(mean(as.numeric(cm$weighted_explicit)*-1, na.rm=T), 2)` (SD = `r round(sd(as.numeric(cm$weighted_explicit), na.rm=T), 2)`).
+
 ```{r teachers-summary}
+df <- rbind(read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2013.sav'),
+            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2012.sav'),
+            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2010.sav'))
 
@@
 As such, our teacher analysis is limited to just `r dim(teacher_estimates)[1]` counties.
-Additionally, because we do not know of any state-level demographic estimates for teachers, we are unable to perform post-stratification for these data.
-
-## Measures
-The primary outcome in this analysis is a count of the number of students by race (black and white) who received one of several types of disciplinary action.
-We report here rates of out-of-school suspension, in-school suspension, school-related arrests, law enforcement referrals, and total number of expulsions of any type.
-
-For both the post-stratification procedure (described below) and the actual statistical models used for inference, we used the same set of covariates.
-These covariates were at the state-level for post-stratification and were at the county-level for the statistical models used for final estimation and inference.
-Specifically, our population based covariates were the total population count, the proportion of the population that is black, the proportion of the population that is white and the ratio of black-to-white persons.
-We also used socioeconomic covariates.
-We used the percent of individuals aged 16 or over who were in the labor force but unemployed, the median household income, and the percentage of all families whose income is below the poverty line.
-
-Finally, for the implicit and explicit bias measures, we relied on two primary variables from the Project Implicit data.
-Implicit bias was assessed via an Implicit Association Test (IAT).
-This test uses a speeded dual-categorization task in which individuals must quickly categorize black and white faces and "good" and "bad" words with key presses.
-The difference in how quickly and accurately participants are able to pair white faces with "good" words and black faces with "bad" words in comparison to the inverse is thought to reflect implicit associations between the two races and positive and negative affective reactions.
-This association is reflected in the IAT D-score, which we used as a measure of implicit bias.
-Our measure of explicit bias is the difference between reported warmth towards whites (i.e.
-*how warm or cold do you feel towards Whites?
-0=very cold, 10=very warm*) and reported warmth toward blacks.
-
-## Data analysis
-Data analysis proceded in two steps.
-We first estimated county-level implict and explicit bias using multilevel regression and post-stratification.
-Post-stratification is a valuable procedure in obtaining accurate geographical population-based estimates because it allows a non-representative sample (e.g Project Implicit) to more closely resemble the true population, and it regularizes extreme observations with little data to support them (e.g.
-a county with only a handful of respondents with especially high or low scores) [@park2004bayesian; @gelman1997poststratification].
-Following past work [@leitner2016blacks], we identified age as one dimension along which IAT respondents differed from the general population in ways that could bias our conclusions [@gonsalkorale2009aging].
-Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories (15-24, 25-34, 35-54, 55-75, and 75+).
-We next fit multilevel models estimating bias (implicit and explicit biases seperately) as a function of our state-level covariates (the "fixed" effects), and allowed the estimates to vary by age bin, county, and state (the "random" effects).
-Next, we determined the population of whites in each county in these age groups using the American Community Survey's 5-year estimates ending in 2014.
-Finally, we used our estimated models to predict the expected response for each age bin, in each county.
-Our final county-level estimates are the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county.
-As a result of this procedure, we can be confident that our county-level estimates should more closely approximate what our estimates would look like if the Project Implicit data were truly representative along the age dimension in all counties.
-
-After obtaining these estimates, we use them as predictors in bayesian multilevel logistic regressions. Formally, the likelihood for a given observation is written as a binomial function:
-
-$$
-{n\choose y}\pi^y(1-\pi)^{n-y}
-$$
-
-Where $y$ is the observed count of incidents (e.g.
-number of black or white students suspended), $n$ is the number of at-risk students (e.g.
-total number of black or white students), and $\pi = g^{-1}(\eta)$ is the probability of the incident ocurring.
-For this analysis, the linear predictor takes the form of a multilevel model with a set of effects that vary over county:
-
-$$
-\eta = \alpha + X\beta + \gamma_{county}
-$$
-
-Where $\alpha$ is an intercept that is constant across observations, $\beta$ represents a set of effects that are also constant across observations (i.e.
-fixed effects), and $\gamma_{county}$ represents intercepts and effects of ethnicity that vary across the counties (i.e.
-random efffects).
-
-In addition to the covariates described above, we also include effects of race, implicit bias, explicit bias, and the two-way interactions between implicit bias and race and explicit bias and race.
-We fit separate models for each of the outcomes.
-
-Because of the computational demands of fitting such a high-dimensional model to such a large dataset (the full model for each metric would consist of over 6k parameters to approximately 170k observations), we used a consensus monte carlo algorithm to obtain approximate posterior distributions for the parameters of interest [@scott2016bayes].
-The approximate posteriors derived from this algorithm have been shown to be nearly indistinguishable from the true posterior, a result we verified using a small subset of our own data.
-
-All numerical predictor variables were standardized at the appropriate level (county, state) before model estimation to help with estimation efficiency and interpretability.
-We set priors for the intercept and coefficients in the bayesian models to be weakly informative normal distributions centered on zero with a standard deviation of five.
-All other parameters were left to default values.
-Data analysis was done in R [@rcitation] version 3.3.2 running under OS X 10.11.6.
-Post-stratification was done with lme4, version 1.1.14 [@bates2015fitting].
-Final model fitting was done on the university cluster running Springdale Linux, release 6.9 using rstanarm, version 2.17.2 [@rstanarm2016].
-We used the implementation of the consensus monte carlo algorithm found in parallelMCMCcombine, version 1.0 [@miroshnikov2014parallel].
-Figures were made with ggplot2, version 2.2.1 [@wickham2009ggplot2], with data manipulation done using dplyr version 0.7.2 [@wickham2017dplyr] and tidyr, version 0.7.1 [@wickham2017tidyr].
-A full report of session information can be found on the OSF page (....)
-
-# Results
-
-# Discussion
-
-\newpage
-
-# References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-```
-
-\newpage
-# Appendix
-## Preregistered analysis
-```{r summary_counts}
-
-###############
-schools_dat %>%
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
-  gather(group, total_number, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>%
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='NR_', 
-                          biracial='TR_', white='WH_')) %>%
-  filter(total_number>=0) %>%
-  group_by(COMBOKEY, group) %>%
-  mutate(total_number = sum(total_number)) %>%
-  ungroup() %>%
-  select(-prefix, -gender) %>%
-  distinct() -> ps_enrollment
-
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
-         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>% 
-  filter(SCH_GRADE_PS=='YES') %>%
-  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>% 
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='OS_', 
-                          biracial='TR_', white='WH_')) %>%
-  select(-prefix) -> ps_susp
-
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
-         SCH_GRADE_PS, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>% 
-  filter(SCH_GRADE_PS=='YES') %>%
-  gather(group, number_2, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>%
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='OS_', 
-                          biracial='TR_', white='WH_')) %>%
-  select(COMBOKEY, group, gender, number_2) %>%
-  right_join(ps_susp) %>%
-  filter(number>=0 & number_2 >= 0) %>%
-  mutate(number = number+number_2) -> ps_susp
-
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
-         SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>% 
-  gather(group, number, SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>% 
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='XP_', 
-                          biracial='TR_', white='WH_')) %>%
-  filter(number>=0) %>%
-  select(-prefix) -> ps_exp
-
-ps_susp %>% group_by(group) %>% summarise(n_students=sum(number)) -> susp_summ
-ps_exp %>% group_by(group) %>% summarise(n_students=sum(number)) -> exp_summ
-ps_enrollment %>% group_by(group) %>% summarise(n_students=sum(total_number)) -> enr_summ
-
-#s,
-
-```
-In our preanalysis plan, we specified our analyses to focus on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.
-However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes.
-In particular, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others.
-Additionally, the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]` total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely.
-We additionally discovered that counts of one expulsion category (expulsion under zero-tolerance policies) overlapped with counts in other categories, and so excluded this category.
-To remain consistent with previous studies, we opted to combine the remaining two expulsion categories to yield one overall count of the number of students expelled.
-
-We also preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e.
-*how warm or cold do you feel towards Blacks?
-0=very cold, 10=very warm*).
-However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so we in the main text, we report models using this metric of explicit bias.
-Additionally, we preregistered analyses with poststratified estimates (as presented in the main text) along with raw, county-based means.
-Finally, we had not known about the issues with juvenile justice facilities, or with the school districts with reporting errors.
-Here, we present the results of the preregistered analyses exactly.
-
-#Simple county means
-
-#Post stratified estimates
-
-#Teacher analyses
+Additionally, because we do not know of any state-level demographic estimates for teachers, we were unable to perform post-stratification for these data.
+Nonetheless, across these limited counties, the overall estimate of implicit and explicit bias are similar to those for the whole dataset ($M_{implicit}$ = `r round(mean(as.numeric(teacher_estimates$teacher_bias), na.rm=T), 2)`, $SD_{implicit}$ = `r round(sd(as.numeric(teacher_estimates$teacher_bias), na.rm=T), 2)`; $M_{explicit}$ = `r round(mean(as.numeric(teacher_estimates$teacher_explicit), na.rm=T), 2)`, $SD_{explicit}$ = `r round(sd(as.numeric(teacher_estimates$teacher_explicit), na.rm=T), 2)`)
 

### 2018-02-26 written_docs/draft.tex
There's some content here.

\documentclass[english,floatsintext,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Race-based disparities in allocation of academic disciplinary actions are associated with county-level rates of bias},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Race-based disparities in allocation of academic disciplinary actions
are associated with county-level rates of bias}

  \shorttitle{Discipline Disparities}

  \author{Travis Riddle\textsuperscript{1}~\& Stacey Sinclair\textsuperscript{1}}

  \def\affdep{{"", ""}}%
  \def\affcity{{"", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Princeton University  }

  \authornote{
    \newcounter{author}
    Enter author note here.

                      Correspondence concerning this article should be addressed to Travis Riddle. E-mail: \href{mailto:triddle@princeton.edu}{\nolinkurl{triddle@princeton.edu}}
                          }

  \abstract{Enter abstract here.}
  \keywords{keywords \\

    \indent Word count: X
  }

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

\section{Introduction}\label{introduction}

\section{Methods}\label{methods}

\subsection{Data Sources}\label{data-sources}

We used three distinct data sources for the work described here. The
academic disciplinary data is part of the Civil Rights Data Collection
(CRDC) through the US Department of Education. The dataset we used comes
from the 2013-2014 academic year and has data on \enquote{all public
local and educational agencies and schools, including long-term secure
juvenile justice facilities, charter schools, alternative schools, and
schools serving students with disabilities.} In total, the CRDC data
represents 95507 institutions enrolling approximately 50 million
students, of which approximately 25 million are white and 7.8 million
are Black\footnote{We note that there are a number of differences
  between the analyses we registered and those presented in the main
  text. Our general conclusions are largely the same for both sets of
  analyses. We opted to report the modified analyses for reasons of
  clarity and to remain congruent with previous research on the same
  topics. The registered analyses can be found, in full, in the
  appendix.}. Previous work using these data have identified a number of
districts whose data are in error, and have excluded juvenile justice
facilities, as these institutions constitute dramatically different
educational environments, where the meaning of disciplinary actions may
be quite different (Losen et al., 2015). After these exclusions are
applied, the final sample used for modeling consists of 90002
institutions, enrolling 32 million black or white students, of which 25
million are white and 7 million are black.

We obtained county-level demographic information for use as covariates
and state-level demographic information for use in post-stratification
from the US Census Bureau. For both county-level and state-level
demographics, we use 5-year estimates for the period ending in 2014 from
the American Community Survey, which surveys around 295k households per
month.

We used measurements of implicit and explicit bias available from data
collected through Project Implicit (Xu, Nosek, \& Greenwald, 2014). From
this data, we used only respondents who had geographic information that
would allow us to place them in a United States county, identified as
White, and visited the site before 2015. This consisted of approximately
1.1 million total respondents from 3091 counties.

A subset of years in the Project Implicit data also collected
occupational information from respondents. As identified in our
pre-analysis plan, we took advantage of the presence of primary and
secondary educators in these data to test whether any associations
between bias and race-based differences in the rates of disciplinary
action were stronger among these respondents. Filtering for only white
individuals who identified as primary, secondary, special education, and
other teachers and instructors (occupation codes 25-2000 and 25-3000)
reduced the dataset to 63552 respondents. In order to assure that our
estimates were reasonably stable, we limited analysis to only counties
that had at least 50 respondents. As such, our teacher analysis is
limited to just 287 counties. Additionally, because we do not know of
any state-level demographic estimates for teachers, we are unable to
perform post-stratification for these data.

\subsection{Measures}\label{measures}

The primary outcome in this analysis is a count of the number of
students by race (black and white) who received one of several types of
disciplinary action. We report here rates of out-of-school suspension,
in-school suspension, school-related arrests, law enforcement referrals,
and total number of expulsions of any type.

For both the post-stratification procedure (described below) and the
actual statistical models used for inference, we used the same set of
covariates. These covariates were at the state-level for
post-stratification and were at the county-level for the statistical
models used for final estimation and inference. Specifically, our
population based covariates were the total population count, the
proportion of the population that is black, the proportion of the
population that is white and the ratio of black-to-white persons. We
also used socioeconomic covariates. We used the percent of individuals
aged 16 or over who were in the labor force but unemployed, the median
household income, and the percentage of all families whose income is
below the poverty line.

Finally, for the implicit and explicit bias measures, we relied on two
primary variables from the Project Implicit data. Implicit bias was
assessed via an Implicit Association Test (IAT). This test uses a
speeded dual-categorization task in which individuals must quickly
categorize black and white faces and \enquote{good} and \enquote{bad}
words with key presses. The difference in how quickly and accurately
participants are able to pair white faces with \enquote{good} words and
black faces with \enquote{bad} words in comparison to the inverse is
thought to reflect implicit associations between the two races and
positive and negative affective reactions. This association is indicated
in the IAT D-score, which we used as a measure of implicit bias. Our
measure of explicit bias is the difference between reported warmth
towards whites (i.e. \emph{how warm or cold do you feel towards Whites?
0=very cold, 10=very warm}) and reported warmth toward blacks.

\subsection{Data analysis}\label{data-analysis}

Data analysis proceded in two steps. We first estimated county-level
implict and explicit bias using multilevel regression and
post-stratification. Post-stratification is a valuable procedure in
obtaining accurate geographical population-based estimates because it
allows a non-representative sample (e.g Project Implicit) to more
closely resemble the true population, and it regularizes extreme
observations with little data to support them (e.g.~a county with only a
handful of respondents with especially high or low scores) (Gelman \&
Little, 1997; Park, Gelman, \& Bafumi, 2004). Following past work
(Leitner, Hehman, Ayduk, \& Mendoza-Denton, 2016), we identified age as
one dimension along which IAT respondents differed from the general
population in ways that could bias our conclusions (Gonsalkorale,
Sherman, \& Klauer, 2009). Our post-stratification weighting scheme is
as follows: We first grouped respondents into five age group categories
(15-24, 25-34, 35-54, 55-75, and 75+). We next fit multilevel models
estimating bias (implicit and explicit biases seperately) as a function
of our state-level covariates (the \enquote{fixed} effects), and allowed
the estimates to vary by age bin, county, and state (the
\enquote{random} effects). Next, we determined the population of whites
in each county in these age groups using the American Community Survey's
5-year estimates ending in 2014. Finally, we used our estimated models
to predict the expected response for each age bin, in each county. Our
final county-level estimates are the average of the values predicted for
the 5 age bins, weighted by the population size of that bin in that
county. As a result of this procedure, we can be confident that our
county-level estimates should more closely approximate what our
estimates would look like if the Project Implicit data were truly
representative along the age dimension in all counties.

After obtaining these estimates, we use them as predictors in bayesian
multilevel logistic regressions. Formally, the likelihood for a given
observation is written as a binomial function:

\[
{n\choose y}\pi^y(1-\pi)^{n-y}
\]

Where \(y\) is the observed count of incidents (e.g.~number of black or
white students suspended), \(n\) is the number of at-risk students
(e.g.~total number of black or white students), and
\(\pi = g^{-1}(\eta)\) is the probability of the incident ocurring. For
this analysis, the linear predictor takes the form of a multilevel model
with a set of effects that vary over county:

\[
\eta = \alpha + X\beta + \gamma_{county}
\]

Where \(\alpha\) is an intercept that is constant across observations,
\(\beta\) represents a set of effects that are also constant across
observations (i.e.~fixed effects), and \(\gamma_{county}\) represents
intercepts and effects of ethnicity that vary across the counties
(i.e.~random efffects).

In addition to the covariates described above, we also include effects
of race, implicit bias, explicit bias, and the two-way interactions
between implicit bias and race and explicit bias and race. We fit
separate models for each of the outcomes.

Because of the computational demands of fitting such a high-dimensional
model to such a large dataset (the full model for each metric would
consist of over 6k parameters to approximately 170k observations), we
used a consensus monte carlo algorithm to obtain approximate posterior
distributions for the parameters of interest (Scott et al., 2016). The
approximate posteriors derived from this algorithm have been shown to be
nearly indistinguishable from the true posterior, a result we verified
using a small subset of our own data.

All numerical predictor variables were standardized at the appropriate
level (county, state) before model estimation to help with estimation
efficiency and interpretability. We set priors for the intercept and
coefficients in the bayesian models to be weakly informative normal
distributions centered on zero with a standard deviation of five. All
other parameters were left to default values. Data analysis was done in
R (R Core Team, 2016) version 3.3.2 running under OS X 10.11.6.
Post-stratification was done with lme4, version 1.1.14 (Bates, Mächler,
Bolker, \& Walker, 2015). Final model fitting was done on the university
cluster running Springdale Linux, release 6.9 using rstanarm, version
2.17.2 (Stan Development Team, 2016). We used the implementation of the
consensus monte carlo algorithm found in parallelMCMCcombine, version
1.0 (Miroshnikov \& Conlon, 2014). Figures were made with ggplot2,
version 2.2.1 (Wickham, 2009), with data manipulation done using dplyr
version 0.7.2 (Wickham, Francois, Henry, \& Müller, 2017) and tidyr,
version 0.7.1 (Wickham \& Henry, 2017). A full report of session
information can be found on the OSF page (\ldots{}.)

\section{Results}\label{results}

\subsection{Project Implicit
Estimates}\label{project-implicit-estimates}

We first report the results of estimating the implicit and explicit
biases in from Project Implicit data. Overall, the individuals in
project implicit show a pro-white bias in both implicit (\emph{mean} =
0.40, \emph{sd} = 0.41), and explicit measures (\emph{mean} = 0.88
\emph{sd} = 1.83). When aggregated at the county level and adjusted with
poststratification, the summary statistics across counties are similar
in terms of their location, but as expected, the variability is much
diminished (\(mean_{implicit}\) = 0.40, \(sd_{implicit}\) = 0.02;
\(mean_{explicit}\) = 0.79, \(sd_{explicit}\) = 0.15, where on both
scales 0 = no bias, and positive numbers indicate a pro-white bias).

\subsection{Disciplinary action
frequency}\label{disciplinary-action-frequency}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:disc-count}Count of students by race receiving each type of disciplinary action}
\begin{tabular}{lll}
\toprule
group & \multicolumn{1}{c}{metric} & \multicolumn{1}{c}{students}\\
\midrule
black & expulsions & 36,755.00\\
black & school arrests & 21,456.00\\
black & in-school suspension & 829,706.00\\
black & law enforcement referral & 53,306.00\\
black & out-of-school suspension & 1,850,492.00\\
white & expulsions & 55,832.00\\
white & school arrests & 21,420.00\\
white & in-school suspension & 1,054,172.00\\
white & law enforcement referral & 81,565.00\\
white & out-of-school suspension & 1,768,326.00\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

Table \ref{tab:disc-count} shows the number of students of each race who
were reported having received each of the actions under consideration.
The counts range from a low of just 21420 white students arrested to a
high of 1850492 black students receiving an out-of-school suspension.
Considering the vast differences in the overall number of black and
white students, this simple count already illustrates that black
students are disciplined at rates far higher than their white
counterparts.

\subsection{Associations across
county}\label{associations-across-county}

Figure \ref{fig:overall-associations} shows the estimate of primary
interest for each of the models. The estimates displayed are the
coefficients for the interaction between race and each of the two bias
measurements. Given that African Americans are the baseline group,
negative values for this coefficient indicate that as the bias in a
county increases, the gap between the probability of a black student
being disciplined and the probability of a white student being
disciplined grows.

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/overall-associations-1.pdf}
\caption{\label{fig:overall-associations}Association between each metric and
county-level estimates of explicit and implicit bias. Negative values
indicate that the rate of increase (or decrease) for blacks is faster
(or slower) than for whites. Point is the mean of the posterior and
error bars represent 95\% bayesian uncertainty intervals.}
\end{figure}

Several patterns are apparent from this figure. First, with the
exceptions of implicit bias and expulsions and law enforcement
referrals, all estimates are directionally consistent with higher levels
of bias leading to larger differences between groups. Second, these
effects are especially consistent for the two types of suspensions. The
largest effect estimated is between implicit bias and out-of-school
suspensions. The difference in the slope of the association between
implicit bias and the log of the odds for out-of-school suspensions
between white and black students is estimated to be -0.25, with 95\% of
the posterior distribution between -0.31 and -0.20 and a proportion
\textgreater{}.99 of the posterior distribution consistent with a
negative effect.

Although not nearly as large of a difference, we have similar certainty
with respect to the association between out-of-school suspensions and
explicit bias. The difference in the slope of the association between
explicit bias and the log of the odds for out-of-school suspensions
between white and black students is estimated to be -0.08, with 95\% of
the posterior distribution between -0.14 and -0.03 and a proportion
\textgreater{}.99 of the posterior distribution consistent with a
negative effect.

The estimated associations for in-school suspensions are smaller still,
but are generally consistent with effects of the same direction. For
implicit bias, the relevant parameter is estimated at -0.02 {[}-0.06,
0.01{]} \(p_{neg}\) = 0.93, and for explicit bias, the effect is
slightly larger, and a positive effect is essentially not credible,
given the data and model -0.05 {[}-0.08, -0.02{]} \(p_{neg}\)
\textgreater{}.99.

Other outcomes are estimated with less precision, or with patterns that
are inconsistent between implicit and explicit bias. For instance,
examining the associations for expulsions, the effect of explicit biases
are generally in the expected direction (\emph{est} = -0.07, {[}-0.15,
0.02{]}, \(p_{neg}\) = .94), but the effect for implicit bias is
estimated to be close to zero, with a enough uncertainty (\emph{est} =
0.04, {[}-0.05, 0.12{]}, \(p_{neg}\) = .19) to make it difficult to
claim an effect of one direction or the other. The model indicates
similar uncertainty with respect to school-related arrests and both
estimates of bias (\(est_{explicit}\) = -0.02, {[}-0.11, 0.07{]},
\(p_{neg}\) = .67; \(est_{implicit}\) = -0.02, {[}-0.12, 0.08{]},
\(p_{neg}\) = .66).

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/detail-fig-exp-1.pdf}
\caption{\label{fig:detail-fig-exp}Association between bias and expulsions.
Top: Association between bias and the estimated probability of
expulsion. Line is the mean of the posterior. Bands indicate 95\%
uncertainty intervals; Bottom: Association between bias and the relative
risk ratio for black students to white students. Points represent
counties, whose size are scaled to the number of schools in that
county.}
\end{figure}

To better illustrate the nature of these relationships, figure
\ref{fig:detail-fig-exp} shows the estimated probabilities of expulsion
for black and white students as a function of bias, along with the
relative risk ratio for black students. The relative risk ratio is the
ratio of the probability that a black student will be expelled to the
probability that a white student will get expelled. Values over 1
reflect higher levels of punishment for black students. As previously
indicated in table \ref{tab:disc-count}, the top part of this figure
makes plain the higher probability of expulsion for black children. In a
county at the mean of the distribution of bias, approximately 0.15\% of
black students are expected to be expelled {[}0.14, 0.17{]}. The
corresponding rate for white students is much lower, with about 0.06\%
expected to be expelled {[}0.06, 0.07{]}. Moving to a county one
standard deviation above the mean of explicit bias has the effect of
increasing the estimated percentage of black students expected to be
expelled to 0.16\% {[}0.13, 0.18{]}, while the percentage of white
students expected to be expelled would decline to 0.06\% {[}0.05,
0.07{]}. The same movement for implicit bias would slightly increase the
expected expulsions for black students ( 0.15\% {[}0.13, 0.18), and
increase the expected expulsions for white students a very small amount
more to (0.06\% {[}0.06, 0.07). In real terms, in a county at the mean
of the distributions of bias, for every white student expelled, we
should expect 2.45 {[}2.25, 2.45{]} black students to be expelled. If we
move to a county one standard deviation above the mean of explicit bias,
the ratio of black to white students expelled increases to 2.63 {[}2.33,
2.63{]}, while the same movement for implicit bias slightly decreases
the ratio to 2.36 {[}2.13, 2.36{]}.

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/detail-fig-susp-1.pdf}
\caption{\label{fig:detail-fig-susp}Association between bias and
out-of-school suspensions Top: Association between bias and the
estimated probability of suspension Line is the mean of the posterior.
Bands indicate 95\% uncertainty intervals; Bottom: Association between
bias and the relative risk ratio for black students to white students.
Points represent counties, whose size are scaled to the number of
schools in that county.}
\end{figure}

Figure \ref{fig:detail-fig-susp} shows similar patterns, but for
out-of-school suspensions. In a county at the mean of the distribution
of bias, approximately 11.50\% of black students are expected to be
suspended {[}11, 12{]}. The corresponding rate for white students is
just over half that for black students, with about 6.40\% expected to be
suspended {[}6.20, 6.50{]}. Moving to a county one standard deviation
above the mean of explicit bias has the effect of slightly decreasing
the estimated percentage of black students expected to be suspended to
11.10\% {[}10.30, 11.80{]}, while the percentage of white students
expected to be suspended would decrease at a faster rate to 5.60\%
{[}5.40, 5.90{]}. The same movement for implicit bias would dramatically
increase the expected suspensions for black students (15.40\% {[}14.40,
16.40), and increase the expected suspensions for white students a much
smaller amount to (6.90\% {[}6.60, 7.30). In real terms, in a county at
the mean of the distributions of bias, for every white student expelled,
we should expect 1.80 {[}1.73, 1.80{]} black students to be expelled. If
we move to a county one standard deviation above the mean of explicit
bias, the ratio of black to white students expelled increases to 1.96
{[}1.84, 1.96{]}, while the same movement for implicit bias slightly
decreases the ratio to 2.23 {[}2.11, 2.23{]}.`

\section{Discussion}\label{discussion}

\newpage

\section{References}\label{references}

\newpage

\section{Appendix}\label{appendix}

\subsection{Preregistered analysis}\label{preregistered-analysis}

In our preanalysis plan, we specified our analyses to focus on 13
actions - corporal punishment, in-school suspension, out-of-school
suspension, expulsion with educational services, expulsion without
educational services, expulsion under zero-tolerance policies, referral
to law enforcement, school-related arrests, mechanical restraint,
physical restraint, seclusion, preschool suspension, and preschool
expulsion. However, upon further study, we discovered reasons we thought
justified excluding a number of these outcomes. In particular,
seclusion, physical restraint, and mechanical restraint are not
disciplinary actions, but are rather used as means to restrain students
who are at risk of harming themselves or others. Additionally, the
number of preschool students who are expelled or suspended is
vanishingly small (131 total expulsions and 6751 total suspensions out
of over 1.4 million enrolled preschool students), making reliably
estimating any association across counties exceedingly unlikely. We
additionally discovered that counts of one expulsion category (expulsion
under zero-tolerance policies) overlapped with counts in other
categories, and so excluded this category. To remain consistent with
previous studies, we opted to combine the remaining two expulsion
categories to yield one overall count of the number of students
expelled.

We also preregistered our explicit bias as a simple feeling thermometer
towards balcks (i.e. \emph{how warm or cold do you feel towards Blacks?
0=very cold, 10=very warm}). However, past research (Hehman, Flake, \&
Calanchini, 2017; Leitner et al., 2016) has used the difference in
reported warmth towards whites and blacks, and so in the main text, we
report models using this metric of explicit bias. Additionally, we
preregistered analyses with poststratified estimates (as presented in
the main text) along with raw, county-based means. Finally, we had not
known about the issues with juvenile justice facilities, or with the
school districts with reporting errors. Here, we present the results of
the preregistered analyses exactly.

\section{Simple county means}\label{simple-county-means}

\section{Post stratified estimates}\label{post-stratified-estimates}

\section{Teacher analyses}\label{teacher-analyses}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-bates2015fitting}{}
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting linear
mixed-effects models using lme4. \emph{Journal of Statistical Software},
\emph{67}(1), 1--48.
doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}

\hypertarget{ref-gelman1997poststratification}{}
Gelman, A., \& Little, T. C. (1997). Poststratification into many
categories using hierarchical logistic regression. \emph{Survey
Methodology}, \emph{23}(2), 127--35.

\hypertarget{ref-gonsalkorale2009aging}{}
Gonsalkorale, K., Sherman, J. W., \& Klauer, K. C. (2009). Aging and
prejudice: Diminished regulation of automatic race bias among older
adults. \emph{Journal of Experimental Social Psychology}, \emph{45}(2),
410--414.

\hypertarget{ref-hehman2017disproportionate}{}
Hehman, E., Flake, J. K., \& Calanchini, J. (2017). Disproportionate use
of lethal force in policing is associated with regional racial biases of
residents. \emph{Social Psychological and Personality Science},
1948550617711229.

\hypertarget{ref-leitner2016blacks}{}
Leitner, J. B., Hehman, E., Ayduk, O., \& Mendoza-Denton, R. (2016).
Blacks' death rate due to circulatory diseases is positively related to
whites' explicit racial bias: A nationwide investigation using project
implicit. \emph{Psychological Science}, \emph{27}(10), 1299--1311.

\hypertarget{ref-losen2015we}{}
Losen, D. J., Hodson, C. L., Keith, I., Michael, A., Morrison, K., \&
Belway, S. (2015). Are we closing the school discipline gap?

\hypertarget{ref-miroshnikov2014parallel}{}
Miroshnikov, A., \& Conlon, E. (2014). \emph{ParallelMCMCcombine:
Methods for combining independent subset markov chain monte carlo (mcmc)
posterior samples to estimate a posterior density given the full data
set}. Retrieved from
\url{https://CRAN.R-project.org/package=parallelMCMCcombine}

\hypertarget{ref-park2004bayesian}{}
Park, D. K., Gelman, A., \& Bafumi, J. (2004). Bayesian multilevel
estimation with poststratification: State-level estimates from national
polls. \emph{Political Analysis}, \emph{12}(4), 375--385.

\hypertarget{ref-rcitation}{}
R Core Team. (2016). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\hypertarget{ref-scott2016bayes}{}
Scott, S. L., Blocker, A. W., Bonassi, F. V., Chipman, H. A., George, E.
I., \& McCulloch, R. E. (2016). Bayes and big data: The consensus monte
carlo algorithm. \emph{International Journal of Management Science and
Engineering Management}, \emph{11}(2), 78--88.

\hypertarget{ref-rstanarm2016}{}
Stan Development Team. (2016). Rstanarm: Bayesian applied regression
modeling via Stan. Retrieved from \url{http://mc-stan.org/}

\hypertarget{ref-wickham2009ggplot2}{}
Wickham, H. (2009). \emph{Ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}

\hypertarget{ref-wickham2017tidyr}{}
Wickham, H., \& Henry, L. (2017). \emph{Tidyr: Easily tidy data with
'spread()' and 'gather()' functions}. Retrieved from
\url{https://CRAN.R-project.org/package=tidyr}

\hypertarget{ref-wickham2017dplyr}{}
Wickham, H., Francois, R., Henry, L., \& Müller, K. (2017). \emph{Dplyr:
A grammar of data manipulation}. Retrieved from
\url{https://CRAN.R-project.org/package=dplyr}

\hypertarget{ref-xu2014psychology}{}
Xu, K., Nosek, B., \& Greenwald, A. (2014). Psychology data from the
race implicit association test on the project implicit demo website.
\emph{Journal of Open Psychology Data}, \emph{2}(1).

\end{document}

### 2018-02-26 talk_plots.R

@@ -1041,8 +1041,7 @@ ggplot(plot.dat, aes(x=weighted_bias, y=est, group=group)) +
         axis.title.x = element_text(size=14),
         axis.text = element_text(size=12),
         legend.text = element_text(size=12),
-        strip.text = element_text(size=12))
-  -> plot
+        strip.text = element_text(size=12)) -> p1
 
 posterior_distribution %>%
   select(group, weighted_warmth, sample, value) %>%
@@ -1119,7 +1118,7 @@ ggplot(plot.dat, aes(x=weighted_bias, y=est)) +
         legend.text = element_text(size=14)) +
   facet_wrap(~bias)
 
-plot$data %>% filter(weighted_bias%in%c(0,1)) %>% mutate(students = est*8000000)
+p1$data %>% filter(weighted_bias%in%c(0,1)) %>% mutate(students = est*8000000)
 
 i <- 'expulsion_combined'
 pth <- paste(base_path, i, '/', sep='')

### 2018-03-05 aggregate_data.R
additional covariates

@@ -313,9 +313,10 @@ schools_loc$county_name[which(schools_loc$county_id=='VA-510')] <- 'Alexandria c
 county_means %>%
   select(-X) %>%
   right_join(schools_loc) %>%
-  select(county_id, county_name, state_abb, bias, explicit, explicit_diff, 
-         weighted_bias, weighted_explicit, weighted_explicit_diff, 
-         COMBOKEY, group, number, total_number, metric, LEAID) -> tempout
+  select(county_id, county_name, state_abb, state_fips, county_fips, bias, 
+         explicit, explicit_diff, weighted_bias, weighted_explicit, 
+         weighted_explicit_diff, COMBOKEY, group, number, total_number, metric, 
+         LEAID) -> tempout
 
 ################
 # Exclude schools #
@@ -364,32 +365,67 @@ tempout %>%
 
 df_acs <- read.csv('/Users/travis/Documents/gits/Data/ACS/county_ethnicity/ACS_14_5YR_B02001_with_ann.csv',skip = 1)
 
-covs1 <- df_acs[,c(3, 4, 6, 8)]
-names(covs1) <- c('county', 'total_pop', 'white_pop', 'black_pop')
+covs_pop <- df_acs[,c(3, 4, 6, 8)]
+names(covs_pop) <- c('county', 'total_pop', 'white_pop', 'black_pop')
 
 df_acs <- read.csv('/Users/travis/Documents/gits/Data/ACS/county_poverty_emp/ACS_14_5YR_DP03_with_ann.csv',skip = 1)
 
-covs2 <- df_acs[,c(3, 22, 248, 478)]
-names(covs2) <- c('county', 'unemp_rate', 'med_income', 'poverty_rate')
+covs_emp <- df_acs[,c(3, 22, 248, 478)]
+names(covs_emp) <- c('county', 'unemp_rate', 'med_income', 'poverty_rate')
 
 df_acs <- read.csv('/Users/travis/Documents/gits/Data/ACS/county_education/ACS_14_5YR_S1501_with_ann.csv',skip=1)
-covs3 <- df_acs[,c(3, 28)]
-names(covs3) <- c('county', 'col_grads')
+covs_ed <- df_acs[,c(3, 28)]
+names(covs_ed) <- c('county', 'col_grads')
 
-covs1 %>%
-  left_join(covs2) %>%
-  left_join(covs3) %>%
[…]

### 2018-03-05 iat_analysis.R

@@ -22,8 +22,10 @@ df8 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2004.sav'
 df9 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2002-2003.sav')
 df10 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2014.sav')
 df_acs <- read.csv('/Users/travis/Documents/gits/Data/ACS/county_age/ACS_14_5YR_DP05_with_ann.csv', skip=1)
+df_haley <- read.csv('/Users/travis/Documents/gits/Data/Haley_countylinks/_Master Spreadsheet.csv', stringsAsFactors = F)
 df_county_linking_info <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_linking_table.csv', stringsAsFactors = F)
 df_county_linking_info$county_name[1904] <- 'Doña Ana'
+df_haley[1791,6] <- 'Doña Ana County, New Mexico'
 df_states <- data.frame(state=c(state.name, 'District of Columbia'),
                         state_abb=c(state.abb, 'DC'))
 df_acs_eth <- read.csv('/Users/travis/Documents/gits/Data/ACS/state_ethnicity/ACS_14_5YR_B02001_with_ann.csv',
@@ -32,6 +34,12 @@ df_acs_ed <- read.csv('/Users/travis/Documents/gits/Data/ACS/state_education/ACS
                       skip = 1, stringsAsFactors = F)
 df_acs_pov_emp <- read.csv('/Users/travis/Documents/gits/Data/ACS/state_poverty_emp/ACS_14_5YR_DP03_with_ann.csv',
                            skip = 1, stringsAsFactors = F)
+df_acs_hous <- read.csv('/Users/travis/Documents/gits/Data/ACS/state_housing/DEC_10_SF1_GCTPH1.US04PR_with_ann.csv',
+                        skip=1, stringsAsFactors = F)
+df_acs_mob <- read.csv('/Users/travis/Documents/gits/Data/ACS/state_mobility/ACS_14_5YR_S0701_with_ann.csv',
+                       skip=1, stringsAsFactors = F)
+df_fbi <- read.csv('/Users/travis/Documents/gits/Data/FBI/state_crimes/CrimeTrendsInOneVar.csv')
+
 
 #educators <- c('25-2000', '25-3000', '25-4000', '25-9000')
[…]

### 2018-03-05 written_docs/draft.Rmd
partial update w/stacey feedback

@@
 ## Data Sources
+### Disciplinary actions
 ```{r summary-data}
@@
 ```
-We used three distinct data sources for the work described here.
-The academic disciplinary data is part of the Civil Rights Data Collection (CRDC) through the US Department of Education.
+To assess rates of disciplinary action, we used data from the Civil Rights Data Collection (CRDC) conducted by the US Department of Education.
 The dataset we used comes from the 2013-2014 academic year and has data on "all public local and educational agencies and schools, including long-term secure juvenile justice facilities, charter schools, alternative schools, and schools serving students with disabilities." In total, the CRDC data represents `r dim(schools_dat)[1]` institutions enrolling approximately 50 million students, of which approximately 25 million are white and 7.8 million are Black[^1].
@@
 After these exclusions are applied, the final sample used for modeling consists of `r enrollment_totals$total_schools[1]` institutions, enrolling `r round(sum(enrollment_totals$ns)/1000000, 1)` million black or white students, of which `r  round(enrollment_totals$ns[2]/1000000)` million are white and `r  round(enrollment_totals$ns[1]/1000000)` million are black.
+From these data, we focus on the number of students by race (black and white) who received one of several types of disciplinary action.
+We report here rates of out-of-school suspension, in-school suspension, school-related arrests, law enforcement referrals, and total number of expulsions of any type.
 
@@
 
-We obtained county-level demographic information for use as covariates and state-level demographic information for use in post-stratification from the US Census Bureau.
-For both county-level and state-level demographics, we use 5-year estimates for the period ending in 2014 from the American Community Survey, which surveys around 295k households per month.
-
+### Racial Bias
 ```{r iat-data}
@@
 We used measurements of implicit and explicit bias available from data collected through Project Implicit [@xu2014psychology].
+For a full description of the implicit and explicit bias measures available in these data, refer to @xu2014psychology and @leitner2016blacks.
 From this data, we used only respondents who had geographic information that would allow us to place them in a United States county, identified as White, and visited the site before 2015.
@@
 
+Because Project Implicit is a nonrandom sample, we used multilevel regression and post-stratification to obtain accurate geographical population-based estimates of implicit and explicit bias.
+Multilevel regression and post-stratification corrects for biased sampling and regularizes extreme observations with little data to support them (e.g.
+a county with only a handful of respondents with especially high or low scores) [@park2004bayesian; @gelman1997poststratification].
+Following past work [@leitner2016blacks], we identified age as one dimension along which IAT respondents differed from the general population in ways that could bias our conclusions [@gonsalkorale2009aging].
+Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories (15-24, 25-34, 35-54, 55-75, and 75+).
+We next fit multilevel models estimating bias (implicit and explicit biases seperately) as a function of our state-level covariates (the "fixed" effects: proportion of population that is white; proportion of the population that is black, the ratio of black to white people in the population, the percentage of college graduates, the unemployment rate, the median household income, and the poverty rate), and allowed the estimates to vary by age bin, county, and state (the "random" effects).
+Next, we determined the population of whites in each county in these age groups using the American Community Survey's 5-year estimates ending in 2014.
+Finally, we used our estimated models to predict the expected response for each age bin, in each county.
+Our final county-level estimates are the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county.
+As a result of this procedure, we can be confident that our county-level estimates should more closely approximate what our estimates would look like if the Project Implicit data were truly representative along the age dimension in all counties.
+
+### Covariates
+All county-level demographic information for use as covariates and state-level demographic information for use in post-stratification were taken from the US Census Bureau.
+For both county-level and state-level demographics, we use 5-year estimates for the period ending in 2014 from the American Community Survey, which surveys around 295k households per month.
+We additionally took an estimate of violent crime from the FBI Uniform Crime Reporting program, as made available through the National Archive of Criminal Justice Data.
+
 ## Measures
-The primary outcome in this analysis is a count of the number of students by race (black and white) who received one of several types of disciplinary action.
-We report here rates of out-of-school suspension, in-school suspension, school-related arrests, law enforcement referrals, and total number of expulsions of any type.
 
@@
 
-Finally, for the implicit and explicit bias measures, we relied on two primary variables from the Project Implicit data.
-Implicit bias was assessed via an Implicit Association Test (IAT).
-This test uses a speeded dual-categorization task in which individuals must quickly categorize black and white faces and "good" and "bad" words with key presses.
-The difference in how quickly and accurately participants are able to pair white faces with "good" words and black faces with "bad" words in comparison to the inverse is thought to reflect implicit associations between the two races and positive and negative affective reactions.
-This association is indicated in the IAT D-score, which we used as a measure of implicit bias.
-Our measure of explicit bias is the difference between reported warmth towards whites (i.e.
-*how warm or cold do you feel towards Whites?
-0=very cold, 10=very warm*) and reported warmth toward blacks.
-
 ## Data analysis
 Data analysis proceded in two steps.
-We first estimated county-level implict and explicit bias using multilevel regression and post-stratification.
-Post-stratification is a valuable procedure in obtaining accurate geographical population-based estimates because it allows a non-representative sample (e.g Project Implicit) to more closely resemble the true population, and it regularizes extreme observations with little data to support them (e.g.
-a county with only a handful of respondents with especially high or low scores) [@park2004bayesian; @gelman1997poststratification].
-Following past work [@leitner2016blacks], we identified age as one dimension along which IAT respondents differed from the general population in ways that could bias our conclusions [@gonsalkorale2009aging].
-Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories (15-24, 25-34, 35-54, 55-75, and 75+).
-We next fit multilevel models estimating bias (implicit and explicit biases seperately) as a function of our state-level covariates (the "fixed" effects), and allowed the estimates to vary by age bin, county, and state (the "random" effects).
-Next, we determined the population of whites in each county in these age groups using the American Community Survey's 5-year estimates ending in 2014.
-Finally, we used our estimated models to predict the expected response for each age bin, in each county.
-Our final county-level estimates are the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county.
-As a result of this procedure, we can be confident that our county-level estimates should more closely approximate what our estimates would look like if the Project Implicit data were truly representative along the age dimension in all counties.
 
@@
 #the p_neg for oos_susp & implicit bias is 100%, so when reporting the number in-text, i just replace it with 1 to prevent it from being rendered as NA in the document.
-base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/mw_uclaexcl_diff/'
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/mw_uclaexcl_diff_extracovs/'
 plot.dat <- data.frame(group = NA, weighted_bias = NA, est = NA, lower = NA, upper=NA, metric = NA)
@@
   j <- 1
-  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+  posterior_combo <- array(0, dim=c(16,4000,length(files)))
   raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
@@
       filter(!is.na(county_id)) -> raw_rate
-    posterior_combo[,,j] <- t(df[,c(1:14)])
+    posterior_combo[,,j] <- t(df[,c(1:16)])
     j <- j+1
@@
 
-  for (p in 1:14){
+  for (p in 1:16){
     m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
   }
-  for (p in 1:14){
+  for (p in 1:16){
     m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
   }
-  for (p in 1:14){
+  for (p in 1:16){
     m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
   }
-  for (p in 1:14){
+  for (p in 1:16){
     m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
@@
 
-  data.frame(est=mean(p_cons[13,]),
-             lower=quantile(p_cons[13,], .025),
-             upper=quantile(p_cons[13,], .975),
-             p_neg = prop.table(table(p_cons[13,]<0))[2],
+  data.frame(est=mean(p_cons[15,]),
+             lower=quantile(p_cons[15,], .025),
+             upper=quantile(p_cons[15,], .975),
+             p_neg = prop.table(table(p_cons[15,]<0))[2],
              metric=i) %>%
@@
 #weighted warmth w/exclusions & explicit difference
-plot.dat <- data.frame(group = NA, weighted_warmth = NA, est = NA, lower = NA, upper=NA, metric = NA)
-plot.dat2 <- data.frame(weighted_warmth = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat <- data.frame(group = NA, weighted_explicit_diff = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(weighted_explicit_diff = NA, est = NA, lower=NA, upper=NA, metric=NA)
 plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg = NA, metric=NA)
@@
   j <- 1
-  posterior_combo <- array(0, dim=c(14,4000,length(files)))
-  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, weighted_warmth=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  posterior_combo <- array(0, dim=c(16,4000,length(files)))
+  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, weighted_explicit_diff=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
   for (k in files){
@@
     n_schools %>%
-      left_join(m$data[, c('county_id', 'weighted_warmth')]) %>%
+      left_join(m$data[, c('county_id', 'weighted_explicit_diff')]) %>%
       distinct() %>%
@@
       filter(!is.na(county_id)) -> raw_rate_warmth
-    posterior_combo[,,j] <- t(df[,c(1:14)])
+    posterior_combo[,,j] <- t(df[,c(1:16)])
     j <- j+1
@@
 
-  data.frame(est=mean(p_cons[14,]),
-             lower=quantile(p_cons[14,], .025),
-             upper=quantile(p_cons[14,], .975),
-             p_neg = prop.table(table(p_cons[14,]<0))[2],
+  data.frame(est=mean(p_cons[16,]),
+             lower=quantile(p_cons[16,], .025),
+             upper=quantile(p_cons[16,], .975),
+             p_neg = prop.table(table(p_cons[16,]<0))[2],
              metric=i) %>%

### 2018-03-05 data_exploration.r
delete old files

@@ -1,1399 +1,0 @@
-library(dplyr)
-library(tidyr)
-library(forcats)
-library(ggplot2)
-library(rstanarm)
-library(lme4)
-
-options(mc.cores = parallel::detectCores())
-
-district_content <- read.csv('../Data/crdc201314csv/CRDC2013_14_LEA_content.csv')
-df_district <- read.csv('../Data/crdc201314csv/CRDC2013_14_LEA.csv')
-school_content <- read.csv('../Data/crdc201314csv/CRDC2013_14_SCH_content.csv')
-df_school <- read.csv('../Data/crdc201314csv/CRDC2013_14_SCH.csv')
-
-
-# Get enrollment figures
-df_school %>%
-  select(COMBOKEY, LEA_STATE:SCH_NAME, SCH_ENR_HI_M:TOT_ENR_F) %>%
-  gather(group, total_number, SCH_ENR_HI_M:TOT_ENR_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>%
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='NR_', 
-                          biracial='TR_', white='WH_')) %>%
-  select(-prefix) -> enrollment
-
-
-###### corporal punishment
-df_school %>% 
-  select(COMBOKEY, LEA_STATE:SCH_NAME, SCH_CORPINSTANCES_IND:TOT_DISCWODIS_CORP_F) %>% 
-  filter(SCH_CORPINSTANCES_IND=='YES') %>%
-  gather(group, number, SCH_DISCWODIS_CORP_HI_M:TOT_DISCWODIS_CORP_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>%
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='RP_', 
-                          biracial='TR_', white='WH_')) %>%
-  select(-prefix) -> corporal
-
-corporal %>%
-  left_join(enrollment) %>%
-  group_by(COMBOKEY, LEA_STATE, LEA_NAME, SCH_NAME, group) %>%
-  summarise(number = sum(number),
-            total_number = sum(total_number)) %>%
[…]

### 2018-03-05 data_subsetter.r

@@ -1,10 +1,0 @@
-library(dplyr)
-
-df <- read.csv('../output/full_model_data.csv')
-
-df %>%
-  sample_n(20000) %>%
-  group_by(COMBOKEY) %>% 
-  mutate(group = sample(c(1:5), 1)) -> subdat
-
-write.csv(subdat, file='big_bayes_testingdata.csv', row.names = F)

### 2018-03-13 consensusMCMC_model_prep.R
Adjust to handle one data file + write models for extra covs

@@ -196,31 +196,31 @@ library(rstanarm)
 
 options(mc.cores = parallel::detectCores())
 
-df <- read.csv('/home/triddle/educational_disparities/cluster/data/subcounty_grouping_diff.csv')
+df <- read.csv('/home/triddle/educational_disparities/cluster/data/county_grouping.csv')
 
-df %>% filter(metric=='"
+df %>% filter(exclude==FALSE) %>% filter(metric=='"
 
-string2 <- "') %>% select(county_id, bias:weighted_warmth, total_pop, unemp_rate:b.w.ratio) %>% distinct() %>% mutate_at(vars(-county_id), scale) -> scaled_county_level
+string2 <- "') %>% select(county_id, bias:weighted_explicit_diff) %>% distinct() %>% mutate_at(vars(-county_id), scale) -> scaled_county_level
 
 df %>% filter(grouping=="
 
-string3 <- ") %>% filter(metric=='"
-string4 <- "') %>% select(county_id, COMBOKEY:metric) %>% left_join(scaled_county_level) -> mod.dat
+string3 <- ") %>% filter(exclude==FALSE) %>% filter(metric=='"
+string4 <- "') %>% select(county_id, total_pop:poverty_rate, crime_rate:b.w.ratio, COMBOKEY:metric) %>% left_join(scaled_county_level) -> mod.dat
 
 m <- stan_glmer(cbind(number, total_number-number) ~ group + weighted_bias + 
-weighted_warmth + group:weighted_bias + group:weighted_warmth + total_pop + 
+weighted_explicit_diff + group:weighted_bias + group:weighted_explicit_diff + total_pop + 
 unemp_rate + med_income + poverty_rate + col_grads + white_prop + 
-black_prop + b.w.ratio + 
+black_prop + b.w.ratio + mobility + crime_rate + density + 
 (group|county_id), data=mod.dat, prior = normal(0,5), 
 prior_intercept = normal(0,5), family=binomial, adapt_delta=.99)
 
-save(m, file='/tigress/triddle/educational_disparities/mw_uclaexcl_diff/"
+save(m, file='/tigress/triddle/educational_disparities/mw_uclaexcl_diff_extra_covs/"
 
 
-for(i in rownames(table(dfss$metric))){
-  for (j in 1:12){
+for(i in rownames(table(dfs$metric))){
+  for (j in 1:14){
[…]

### 2018-03-13 written_docs/draft.Rmd
Intro! plus rewrite of results & methods

--- 2018-02-05 written_docs/draft.Rmd
@@
 # Introduction
+In comparison to White Americans, African Americans exhibit poorer educational outcomes across a range of metrics.
+One outcome of particular concern is the gap in disciplinary actions [@kinsler2011understanding; @okonofua2016vicious].
+Research using administrative datasets and longitudinal samples clearly show that African American students are far more likely to be suspended or expelled [@aud2011america; @yeager2017loss], and conditional on an office referral, are more likely to receive stiffer punishments [@gregory1995crime].
+Concerns about these disparities are exacerbated by their associations with long-term life outcomes, including prospects for employment [@pager2009sequencing], mental and physical health [@pascoe2009perceived], and involvement in the criminal justice system [@hirschfield2009another].
+In this paper, we aim to establish that these disciplinary gaps are related to environmental levels of bias in predictable ways.
+
+It is perhaps not surprising that levels of bias would be associated with racial disciplinary gaps.
+This straightforward statement, however, is contextualized in a number of ways.
+First, the operationalization of bias should play a role in the association.
+As we describe below, levels of implicit and explicit biases should be differentially associated with disciplinary actions.
+This differential association should be determined, in part, by the type of disciplinary action under examination.
+For a given student committing a specific offense, the educational institution may opt for one of several disciplinary actions.
+These actions vary in severity, institutional costs, and the degree of formal administrative involvement.
+
+To illustrate this problem in the context of previous research, we now describe research on the associations between implicit bias and explicit bias and decision making.
+This sets the stage for understanding what types of decisions administrators and educators make when electing a specific disciplinary action.
+We then describe the types of disciplinary actions that are taken and the characteristics of these actions in terms of severity, institutional costs, and formal administrative involvement.
+Finally, in crossing this information, we concretely specify the set of associations expected when examining bias and disciplinary actions in American schools.
+
+## Bias and decision-making
+Previous research on biased attitudes and behavior has shown that although implicit and explicit attitudes are typically correlated [@mcconnell2001relations; @hofmann2005meta], they tend to behaviorally manifest slightly differently.
+As a default, a naiive expectation is that explicit attitudes should guide behavior, since if people tell us they like something, we should expect that to generally correspond to their behavior.
+However, under some circumstances, implicit biases are more strongly associated with behavioral outcomes.
+@nosek2011implicit suggest that, in comparison to explicit attitudes, implicit attitudes are more strongly associated with behavior when the individual lacks motivated to explicitly respond differently than their implicit attitude would suggest, when they do not have the opportunity or ability to modify their initial response inclination, or when they do not have awareness of the implicit response, the behavior, or the link between them.
+In contrast, the more motivated, more able, or more aware the person is over their behavioral response, the more explicit biases should play a role.
+One key situational factor that determines the response is situational uncertainty.
+In uncertain situations, the respondent is not likely to have strong motivations to change their initial response, otherwise the situation would not be uncertain.
+Research supports this idea, as when the situation is characterized by greater uncertainty, implicit biases are known to be stronger determinants of behavior [@friese2009and; @gawronski2003implicit; @galdi2008automatic].
+In short, when taking (or not taking) an action can be attributed to multiple reasons due to situational ambiguity, then implicit biases should result in higher rates of discriminatory behavior.
+
+Accordingly, this research suggests that disciplinary actions that are implemented quickly, have relatively little administrative oversight, and require little in the way of insitutional resources should be characterized by strong associations with implicit bias.
+On the other hand, any disciplinary action that requires higher levels of oversight, resources, or lengthy time investments to complete should be more strongly associated with explicit bias.
+
+## Types of disciplinary actions
+We report here on five different types of disciplinary actions: in-school suspensions, out-of-school suspensions, law enforcement referrals, school-related arrests, and expulsions.
+As described above, these actions vary in severity, required resources, and administrative oversight.
+
+The most extreme outcome, expulsion, is defined as when a student is prohibited from ever returning to the educational institution.
+The institution may or may not set up alternative educational services for the student.
+Because of the the severe nature of this punishment, it also requires considerable administrative oversight, and is often conferred for behaviors that are relatively unambiguous (e.g.
+bringing a gun to school).
+We believe similar cases can be made for law enforcement referrals and school-related arrests insofar as these actions tend to be more severe, and require resources, or occasionally the involvement of outside agencies.
+
+Somewhat less severe than expulsions, but one of the most frequently utilized disciplinary actions, out-of-school suspensions occupy something of a middle ground.
+There is a substantial body of research documenting the negative associations between this particular disciplinary action and life outcome [@cuellar2015school; @matjasko2011effective], as well as their disproportionate use on stigmatized students [@haight2014ecological; @kinsler2011understanding].
+Out-of-school suspensions are administered for a wide range of behaviors.
+Some of these behaviors are apt to be determined with a fair degree of subjectivity (e.g.
+defiance) [@gregory2008discipline], while others may be more unambiguous (e.g.
+fighting, drug use).
+In comparison to expulsions and arrests, this type of action demands less administrative oversight.
+
+In-school suspensions, wherein the student is removed from the classroom but remains under the direct supervision of school personnel, are also administered for a wide range of behaviors.
+However, in-school suspensions have been highlighted as a promising alternative to more exclusionary types of disciplinary actions (e.g.
+out-of-school suspensions, expulsions) partly because students remain under the supervision of school personnel, which keeps them in a structured, educational setting [@anyon2014persistent].
+Like out-of-school suspensions, this type of action requires relatively little administrative oversight.
+But unlike out-of-school suspension, removing a student from a classroom to a *different* educational setting requires additional space and personnel, and so this represents an inherent investment in the student.
+
+## Concrete predictions of association
+Given these disciplinary outcomes and what is known about implicit and explicit attitudes, what should we expect in examining associations between them?
+First, it's clear that we should expect African American students to receive higher levels of disciplinary actions.
+Second, we should expect the disparity in disciplinary actions to generally be positively associated with explicit bias.
+Finally, we should expect implicit bias to only show associations in instances where there is little administrative overhead, and where the behaviors to-be-punished are often ambiguous.
+By this logic, out-of-school suspensions are the only metric that meet these criteria.
+
+## The present study
+The present study uses a national dataset of disciplinary outcomes to evaluate the degree to which implicit and explicit bias are associated with each of these outcomes.
+After an initial, registered analytic approach, we subsequently made several adjustments to strengthen our analyses and streamlined them with respect to previous research.
+In general, the data support the conclusion that implicit biases are specifically associated with out-of-school suspensions, while explicit biases tend to be associated with disciplinary actions of all types.
 
@@
 library(tidyr)
-county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv', 
-                         colClasses = 'character')
+library(ggplot2)
+county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv')
 schools_dat <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')
@@
 ```
-We used three distinct data sources for the work described here.
-The academic disciplinary data is part of the Civil Rights Data Collection (CRDC) through the US Department of Education.
+To assess rates of disciplinary action, we used data from the Civil Rights Data Collection (CRDC) conducted by the US Department of Education.
 The dataset we used comes from the 2013-2014 academic year and has data on "all public local and educational agencies and schools, including long-term secure juvenile justice facilities, charter schools, alternative schools, and schools serving students with disabilities." In total, the CRDC data represents `r dim(schools_dat)[1]` institutions enrolling approximately 50 million students, of which approximately 25 million are white and 7.8 million are Black[^1].
 Previous work using these data have identified a number of districts whose data are in error, and have excluded juvenile justice facilities, as these institutions constitute dramatically different educational environments, where the meaning of disciplinary actions may be quite different [@losen2015we].
-After these exclusions are applied, the final sample used for modeling consists of `r enrollment_totals$total_schools[1]` institutions, enrolling `r round(sum(enrollment_totals$ns)/1000000, 1)` million black or white students, of which `r  round(enrollment_totals$ns[2]/1000000, 1)` million are white and `r  round(enrollment_totals$ns[1]/1000000, 1)` million are black.
+After these exclusions are applied, the final sample used for modeling consists of `r enrollment_totals$total_schools[1]` institutions, enrolling `r round(sum(enrollment_totals$ns)/1000000, 1)` million black or white students, of which `r  round(enrollment_totals$ns[2]/1000000,1)` million are white and `r  round(enrollment_totals$ns[1]/1000000, 1)` million are black.
+From these data, we focus on the number of students by race (black and white) who received one of several types of disciplinary action.
+We report here rates of out-of-school suspension, in-school suspension, school-related arrests, law enforcement referrals, and total number of expulsions of any type.
 
@@
 We used measurements of implicit and explicit bias available from data collected through Project Implicit [@xu2014psychology].
-From this data, we used only respondents who had geographic information that would allow us to place them in a United States county, identified as White, and visited the site before 2015.
+For a full description of the implicit and explicit bias measures available in these data, refer to @xu2014psychology and @leitner2016blacks.
+We note only that we used the IAT D-score as a measure of implicit bias, and the difference between reported warmth towards whites and warmth towards black (both measured from 0=very cold, 10=very warm) as a measure of explicit bias.
+Additionally, we used only respondents who had geographic information that would allow us to place them in a United States county, identified as White, and visited the site before 2015.
 This consisted of approximately 1.1 million total respondents from `r dim(county_means)[1]` counties.
 
+### Covariates
+Each county-level variable used as a covariates in the final model and the corresponding state-level variable used as a predictor in the post-stratification scheme were taken from the from the same source.
+Population size and proportions, socioeconomic indicators, and mobility were all taken from the American Community Survey (ACS) 5-year estimates for the time period ending in 2014.
+Urban-rural indicators were taken from the 2010 US Census, and crime rates were taken from the FBI Uniform Crime Reporting program, as made available through the National Archive of Criminal Justice Data for each year from 2010-2014.
+Each of these variables is described below.
+
+#### Population size and proportions
+We obtained the total population, the proportion of the population that is white, the proportion of the population that is black, and the ratio of black-to-white people in the population from ACS table B02001.
+
+#### Socioeconomic indicators
+We obtained estimates for the percentage of population with a Bachelor's degree or higher, the percentage of the population aged 16 or over in the labor force that is unemployed, the median household income, and the percentage of families and people whose income in the last year was below the poverty line from the ACS table DP03.
+
+#### Urban-rural indicator
+We obtained estimates of housing density per square mile of land area from Census table GCT-PH1.
+
+#### Mobility
+We obtained estimates of population mobility by summing the percentage of African Americans who moved from a different county, state, or country into the county of interest (county-level covariate) or who moved from a different state or country into the state of interest (state-level covariate).
+We took these metrics from the ACS table S0701.
+
+#### Crime
+We computed estimates of the number of violent crimes per person by taking the number of crimes reported divided by the population size for each year, and averaging the resulting proportions across the 5 years of data.
+
+# Results
+
+## Data analysis
+Because Project Implicit is a nonrandom sample, we used multilevel regression and post-stratification to obtain accurate geographical population-based estimates of implicit and explicit bias.
+Multilevel regression and post-stratification corrects for biased sampling and regularizes extreme observations with little data to support them (e.g.
+a county with only a handful of respondents with especially high or low scores) [@park2004bayesian; @gelman1997poststratification].
+Following past work [@leitner2016blacks], we identified age as one dimension along which IAT respondents differed from the general population in ways that could bias our conclusions [@gonsalkorale2009aging].
+Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories (15-24, 25-34, 35-54, 55-75, and 75+).
+We next fit multilevel models estimating bias (implicit and explicit biases seperately) as a function of our state-level covariates (the "fixed" effects: proportion of population that is white; proportion of the population that is black, the ratio of black to white people in the population, the percentage of college graduates, the unemployment rate, the median household income, the poverty rate, housing density, mobility rate, & crime rate; each of these is described in more detail below), and allowed the estimates to vary by age bin, county, and state (the "random" effects).
+Next, we determined the population of whites in each county in these age groups using the American Community Survey's 5-year estimates ending in 2014.
+Finally, we used our estimated models to predict the expected response for each age bin, in each county.
+Our final county-level estimates are the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county.
+As a result of this procedure, we can be confident that our county-level estimates should more closely approximate what our estimates would look like if the Project Implicit data were truly representative along the age dimension in all counties.
+
+We analyzed these data using a series of bayesian multilevel logistic regressions.
+We modeled the probability that a student would be expelled as a function of a set of effects that are constant across observations (i.e.
+fixed effects: race (dummy coded), implicit bias, explicit bias, an interaction term between race and implicit bias, an interaction term between race and explicit bias, and all covariates described above) and a set of effects that vary across counties (i.e.
+random effects: overall intercept & race).
+We fit separate models for each of the outcomes.
+All numerical predictors were standardized at the appropriate level (county, state) before model estimation (for post-stratification as well as for final inference) to help with estimation efficiency and interpretability.
+We also set priors for the intercept and coefficients in the bayesian model to be weakly informative normal distributions centered on zero with a standard deviation of five.
+This corresponds to a prior belief of all parameters taking values between -15 and 15.
+Values outside this range are extremely unlikely given that all variables were standardized prior to estimation.
+All other parameters were left to default values.
+
+Because of the computational demands of fitting such a high-dimensional model to such a large dataset (the full model for each metric would consist of over 6k parameters to approximately 170k observations), we used a consensus monte carlo algorithm to obtain approximate posterior distributions for the parameters of interest [@scott2016bayes].
+The approximate posteriors derived from this algorithm have been shown to be nearly indistinguishable from the true posterior, a result we verified using a small subset of our own data.
+
+## Project Implicit Estimates
+```{r exp-diff}
+individual_data %>%
+  mutate(explicit_bias_diff = twhite_0to10-tblack_0to10) -> individual_data
+```
+We first report the results of estimating the implicit and explicit biases in from Project Implicit data.
+When examining the unstandardized county-level estimates adjusted with poststratification, we see that there is a pro-white bias in both implicit ($mean$ = `r round(mean(as.numeric(county_means$weighted_bias), na.rm=T), 2)`, $sd$ = `r round(sd(as.numeric(county_means$weighted_bias), na.rm=T), 2)`) and explicit measures ($mean$ = `r round(mean(as.numeric(county_means$weighted_explicit_diff), na.rm=T), 2)`, $sd$ = `r round(sd(as.numeric(county_means$weighted_explicit_diff), na.rm=T), 2)`, where on both scales 0 = no bias, and positive numbers indicate a pro-white bias).
+
+## Disciplinary action frequency
+
+```{r disc-count, results='asis'}
+model.data %>% 
+  filter(exclude==FALSE) %>%
+  group_by(group, metric) %>% 
+  summarise(percentage = sum(number)/sum(total_number)) %>%
+  mutate(percentage = percentage*100) %>%
+  mutate(percentage = paste(round(percentage, 2), '%', sep='')) %>%
+  spread(group, percentage) %>%
+  mutate(metric = fct_recode(metric, 
+                             'expulsions' = 'expulsion_combined',
+                             'school arrests' = 'in_school_arrest',
+                             'in-school suspension' = 'inschool_susp',
+                             'law enforcement referral' = 'law_enforcement',
+                             'out-of-school suspension' = 'oos_susp')) -> counts_table
+
+papaja::apa_table(counts_table, caption = 'Percentage of students of each race receiving each type of disciplinary action')
+```
+
+Table \@ref(tab:disc-count) shows the percentage of students of each race who were reported having received each of the actions under consideration.
+Our statistical models (described below) show that these differences are extremely unlikely to be due to chance.
+
+## Associations across counties
+Figure \@ref(fig:overall-associations) shows the estimate of primary interest for each of the models.
+The estimates displayed are the coefficients for the interaction between race and each of the two bias measurements.
+Given that African Americans are the baseline group, negative values for this coefficient indicate that as one moves into counties with higher levels of bias, the gap between the probability of a black student being disciplined and the probability of a white student being disciplined grows.
+Table \@ref(tab:reg-coefs) displays the estimated coefficients and 95% uncertainty intervals for each of the fixed effects for each model.
+
+```{r overall-associations, fig.cap="Association between each metric and county-level estimates of explicit and implicit bias.
+Negative values indicate that the rate of increase (or decrease) for blacks is faster (or slower) than for whites.
+Point is the mean of the posterior and error bars represent 95% bayesian uncertainty intervals."}
+#the p_neg for oos_susp & implicit bias is 100%, so when reporting the number in-text, i just replace it with 1 to prevent it from being rendered as NA in the document.
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/mw_uclaexcl_diff_extracovs/'
+plot.dat <- data.frame(group = NA, weighted_bias = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(weighted_bias = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg=NA, metric=NA)
+
+# weighted bias w/exclusions & explicit difference
+metrics <- list.files(base_path)
+#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(17,4000,length(files)))
+  raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    dat <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'weighted_bias')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate) %>%
+      filter(!is.na(county_id)) -> raw_rate
+    posterior_combo[,,j] <- t(dat[,c(1:17)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:17){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[16,]),
+             lower=quantile(p_cons[16,], .025),
+             upper=quantile(p_cons[16,], .975),
+             p_neg = prop.table(table(p_cons[16,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Expulsion' = 'expulsion_combined')
+plot.dat3_all <-plot.dat3
+plot.dat3_all$bias <- 'Implicit'
+
+#weighted warmth w/exclusions & explicit difference
+plot.dat <- data.frame(group = NA, weighted_explicit_diff = NA, est = NA, lower = NA, upper=NA, metric = NA)
+plot.dat2 <- data.frame(weighted_explicit_diff = NA, est = NA, lower=NA, upper=NA, metric=NA)
+plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg = NA, metric=NA)
+
+for(i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(17,4000,length(files)))
+  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, weighted_explicit_diff=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    dat <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'weighted_explicit_diff')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate_warmth) %>%
+      filter(!is.na(county_id)) -> raw_rate_warmth
+    posterior_combo[,,j] <- t(dat[,c(1:17)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:17){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(est=mean(p_cons[17,]),
+             lower=quantile(p_cons[17,], .025),
+             upper=quantile(p_cons[17,], .975),
+             p_neg = prop.table(table(p_cons[17,]<0))[2],
+             metric=i) %>%
+    rbind(plot.dat3) %>%
+    filter(!is.na(est)) %>%
+    mutate(metric=reorder(metric, est)) -> plot.dat3
+}
+
+plot.dat3$bias <- 'Explicit'
+plot.dat3$metric <- fct_recode(plot.dat3$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Expulsion' = 'expulsion_combined')
+plot.dat3_all <- rbind(plot.dat3_all, plot.dat3)
+ggplot(plot.dat3_all, aes(x=metric, y=est, group=bias, color=bias, shape=bias)) +
+  geom_point(position=position_dodge(width=.5), size=2) +
+  geom_errorbar(aes(ymin=lower, ymax=upper), 
+                width=.5, position=position_dodge(width=.5)) +
+  theme_classic() +
+  coord_flip() +
+  geom_hline(yintercept=0) +
+  ylab('Black-White Difference in log-odds slope') +
+  theme(axis.title.y = element_blank(),
+        legend.title = element_blank())
+```
+
+Several patterns are apparent from this figure.
+First, in all cases but school-related arrests, county-level explicit biases are reliably associated with disciplinary disparities, such that counties with more bias are expected to have a larger racial gap.
+The effect with the largest magnitude of these is out-of-school suspensions ($M_{slope}$ = `r round(plot.dat3_all$est[6], 2)`, [`r round(plot.dat3_all$lower[6], 2)`, `r round(plot.dat3_all$upper[6], 2)`]), followed by expulsions ($M_{slope}$ = `r round(plot.dat3_all$est[10], 2)`, [`r round(plot.dat3_all$lower[10], 2)`, `r round(plot.dat3_all$upper[10], 2)`]), law enforcement referrals ($M_{slope}$ = `r round(plot.dat3_all$est[7], 2)`, [`r round(plot.dat3_all$lower[7], 2)`, `r round(plot.dat3_all$upper[7], 2)`]), and in-school suspensions ($M_{slope}$ = `r round(plot.dat3_all$est[8], 2)`, [`r round(plot.dat3_all$lower[8], 2)`, `r round(plot.dat3_all$upper[8], 2)`]).
+The model indicates that the evidence for school-related arrests is inconclusive ($M_{slope}$ = `r round(plot.dat3_all$est[9], 2)`, [`r round(plot.dat3_all$lower[9], 2)`, `r round(plot.dat3_all$upper[9], 2)`]).
+
+Second, the effect with the largest magnitude of all is that between implicit biases and out-of-school suspension ($M_{slope}$ = `r round(plot.dat3_all$est[1], 2)`, [`r round(plot.dat3_all$lower[1], 2)`, `r round(plot.dat3_all$upper[1], 2)`]).
+All other estimated parameters are inconclusive.
+
+```{r, reg-coefs, results='asis'}
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/mw_uclaexcl_diff_extracovs/'
+plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA)
+
+# weighted bias w/exclusions & explicit difference
+metrics <- list.files(base_path)
+#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(17,4000,length(files)))
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    dat <- as.matrix(m)
+    posterior_combo[,,j] <- t(dat[,c(1:17)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:17){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(coef=colnames(dat)[1:17],
+             est=rowMeans(p_cons),
+             lower=apply(p_cons,1,quantile, probs=.025),
+             upper=apply(p_cons,1,quantile, probs=.975),
+             p_neg = rowSums(p_cons<0)/4000,
+             metric=rep(i, 17)) %>%
+    rbind(plot.dat) %>%
+    filter(!is.na(est)) -> plot.dat
+}
+
+plot.dat$metric <- fct_recode(plot.dat$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Expulsion' = 'expulsion_combined')
+plot.dat$coef <- fct_recode(plot.dat$coef,
+                            'Intercept' = '(Intercept)',
+                            'total population' = 'total_pop',
+                            'propotion black' = 'black_prop',
+                            'proportion white' = 'white_prop',
+                            'black-white ratio' = 'b.w.ratio',
+                            'college grads' = 'col_grads',
+                            'income' = 'med_income',
+                            'poverty' = 'poverty_rate',
+                            'unemployment' = 'unemp_rate',
+                            'crime' = 'crime_rate',
+                            'housing density' = 'density',
+                            'mobility' = 'mobility',
+                            'race: white' = 'groupwhite',
+                            'implicit bias' = 'weighted_bias',
+                            'explicit bias' = 'weighted_explicit_diff',
+                            'implicit bias*race: white' = 'groupwhite:weighted_bias',
+                            'explicit bias*race: white' = 'groupwhite:weighted_explicit_diff')
+
+plot.dat %>%
+  mutate(estimate=paste(round(est, 2), ', [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
+  select(metric, estimate, coef) %>%
+  spread(metric, estimate) -> tabledata
+
+tabledata <- tabledata[c(1,13,3,17,2,4,10,12,14,5,6,11,7,15,16,8,9),]
+names(tabledata)[1] <- ''
+
+papaja::apa_table(tabledata, caption = 'Regression coefficient estimates for the population-level (i.e.
+fixed) effects, along with 95\\% uncertainty intervals for each of the disciplinary metrics.', small = T)
+```
+
+The findings for other disciplinary metrics suggest a differential effect of explicit and implicit bias.
+For example, there is a reliable association between in-school suspensions and explicit bias.
+The difference in the slope of the association between explicit bias and the log of the odds for in-school suspensions between white and black students is estimated to be `r round(plot.dat3_all$est[6], 2)`, with 95% of the posterior distribution between `r round(plot.dat3_all$lower[6], 2)` and `r round(plot.dat3_all$upper[6], 2)` and a proportion `r afex::round_ps(1)` of the posterior distribution consistent with a negative effect.
+Similar effects can be seen for explicit bias and law enforcement referrals (`r round(plot.dat3_all$est[7], 2)`, [`r round(plot.dat3_all$lower[7],2)`, `r round(plot.dat3_all$upper[7],2)`], $p_{neg}$ `r afex::round_ps(plot.dat3_all$p_neg[7])`), and explicit bias and expulsions (`r round(plot.dat3_all$est[10], 2)`, [`r round(plot.dat3_all$lower[10],2)`, `r round(plot.dat3_all$upper[10],2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[10])`).
+In contrast, zero is included as a plausible value for the associations between implicit bias and in-school suspensions (`r round(plot.dat3_all$est[3], 2)`, [`r round(plot.dat3_all$lower[3],2)`, `r round(plot.dat3_all$upper[3],2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[3])`), law enforcement referrals (`r round(plot.dat3_all$est[2], 2)`, [`r round(plot.dat3_all$lower[2],2)`, `r round(plot.dat3_all$upper[2],2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[2])`), and expulsions (`r round(plot.dat3_all$est[5], 2)`, [`r round(plot.dat3_all$lower[5],2)`, `r round(plot.dat3_all$upper[5],2)`], $p_{neg}$ = `r afex::round_ps(plot.dat3_all$p_neg[5])`).
+
+```{r detail-risks, fig.cap="Association between bias and relative risk ratio for black students to white students.
+Line is the mean of the posterior.
+Bands indicate 95% uncertainty intervals.
+Points represent counties, whose sizes are scaled to the number of schools in that county.
+Note that the y axis is cut at 8 for legibility despite data extending beyond."}
+plot.dat <- data.frame(group=NA, weighted_bias=NA, est=NA, 
+                       lower=NA, upper=NA, bias=NA, metric=NA)
+plot.dat2 <- data.frame(weighted_bias=NA, est=NA, 
+                        lower=NA, upper=NA, bias=NA, metric=NA)
+raw_or <- data.frame(county_id=NA, nschools=NA, bias=NA, 
+                     value=NA, black=NA, white=NA, odds_r=NA, metric=NA)
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(17,4000,length(files)))
+  for (k in files){
+  raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, 
+                         weighted_explicit_diff=NA, group=NA, nincidents=NA, 
+                         nstudents=NA, rate=NA)
+    #print(i)
+    load(paste(pth,k,sep=''))
+    df <- as.matrix(m)
+    m$data %>% 
+      select(county_id, COMBOKEY) %>% 
+      distinct() %>% 
+      group_by(county_id) %>% 
+      summarise(nschools=n()) -> n_schools
+    m$data %>%
+      group_by(county_id, group) %>%
+      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
+      mutate(rate=nincidents/nstudents) -> rate
+    n_schools %>%
+      left_join(m$data[, c('county_id', 'weighted_bias', 'weighted_explicit_diff')]) %>%
+      distinct() %>%
+      left_join(rate) %>%
+      rbind(raw_rate) %>%
+      filter(!is.na(county_id)) -> raw_rate
+    posterior_combo[,,j] <- t(df[,c(1:17)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:17){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:17){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  newdat <- expand.grid(group = rownames(table(m$data$group)),
+                        total_pop=0,
+                        unemp_rate=0,
+                        med_income=0,
+                        poverty_rate=0,
+                        col_grads=0,
+                        white_prop=0,
+                        black_prop=0,
+                        b.w.ratio=0,
+                        mobility=0,
+                        crime_rate=0,
+                        density=0,
+                        weighted_bias = seq(-2.5,2.5,.25),
+                        weighted_explicit_diff = 0,
+                        number=0,
+                        total_number=1)
+  y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)
+
+  cbind(newdat, t(y_hat)) %>% 
+    mutate(index=row_number()) %>%
+    gather(sample, value, `1`:`4000`) -> posterior_distribution
+
+  posterior_distribution %>%
+    select(group, weighted_bias, sample, value, index) %>%
+    group_by(group, weighted_bias) %>%
+    summarise(est = mean(value),
+              lower = quantile(value, .025),
+              upper = quantile(value, .975)) %>%
+    ungroup() %>%
+    mutate(bias = 'Implicit',
+           metric = i) %>%
+    rbind(plot.dat) -> plot.dat
+
+  posterior_distribution %>%
+    select(group, weighted_bias, sample, value) %>%
+    spread(group, value) %>%
+    mutate(ratio = black/white) %>%
+    group_by(weighted_bias) %>%
+    summarise(est = mean(ratio),
+              lower = quantile(ratio, .025),
+              upper = quantile(ratio, .975)) %>%
+    ungroup() %>%
+    mutate(bias = 'Implicit',
+           metric=i)  %>%
+    rbind(plot.dat2) -> plot.dat2
+
+  newdat <- expand.grid(group = rownames(table(m$data$group)),
+                        total_pop=0,
+                        unemp_rate=0,
+                        med_income=0,
+                        poverty_rate=0,
+                        col_grads=0,
+                        white_prop=0,
+                        black_prop=0,
+                        b.w.ratio=0,
+                        mobility=0,
+                        crime_rate=0,
+                        density=0,
+                        weighted_bias = 0,
+                        weighted_explicit_diff = seq(-2.5,2.5,.25),
+                        number=0,
+                        total_number=1)
+  y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)
+
+  cbind(newdat, t(y_hat)) %>% 
+    mutate(index=row_number()) %>%
+    gather(sample, value, `1`:`4000`) -> posterior_distribution
+
+  posterior_distribution %>%
+    select(group, weighted_explicit_diff, sample, value, index) %>%
+    mutate(weighted_bias = weighted_explicit_diff) %>%
+    group_by(group, weighted_bias) %>%
+    summarise(est = mean(value),
+              lower = quantile(value, .025),
+              upper = quantile(value, .975)) %>%
+    ungroup() %>%
+    mutate(bias = 'Explicit',
+           metric=i) %>%
+    rbind(plot.dat) -> plot.dat
+
+  posterior_distribution %>%
+    select(group, weighted_explicit_diff, sample, value) %>%
+    spread(group, value) %>%
+    mutate(ratio = black/white,
+           weighted_bias = weighted_explicit_diff) %>%
+    group_by(weighted_bias) %>%
+    summarise(est = mean(ratio),
+              lower = quantile(ratio, .025),
+              upper = quantile(ratio, .975)) %>%
+    ungroup() %>%
+    mutate(bias = 'Explicit',
+           metric=i) %>%
+    rbind(plot.dat2) -> plot.dat2
+
+  raw_rate %>%
+    mutate(Implicit = weighted_bias,
+           Explicit = weighted_explicit_diff) %>%
+    select(-weighted_bias, -weighted_explicit_diff) %>%
+    gather(bias, value, Implicit:Explicit) %>%
+    select(county_id, nschools, bias, group, rate, value) %>%
+    spread(group, rate) %>%
+    mutate(odds_r = black/white,
+           metric=i) %>%
+    rbind(raw_or) %>%
+    filter(!is.na(county_id)) -> raw_or
+}
+
+plot.dat$metric <- fct_recode(plot.dat$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Expulsion' = 'expulsion_combined')
+
+plot.dat2$metric <- fct_recode(plot.dat2$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Expulsion' = 'expulsion_combined')
+
+raw_or$metric <- fct_recode(raw_or$metric, 
+                            'Out-of-School Suspension' = 'oos_susp',
+                            'In-School Suspension' = 'inschool_susp',
+                            'School-Related Arrest' = 'in_school_arrest',
+                            'Law Enf. Referral' = 'law_enforcement',
+                            'Expulsion' = 'expulsion_combined')
+
+plot.dat %>%
+  filter(!is.na(weighted_bias)) %>%
+  ggplot(aes(x=weighted_bias, y=est, group=group, linetype=group)) +
+    geom_line(aes(color=group)) +
+    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.15) +
+    theme_classic() +
+    ylab('Probability of \n disciplinary action') +
+    xlab('Standardized Bias') + 
+    facet_grid(metric~bias, scales='free') +
+    xlim(-3, 3) +
+    theme(legend.title = element_blank()) -> p1
+
+plot.dat2 %>% 
+  filter(!is.na(weighted_bias)) %>%
+  ggplot(aes(x=weighted_bias, y=est)) +
+    geom_point(data=raw_or, aes(x=value, y=odds_r, size=nschools), alpha=.05) +
+    geom_line() +
+    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.25) +
+    theme_classic() +
+    scale_size_continuous(name='Number\nof schools') + 
+    #scale_color_manual(values=c('grey', 'darkorange'), guide=F) + 
+    scale_alpha_manual(values=c(.2, 1), guide=F) +
+    ylim(0,8) +
+    xlim(-3, 3) + 
+    xlab('Standardized Bias') +
+    ylab('Relative \n Risk Ratio') +
+    geom_hline(yintercept=1) +
+    facet_grid(metric~bias) -> p2
+
+p1$data %>% filter(weighted_bias %in% c(0, 1)) -> rates
+p2$data %>% filter(weighted_bias %in% c(0, 1)) -> risks
+
+#grid.arrange(p1, p2)
+p2
+```
+
+To better illustrate the nature of these relationships, figure \@ref(fig:detail-risks) shows the estimated relative risk ratios for each type of bias measurement and each disciplinary metric.
+Examining the risk ratio for out-of-school suspensions, in a hypothetical county with 0 standardized bias on both implicit and explicit, the model predicts that for every white student expelled, we should expect `r round(risks$est[1], 2)` [`r round(risks$lower[1], 2)`, `r round(risks$upper[1], 2)`] black students to be expelled.
+If we move to a county one standard deviation above the mean of explicit bias, the ratio of black to white students expelled increases to `r round(risks$est[4], 2)` [`r round(risks$lower[4], 2)`, `r round(risks$upper[4], 2)`], while the same movement for implicit bias increase the ratio to `r round(risks$est[2], 2)` [`r round(risks$lower[2], 2)`, `r round(risks$upper[2], 2)`].`
+
+```{r detail-rates, fig.cap="Association between bias and probability of receiving disciplinary action for black and white students.
+Lines are the means of the posterior.
+Bands indicate 95% uncertainty intervals."}
+
+p1
+
+```
+
+Figure \@ref(fig:detail-rates) unpacks these risk ratios by detailing the probabilities of receiving disciplinary action as a function of race, bias, and type of action.
+In a hypotheticl county with 0 standardized bias on both implicit and explicit, the model predicts that `r round(rates$est*100, 1)[1]`% [`r round(rates$lower*100, 1)[1]`, `r round(rates$upper*100, 1)[1]`] of black students will receive an out-of-school suspension .
+The corresponding rate for white students is just over half that for black students, with about `r round(rates$est*100, 1)[3]`% [`r round(rates$lower*100, 1)[3]`, `r round(rates$upper*100, 1)[3]`] expected to be suspended.
+Holding all other variables constant, moving to a hypothetical county one standard deviation above the mean of explicit bias has the effect of slightly decreasing the estimated percentage of black students expected to be suspended to `r round(rates$est*100, 1)[2]`% [`r round(rates$lower*100, 1)[2]`, `r round(rates$upper*100, 1)[2]`], while the percentage of white students expected to be suspended would decrease at a faster rate to `r round(rates$est*100, 1)[4]`% [`r round(rates$lower*100, 1)[4]`, `r round(rates$upper*100, 1)[4]`].
+The same movement for implicit bias would dramatically increase the expected suspensions for black students (`r round(rates$est*100, 1)[6]`% [`r round(rates$lower*100, 1)[6]`, `r round(rates$upper*100, 1)[6]`), and increase the expected suspensions for white students a much smaller amount to (`r round(rates$est*100, 1)[8]`% [`r round(rates$lower*100, 1)[8]`, `r round(rates$upper*100, 1)[8]`).
+
+# Discussion
+
+\newpage
+
+# References
+```{r create_r-references}
+r_refs(file = "r-references.bib")
+```
+
+\newpage
+# Appendix
+## Session details
+Data analysis was done in R [@rcitation] version 3.3.2 running under OS X 10.11.6.
+Post-stratification was done with lme4, version 1.1.14 [@bates2015fitting].
+Final model fitting was done on the university cluster running Springdale Linux, release 6.9 using rstanarm, version 2.17.2 [@rstanarm2016].
+We used the implementation of the consensus monte carlo algorithm found in parallelMCMCcombine, version 1.0 [@miroshnikov2014parallel].
+Figures were made with ggplot2, version 2.2.1 [@wickham2009ggplot2], with data manipulation done using dplyr version 0.7.2 [@wickham2017dplyr] and tidyr, version 0.7.1 [@wickham2017tidyr].
+A full report of session information can be found on the OSF page (....)
+
+## Preregistered analysis
+```{r summary_counts}
+
+###############
+schools_dat %>%
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
+  gather(group, total_number, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>%
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='NR_', 
+                          biracial='TR_', white='WH_')) %>%
+  filter(total_number>=0) %>%
+  group_by(COMBOKEY, group) %>%
+  mutate(total_number = sum(total_number)) %>%
+  ungroup() %>%
+  select(-prefix, -gender) %>%
+  distinct() -> ps_enrollment
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
+         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>% 
+  filter(SCH_GRADE_PS=='YES') %>%
+  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>% 
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='OS_', 
+                          biracial='TR_', white='WH_')) %>%
+  select(-prefix) -> ps_susp
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
+         SCH_GRADE_PS, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>% 
+  filter(SCH_GRADE_PS=='YES') %>%
+  gather(group, number_2, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>%
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='OS_', 
+                          biracial='TR_', white='WH_')) %>%
+  select(COMBOKEY, group, gender, number_2) %>%
+  right_join(ps_susp) %>%
+  filter(number>=0 & number_2 >= 0) %>%
+  mutate(number = number+number_2) -> ps_susp
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
+         SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>% 
+  gather(group, number, SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>% 
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='XP_', 
+                          biracial='TR_', white='WH_')) %>%
+  filter(number>=0) %>%
+  select(-prefix) -> ps_exp
+
+ps_susp %>% group_by(group) %>% summarise(n_students=sum(number)) -> susp_summ
+ps_exp %>% group_by(group) %>% summarise(n_students=sum(number)) -> exp_summ
+ps_enrollment %>% group_by(group) %>% summarise(n_students=sum(total_number)) -> enr_summ
+
+schools_dat %>%
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_ENR_HI_M:TOT_ENR_F) %>%
+  gather(group, total_number, SCH_ENR_HI_M:TOT_ENR_F) %>%
+  separate(group, into=c('group', 'gender'), -2) %>%
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='NR_', 
+                          biracial='TR_', white='WH_')) %>%
+  group_by(COMBOKEY, group) %>%
+  mutate(total_number = sum(total_number)) %>%
+  ungroup() %>%
+  select(-prefix, -gender) %>%
+  distinct() -> enrollment
+
+schools_dat %>% 
+  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
+         SCH_DISCWODIS_CORP_HI_M:TOT_DISCWODIS_CORP_F, 
+         SCH_DISCWDIS_CORP_IDEA_HI_M:TOT_DISCWDIS_CORP_IDEA_F) %>% 
+  gather(group, number, SCH_DISCWODIS_CORP_HI_M:TOT_DISCWDIS_CORP_IDEA_F) %>%
+  mutate(g=group) %>%
+  separate(group, into=c('group', 'gender'), -2) %>% 
+  separate(group, into=c('prefix', 'group'), -4) %>%
+  separate(prefix, into=c('prefix', 'disability'), -3) %>%
+  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
+                          hispanic='HI_', pac_isl='HP_', total='RP_', 
+                          total='EA_', biracial='TR_', white='WH_')) %>%
+  mutate(disability=fct_recode(disability, not_disabled='P_', disabled='ID',
+                               not_disabled='CO', disabled='A_')) %>%
+  select(-prefix) -> corp
+
+corp %>%
+  filter(group=='black'|group=='white') %>%
+  group_by(COMBOKEY, LEA_STATE, LEA_NAME, LEAID,
+           CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
+  summarise(number = sum(number, na.rm=T)) %>%
+  left_join(enrollment) %>%
+  mutate(proportion = number/total_number) -> corp_stats
+
+corp %>% 
+  select(COMBOKEY, LEA_STATE, number) %>% 
+  group_by(COMBOKEY, LEA_STATE) %>% 
+  summarise(number=sum(number, na.rm=T)) -> corp_stats
+
+schools_dat$south <- schools_dat$LEA_STATE %in% c('AL', 'AR', 'FL', 'GA', 'LA', 'MO', 'MS', 'OK', 'TN', 'TX')
+corp_stats %>% filter(number>0) -> uses_corp
+uses_corp$south <- uses_corp$LEA_STATE %in% c('AL', 'AR', 'FL', 'GA', 'LA', 'MO', 'MS', 'OK', 'TN', 'TX')
+
+```
+Here, we describe, in detail, the differences between the registered analysis plan and that presented in the main text, along with our reasons for making the alterations.
+
+### Data exclusions
+In our preanalysis plan, we specified our analyses to focus on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.
+However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes.
+
+First, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others.
+
+Next, we discovered that the administration of corporal punishment was extremely irregular across counties, with almost all cases ocurring in the South.
+Specifically, Alabama, Arkansas, Florida, Georgia, Louisiana, Missouri, Mississippi, Oklahoma, Tennessee, and Texas accounted for `r round((table(uses_corp$south)[2]/dim(uses_corp)[1])*100, 1)`% of all educational institutions that administered corporal punishment at least once, despite containing only `r round((table(schools_dat$south)[2]/dim(schools_dat)[1])*100, 1)`% of the educational institutions in the full dataset.
+It is not clear whether this uneven distribution is due to legal prohibitions, cultural differences, or some combination of the two.
+Because this makes interpretation of the parameters uncertain, this analysis is also placed in the appendix.
+
+We also found that the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]` total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely.
+We additionally discovered that counts of one expulsion category (expulsion under zero-tolerance policies) overlapped with counts in other categories, and so excluded this category from the main text.
+We also decided to combine the remaining two expulsion categories to yield one overall count of the number of students expelled.
+We did this to remain consistent with previous research, and to obtain district-level rates of expulsion that were slightly higher, and therefore more amenable to exploring variation across counties.
+
+When preparing the preregistration, we also had not known about the issues with juvenile justice facilities, or with the school districts with reporting errors, and so excluded these schools from analyses in the main text.
+
+### Changes in measurement & estimation of bias
+We preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e.
+*how warm or cold do you feel towards Blacks?
+0=very cold, 10=very warm*).
+However, the majority of past research (see for example @leitner2016blacks and @hehman2017disproportionate)  has used the difference in reported warmth towards whites and blacks, and so in the main text, we report models using this metric of explicit bias.
+
+Our main text presents analyses with poststratified estimates of bias.
+Our registered analysis also described analyses with raw, county-based means.
+Unlike previous papers that used poststratification, we see substantial differences in results between the two estimation methods.
+Because post-stratification is known to yield better estimates, we suspect that the difference in results is largely driven by extreme county-level observations with little data to support these extreme estimates.
+Nevertheless, in the interest of completion, we present the model using raw county-level means as well.
+
+We also registered an exploratory analysis using the responses of people who visited Project Implicit and identified themselves as teachers.
+Unfortunately, without population-level data on the demographic characteristics of teachers, we are not able to post-stratify these estimates.
+Additionally, there were relatively few teachers in the Project Implicit from which to derive estimates.
+Ultimately, we have little confidence that this analysis tells us anything about whether the biases of teachers specifically contribute to disparities in disciplinary actions, and so this analysis is also limited to the appendix.
+
+### Additional covariates
+We included three covariates that were not specified in the registered analysis.
+Specifically, the mobility, housing density, and crime rate covariates.
+These covariates were included in order to cover more of the covariates that were included in previous papers [@leitner2016blacks].
+
+## Registered analysis results
+Here, we present the results of the preregistered analyses exactly.
+Though we described a number of changes above, it is not practical to run all possible combination of changes, as each set of models takes four to five days to just to estimate, and there are at least several dozen possible combinations of decisions to test, depending on how fine-grained one wishes to be with the decisions.
+Accordingly, we limit the presentation of results below to just the three models that we registered.
+
+### County-level estimates
+```{r raw-county-estimates, fig.cap='Difference in raw county means and post-stratifed estimates of implicit bias as a function of the number of observations in each county.
+The X axis is on the log scale.'}
+individual_data %>%
+  group_by(county_id) %>%
+  summarise(n_obs=n()) %>%
+  left_join(county_means) -> cm
+
+cm %>%
+  mutate(post_diff = as.numeric(bias) - as.numeric(weighted_bias)) %>%
+  ggplot(aes(x=n_obs, y=post_diff)) + 
+  geom_point() + 
+  scale_x_log10(breaks=c(1, 10, 100, 1000, 10000)) +
+  theme_classic() +
+  ylab('Raw Bias - Weighted Bias difference') +
+  xlab('Number of Observations')
+
+```
+
+Examining the models fit with simple county means, we find that the county-level estimates vary more than with the post-stratified estimates.
+This pattern is expected, as one desireable feature of post-stratification is that it regularizes extreme observations without a large amount of data to support them.
+This phenomena is illustrated in figure \@ref(fig:raw-county-estimates), which shows the distribution of differences between the raw county means and the post-stratified estimates of implicit bias as a function of the number of observations in each county.
+More specifically, while the mean of the county-level means does not meaningfully differ from the post-stratified estimates, the standard deviation is much larger ($M_{implicit}$ = `r round(mean(as.numeric(cm$bias), na.rm=T), 2)`, $SD_{implicit}$ = `r round(sd(as.numeric(cm$bias), na.rm=T), 2)`; $M_{explicit}$ = `r round(mean(as.numeric(cm$explicit), na.rm=T), 2)`, $SD_{explicit}$ = `r round(sd(as.numeric(cm$explicit), na.rm=T), 2)`)[^2]
+
+[^2]: This explicit estimate is based on the simple warmth towards blacks question, and not the difference between warmth towards whites and warmth twoards blacks.
+
+The county-level implicit bias estimates for the poststratified model are as reported in the main text.
+The estimates for explicit bias are slightly different, as they are not computed based on the difference in reporting warmth toward whites and warmth toward blacks.
+Specifically, the average amount of bias at the county-level is `r round(mean(as.numeric(cm$weighted_explicit)*-1, na.rm=T), 2)` (SD = `r round(sd(as.numeric(cm$weighted_explicit), na.rm=T), 2)`).
+
 ```{r teachers-summary}
+df <- rbind(read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2013.sav'),
+            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2012.sav'),
+            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2010.sav'))
 
@@
 As such, our teacher analysis is limited to just `r dim(teacher_estimates)[1]` counties.
-Additionally, because we do not know of any state-level demographic estimates for teachers, we are unable to perform post-stratification for these data.
+Additionally, because we do not know of any state-level demographic estimates for teachers, we were unable to perform post-stratification for these data.
+Nonetheless, across these limited counties, the overall estimate of implicit and explicit bias are similar to those for the whole dataset ($M_{implicit}$ = `r round(mean(as.numeric(teacher_estimates$teacher_bias), na.rm=T), 2)`, $SD_{implicit}$ = `r round(sd(as.numeric(teacher_estimates$teacher_bias), na.rm=T), 2)`; $M_{explicit}$ = `r round(mean(as.numeric(teacher_estimates$teacher_explicit), na.rm=T), 2)`, $SD_{explicit}$ = `r round(sd(as.numeric(teacher_estimates$teacher_explicit), na.rm=T), 2)`)
 
-## Measures
-The primary outcome in this analysis is a count of the number of students by race (black and white) who received one of several types of disciplinary action.
-We report here rates of out-of-school suspension, in-school suspension, school-related arrests, law enforcement referrals, and total number of expulsions of any type.
+### Model estimates
 
-For both the post-stratification procedure (described below) and the actual statistical models used for inference, we used the same set of covariates.
-These covariates were at the state-level for post-stratification and were at the county-level for the statistical models used for final estimation and inference.
-Specifically, our population based covariates were the total population count, the proportion of the population that is black, the proportion of the population that is white and the ratio of black-to-white persons.
-We also used socioeconomic covariates.
-We used the percent of individuals aged 16 or over who were in the labor force but unemployed, the median household income, and the percentage of all families whose income is below the poverty line.
+Table \@ref(tab:mod-raw) presents results from the model fit to the raw, county-level means.
+The table presents parameter estimates (operationalized as the mean of the posterior) and 95% uncertainty intervals for each of the population-level effects for each of the outcomes.
+Table \@ref(tab:mod-post) and table \@ref(tab:mod-teach) present the corresponding results for the post-stratified and teacher models, respectively.
 
-Finally, for the implicit and explicit bias measures, we relied on two primary variables from the Project Implicit data.
-Implicit bias was assessed via an Implicit Association Test (IAT).
-This test uses a speeded dual-categorization task in which individuals must quickly categorize black and white faces and "good" and "bad" words with key presses.
-The difference in how quickly and accurately participants are able to pair white faces with "good" words and black faces with "bad" words in comparison to the inverse is thought to reflect implicit associations between the two races and positive and negative affective reactions.
-This association is reflected in the IAT D-score, which we used as a measure of implicit bias.
-Our measure of explicit bias is the difference between reported warmth towards whites (i.e.
-*how warm or cold do you feel towards Whites?
-0=very cold, 10=very warm*) and reported warmth toward blacks.
+```{r mod-raw, results='asis'}
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_raw/'
+plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA, model=NA)
 
-## Data analysis
-Data analysis proceded in two steps.
-We first estimated county-level implict and explicit bias using multilevel regression and post-stratification.
-Post-stratification is a valuable procedure in obtaining accurate geographical population-based estimates because it allows a non-representative sample (e.g Project Implicit) to more closely resemble the true population, and it regularizes extreme observations with little data to support them (e.g.
-a county with only a handful of respondents with especially high or low scores) [@park2004bayesian; @gelman1997poststratification].
-Following past work [@leitner2016blacks], we identified age as one dimension along which IAT respondents differed from the general population in ways that could bias our conclusions [@gonsalkorale2009aging].
-Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories (15-24, 25-34, 35-54, 55-75, and 75+).
-We next fit multilevel models estimating bias (implicit and explicit biases seperately) as a function of our state-level covariates (the "fixed" effects), and allowed the estimates to vary by age bin, county, and state (the "random" effects).
-Next, we determined the population of whites in each county in these age groups using the American Community Survey's 5-year estimates ending in 2014.
-Finally, we used our estimated models to predict the expected response for each age bin, in each county.
-Our final county-level estimates are the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county.
-As a result of this procedure, we can be confident that our county-level estimates should more closely approximate what our estimates would look like if the Project Implicit data were truly representative along the age dimension in all counties.
+metrics <- list.files(base_path)
+#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
 
-After obtaining these estimates, we use them as predictors in bayesian multilevel logistic regressions. Formally, the likelihood for a given observation is written as a binomial function:
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
 
-$$
-{n\choose y}\pi^y(1-\pi)^{n-y}
-$$
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    dat <- as.matrix(m)
+    posterior_combo[,,j] <- t(dat[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
 
-Where $y$ is the observed count of incidents (e.g.
-number of black or white students suspended), $n$ is the number of at-risk students (e.g.
-total number of black or white students), and $\pi = g^{-1}(\eta)$ is the probability of the incident ocurring.
-For this analysis, the linear predictor takes the form of a multilevel model with a set of effects that vary over county:
-
-$$
-\eta = \alpha + X\beta + \gamma_{county}
-$$
-
-Where $\alpha$ is an intercept that is constant across observations, $\beta$ represents a set of effects that are also constant across observations (i.e.
-fixed effects), and $\gamma_{county}$ represents intercepts and effects of ethnicity that vary across the counties (i.e.
-random efffects).
-
-In addition to the covariates described above, we also include effects of race, implicit bias, explicit bias, and the two-way interactions between implicit bias and race and explicit bias and race.
-We fit separate models for each of the outcomes.
-
-Because of the computational demands of fitting such a high-dimensional model to such a large dataset (the full model for each metric would consist of over 6k parameters to approximately 170k observations), we used a consensus monte carlo algorithm to obtain approximate posterior distributions for the parameters of interest [@scott2016bayes].
-The approximate posteriors derived from this algorithm have been shown to be nearly indistinguishable from the true posterior, a result we verified using a small subset of our own data.
-
-All numerical predictor variables were standardized at the appropriate level (county, state) before model estimation to help with estimation efficiency and interpretability.
-We set priors for the intercept and coefficients in the bayesian models to be weakly informative normal distributions centered on zero with a standard deviation of five.
-All other parameters were left to default values.
-Data analysis was done in R [@rcitation] version 3.3.2 running under OS X 10.11.6.
-Post-stratification was done with lme4, version 1.1.14 [@bates2015fitting].
-Final model fitting was done on the university cluster running Springdale Linux, release 6.9 using rstanarm, version 2.17.2 [@rstanarm2016].
-We used the implementation of the consensus monte carlo algorithm found in parallelMCMCcombine, version 1.0 [@miroshnikov2014parallel].
-Figures were made with ggplot2, version 2.2.1 [@wickham2009ggplot2], with data manipulation done using dplyr version 0.7.2 [@wickham2017dplyr] and tidyr, version 0.7.1 [@wickham2017tidyr].
-A full report of session information can be found on the OSF page (....)
-
-# Results
-
-# Discussion
-
-\newpage
-
-# References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-```
-
-\newpage
-# Appendix
-## Preregistered analysis
-```{r summary_counts}
-
-###############
-schools_dat %>%
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
-  gather(group, total_number, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>%
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='NR_', 
-                          biracial='TR_', white='WH_')) %>%
-  filter(total_number>=0) %>%
-  group_by(COMBOKEY, group) %>%
-  mutate(total_number = sum(total_number)) %>%
-  ungroup() %>%
-  select(-prefix, -gender) %>%
-  distinct() -> ps_enrollment
-
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
-         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>% 
-  filter(SCH_GRADE_PS=='YES') %>%
-  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>% 
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='OS_', 
-                          biracial='TR_', white='WH_')) %>%
-  select(-prefix) -> ps_susp
-
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
-         SCH_GRADE_PS, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>% 
-  filter(SCH_GRADE_PS=='YES') %>%
-  gather(group, number_2, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>%
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='OS_', 
-                          biracial='TR_', white='WH_')) %>%
-  select(COMBOKEY, group, gender, number_2) %>%
-  right_join(ps_susp) %>%
-  filter(number>=0 & number_2 >= 0) %>%
-  mutate(number = number+number_2) -> ps_susp
-
-schools_dat %>% 
-  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
-         SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>% 
-  gather(group, number, SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>%
-  separate(group, into=c('group', 'gender'), -2) %>% 
-  separate(group, into=c('prefix', 'group'), -4) %>%
-  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
-                          hispanic='HI_', pac_isl='HP_', total='XP_', 
-                          biracial='TR_', white='WH_')) %>%
-  filter(number>=0) %>%
-  select(-prefix) -> ps_exp
-
-ps_susp %>% group_by(group) %>% summarise(n_students=sum(number)) -> susp_summ
-ps_exp %>% group_by(group) %>% summarise(n_students=sum(number)) -> exp_summ
-ps_enrollment %>% group_by(group) %>% summarise(n_students=sum(total_number)) -> enr_summ
-
-#s,
-
-```
-In our preanalysis plan, we specified our analyses to focus on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.
-However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes.
-In particular, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others.
-Additionally, the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]` total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely.
-We additionally discovered that counts of one expulsion category (expulsion under zero-tolerance policies) overlapped with counts in other categories, and so excluded this category.
-To remain consistent with previous studies, we opted to combine the remaining two expulsion categories to yield one overall count of the number of students expelled.
-
-We also preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e.
-*how warm or cold do you feel towards Blacks?
-0=very cold, 10=very warm*).
-However, past research [@leitner2016blacks; @hehman2017disproportionate] has used the difference in reported warmth towards whites and blacks, and so in the main text, we report models using this metric of explicit bias.
-Additionally, we preregistered analyses with poststratified estimates (as presented in the main text) along with raw, county-based means.
-Finally, we had not known about the issues with juvenile justice facilities, or with the school districts with reporting errors.
-Here, we present the results of the preregistered analyses exactly.
-
-#Simple county means
-```{r cross_val_county_means}
-"%ni%" <- Negate("%in%")
-df <- read.csv('/Users/travis/Documents/gits/educational_disparities/cluster/data/county_grouping.csv')
-f <- list.files('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_raw/corporal/')
-
-test_summary_dat <- data.frame(file=f, 
-                               r_2 = NA,
-                               r_2_lower=NA,
-                               r_2_upper=NA)
-j=1
-for (i in f){
-  print(i)
-  load(paste('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_raw/inschool_susp/', i, sep=''))
-  model <- m
-  testing_model <- sample(f[f %ni% i], 1)
-  load(paste('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_raw/inschool_susp/', testing_model, sep=''))
-  m$data$bias <- m$data$bias[,1]
-  m$data$warmth <- m$data$warmth[,1]
-  m$data$weighted_bias <- m$data$weighted_bias[,1]
-  m$data$weighted_warmth <- m$data$weighted_warmth[,1]
-  m$data$total_pop <- m$data$total_pop[,1]
-  m$data$unemp_rate <- m$data$unemp_rate[,1]
-  m$data$med_income <- m$data$med_income[,1]
-  m$data$poverty_rate <- m$data$poverty_rate[,1]
-  m$data$col_grads <- m$data$col_grads[,1]
-  m$data$white_prop <- m$data$white_prop[,1]
-  m$data$black_prop <- m$data$black_prop[,1]
-  m$data$b.w.ratio <- m$data$b.w.ratio[,1]
-  m$data %>%
-    filter(!is.na(bias)) %>%
-    filter(!is.na(warmth)) %>%
-    filter(!is.na(weighted_bias)) %>%
-    filter(!is.na(weighted_warmth)) %>%
-    filter(!is.na(total_pop)) -> testing_data
-  y_hat <- data.frame(t(posterior_predict(m, draws = 250, newdata = testing_data)))
-  y_hat$index <- seq(1, length(y_hat$X1))
-  testing_data <- cbind(testing_data, y_hat)
-  testing_data %>%
-    gather(sample, pred, X1:X250) %>%
-    group_by(sample) %>%
-    summarise(r = cor(number, pred)^2) %>%
-    summarise(r_2 = mean(r),
-              r_2_lower = quantile(r, .25),
-              r_2_upper = quantile(r, .75)) -> r2
-  test_summary_dat[j,2:4] <- r2
-  j = j+1
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
   }
 
-test_summary_datweighted <- data.frame(file=f, 
-                               r_2 = NA,
-                               r_2_lower=NA,
-                               r_2_upper=NA)
-j=1
-for (i in f){
-  print(i)
-  load(paste('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_weighted/older_run/inschool_susp/', i, sep=''))
-  model <- m
-  testing_model <- sample(f[f %ni% i], 1)
-  load(paste('/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_weighted/older_run/inschool_susp/', testing_model, sep=''))
-  m$data$bias <- m$data$bias[,1]
-  m$data$warmth <- m$data$warmth[,1]
-  m$data$weighted_bias <- m$data$weighted_bias[,1]
-  m$data$weighted_warmth <- m$data$weighted_warmth[,1]
-  m$data$total_pop <- m$data$total_pop[,1]
-  m$data$unemp_rate <- m$data$unemp_rate[,1]
-  m$data$med_income <- m$data$med_income[,1]
-  m$data$poverty_rate <- m$data$poverty_rate[,1]
-  m$data$col_grads <- m$data$col_grads[,1]
-  m$data$white_prop <- m$data$white_prop[,1]
-  m$data$black_prop <- m$data$black_prop[,1]
-  m$data$b.w.ratio <- m$data$b.w.ratio[,1]
-  m$data %>%
-    filter(!is.na(bias)) %>%
-    filter(!is.na(warmth)) %>%
-    filter(!is.na(weighted_bias)) %>%
-    filter(!is.na(weighted_warmth)) %>%
-    filter(!is.na(total_pop)) -> testing_data
-  y_hat <- data.frame(t(posterior_predict(m, draws = 250, newdata = testing_data)))
-  y_hat$index <- seq(1, length(y_hat$X1))
-  testing_data <- cbind(testing_data, y_hat)
-  testing_data %>%
-    gather(sample, pred, X1:X250) %>%
-    group_by(sample) %>%
-    summarise(r = cor(number, pred)^2) %>%
-    summarise(r_2 = mean(r),
-              r_2_lower = quantile(r, .25),
-              r_2_upper = quantile(r, .75)) -> r2
-  test_summary_datweighted[j,2:4] <- r2
-  j = j+1
+  data.frame(coef=colnames(dat)[1:14],
+             est=rowMeans(p_cons),
+             lower=apply(p_cons,1,quantile, probs=.025),
+             upper=apply(p_cons,1,quantile, probs=.975),
+             p_neg = rowSums(p_cons<0)/4000,
+             metric=rep(i, 14),
+             model = rep('A', 14)) %>%
+    rbind(plot.dat) %>%
+    filter(!is.na(est)) -> plot.dat
 }
+plot.dat$metric <- fct_recode(plot.dat$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Corporal Punishment' = 'corporal',
+                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
+                               'Expulsion with Educational Services' = 'expulsion_w_ed',
+                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
+                               'Mechanical Restraint' = 'mechanical_restraint',
+                               'Physical Restraint' = 'physical_restraint',
+                               'Preschool Expulsion' = 'preschool_expulsion',
+                               'Preschool Suspension' = 'preschool_susp',
+                               'Seclusion' = 'seclusion')
+plot.dat %>%
+  mutate(estimate=paste(round(est, 2), ', [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
+  select(metric, estimate, coef) %>%
+  spread(metric, estimate) -> tabledata_raw
+
+tabledata_raw$coef <- fct_recode(tabledata_raw$coef,
+                            'Intercept' = '(Intercept)',
+                            'total population' = 'total_pop',
+                            'propotion black' = 'black_prop',
+                            'proportion white' = 'white_prop',
+                            'black-white ratio' = 'b.w.ratio',
+                            'college grads' = 'col_grads',
+                            'income' = 'med_income',
+                            'poverty' = 'poverty_rate',
+                            'unemployment' = 'unemp_rate',
+                            'race: white' = 'groupwhite',
+                            'implicit bias' = 'bias',
+                            'explicit bias' = 'warmth',
+                            'implicit bias*race: white' = 'groupwhite:bias',
+                            'explicit bias*race: white' = 'groupwhite:warmth')
+tabledata_raw <- tabledata_raw[c(1,11,4,14,2,5,9,10,12,6,3,13,7,8),]
+names(tabledata_raw)[1] <- ''
+
+papaja::apa_table(tabledata_raw, caption = 'Regression coefficient estimates for the population-level (i.e.
+fixed) effects, along with 95\\% uncertainty intervals for each of the disciplinary metrics.
+Model uses raw, county-level means as estimates of implicit and explicit bias.', landscape=T, small=T)
 
@@
 
-#Post stratified estimates
+```{r mod-post, results='asis'}
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/metrics_weighted/'
+plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA, model=NA)
 
-#Teacher analyses
+metrics <- list.files(base_path)
+#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    dat <- as.matrix(m)
+    posterior_combo[,,j] <- t(dat[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(coef=colnames(dat)[1:14],
+             est=rowMeans(p_cons),
+             lower=apply(p_cons,1,quantile, probs=.025),
+             upper=apply(p_cons,1,quantile, probs=.975),
+             p_neg = rowSums(p_cons<0)/4000,
+             metric=rep(i, 14),
+             model = rep('A', 14)) %>%
+    rbind(plot.dat) %>%
+    filter(!is.na(est)) -> plot.dat
+}
+plot.dat$metric <- fct_recode(plot.dat$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Corporal Punishment' = 'corporal',
+                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
+                               'Expulsion with Educational Services' = 'expulsion_w_ed',
+                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
+                               'Mechanical Restraint' = 'mechanical_restraint',
+                               'Physical Restraint' = 'physical_restraint',
+                               'Preschool Expulsion' = 'preschool_expulsion',
+                               'Preschool Suspension' = 'preschool_susp',
+                               'Seclusion' = 'seclusion')
+plot.dat %>%
+  mutate(estimate=paste(round(est, 2), ', [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
+  select(metric, estimate, coef) %>%
+  spread(metric, estimate) -> tabledata_post
+
+tabledata_post$coef <- fct_recode(tabledata_post$coef,
+                            'Intercept' = '(Intercept)',
+                            'total population' = 'total_pop',
+                            'propotion black' = 'black_prop',
+                            'proportion white' = 'white_prop',
+                            'black-white ratio' = 'b.w.ratio',
+                            'college grads' = 'col_grads',
+                            'income' = 'med_income',
+                            'poverty' = 'poverty_rate',
+                            'unemployment' = 'unemp_rate',
+                            'race: white' = 'groupwhite',
+                            'implicit bias' = 'weighted_bias',
+                            'explicit bias' = 'weighted_warmth',
+                            'implicit bias*race: white' = 'groupwhite:weighted_bias',
+                            'explicit bias*race: white' = 'groupwhite:weighted_warmth')
+
+tabledata_post <- tabledata_post[c(1,11,4,14,2,5,9,10,12,6,3,13,7,8),]
+names(tabledata_post)[1] <- ''
+
+papaja::apa_table(tabledata_post, caption = 'Regression coefficient estimates for the population-level (i.e.
+fixed) effects, along with 95\\% uncertainty intervals for each of the disciplinary metrics.
+Model uses poststratified estimates of implicit and explicit bias.', landscape=T, small=T)
+
+```
+
+```{r mod-teach, results='asis'}
+base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/teacher_metrics/'
+plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA, model=NA)
+
+metrics <- list.files(base_path)
+#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
+for (i in metrics){
+  pth <- paste(base_path, i, '/', sep='')
+
+  files <- list.files(pth)
+  j <- 1
+  posterior_combo <- array(0, dim=c(14,4000,length(files)))
+
+  for (k in files){
+    #print(i)
+    load(paste(pth,k,sep=''))
+    dat <- as.matrix(m)
+    posterior_combo[,,j] <- t(dat[,c(1:14)])
+    j <- j+1
+  }
+  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)
+
+  for (p in 1:14){
+    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
+  }
+  for (p in 1:14){
+    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
+  }
+
+  data.frame(coef=colnames(dat)[1:14],
+             est=rowMeans(p_cons),
+             lower=apply(p_cons,1,quantile, probs=.025),
+             upper=apply(p_cons,1,quantile, probs=.975),
+             p_neg = rowSums(p_cons<0)/4000,
+             metric=rep(i, 14),
+             model = rep('A', 14)) %>%
+    rbind(plot.dat) %>%
+    filter(!is.na(est)) -> plot.dat
+}
+plot.dat$metric <- fct_recode(plot.dat$metric, 
+                               'Out-of-School Suspension' = 'oos_susp',
+                               'In-School Suspension' = 'inschool_susp',
+                               'School-Related Arrest' = 'in_school_arrest',
+                               'Law Enf. Referral' = 'law_enforcement',
+                               'Corporal Punishment' = 'corporal',
+                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
+                               'Expulsion with Educational Services' = 'expulsion_w_ed',
+                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
+                               'Mechanical Restraint' = 'mechanical_restraint',
+                               'Physical Restraint' = 'physical_restraint',
+                               'Preschool Expulsion' = 'preschool_expulsion',
+                               'Preschool Suspension' = 'preschool_susp',
+                               'Seclusion' = 'seclusion')
+plot.dat %>%
+  mutate(estimate=paste(round(est, 2), ', [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
+  select(metric, estimate, coef) %>%
+  spread(metric, estimate) -> tabledata_teachers
+
+tabledata_teachers$coef <- fct_recode(tabledata_teachers$coef,
+                            'Intercept' = '(Intercept)',
+                            'total population' = 'total_pop',
+                            'propotion black' = 'black_prop',
+                            'proportion white' = 'white_prop',
+                            'black-white ratio' = 'b.w.ratio',
+                            'college grads' = 'col_grads',
+                            'income' = 'med_income',
+                            'poverty' = 'poverty_rate',
+                            'unemployment' = 'unemp_rate',
+                            'race: white' = 'groupwhite',
+                            'implicit bias' = 'county_bias',
+                            'explicit bias' = 'county_warmth',
+                            'implicit bias*race: white' = 'groupwhite:county_bias',
+                            'explicit bias*race: white' = 'groupwhite:county_warmth')
+
+tabledata_teachers <- tabledata_teachers[c(1,11,4,14,2,5,9,10,12,6,3,13,7,8),]
+names(tabledata_teachers)[1] <- ''
+
+papaja::apa_table(tabledata_teachers, caption = 'Regression coefficient estimates for the population-level (i.e.
+fixed) effects, along with 95\\% uncertainty intervals for each of the disciplinary metrics.
+Model uses only data from Project Implicit respondents who identified as teachers.', landscape=T, small=T)
+
+```
 

### 2018-03-22 written_docs/draft.Rmd
edit the plotting code

@@
 plot.dat3_all <- rbind(plot.dat3_all, plot.dat3)
-ggplot(plot.dat3_all, aes(x=metric, y=est, group=bias, color=bias, shape=bias)) +
-  geom_point(position=position_dodge(width=.5), size=2) +
+p <- ggplot(plot.dat3_all, aes(x=metric, y=est, group=bias, color=bias, shape=bias)) +
+  geom_point(position=position_dodge(width=.5), size=3) +
   geom_errorbar(aes(ymin=lower, ymax=upper), 
@@
   theme(axis.title.y = element_blank(),
-        legend.title = element_blank())
+        legend.title = element_blank(),
+        axis.text = element_text(size=12),
+        axis.title.x = element_text(size=14, face='bold'),
+        legend.text = element_text(size=12))
+ggsave('/Users/travis/Documents/gits/educational_disparities/figs/slope_ests.jpeg', p, width = 11, height = 6)
 ```
@@
 Note that the y axis is cut at 8 for legibility despite data extending beyond."}
-plot.dat <- data.frame(group=NA, weighted_bias=NA, est=NA, 
-                       lower=NA, upper=NA, bias=NA, metric=NA)
-plot.dat2 <- data.frame(weighted_bias=NA, est=NA, 
-                        lower=NA, upper=NA, bias=NA, metric=NA)
+plot.dat <- data.frame(group=NA, weighted_bias=NA, est=NA, lower=NA, 
+                       upper=NA, bias=NA, metric=NA, p_neg_black=NA, p_neg_white=NA)
+plot.dat2 <- data.frame(weighted_bias=NA, est=NA, lower=NA, 
+                        upper=NA, bias=NA, metric=NA, p_neg=NA)
 raw_or <- data.frame(county_id=NA, nschools=NA, bias=NA, 
@@
 
+  p_neg_blackimp <- sum(p_cons[3,]>0)/4000
+  p_neg_blackexp <- sum(p_cons[4,]>0)/4000
+  p_neg_whiteimp <- sum((p_cons[3,]+p_cons[16,])>0)/4000
+  p_neg_whiteexp <- sum((p_cons[4,]+p_cons[17,])>0)/4000
+
   posterior_distribution %>%
@@
     mutate(bias = 'Implicit',
-           metric = i) %>%
+           metric = i,
+           p_neg_black=p_neg_blackimp,
+           p_neg_white=p_neg_whiteimp) %>%
     rbind(plot.dat) -> plot.dat
+
+  p_neg <- sum(p_cons[16,]<0)/4000
 
@@
     mutate(bias = 'Implicit',
-           metric=i)  %>%
+           metric=i,
+           p_neg = p_neg)  %>%
     rbind(plot.dat2) -> plot.dat2
@@
     mutate(bias = 'Explicit',
-           metric=i) %>%
+           metric=i,
+           p_neg_black=p_neg_blackexp,
+           p_neg_white=p_neg_whiteexp) %>%
     rbind(plot.dat) -> plot.dat
+
+  p_neg <- sum(p_cons[17,]<0)/4000
 
@@
     mutate(bias = 'Explicit',
-           metric=i) %>%
+           metric=i,
+           p_neg=p_neg) %>%
     rbind(plot.dat2) -> plot.dat2
@@
 plot.dat$metric <- fct_recode(plot.dat$metric, 
-                               'Out-of-School Suspension' = 'oos_susp',
-                               'In-School Suspension' = 'inschool_susp',
-                               'School-Related Arrest' = 'in_school_arrest',
-                               'Law Enf. Referral' = 'law_enforcement',
+                               'Out-of-School \n Suspension' = 'oos_susp',
+                               'In-School \n Suspension' = 'inschool_susp',
+                               'School-Related \n Arrest' = 'in_school_arrest',
+                               'Law Enf. \n Referral' = 'law_enforcement',
                                'Expulsion' = 'expulsion_combined')
@@
 plot.dat2$metric <- fct_recode(plot.dat2$metric, 
-                               'Out-of-School Suspension' = 'oos_susp',
-                               'In-School Suspension' = 'inschool_susp',
-                               'School-Related Arrest' = 'in_school_arrest',
-                               'Law Enf. Referral' = 'law_enforcement',
+                               'Out-of-School \n Suspension' = 'oos_susp',
+                               'In-School \n Suspension' = 'inschool_susp',
+                               'School-Related \n Arrest' = 'in_school_arrest',
+                               'Law Enf. \n Referral' = 'law_enforcement',
                                'Expulsion' = 'expulsion_combined')
@@
 raw_or$metric <- fct_recode(raw_or$metric, 
-                            'Out-of-School Suspension' = 'oos_susp',
-                            'In-School Suspension' = 'inschool_susp',
-                            'School-Related Arrest' = 'in_school_arrest',
-                            'Law Enf. Referral' = 'law_enforcement',
+                            'Out-of-School \n Suspension' = 'oos_susp',
+                            'In-School \n Suspension' = 'inschool_susp',
+                            'School-Related \n Arrest' = 'in_school_arrest',
+                            'Law Enf. \n Referral' = 'law_enforcement',
                             'Expulsion' = 'expulsion_combined')
+
+plot.dat <- plot.dat[!is.na(plot.dat$weighted_bias),]
+
+plot.dat %>%
+  filter(weighted_bias==2.5) %>%
+  mutate(text_y = est,
+         text_x = 2.75) %>%
+  select(group, text_y, text_x, metric, p_neg_black, p_neg_white, bias) %>%
+  distinct() %>%
+  gather(group1, p_neg, p_neg_black, p_neg_white) %>%
+  mutate(group1 = str_sub(group1, 7, -1)) %>%
+  filter(group==group1) %>%
+  select(-group1) -> text_data
+
+text_data$text_comp <- '=='
+text_data$text_comp[text_data$p_neg>.99] <- '>'
+text_data$text_comp[text_data$p_neg<.01] <- '<'
+text_data$p_neg[text_data$p_neg>.99] <- .99
+text_data$p_neg[text_data$p_neg<.01] <- .01
 
@@
     theme_classic() +
+    geom_text(data=text_data, parse=T, size=2.5, 
+              aes(x=2.8, y=text_y, vjust=0,
+                  label=paste0('p[pos]', text_comp, round(p_neg,2)))) +
     ylab('Probability of \n disciplinary action') +
@@
     facet_grid(metric~bias, scales='free') +
-    xlim(-3, 3) +
+    xlim(-2.75, 3.25) +
     theme(legend.title = element_blank()) -> p1
+
+plot.dat2$text_x <- -2.75
+plot.dat2$text_y <- 5
+plot.dat2$text_comp <- '=='
+plot.dat2$text_comp[plot.dat2$p_neg>.99] <- '>'
+plot.dat2 %>%
+  select(text_x, text_y, text_comp, p_neg, metric, bias) %>%
+  distinct() %>%
+  mutate(p_neg = round(p_neg, 2)) %>%
+  filter(!is.na(p_neg)) -> text_data
+text_data$p_neg[text_data$p_neg==1] <- .99
 
@@
   ggplot(aes(x=weighted_bias, y=est)) +
-    geom_point(data=raw_or, aes(x=value, y=odds_r, size=nschools), alpha=.05) +
+    geom_point(data=raw_or, aes(x=value, y=odds_r, size=nschools), alpha=.1) +
     geom_line() +
@@
     scale_size_continuous(name='Number\nof schools') + 
+    geom_text(data=text_data, parse=T, size=3,
+              aes(x=-2.75, y=4, label=paste0('p[pos]', text_comp, round(p_neg,2)))) +
     #scale_color_manual(values=c('grey', 'darkorange'), guide=F) + 
     scale_alpha_manual(values=c(.2, 1), guide=F) +
-    ylim(0,8) +
+    ylim(0,6) +
     xlim(-3, 3) + 
@@
     geom_hline(yintercept=1) +
-    facet_grid(metric~bias) -> p2
+    facet_grid(metric~bias) +
+    theme(legend.position = 'top',
+          strip.text.y=element_text(size=6)) -> p2
+ggsave('/Users/travis/Documents/gits/educational_disparities/figs/rel_risk.jpeg', p2, width = 11, height = 6)
 
@@
 
-p1
+ggsave('/Users/travis/Documents/gits/educational_disparities/figs/rates.jpeg', p1, width = 11, height = 6)
 

### 2018-03-24 aggregate_data.R
remove the mistaken exclusions

@@ -65,7 +65,7 @@ tempout$number[which(tempout$number<0)] <- NA
 tempout %>% 
   ungroup() %>%
   filter(is.finite(proportion)) %>%
-  filter(total_number>number) %>%
+  filter(total_number>=number) %>%
   mutate(metric='inschool_susp') %>%
   mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
   distinct() -> subdat
@@ -118,7 +118,7 @@ tempout %>%
   ungroup() %>%
   filter(is.finite(proportion)) %>%
   filter(number>=0) %>%
-  filter(total_number>number) %>%
+  filter(total_number>=number) %>%
   mutate(metric='oos_susp') %>%
   mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
   distinct() %>%
@@ -174,7 +174,7 @@ exp_w_ed %>%
 tempout %>% 
   ungroup() %>%
   filter(is.finite(proportion)) %>%
-  filter(total_number>number) %>%
+  filter(total_number>=number) %>%
   mutate(metric='expulsion_combined') %>%
   mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
   distinct() %>%
@@ -208,7 +208,7 @@ tempout %>%
   ungroup() %>%
   filter(is.finite(proportion)) %>%
   filter(number>=0) %>%
-  filter(total_number>number) %>%
+  filter(total_number>=number) %>%
   ungroup() %>%
   mutate(metric='law_enforcement') %>%
   mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
@@ -245,7 +245,7 @@ tempout %>%
   ungroup() %>%
   filter(is.finite(proportion)) %>%
   filter(number>=0) %>%
-  filter(total_number>number) %>%
+  filter(total_number>=number) %>%
   mutate(metric='in_school_arrest') %>%
   mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
   distinct() %>%
@@ -329,35 +329,33 @@ df_school %>%
   mutate(LEAID = droplevels(LEAID))-> exclude
 
 # these are LEAID numbers of schools to exclude
-error_elem <- c('06CC088', '06CC027', '0623340', '0625470', '0629370', '0633600', 
-                '0634410', '0637050', '1201710', '1500030', '2802580', '3620580', 
-                '4218990', '5308700', '1201860', '0500394') #taken from elementary spreadsheet at:
[…]

### 2018-03-24 consensusMCMC_model_prep.R
15 shards

@@ -13,7 +13,7 @@ dft %>%
   distinct() -> teacher_counties
 
 counties %>% 
-  mutate(grouping=sample(1:14, n(), replace=T)) -> county_assignments
+  mutate(grouping=sample(1:15, n(), replace=T)) -> county_assignments
 
 teacher_counties %>%
   mutate(grouping=sample(1:12, n(), replace=T)) -> teacher_county_assignments
@@ -218,9 +218,9 @@ save(m, file='/tigress/triddle/educational_disparities/mw_uclaexcl_diff_extra_co
 
 
 for(i in rownames(table(dfs$metric))){
-  for (j in 1:14){
+  for (j in 1:15){
     fs <- paste(string1, i, string2, j, string3, i, string4, i, "/m", j, ".rdata')", sep='')
-    fileConn <- file(paste("cluster/model_scripts/mw_uclaexcl_diff_extracovs/",i,"/m", j, ".R", sep=''))
+    fileConn <- file(paste("/Users/travis/Documents/gits/educational_disparities/cluster/model_scripts/mw_uclaexcl_diff_extracovs/",i,"/m", j, ".R", sep=''))
     writeLines(fs, fileConn)
     close(fileConn)
   }

### 2018-03-24 aggregate_data.R
exclude if over reporting on any race

@@ -52,20 +52,24 @@ df_school %>%
                                not_disabled='S_', disabled='A_')) %>%
   select(-prefix) -> susp_inschool
 
+susp_inschool$number[which(susp_inschool$number<0)] <- NA
+
 susp_inschool %>%
-  filter(group=='black'|group=='white') %>%
+  #filter(group=='black'|group=='white') %>%
   group_by(COMBOKEY, LEA_STATE, LEA_NAME, LEAID,
            CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
   summarise(number = sum(number, na.rm=T)) %>%
   left_join(enrollment) %>%
   mutate(proportion = number/total_number) -> tempout
 
-tempout$number[which(tempout$number<0)] <- NA
-
 tempout %>% 
   ungroup() %>%
-  filter(is.finite(proportion)) %>%
-  filter(total_number>=number) %>%
+  mutate(impossible = number>total_number) %>%
+  group_by(COMBOKEY) %>%
+  mutate(impossible_school = sum(impossible)>0) %>%
+  ungroup() %>%
+  filter(impossible_school==F) %>%
+  filter(group=='black'|group=='white') %>%
   mutate(metric='inschool_susp') %>%
   mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
   distinct() -> subdat
@@ -104,21 +108,24 @@ df_school %>%
   filter(number>=0 & number_2 >= 0) %>%
   mutate(number = number+number_2) -> oos_susp
 
+oos_susp$number[which(oos_susp$number<0)] <- NA
+
 oos_susp %>%
-  filter(group=='black'|group=='white') %>%
+  #filter(group=='black'|group=='white') %>%
   group_by(COMBOKEY, LEA_STATE, LEA_NAME, LEAID,
            CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
   summarise(number = sum(number, na.rm=T)) %>%
   left_join(enrollment) %>%
   mutate(proportion = number/total_number) -> tempout
 
-tempout$number[which(tempout$number<0)] <- NA
-
 tempout %>% 
   ungroup() %>%
-  filter(is.finite(proportion)) %>%
-  filter(number>=0) %>%
-  filter(total_number>=number) %>%
+  mutate(impossible = number>total_number) %>%
+  group_by(COMBOKEY) %>%
[…]

### 2018-03-24 consensusMCMC_model_prep.R
change path names to write for the extra exclusions

@@ -214,13 +214,13 @@ black_prop + b.w.ratio + mobility + crime_rate + density +
 (group|county_id), data=mod.dat, prior = normal(0,5), 
 prior_intercept = normal(0,5), family=binomial, adapt_delta=.99)
 
-save(m, file='/tigress/triddle/educational_disparities/mw_uclaexcl_diff_extra_covs/"
+save(m, file='/tigress/triddle/educational_disparities/mw_uclaexcl_diff_extra_covs_extra_excl/"
 
 
 for(i in rownames(table(dfs$metric))){
   for (j in 1:15){
     fs <- paste(string1, i, string2, j, string3, i, string4, i, "/m", j, ".rdata')", sep='')
-    fileConn <- file(paste("/Users/travis/Documents/gits/educational_disparities/cluster/model_scripts/mw_uclaexcl_diff_extracovs/",i,"/m", j, ".R", sep=''))
+    fileConn <- file(paste("/Users/travis/Documents/gits/educational_disparities/cluster/model_scripts/mw_uclaexcl_diff_extracovs_extra_excl/",i,"/m", j, ".R", sep=''))
     writeLines(fs, fileConn)
     close(fileConn)
   }

### 2018-05-18 written_docs/Riddle_Sinclair_PNAS/Riddle_Sinclair_PNAS.Rmd
PNAS submitted

---
title: Race-based disparities in academic disciplinary actions are associated
  with county-level rates of racial bias

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author:
  - name: Travis Riddle
    affiliation: a
  - name: Stacey Sinclair
    affiliation: a,b
address:
  - code: a
    address: Princeton University, Department of Psychology, Princeton, NJ, 08544
  - code: b
    address: African American Studies

corresponding_author:
  - code: 1
    text: "To whom correspondence should be addressed. E-mail: triddle@princeton.edu"

# For footer text
lead_author_surname: Riddle

author_contributions: |
  TR and SS designed the research and wrote the paper. TR analyzed the data.

## Remove this if not required
conflict_of_interest: |
  Authors have no known conflicts of interest

abstract: |
  There are substantial gaps in educational outcomes between black and white students in the United States. Recently, increased attention has focused on differences in the rates at which black and white students are disciplined, finding that black students are more likely to be seen as problematic and more likely to be punished than white students are for the same offense. Although these disparities suggest that racial biases are a contributor, no previous research has shown associations with psychological measurements of bias and disciplinary outcomes. We show that county-level estimates of explicit racial bias, as measured using data from approximately 1.3 million visitors to the Project Implicit website, are associated with racial disciplinary disparities across approximately 93k schools in the United States, covering around 32 million white and black students. These associations do not extend to explicit or implicit sexuality biases, showing the specificity of the effect. These findings suggest that reducing racial disparities in education may require efforts to reduce explicit racial bias.

significance: |
  Black students in the United States are subject to disciplinary action at rates much higher than their white counterparts. These disciplinary actions put students at higher risk for negative life outcomes, including involvement in the criminal justice system. Using government data covering over 32 million students at nearly 93k schools, our research demonstrates that the disciplinary gap between black and white students across 5 different types of disciplinary actions is associated with county-level rates of explicit racial bias, and suggests solving racial disparities in education must involve combating explicit racial bias.

acknowledgements: |
  Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.

keywords:
  - one
  - two
  - optional
  - optional
  - optional

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: ref.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: true

output: rticles::pnas_article
---

```{r load-libs, echo=FALSE, message=FALSE, results='hide'}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, 
                      message = FALSE, cache.lazy = FALSE)
library(dplyr)
library(forcats)
library(haven)
library(tidyr)
library(ggplot2)
library(stringr)
```

In comparison to White Americans, Black Americans exhibit poorer educational outcomes across a range of metrics. One outcome of particular concern is the gap in disciplinary actions [@kinsler2011understanding; @okonofua2016vicious]. Research using administrative datasets and longitudinal samples clearly show that Black American students are far more likely to be suspended or expelled [@aud2011america; @yeager2017loss], and conditional on an office referral, are more likely to receive stiffer punishments [@gregory1995crime; @skiba2011race]. These disparities are particularly concerning as they are associated with long-term life outcomes, including prospects for employment [@pager2009sequencing] and involvement in the criminal justice system [@hirschfield2009another].

As complex social phenomena, these racial differences in disciplinary outcomes are multiply determined [@okonofua2016vicious]. There are clear structural and socieconomic contributors such as the practice of segregating students into achievement-based "tracks" of low and high performers [@togut2011gestalt] or racial differences in socioeconomic status [@skiba2002color]. However, racial bias is also thought to contribute to the problem. For instance, in a controlled study, @okonofua2015two found that in comparison to white students, teachers were more likely to view the same behavior from black students as being indicative of a long term problem and deserving of suspension. Similarly, using discipline data from an urban high school, @gregory2008discipline showed that black students were especially likely to be referred to the office for disciplinary action on the basis of defiant behavior - a relatively subjective category of misbehavior in comparison to others they examined, including truancy or fighting. Overall, there has been consistent evidence that black students' behaviors are both perceived as more problematic and are punished more harshly as compared to white students'. However, to our knowledge, there has been no work assessing whether racial bias is directly associated with disciplinary disparities. Additionally, there has been no work assessing how racial bias at the community level is associated with educational disparities.

Psychological measurements of racial bias typically occur through one of two ways. Either individuals are asked to self-report their relative attitudes toward different racial groups, or via methods designed to assess automatic associations with people of different races. The latter is thought to reflect the associations learned by early and frequent exposure to environmental stimuli, whereas the former reflects top-down modulation of those associations [@rudman2004sources]. Recently, researchers have begun aggregating these measures up to geographical regions such as counties or states, finding that regional-level measures of implicit and explicit racial bias are associated with racial disparities in key social outcomes, although the relative contributions are not consistent across studies [@leitner2016blacks; @orchard2017county; @hehman2017disproportionate]. For example, @leitner2016blacks found that Black Americans had reduced access to health care and increased rates of death due to circulatory disease in comparison to whites in counties with higher levels of explicit racial bias against blacks. They found no such associations for implicit racial bias. In contrast, @hehman2017disproportionate, found that the disproportionate use of lethal force by police on Black Americans was associated with regional implicit biases, but were not associated with explicit biases. As such, it is important to assess both types of bias when seeking to understand the relationship between regional-level bias and behavioral outcomes.

Regional levels of bias could be associated with the size of racial student disciplinary disparities for a number of reasons. First, being in an area characterized by higher racial bias likely means encountering individuals who have negative feelings and beliefs about one’s group. The actions of such individuals within and/or outside of an educational setting could contribute to disciplinary disparities. For example, if teachers and administrators are biased, then they may be more likely to make decisions that are unfavorable to black students, such as deciding that a given misbehavior is worthy of official disciplinary action. Biased administrators or local voters might also support school or district policies thought to disproportionally punish students of color, such as zero-tolerance policies or implementation of random drug sweeps [@verdugo2002race]. Second, Black students who live in high bias areas may be more apt to engage in behavior that warrants disciplinary action due to behavior confirmation of the negative expectancies held by area peers and adults [@cadinu2003stereotype; @darley1980expectancy; @major2005social]. Third, the norms and structural factors (e.g., laws, policies) that characterize regions higher in bias may constrain even those individuals who are not biased themselves into engaging in or suborning actions that negatively impact students of color [@jetten1996intergroup; @pettigrew1958personality]. Fourth, biases assessed at the regional level might reflect affordances of the local environment (e.g., confederate statues, biased media) that undergird these biases and prime behaviors that contribute to disciplinary disparities [@payne2017bias]. Overall, these reasons, and the likely possibility that they work in concert to inform behavior [@markus2010cultures], substantiate the possibility that there will be a relationship between regional bias and disciplinary outcomes.

Most previous research has focused on out-of-school suspensions - likely because they are the most frequently used and are regularly found to be associated with negative outcomes [@cuellar2015school; @matjasko2011effective]. However, other disciplinary outcomes, though used less often, may be at least as damaging to students as out-of-school suspensions [@rausch2005academic]. For instance school arrests have been associated with increased likelihood of engaging in anti-social behavior [@hemphill2006effect], and with increased risk of dropping out [@hirschfield2009another]. In addition, though alternative forms of discipline (e.g., in-school suspension) are intended to insulate students from the negative consequences of exclusionary discipline, the criteria by which students are assigned the former kind of discipline often remain vulnerable to bias [@anyon2014persistent]. As such, examining the presence and basis of disparities in the application of a wide range of disciplinary actions is warranted.

```{r iat-data, echo=FALSE, results='hide', message=FALSE, cache=TRUE}

df <- rbind(read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2013.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2012.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2010.sav'))

df2 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2011.sav')
df3 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2009.sav')
df4 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2008.sav')
df5 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2007.sav')
df6 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2006.sav')
df7 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2005.sav')
df8 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2004.sav')
df9 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2002-2003.sav')
df10 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2014.sav')

"%ni%" <- Negate("%in%")

df %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  filter(raceomb==6) -> subdat

df2 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all,  
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df3 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df4 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df5 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df6 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df7 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df8 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df9 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(raceomb = ethnic) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df10 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

subdat %>%
  mutate(explicit_bias=tblack_0to10) %>%
  #filter(!is.na(twhite_0to10)) %>%
  mutate(explicit_bias = twhite_0to10 - tblack_0to10) %>%
  mutate(age_bin = cut(age, breaks=c(14, 24, 34, 54, 75, 120))) %>%
  mutate(race = fct_recode(as.character(raceomb), 
                           'Black'='5', 'White'='6')) %>%
  filter(!is.na(age_bin)) %>%
  filter(race=='White') %>%
  filter(STATE %ni% 
           c('AA', 'AE', 'AP', 'AS', 'FM', 'GU', 'MH', 'MP', 'PR', 'VI')) %>%
  mutate(county_id = paste(STATE, CountyNo, sep='-')) -> individual_data

```

The present analyses combine regionally-coded implicit and explicit racial bias measures from approximately `r round(dim(individual_data)[1]/1000000, 1)` million respondents who visited Project Implicit [@xu2014psychology] with the most recent available data from the Civil Rights Data Collection (CRDC) conducted by the US Department of Education, a mandated census of disciplinary action in all US public k-12 schools. The CRDC allowed well-powered examinations of five different disciplinary metrics: in-school suspensions, out-of-school suspensions, law enforcement referrals, school-related arrests, and expulsions. We designed the analyses to determine whether, and if so the extent to which regional estimates of pro-White/anti-Black implicit and explicit bias are associated with black-white outcomes in disciplinary gaps. We also use regional estimates of sexuality bias from Project Implicit to determine whether demonstrated patterns are distinct to racial bias, or an epiphenomenal associate of bias measures in general.

## Results

### Project Implicit estimates
```{r exp-diff, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv')
schools_dat <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')
model.data <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/full_model_data.csv')

individual_data %>%
  mutate(explicit_bias_diff = twhite_0to10-tblack_0to10) -> individual_data
```

We first report the results of estimating the implicit and explicit biases from Project Implicit data. When examining the unstandardized county-level estimates adjusted with poststratification, we see that there is a pro-white bias in both implicit ($mean$ = `r round(mean(as.numeric(county_means$weighted_bias), na.rm=T), 2)`, $sd$ = `r round(sd(as.numeric(county_means$weighted_bias), na.rm=T), 2)`) and explicit measures ($mean$ = `r round(mean(as.numeric(county_means$weighted_explicit_diff), na.rm=T), 2)`, $sd$ = `r round(sd(as.numeric(county_means$weighted_explicit_diff), na.rm=T), 2)`), where on both scales 0 = no bias, and positive numbers indicate a pro-white bias.

## Disciplinary action frequency
```{r disc-count, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
model.data %>% 
  filter(exclude==FALSE) %>%
  group_by(group, metric) %>% 
  summarise(percentage = sum(number)/sum(total_number)) %>%
  mutate(percentage = percentage*100) %>%
  mutate(percentage = paste(round(percentage, 2), '%', sep='')) %>%
  spread(group, percentage) %>%
  mutate(metric = fct_recode(metric, 
                             'expulsions' = 'expulsion_combined',
                             'school arrests' = 'in_school_arrest',
                             'in-school suspension' = 'inschool_susp',
                             'law enforcement referral' = 'law_enforcement',
                             'out-of-school suspension' = 'oos_susp')) %>%
  arrange(black) %>% data.frame() -> counts_table

stargazer::stargazer(counts_table, summary = F, label='tab:disc-count', title = 'Percentage of students of each race receiving each type of disciplinary action', header = F, rownames=F)
```

Table \ref{tab:disc-count} shows the percentage of students of each race who were reported having received each of the actions under consideration. Additionally, figure \ref{fig:suspension-disp-map} shows the relative risk ratio for out-of-school suspensions across counties. Our statistical models (described below) show that the racial differences seen here are extremely unlikely to be due to chance.

```{r suspension-disp-map, results='hide', echo=F, message=F, warning=F, fig.cap='Relative Risk Ratio for out-of-school suspensions for each county in the continental United States. Relative risk ratio is computed as the percentage of black students suspended to the percentage of white students suspended, as reported in the CRDC. Higher values indicate more black students suspended, relative to white students. Counties with the value NA either have no schools, have zero black students enrolled, or have zero white students enrolled. Interactive maps for all disciplinary outcomes and bias estimates can be found at https://osf.io/pu79a/'}
library(rgdal)
library(viridis)

#shapefiles obtained from 2010 US census: https://www.census.gov/geo/maps-data/data/cbf/cbf_counties.html
US.counties <- readOGR(dsn='/Users/travis/Documents/gits/Data/gz_2010_us_050_00_5m/',
                       layer='gz_2010_us_050_00_5m')
US.states <- readOGR(dsn='/Users/travis/Documents/gits/Data/gz_2010_us_040_00_5m/',
                       layer='gz_2010_us_040_00_5m')
US.counties <- US.counties[!(US.counties$STATE %in% c("02","15","72")),]
US.states <- US.states[!(US.states$STATE %in% c('02', '15', '72')),]
county.data <- US.counties@data
state.data <- US.states@data
county.data$id <- rownames(county.data)
state.data$id <- rownames(state.data)
model.data %>%
  mutate(GEO_ID = id_from_census) %>%
  dplyr::select(county, state_code, state_name, GEO_ID, county_id, 
                COMBOKEY, group, number, total_number, metric, exclude) %>%
  filter(exclude==FALSE) %>%
  filter(metric=='oos_susp') %>%
  group_by(GEO_ID, group) %>%
  summarise(number=sum(number),
            total_number = sum(total_number)) %>%
  mutate(prop = number/total_number) %>%
  dplyr::select(GEO_ID, group, prop) %>%
  spread(group, prop) %>%
  #mutate(rel_risk = black/white) %>%
  mutate(rel_risk = case_when(black == 0 & white == 0 ~ 1, 
                             TRUE ~ black/white)) %>%
  mutate(risk_cat = cut(rel_risk, breaks = c(0, .5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, Inf), include.lowest = T)) %>%
  mutate(risk_cat = fct_recode(risk_cat, 
                               '0 - .5' = '[0,0.5]',
                               '.51 - 1' = '(0.5,1]',
                               '1.01 - 1.5' = '(1,1.5]',
                               '1.51 - 2' = '(1.5,2]',
                               '2.01 - 2.5' = '(2,2.5]',
                               '2.51 - 3' = '(2.5,3]',
                               '3.01 - 3.5' = '(3,3.5]',
                               '3.51 - 4' = '(3.5,4]',
                               '4.01 - 4.5' = '(4,4.5]',
                               '4.51 - Inf' = '(4.5,Inf]')) %>%
  mutate(risk_cat_num = as.numeric(risk_cat)) -> prop_diff

map.df <- broom::tidy(US.counties)
state.df <- broom::tidy(US.states)
map.df %>% 
  left_join(county.data) %>%
  left_join(prop_diff) %>%
  filter(!is.na(GEO_ID)) -> plot.dat

map <- ggplot(plot.dat, aes(x=long, y=lat, group=group)) +
  scale_fill_viridis(name = 'Relative Risk', discrete=T) +
  geom_polygon(aes(fill=risk_cat)) + 
  geom_polygon(data=state.df, aes(x=long, y=lat, group=group, fill=NA), color='gray45', size=.5) + 
  coord_map() +
  theme_void() +
  #ggtitle('Relative Risk Ratio for Out-of-school Suspensions') + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.title=element_text(size=10)) +
  guides(fill = guide_legend(reverse=T))
ggsave('/Users/travis/Documents/gits/educational_disparities/figs/map.jpeg', map, width = 11, height = 6)
map
```

## Associations across counties
Figure \[fig:overall-associations\]  shows the estimate of primary interest for each of the models. The estimates displayed are the coefficients for the interaction between race and each of the two bias measurements. Given that Black Americans are the baseline group, negative values for this coefficient indicate that as one moves into counties with higher levels of bias, the gap between the probability of a black student being disciplined and the probability of a white student being disciplined grows. Table \[tab:reg-coefs\] displays the estimated coefficients and 95% uncertainty intervals for each of the fixed effects for each model.

```{r overall-associations, results='hide', echo=F, message=F, warning=F, fig.cap='Association between each metric and county-level estimates of explicit and implicit bias. Negative values indicate that the rate of increase (or decrease) for blacks is faster (or slower) than for whites. Point is the mean of the posterior and error bars represent 95% bayesian uncertainty intervals'}
#the p_neg for oos_susp & implicit bias is 100%, so when reporting the number in-text, i just replace it with 1 to prevent it from being rendered as NA in the document.
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/full_models/'
plot.dat <- data.frame(group = NA, weighted_bias = NA, est = NA, lower = NA, upper=NA, metric = NA)
plot.dat2 <- data.frame(weighted_bias = NA, est = NA, lower=NA, upper=NA, metric=NA)
plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg=NA, metric=NA)
outlier_county <- data.frame(weighted_bias=NA, est=NA, lower=NA, upper=NA, bias=NA, metric=NA)

# weighted bias w/exclusions & explicit difference
metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
  for (k in files){
    #print(k)
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    m$data %>% 
      select(county_id, COMBOKEY) %>% 
      distinct() %>% 
      group_by(county_id) %>% 
      summarise(nschools=n()) -> n_schools
    m$data %>%
      group_by(county_id, group) %>%
      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
      mutate(rate=nincidents/nstudents) -> rate
    n_schools %>%
      left_join(m$data[, c('county_id', 'weighted_bias')]) %>%
      distinct() %>%
      left_join(rate) %>%
      rbind(raw_rate) %>%
      filter(!is.na(county_id)) -> raw_rate
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }

    newdat <- expand.grid(group = rownames(table(m$data$group)),
                        total_pop=0,
                        unemp_rate=0,
                        med_income=0,
                        poverty_rate=0,
                        col_grads=0,
                        white_prop=0,
                        black_prop=0,
                        b.w.ratio=0,
                        mobility=0,
                        crime_rate=0,
                        density=0,
                        dissim=0,
                        weighted_bias = 3,
                        weighted_explicit_diff = 0,
                        number=0,
                        total_number=1)
  y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

  cbind(newdat, t(y_hat)) %>% 
    mutate(index=row_number()) %>%
    gather(sample, value, `1`:`4000`) -> posterior_distribution

  posterior_distribution %>%
    select(group, weighted_bias, sample, value) %>%
    spread(group, value) %>%
    mutate(ratio = black/white) %>%
    group_by(weighted_bias) %>%
    summarise(est = mean(ratio),
              lower = quantile(ratio, .025),
              upper = quantile(ratio, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Implicit',
           metric=i) %>%
    rbind(outlier_county) -> outlier_county

  data.frame(est=mean(p_cons[17,]),
             lower=quantile(p_cons[17,], .025),
             upper=quantile(p_cons[17,], .975),
             p_neg = prop.table(table(p_cons[17,]<0))[2],
             metric=i) %>%
    rbind(plot.dat3) %>%
    filter(!is.na(est)) %>%
    mutate(metric=reorder(metric, est)) -> plot.dat3
}

plot.dat3_all <-plot.dat3
plot.dat3_all$bias <- 'Implicit'

#weighted warmth w/exclusions & explicit difference
plot.dat <- data.frame(group = NA, weighted_explicit_diff = NA, est = NA, lower = NA, upper=NA, metric = NA)
plot.dat2 <- data.frame(weighted_explicit_diff = NA, est = NA, lower=NA, upper=NA, metric=NA)
plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg = NA, metric=NA)

for(i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, weighted_explicit_diff=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    m$data %>% 
      select(county_id, COMBOKEY) %>% 
      distinct() %>% 
      group_by(county_id) %>% 
      summarise(nschools=n()) -> n_schools
    m$data %>%
      group_by(county_id, group) %>%
      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
      mutate(rate=nincidents/nstudents) -> rate
    n_schools %>%
      left_join(m$data[, c('county_id', 'weighted_explicit_diff')]) %>%
      distinct() %>%
      left_join(rate) %>%
      rbind(raw_rate_warmth) %>%
      filter(!is.na(county_id)) -> raw_rate_warmth
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }

  data.frame(est=mean(p_cons[18,]),
             lower=quantile(p_cons[18,], .025),
             upper=quantile(p_cons[18,], .975),
             p_neg = prop.table(table(p_cons[18,]<0))[2],
             metric=i) %>%
    rbind(plot.dat3) %>%
    filter(!is.na(est)) %>%
    mutate(metric=reorder(metric, est)) -> plot.dat3
}

plot.dat3$bias <- 'Explicit'
plot.dat3_all <- rbind(plot.dat3_all, plot.dat3)
plot.dat3_all$metric <- factor(plot.dat3_all$metric, c('oos_susp',
                                                        'inschool_susp',
                                                        'law_enforcement',
                                                        'expulsion_combined',
                                                        'in_school_arrest'))

plot.dat3_all$metric <- fct_recode(plot.dat3_all$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
p <- ggplot(plot.dat3_all, aes(x=metric, y=est, group=bias, color=bias, shape=bias)) +
  geom_point(position=position_dodge(width=.5), size=3) +
  geom_errorbar(aes(ymin=lower, ymax=upper), 
                width=.5, position=position_dodge(width=.5)) +
  theme_classic() +
  coord_flip() +
  geom_hline(yintercept=0) +
  ylab('Black-White Difference in log-odds slope') +
  theme(axis.title.y = element_blank(),
        legend.title = element_blank(),
        legend.position='top',
        axis.text = element_text(size=10),
        axis.title.x = element_text(size=12, face='bold'),
        legend.text = element_text(size=10))
ggsave('/Users/travis/Documents/gits/educational_disparities/figs/slope_ests.jpeg', p, width = 11, height = 6)
p
```

Several patterns are apparent from this figure. First, county-level explicit biases are reliably associated with disciplinary disparities, such that counties with more bias are expected to have a larger racial gap. The effect with the largest magnitude of these is school-related arrests ($M_{slope}$ = `r round(plot.dat3_all$est[9], 2)`, [`r round(plot.dat3_all$lower[9], 2)`, `r round(plot.dat3_all$upper[9], 2)`]), followed by expulsions ($M_{slope}$ = `r round(plot.dat3_all$est[10], 2)`, [`r round(plot.dat3_all$lower[10], 2)`, `r round(plot.dat3_all$upper[10], 2)`]), out-of-school suspensions ($M_{slope}$ = `r round(plot.dat3_all$est[6], 2)`, [`r round(plot.dat3_all$lower[6], 2)`, `r round(plot.dat3_all$upper[6], 2)`]), in-school suspensions ($M_{slope}$ = `r round(plot.dat3_all$est[8], 2)`, [`r round(plot.dat3_all$lower[8], 2)`, `r round(plot.dat3_all$upper[8], 2)`]), and law enforcement referrals ($M_{slope}$ = `r round(plot.dat3_all$est[7], 2)`, [`r round(plot.dat3_all$lower[7], 2)`, `r round(plot.dat3_all$upper[7], 2)`]).

In contrast, the estimated effects for implicit bias are more ambiguous. Expulsions ($M_{slope}$ = `r round(plot.dat3_all$est[5], 2)`, [`r round(plot.dat3_all$lower[5], 2)`, `r round(plot.dat3_all$upper[5], 2)`]), school-related arrests ($M_{slope}$ = `r round(plot.dat3_all$est[4], 2)`, [`r round(plot.dat3_all$lower[4], 2)`, `r round(plot.dat3_all$upper[4], 2)`]), and law enforcement referrals ($M_{slope}$ = `r round(plot.dat3_all$est[2], 2)`, [`r round(plot.dat3_all$lower[2], 2)`, `r round(plot.dat3_all$upper[2], 2)`]) are all estimated to be close to zero, or with substantial uncertainty with respect to the direction of the effect. Surprisingly, there is a reliable association for in-school suspensions, with schools higher in implicit bias having smaller in-school suspension gap ($M_{slope}$ = `r round(plot.dat3_all$est[3], 2)`, [`r round(plot.dat3_all$lower[3], 2)`, `r round(plot.dat3_all$upper[3], 2)`]). Though not fully conclusive, the the association between implicit biases and out-of-school suspensions tends to be in a similar, counter-intuitive direction ($M_{slope}$ = `r round(plot.dat3_all$est[1], 2)`, [`r round(plot.dat3_all$lower[1], 2)`, `r round(plot.dat3_all$upper[1], 2)`]). However, it is important to point out that at no point do these associations "cross over" such that white students are actually suspended more frequently than black students. Even at a county with implicit bias 3 standard deviations above the mean, for every white student who receives an out-of-school suspension, about `r round(outlier_county$est[1], 2)` [`r round(outlier_county$lower[1], 2)`, `r round(outlier_county$upper[1], 2)`] black students are expected to be suspended.

```{r, reg-coefs, results='asis', echo=F, message=F, warning=F}
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/full_models/'
plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA)

# weighted bias w/exclusions & explicit difference
metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'in_school_arrest', 'law_enforcement', 'oos_susp')
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  for (k in files){
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }
  data.frame(coef=colnames(dat)[1:30],
             est=rowMeans(p_cons),
             lower=apply(p_cons,1,quantile, probs=.025),
             upper=apply(p_cons,1,quantile, probs=.975),
             p_neg = rowSums(p_cons<0)/4000,
             metric=rep(i, 30)) %>%
    rbind(plot.dat) %>%
    filter(!is.na(est)) -> plot.dat
}

plot.dat$metric <- factor(plot.dat$metric, c('oos_susp', 'inschool_susp',
                                             'law_enforcement', 
                                             'expulsion_combined',
                                             'in_school_arrest'))

plot.dat$metric <- fct_recode(plot.dat$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat$coef <- fct_recode(plot.dat$coef,
                            'Intercept' = '(Intercept)',
                            'total population' = 'total_pop',
                            'proportion black' = 'black_prop',
                            'proportion white' = 'white_prop',
                            'black-white ratio' = 'b.w.ratio',
                            'college grads' = 'col_grads',
                            'income' = 'med_income',
                            'poverty' = 'poverty_rate',
                            'unemployment' = 'unemp_rate',
                            'crime' = 'crime_rate',
                            'housing density' = 'density',
                            'mobility' = 'mobility',
                            'dissimilarity' = 'dissim',
                            'total population*race: white' = 'groupwhite:total_pop',
                            'proportion black*race: white' = 'groupwhite:black_prop',
                            'proportion white*race: white' = 'groupwhite:white_prop',
                            'black-white ratio*race: white' = 'groupwhite:b.w.ratio',
                            'college grads*race: white' = 'groupwhite:col_grads',
                            'income*race: white' = 'groupwhite:med_income',
                            'poverty*race: white' = 'groupwhite:poverty_rate',
                            'unemployment*race: white' = 'groupwhite:unemp_rate',
                            'crime*race: white' = 'groupwhite:crime_rate',
                            'housing density*race: white' = 'groupwhite:density',
                            'mobility*race: white' = 'groupwhite:mobility',
                            'dissimilarity*race: white' = 'groupwhite:dissim',
                            'race: white' = 'groupwhite',
                            'implicit bias' = 'weighted_bias',
                            'explicit bias' = 'weighted_explicit_diff',
                            'implicit bias*race: white' = 'groupwhite:weighted_bias',
                            'explicit bias*race: white' = 'groupwhite:weighted_explicit_diff')

plot.dat %>%
  filter(p_neg<.025 | p_neg>.975) %>%
  mutate(est = paste('\\bf{', round(est, 2), '}', sep='')) -> sigs
plot.dat %>%
  filter(p_neg>=.025) %>%
  filter(p_neg<=.975) %>%
  mutate(est = round(est, 2)) %>%
  rbind(sigs) %>%
  mutate(estimate=paste(est, ', [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
  select(metric, estimate, coef) %>%
  spread(metric, estimate) -> tabledata

tabledata <- tabledata[c(1:7, 23:27, 30, 8, 28, 29, 9:19, 22, 20, 21),]
names(tabledata)[1] <- ''

stargazer::stargazer(tabledata, summary = F, label='tab:reg-coefs', title = 'Regression coefficient estimates for the population-level (i.e. fixed) effects, along with 95\\% uncertainty intervals for each of the disciplinary metrics.', header = F, rownames=F)
```

```{r detail-risks, fig.cap="Association between bias and relative risk ratio for black students to white students. Line is the mean of the posterior. Bands indicate 95% uncertainty intervals. Points represent counties, whose sizes are scaled to the number of schools in that county. Printed p_{pos} represents the posterior probability that the association is positive. Note that the y axis is cut at 8 for legibility despite data extending beyond. Additionally, a number of schools cannot be represented because the ratio of discipline proportions (black/white) is undefined, either because there are zero students of a given race enrolled (e.g. NA/white), or the proportion of white students disciplined is zero (e.g. black/0).", results='hide', echo=F, message=F, warning=F}
plot.dat <- data.frame(group=NA, weighted_bias=NA, est=NA, lower=NA, 
                       upper=NA, bias=NA, metric=NA, p_neg_black=NA, p_neg_white=NA)
plot.dat2 <- data.frame(weighted_bias=NA, est=NA, lower=NA, 
                        upper=NA, bias=NA, metric=NA, p_neg=NA)
raw_or <- data.frame(county_id=NA, nschools=NA, bias=NA, 
                     value=NA, black=NA, white=NA, odds_r=NA, metric=NA)
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  for (k in files){
  raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias=NA, 
                         weighted_explicit_diff=NA, group=NA, nincidents=NA, 
                         nstudents=NA, rate=NA)
    #print(i)
    load(paste(pth,k,sep=''))
    df <- as.matrix(m)
    m$data %>% 
      select(county_id, COMBOKEY) %>% 
      distinct() %>% 
      group_by(county_id) %>% 
      summarise(nschools=n()) -> n_schools
    m$data %>%
      group_by(county_id, group) %>%
      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
      mutate(rate=nincidents/nstudents) -> rate
    n_schools %>%
      left_join(m$data[, c('county_id', 'weighted_bias', 'weighted_explicit_diff')]) %>%
      distinct() %>%
      left_join(rate) %>%
      rbind(raw_rate) %>%
      filter(!is.na(county_id)) -> raw_rate
    posterior_combo[,,j] <- t(df[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }

  newdat <- expand.grid(group = rownames(table(m$data$group)),
                        total_pop=0,
                        unemp_rate=0,
                        med_income=0,
                        poverty_rate=0,
                        col_grads=0,
                        white_prop=0,
                        black_prop=0,
                        b.w.ratio=0,
                        mobility=0,
                        crime_rate=0,
                        density=0,
                        dissim=0,
                        weighted_bias = seq(-2.75,2.75,.25),
                        weighted_explicit_diff = 0,
                        number=0,
                        total_number=1)
  y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

  cbind(newdat, t(y_hat)) %>% 
    mutate(index=row_number()) %>%
    gather(sample, value, `1`:`4000`) -> posterior_distribution

  p_neg_blackimp <- sum(p_cons[3,]>0)/4000
  p_neg_blackexp <- sum(p_cons[4,]>0)/4000
  p_neg_whiteimp <- sum((p_cons[3,]+p_cons[17,])>0)/4000
  p_neg_whiteexp <- sum((p_cons[4,]+p_cons[18,])>0)/4000

  posterior_distribution %>%
    select(group, weighted_bias, sample, value, index) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Implicit',
           metric = i,
           p_neg_black=p_neg_blackimp,
           p_neg_white=p_neg_whiteimp) %>%
    rbind(plot.dat) -> plot.dat

  p_neg <- sum(p_cons[17,]<0)/4000

  posterior_distribution %>%
    select(group, weighted_bias, sample, value) %>%
    spread(group, value) %>%
    mutate(ratio = black/white) %>%
    group_by(weighted_bias) %>%
    summarise(est = mean(ratio),
              lower = quantile(ratio, .025),
              upper = quantile(ratio, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Implicit',
           metric=i,
           p_neg = p_neg)  %>%
    rbind(plot.dat2) -> plot.dat2

  newdat <- expand.grid(group = rownames(table(m$data$group)),
                        total_pop=0,
                        unemp_rate=0,
                        med_income=0,
                        poverty_rate=0,
                        col_grads=0,
                        white_prop=0,
                        black_prop=0,
                        b.w.ratio=0,
                        mobility=0,
                        crime_rate=0,
                        density=0,
                        dissim=0,
                        weighted_bias = 0,
                        weighted_explicit_diff = seq(-2.75,2.75,.25),
                        number=0,
                        total_number=7000000)
  y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

  cbind(newdat, t(y_hat)) %>% 
    mutate(index=row_number()) %>%
    gather(sample, value, `1`:`4000`) -> posterior_distribution

  posterior_distribution %>%
    select(group, weighted_explicit_diff, sample, value, index) %>%
    mutate(weighted_bias = weighted_explicit_diff) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Explicit',
           metric=i,
           p_neg_black=p_neg_blackexp,
           p_neg_white=p_neg_whiteexp) %>%
    rbind(plot.dat) -> plot.dat

  p_neg <- sum(p_cons[18,]<0)/4000

  posterior_distribution %>%
    select(group, weighted_explicit_diff, sample, value) %>%
    spread(group, value) %>%
    mutate(ratio = black/white,
           weighted_bias = weighted_explicit_diff) %>%
    group_by(weighted_bias) %>%
    summarise(est = mean(ratio),
              lower = quantile(ratio, .025),
              upper = quantile(ratio, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Explicit',
           metric=i,
           p_neg=p_neg) %>%
    rbind(plot.dat2) -> plot.dat2

  raw_rate %>%
    mutate(Implicit = weighted_bias,
           Explicit = weighted_explicit_diff) %>%
    select(-weighted_bias, -weighted_explicit_diff) %>%
    gather(bias, value, Implicit:Explicit) %>%
    select(county_id, nschools, bias, group, rate, value) %>%
    spread(group, rate) %>%
    mutate(odds_r = black/white,
           metric=i) %>%
    rbind(raw_or) %>%
    filter(!is.na(county_id)) -> raw_or
}

plot.dat$metric <- factor(plot.dat$metric, c('in_school_arrest', 
                                         'expulsion_combined', 
                                         'law_enforcement',
                                         'inschool_susp', 
                                         'oos_susp'))

plot.dat$metric <- fct_recode(plot.dat$metric, 
                               'Out \n of-School \n Suspension' = 'oos_susp',
                              'In \n School \n Suspension' = 'inschool_susp',
                              'Arrests' = 'in_school_arrest',
                              'Law Enf. \n Referral' = 'law_enforcement',
                              'Expulsion' = 'expulsion_combined')

plot.dat2$metric <- factor(plot.dat2$metric, c('in_school_arrest', 
                                         'expulsion_combined', 
                                         'law_enforcement',
                                         'inschool_susp', 
                                         'oos_susp'))

plot.dat2$metric <- fct_recode(plot.dat2$metric, 
                               'Out \n of-School \n Suspension' = 'oos_susp',
                               'In \n School \n Suspension' = 'inschool_susp',
                               'Arrests' = 'in_school_arrest',
                               'Law Enf. \n Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')

raw_or$metric <- factor(raw_or$metric, c('in_school_arrest', 
                                         'expulsion_combined', 
                                         'law_enforcement',
                                         'inschool_susp', 
                                         'oos_susp'))

raw_or$metric <- fct_recode(raw_or$metric, 
                            'Out \n of-School \n Suspension' = 'oos_susp',
                            'In \n School \n Suspension' = 'inschool_susp',
                            'Arrests' = 'in_school_arrest',
                            'Law Enf. \n Referral' = 'law_enforcement',
                            'Expulsion' = 'expulsion_combined')

plot.dat <- plot.dat[!is.na(plot.dat$weighted_bias),]

plot.dat %>%
  filter(weighted_bias==2.5) %>%
  mutate(text_y = est,
         text_x = 2.75) %>%
  select(group, text_y, text_x, metric, p_neg_black, p_neg_white, bias) %>%
  distinct() %>%
  gather(group1, p_neg, p_neg_black, p_neg_white) %>%
  mutate(group1 = str_sub(group1, 7, -1)) %>%
  filter(group==group1) %>%
  select(-group1) -> text_data

text_data$text_comp <- '=='
text_data$text_comp[text_data$p_neg>.99] <- '>'
text_data$text_comp[text_data$p_neg<.01] <- '<'
text_data$p_neg[text_data$p_neg>.99] <- .99
text_data$p_neg[text_data$p_neg<.01] <- .01

plot.dat %>%
  filter(!is.na(weighted_bias)) %>%
  ggplot(aes(x=weighted_bias, y=est, group=group, linetype=group)) +
    geom_line(aes(color=group)) +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.15) +
    theme_classic() +
    geom_text(data=text_data, parse=T, size=2.5, 
              aes(x=2.8, y=text_y, vjust=0,
                  label=paste0('p[pos]', text_comp, round(p_neg,2)))) +
    ylab('Probability of \n disciplinary action') +
    xlab('Standardized Bias') + 
    facet_grid(metric~bias, scales='free') +
    xlim(-2.75, 3.25) +
    theme(legend.title = element_blank()) -> p1

plot.dat2$text_x <- -2.75
plot.dat2$text_y <- 5
plot.dat2$text_comp <- '=='
plot.dat2$text_comp[plot.dat2$p_neg>.99] <- '>'
plot.dat2 %>%
  select(text_x, text_y, text_comp, p_neg, metric, bias) %>%
  distinct() %>%
  mutate(p_neg = round(p_neg, 2)) %>%
  filter(!is.na(p_neg)) -> text_data
text_data$p_neg[text_data$p_neg==1] <- .99

plot.dat2 %>% 
  filter(!is.na(weighted_bias)) %>%
  ggplot(aes(x=weighted_bias, y=est)) +
    geom_point(data=raw_or, aes(x=value, y=odds_r, size=nschools), alpha=.1) +
    geom_line() +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.25) +
    theme_classic() +
    scale_size_continuous(name='Number\nof schools') + 
    geom_text(data=text_data, parse=T, size=3.25,
              aes(x=-2.5, y=4.25, label=paste0('p[pos]', text_comp, round(p_neg,2)))) +
    #scale_color_manual(values=c('grey', 'darkorange'), guide=F) + 
    scale_alpha_manual(values=c(.2, 1), guide=F) +
    ylim(0,8) +
    xlim(-3, 3) + 
    xlab('Standardized Bias') +
    ylab('Relative \n Risk Ratio') +
    geom_hline(yintercept=1) +
    facet_grid(metric~bias) +
    theme(legend.position = 'top',
          strip.text.y=element_text(size=7),        
          axis.text = element_text(size=10),
          axis.title.x = element_text(size=12, face='bold'),
          legend.text = element_text(size=8),
          legend.title=element_text(size=8)) -> p2
ggsave('/Users/travis/Documents/gits/educational_disparities/figs/rel_risk.jpeg', p2, width = 11, height = 6)
ggsave('/Users/travis/Documents/gits/educational_disparities/figs/rel_risk.pdf', p2, width = 10, height = 6)

p1$data %>% filter(weighted_bias %in% c(0, 1)) -> rates
p2$data %>% filter(weighted_bias %in% c(0, 1)) -> risks

#grid.arrange(p1, p2)
p2
```

To better illustrate the nature of these relationships, figure \@ref(fig:detail-risks) shows the estimated relative risk ratios for each type of bias measurement and each disciplinary metric. This quantity is the ratio of the probability that a black student will receive a given disciplinary action to the probability that a white students will receive a given disciplinary action. Higher values indicate that black students have a higher probability of being punished than white students. Examining the risk ratio for out-of-school suspensions, in a hypothetical county with 0 standardized bias on both implicit and explicit, the model predicts that for every white student suspended, we should expect `r round(risks$est[1], 2)` [`r round(risks$lower[1], 2)`, `r round(risks$upper[1], 2)`] black students to be suspended. If we move to a county one standard deviation above the mean of explicit bias, the ratio of black to white students suspended increases to `r round(risks$est[2], 2)` [`r round(risks$lower[2], 2)`, `r round(risks$upper[2], 2)`], while the same movement for implicit bias decreases the ratio to `r round(risks$est[4], 2)` [`r round(risks$lower[4], 2)`, `r round(risks$upper[4], 2)`].

## Additional analyses: Sexuality bias as predictor
```{r iat-data-sex, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
df_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2004.sav')
#no county data in 2005
#df2 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2005.sav')
df3_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2006.sav')
df4_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2007.sav')
df5_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2008.sav')
df6_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2009.sav')
df7_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2010.sav')
df8_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2011.sav')
df9_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2012.sav')
df10_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2013.sav')
df11_s <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2014.sav')
sex_county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means_sexuality.csv')

df_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age)  -> subdat

df3_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age) %>%
  rbind(subdat) -> subdat

df4_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age) %>%
  rbind(subdat) -> subdat

df5_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age) %>%
  rbind(subdat) -> subdat

df6_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age) %>%
  rbind(subdat) -> subdat

df7_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age) %>%
  rbind(subdat) -> subdat

df8_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age) %>%
  rbind(subdat) -> subdat

df9_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age) %>%
  rbind(subdat) -> subdat

df10_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age) %>%
  rbind(subdat) -> subdat

df11_s %>%
  filter(CountyNo!='' &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.Straight_Good_all, tgaymen, 
         tgaywomen, tstraightmen, tstraightwomen, age) %>%
  rbind(subdat) -> subdat

subdat %>%
  mutate(explicit_bias=((tstraightmen-tgaymen) + (tstraightwomen-tgaywomen))/2) %>%
  mutate(age_bin = cut(age, breaks=c(14, 24, 34, 54, 75, 120))) %>%
  filter(!is.na(age_bin)) %>%
  filter(STATE %ni% 
           c('AA', 'AE', 'AP', 'AS', 'FM', 'GU', 'MH', 'MP', 'PR', 'VI')) %>% #exclude territories, etc
  mutate(county_id = paste(STATE, CountyNo, sep='-')) -> individual_data_sex

```
We sought to test whether the relationships observed above were specific to estimates of racial bias. Accordingly, we ran the same set of analyses with sexuality bias replacing racial bias.

### Project Implicit Estimates
The unstandardized county-level estimates of bias adjusted with poststratification, evidence a pro-straight bias in both implicit ($mean$ = `r round(mean(as.numeric(sex_county_means$weighted_bias), na.rm=T), 2)`, $sd$ = `r round(sd(as.numeric(sex_county_means$weighted_bias), na.rm=T), 2)`) and explicit measures ($mean$ = `r round(mean(as.numeric(sex_county_means$weighted_explicit), na.rm=T), 2)`, $sd$ = `r round(sd(as.numeric(sex_county_means$weighted_explicit), na.rm=T), 2)`), where on both scales 0 = no bias, and positive numbers indicate a pro-straight bias.

### Associations across counties
Figure \@ref(fig:overall-associations-sex) shows the estimate of primary interest for each of the models. The estimates displayed are the coefficients for the interaction between race and each of the two bias measurements. Given that Black Americans are the baseline group, negative values for this coefficient indicate that as one moves into counties with higher levels of bias, the gap between the probability of a black student being disciplined and the probability of a white student being disciplined grows. Table \@ref(tab:reg-coefs-sex) displays the estimated coefficients and 95% uncertainty intervals for each of the fixed effects for each model.

```{r overall-associations-sex, fig.cap="Association between each metric and county-level estimates of explicit and implicit bias. Negative values indicate that the rate of increase (or decrease) for blacks is faster (or slower) than for whites. Point is the mean of the posterior and error bars represent 95% bayesian uncertainty intervals.", results='hide', echo=F, message=F, warning=F}
#the p_neg for oos_susp & implicit bias is 100%, so when reporting the number in-text, i just replace it with 1 to prevent it from being rendered as NA in the document.
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/sex_bias/'
plot.dat <- data.frame(group = NA, weighted_bias = NA, est = NA, lower = NA, upper=NA, metric = NA)
plot.dat2 <- data.frame(weighted_bias = NA, est = NA, lower=NA, upper=NA, metric=NA)
plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg=NA, metric=NA)

# weighted bias w/exclusions & explicit difference
metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  raw_rate <- data.frame(county_id=NA, nschools=NA, weighted_bias_sex=NA, 
                         group=NA, nincidents=NA, nstudents=NA, rate=NA)
  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    m$data %>% 
      select(county_id, COMBOKEY) %>% 
      distinct() %>% 
      group_by(county_id) %>% 
      summarise(nschools=n()) -> n_schools
    m$data %>%
      group_by(county_id, group) %>%
      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
      mutate(rate=nincidents/nstudents) -> rate
    n_schools %>%
      left_join(m$data[, c('county_id', 'weighted_bias_sex')]) %>%
      distinct() %>%
      left_join(rate) %>%
      rbind(raw_rate) %>%
      filter(!is.na(county_id)) -> raw_rate
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }

  data.frame(est=mean(p_cons[17,]),
             lower=quantile(p_cons[17,], .025),
             upper=quantile(p_cons[17,], .975),
             p_neg = prop.table(table(p_cons[17,]<0))[2],
             metric=i) %>%
    rbind(plot.dat3) %>%
    filter(!is.na(est)) %>%
    mutate(metric=reorder(metric, est)) -> plot.dat3
}

plot.dat3_all <-plot.dat3
plot.dat3_all$bias <- 'Implicit'

#weighted warmth w/exclusions & explicit difference
plot.dat <- data.frame(group = NA, weighted_explicit_sex = NA, est = NA, lower = NA, upper=NA, metric = NA)
plot.dat2 <- data.frame(weighted_explicit_sex = NA, est = NA, lower=NA, upper=NA, metric=NA)
plot.dat3 <- data.frame(est=NA, lower=NA, upper=NA, p_neg = NA, metric=NA)

for(i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  raw_rate_warmth <- data.frame(county_id=NA, nschools=NA, weighted_explicit_sex=NA, group=NA, nincidents=NA, nstudents=NA, rate=NA)
  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    m$data %>% 
      select(county_id, COMBOKEY) %>% 
      distinct() %>% 
      group_by(county_id) %>% 
      summarise(nschools=n()) -> n_schools
    m$data %>%
      group_by(county_id, group) %>%
      summarise(nincidents=sum(number), nstudents=sum(total_number)) %>%
      mutate(rate=nincidents/nstudents) -> rate
    n_schools %>%
      left_join(m$data[, c('county_id', 'weighted_explicit_sex')]) %>%
      distinct() %>%
      left_join(rate) %>%
      rbind(raw_rate_warmth) %>%
      filter(!is.na(county_id)) -> raw_rate_warmth
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }

  data.frame(est=mean(p_cons[18,]),
             lower=quantile(p_cons[18,], .025),
             upper=quantile(p_cons[18,], .975),
             p_neg = prop.table(table(p_cons[18,]<0))[2],
             metric=i) %>%
    rbind(plot.dat3) %>%
    filter(!is.na(est)) %>%
    mutate(metric=reorder(metric, est)) -> plot.dat3
}

plot.dat3$bias <- 'Explicit'

plot.dat3_all <- rbind(plot.dat3_all, plot.dat3)
plot.dat3_all$metric <- factor(plot.dat3_all$metric, c('oos_susp', 
                                                       'inschool_susp',
                                                       'law_enforcement', 
                                                       'expulsion_combined',
                                                       'in_school_arrest'))
plot.dat3_all$metric <- fct_recode(plot.dat3_all$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')

pl <- ggplot(plot.dat3_all, aes(x=metric, y=est, group=bias, color=bias, shape=bias)) +
  geom_point(position=position_dodge(width=.5), size=3) +
  geom_errorbar(aes(ymin=lower, ymax=upper), 
                width=.5, position=position_dodge(width=.5)) +
  theme_classic() +
  coord_flip() +
  geom_hline(yintercept=0) +
  ylab('Black-White Difference in log-odds slope') +
  theme(axis.title.y = element_blank(),
        legend.title = element_blank(),
        legend.position='top',
        axis.text = element_text(size=10),
        axis.title.x = element_text(size=12, face='bold'),
        legend.text = element_text(size=10))
ggsave('/Users/travis/Documents/gits/educational_disparities/figs/slope_ests_sex.jpeg', pl, width = 11, height = 6)
pl
```

```{r, reg-coefs-sex, results='asis', echo=F, message=F, warning=F}
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/sex_bias/'
plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA)

# weighted bias w/exclusions & explicit difference
metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'in_school_arrest', 'law_enforcement', 'oos_susp')#1245
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  for (k in files){
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }
  data.frame(coef=colnames(dat)[1:30],
             est=rowMeans(p_cons),
             lower=apply(p_cons,1,quantile, probs=.025),
             upper=apply(p_cons,1,quantile, probs=.975),
             p_neg = rowSums(p_cons<0)/4000,
             metric=rep(i, 30)) %>%
    rbind(plot.dat) %>%
    filter(!is.na(est)) -> plot.dat
}

plot.dat$metric <- factor(plot.dat$metric, c('oos_susp', 'inschool_susp',
                                             'law_enforcement', 
                                             'expulsion_combined',
                                             'in_school_arrest'))

plot.dat$metric <- fct_recode(plot.dat$metric,
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat$coef <- fct_recode(plot.dat$coef,
                            'Intercept' = '(Intercept)',
                            'total population' = 'total_pop',
                            'proportion black' = 'black_prop',
                            'proportion white' = 'white_prop',
                            'black-white ratio' = 'b.w.ratio',
                            'college grads' = 'col_grads',
                            'income' = 'med_income',
                            'poverty' = 'poverty_rate',
                            'unemployment' = 'unemp_rate',
                            'crime' = 'crime_rate',
                            'housing density' = 'density',
                            'mobility' = 'mobility',
                            'dissimilarity' = 'dissim',
                            'total population*race: white' = 'groupwhite:total_pop',
                            'proportion black*race: white' = 'groupwhite:black_prop',
                            'proportion white*race: white' = 'groupwhite:white_prop',
                            'black-white ratio*race: white' = 'groupwhite:b.w.ratio',
                            'college grads*race: white' = 'groupwhite:col_grads',
                            'income*race: white' = 'groupwhite:med_income',
                            'poverty*race: white' = 'groupwhite:poverty_rate',
                            'unemployment*race: white' = 'groupwhite:unemp_rate',
                            'crime*race: white' = 'groupwhite:crime_rate',
                            'housing density*race: white' = 'groupwhite:density',
                            'mobility*race: white' = 'groupwhite:mobility',
                            'dissimilarity*race: white' = 'groupwhite:dissim',
                            'race: white' = 'groupwhite',
                            'implicit bias' = 'weighted_bias_sex',
                            'explicit bias' = 'weighted_explicit_sex',
                            'implicit bias*race: white' = 'groupwhite:weighted_bias_sex',
                            'explicit bias*race: white' = 'groupwhite:weighted_explicit_sex')

plot.dat %>%
  filter(p_neg<.025 | p_neg>.975) %>%
  mutate(est = paste('\\bf{', round(est, 2), '}', sep='')) -> sigs
plot.dat %>%
  filter(p_neg>=.025) %>%
  filter(p_neg<=.975) %>%
  mutate(est = round(est, 2)) %>%
  rbind(sigs) %>%
  mutate(estimate=paste(est, ', [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
  select(metric, estimate, coef) %>%
  spread(metric, estimate) -> tabledata_sex

tabledata_sex <- tabledata_sex[c(1:7, 23:27, 30, 8, 28, 29, 9:19, 22, 20, 21),]
names(tabledata_sex)[1] <- ''

stargazer::stargazer(tabledata_sex, summary = F, label='tab:reg-coefs-sex', title = 'Regression coefficient estimates for the population-level (i.e. fixed) effects, along with 95\\% uncertainty intervals for each of the disciplinary metrics.', header = F, rownames=F)
```

This figure illustrates that in general, county-level explicit and implicit biases in favor of straight individuals are not consistently associated with racial disciplinary disparities. The one exception is for law enforcement referrals, where we see that counties with higher levels of implicit bias are expected to have a larger racial gap ($M_{slope}$ = `r round(plot.dat3_all$est[2], 2)`, [`r round(plot.dat3_all$lower[2], 2)`, `r round(plot.dat3_all$upper[2], 2)`]). For this particular disciplinar action, explicit biases tend to show the opposite association, with counties higher in explicit bias expected to have a smaller racial gap, though the evidence is not as conclusive ($M_{slope}$ = `r round(plot.dat3_all$est[7], 2)`, [`r round(plot.dat3_all$lower[7], 2)`, `r round(plot.dat3_all$upper[7], 2)`]). All other estimates are located near zero, or are inconclusive with respect to the direction of the effect.

# Discussion
These analyses across five types of disciplinary actions are fully consistent with county-level estimates of explicit racial bias being associated with racial disciplinary disparities. Specifically, in counties where the white population is estimated to have higher rates of explicit biases that favor whites, the difference in suspensions, expulsions, law enforcement referrals, and school-related arrests between black and white students is expected to be greater than in those counties where the white population has lower rates of explicit racial biases. In a secondary set of analyses with sexuality bias, we found evidence that the aforementioned pattern cannot be explained by a more general tendency to exhibit explicit biases to stigmatized groups, as opposed to racial biases specifically, since most of the associations with respect to county-level explicit and implicit sexuality bias are directionally inconclusive. Additionally, our analyses cover the vast majority of school-aged students in the United States, and our models include a large set of covariates, suggesting that the relationships observed are not attributable to other indicators that can often co-occur with racial disparities, such as socioeconomic status or population demographics [@williams1999race; @skiba2002color].

Of course, this research is not without limitations and anomalies. The relationship between implicit racial bias and suspension was in an unexpected direction. That is, greater implicit bias is somewhat associated with smaller racial disparities in suspensions. One possibility is that counties higher in implicit bias are not as likely to suspend black students, but rather opt to use other disciplinary tools unmeasured in these data. Another possibility is that people in counties characterized by higher implicit bias are somewhat willing and able to correct their behavioral responses in order to conform to their egalitarian explicit beliefs or egalitarian social norms [@pryor2004dual. Much research at the individual level shows that people are able to counteract the automatic impulses associated with implicit bias when they have the motivation and cognitive bandwidth to do so [@devine2002regulation]. Additionally, because of the correlational nature of the analyses, it is impossible to definitively establish the causal relationship between explicit bias and disciplinary disparities. The conclusion that explicit biases predict disciplinary disparities is consonant with a great deal of research on disciplinary disparities [@mcintosh2014education. However, it is also possible that living in a region in which black students are disciplined to a greater extent than white students exacerbates and/or reinforces the explicit racial biases of community members, or that the relationship between explicit racial biases and disciplinary disparities is bi-directional. Finally, although we used a poststratification scheme to make our estimates more representative with respect to county age distributions, we cannot account for other ways in which Project Implicit data are not representative of the general population. Our analyses trade off the ability to ask these more detailed question with the strengths of statistical power and population coverage offered by the large datasets we used here. A full understanding of the complex issues shaping racial disciplinary disparities will require integrating findings from carefully controlled experiments, targeted surveys, administrative databases, and large-scale observational data.

Nevertheless, our work compliments other research indicating that racial dynamics are an important source for the observed differences in disciplinary rates between black and white students. For instance, students, caregivers, and administrators perceive suspensions and the disproportionate use of them as at least partially racially motivated [@haight2014ecological; @gibson2013caregivers; @ruck2002racial]. Additionally, other work has shown that even after controlling for a range of other factors, race remains associated with the likelihood of receiving disciplinary actions [@anyon2014persistent; @skiba2002color]. Additionally, experimental evidence shows that disciplinary decision-making for teachers differs depending on the race of the student [@okonofua2015two]. The present research adds to this work by showing for the first time associations between disciplinary actions and measurements of racial bias.

Our work also compliments existing studies examining the degree to which implicit and explicit racial biases are associated with racial disparities in key areas, such as health and policing [@leitner2016blacks; @orchard2017county; @hehman2017disproportionate] by extending this type of inquiry to educational outcomes. To properly assess the meaning of these findings, it is imperative that future work focus specifically on what it means to exist in a community that is estimated to have high or low levels of implicit or explicit bias.

As we have already highlighted, students who are subject to the disciplinary actions we examined here are at substantially higher risk for negative life outcomes [@hirschfield2009another]. In dispensing these disciplinary decisions differentially across racial groups, educational agencies are also differentially allocating life prospects. Although implementing training and education focused on implicit bias is popular among educational administrators [@staats2016understanding], and has been recommended by the US Department of Education [@us2014dear], our analysis suggests that effectively combating racial disparities in discipline requires confronting more explicit forms of racial bias, perhaps, for example, through peer influence or interventions designed to change social norms [@paluck2009prejudice].

We offer the research presented here to prompt additional scrutiny with
respect to how and why educational agencies in the United States
differentially administer disciplinary actions, especially when those
actions are known to have dire consequences for student welfare. Although the focus here is on actions taking place within educational agencies, the mechanisms responsible for these disparities likely exist, at least partly, in the larger community. Through understanding and reducing these disciplinary disparities specifically and the biases that exist in the community more broadly, there exists an avenue for education to maximize its promise as the great equalizer it has the potential to be.

## Materials & Methods
### Analytic Approach
Because Project Implicit is a nonrandom sample, we used multilevel regression and post-stratification to obtain accurate geographical population-based estimates of implicit and explicit bias. This procedure corrects for biased sampling and regularizes extreme observations with little data to support them (e.g. a county with only a handful of respondents with especially high or low scores) [@park2004bayesian; @gelman1997poststratification]. Following past work [@leitner2016blacks], we identified age as one dimension along which IAT respondents differed from the general population in ways that could bias our conclusions [@gonsalkorale2009aging]. Our post-stratification weighting scheme is as follows: We first grouped respondents into five age group categories (15-24, 25-34, 35-54, 55-75, and 75+). We next fit multilevel models estimating bias (implicit and explicit biases separately) as a function of our state-level covariates (the "fixed" effects: all identical to those included in the final model, described below), and allowed the estimates to vary by age bin, county, and state (the "random" effects). Next, we determined the population of whites in each county in these age groups using the American Community Survey's 5-year estimates ending in 2014. Finally, we used our estimated models to predict the expected response for each age bin, in each county. Our final county-level estimates are the average of the values predicted for the 5 age bins, weighted by the population size of that bin in that county. As a result of this procedure, we can be confident that our county-level estimates should more closely approximate what our estimates would look like if the Project Implicit data were truly representative along the age dimension in all counties.

We analyzed these data using a series of bayesian multilevel logistic regressions. We modeled the probability that a student would be expelled as a function of a set of effects that are constant across observations (i.e. fixed effects: race (dummy coded), implicit bias, explicit bias, an interaction term between race and implicit bias, an interaction term between race and explicit bias, and all covariates described below, plus each of the covariates interacting with race) and a set of effects that vary across counties (i.e. random effects: overall intercept & race). We fit separate models for each of the outcomes. All numerical predictors were standardized at the appropriate level (county, state) before model estimation (for post-stratification as well as for final inference) to help with estimation efficiency and interpretability. We also set priors for the intercept and coefficients in the bayesian model to be weakly informative normal distributions centered on zero with a standard deviation of five. This corresponds to a prior belief that all parameters take values between -15 and 15 with a probability of more than .99. Realistically, values outside this range are extremely unlikely given that all variables were standardized prior to estimation. Priors on all other parameters (e.g. variance of the county-level intercepts) were left to default settings from the software package used to fit the models.

Because of the computational demands of fitting such a high-dimensional model to such a large dataset (the full model for each metric would consist of over 6k parameters to approximately 170k observations), we used a consensus monte carlo algorithm to obtain approximate posterior distributions for the parameters of interest [@scott2016bayes]. The approximate posteriors derived from this algorithm have been shown to be nearly indistinguishable from the true posterior, a result we verified using a small subset of our own data.
### Data Sources
#### Disciplinary actions
```{r summary-data, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

schools_dat %>%
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_ENR_HI_M:TOT_ENR_F) %>%
  gather(group, total_number, SCH_ENR_HI_M:TOT_ENR_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>%
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='NR_', 
                          biracial='TR_', white='WH_')) %>%
  group_by(COMBOKEY, group) %>%
  mutate(total_number = sum(total_number)) %>%
  ungroup() %>%
  select(-prefix, -gender) %>%
  distinct() -> enrollment_sample

enrollment_sample %>%
  group_by(group) %>%
  summarise(ns = sum(total_number)) -> enrollment_totals

model.data %>% 
  filter(exclude==FALSE) -> kept_schools

model.data %>%
  filter(exclude==FALSE) %>%
  select(COMBOKEY, group, total_number) %>%
  distinct() %>%
  mutate(total_schools = n_distinct(COMBOKEY)) %>%
  group_by(group) %>%
  summarise(ns = sum(total_number),
            n_sch = n_distinct(COMBOKEY),
            total_schools = max(total_schools)) -> enrollment_mod_data

enrollment_sample %>%
  filter(COMBOKEY %in% kept_schools$COMBOKEY) %>%
  group_by(group) %>%
  summarise(ns = sum(total_number),
            n_sch = n_distinct(COMBOKEY)) -> filtered_dat

```

To assess rates of disciplinary action, we used data from the Civil Rights Data Collection (CRDC) conducted by the US Department of Education. The dataset we used comes from the 2013-2014 academic year and has data on "all public local and educational agencies and schools, including long-term secure juvenile justice facilities, charter schools, alternative schools, and schools serving students with disabilities." In total, the CRDC data represents `r dim(schools_dat)[1]` institutions enrolling approximately `r round(enrollment_totals$ns[6]/1000000, 1)` million students, of which approximately `r round(enrollment_totals$ns[8]/1000000, 1)` million are white and `r round(enrollment_totals$ns[3]/1000000, 1)` million are Black[^1]. Previous work using CRDC data have identified a number of districts whose data are in error, and have excluded juvenile justice facilities, as these institutions constitute dramatically different educational environments, where the meaning of disciplinary actions may be quite different [@losen2015we]. We followed similar practices, excluding all juvenile justice facilities. Additionally, we excluded data for a specific disciplinary metric for any schools which reported disciplining more students than it reported enrolling for *any* race for that metric (e.g. a school reported expelling 5 white students when they reported enrolling less than that number). We also excluded any school for all disciplinary actions if they had an overreporting error for 3 or more metrics. After these exclusions are applied, the final sample used for modeling consists of `r filtered_dat$n_sch[1]` institutions, enrolling `r round(filtered_dat$ns[6]/1000000, 1)` million students of which `r round(filtered_dat$ns[8]/1000000,1)` million are white and `r round(filtered_dat$ns[3]/1000000, 1)` million are black.[^2] From these data, we focus on the number of students by race (black and white) who were subjected to each of the disciplinary actions described below.

[^1]: We note that we initially preregistered a number of analyses concerning this work. There are a number of differences between the analyses we registered and those presented in the main text. The registered analyses can be found, in full, in the project's OSF page (https://osf.io/pu79a/).

[^2]: Researchers who compared individual schools' out-of-school suspension rates reported in the 2011-2012 CRDC with suspension rates of the same schools as reported on state websites where available found a number of mismatches [@losen2015we]. We did not have the resources to conduct similar comparisons for all outcomes in all `r filtered_dat$n_sch[1]` institutions; however, informal inspection of the data led us to find one school district for which this was the case. Including versus excluding schools from this district do not change the results in any meaningful way. The analyses presented herein include this district. The results of corresponding analyses without this district can be requested from the first author. It should be noted that the United States Government Accountability Office's recent descriptive overview of the 2013-2014 CRDC also did not exclude cases for which there was a discrepancy between disciplinary rates reported in this dataset and any disciplinary data reported on state websites[@usgao2018discipline]. As such, the analyses here are consistent with the only other published examination of the 2013-2014 CRDC data that we are aware of.

##### Types of disciplinary actions
We report here on five different types of disciplinary actions: in-school suspensions, out-of-school suspensions, law enforcement referrals, school-related arrests, and expulsions.

The most extreme outcome, *expulsion*, is defined as when a student is prohibited from returning to the educational institution for the remainder of the school year or longer. The institution may or may not set up alternative educational services for the student. *Law enforcement referrals* are actions where a student is "reported to any law enforcement agency or official for an incident including a school police unit, for an incident that occurs on school grounds, during school-related events, or while taking school transportation, regardless of whether official action is taken.  Citations, tickets, and court referrals are considered referrals to law enforcement." *School-related arrests* refer to "an arrest of a student for any activity conducted on school grounds, during off-campus school activities (including while taking school transportation)" *Out-of-school suspensions* are actions where the student "is temporarily removed from his or her regular school for at least half a day (but less than the remainder of the school year)". Finally, *in-school suspensions*  are actions where the student is "temporarily removed from his or her regular classroom(s) for at least half a day...,but remains under the direct supervision of school personnel. Direct supervision means school personnel are physically in the same location as the student under their supervision." [@crdc2013survey, pg. 51]

#### Racial Bias
We used measurements of implicit and explicit bias available from data collected through Project Implicit [@xu2014psychology]. For a full description of the implicit and explicit bias measures available in these data, refer to @xu2014psychology and @leitner2016blacks. We note only that we used the IAT D-score as a measure of implicit bias, and the difference between reported warmth towards whites and warmth towards black (both measured from 0=very cold to 10=very warm) as a measure of explicit bias. Additionally, we used only respondents who had geographic information that would allow us to place them in a United States county, identified as White, and visited the site anytime after it went live in 2002 through the end of 2014. This consisted of approximately `r round(dim(individual_data)[1]/1000000, 1)` million total respondents from `r dim(county_means)[1]` counties. `r round(table(is.na(individual_data$D_biep.White_Good_all))[1]/1000000, 2)` million respondents provided data for the IAT, and `r round(table(is.na(individual_data$explicit_bias))[1]/1000000, 2)` million provided explicit bias ratings.

#### Sexuality Bias

We used measurements of implicit and explicit sexuality bias available from data collected through Project Implicit [@xu2014psychology]. We opted for these measures as robustness checks because there were enough observations to closely mimic the racial bias analyses, and they have been used for similar purposes in previous work [@orchard2017county] As in the previous study, we used the IAT D-score as a measure of implicit bias. Our warmth score represented the average of the difference between straight men and gay men and the difference between straight women and lesbian women. We used only respondents who had geographic information that would allow us to place them in a United States county and visited the anytime between when it went live in 2002 through the end of 2014. This consisted of approximately `r round(dim(individual_data_sex)[1]/1000)`K total respondents from `r dim(sex_county_means)[1]` counties. Of these respondents, `r round(table(is.na(individual_data_sex$D_biep.Straight_Good_all))[1]/1000)`K respondents provided data for the IAT, and `r round(table(is.na(individual_data_sex$explicit_bias))[1]/1000)`K respondents provided explicit bias ratings.

#### Covariates
Each county-level variable used as a covariate in the final model and the corresponding state-level variable used as a predictor in the post-stratification scheme (described below) were taken from the same source. Population size and proportions, socioeconomic indicators, mobility, and segregation indices were all taken from the American Community Survey (ACS) 5-year estimates for the time period ending in 2014. Urban-rural indicators were taken from the 2010 US Census, and crime rates were taken from the FBI Uniform Crime Reporting program, as made available through the National Archive of Criminal Justice Data for each year from 2010-2014[^3]. Each of these variables is described below.

[^3]: These years are chosen to for the FBI data to yield estimates that are similar to the 5-year estimates from the ACS data.

##### Population size and proportions
We obtained the total population, the proportion of the population that is white, the proportion of the population that is black, and the ratio of black-to-white people in the population from ACS table B02001.

##### Socioeconomic indicators
We obtained estimates for the percentage of population with a Bachelor's degree or higher, the percentage of the population aged 16 or over in the labor force that is unemployed, the median household income, and the percentage of families and people whose income in the last year was below the poverty line from the ACS table DP03.

##### Urban-rural indicator
We obtained estimates of housing density per square mile of land area from Census table GCT-PH1.

##### Mobility
We obtained estimates of population mobility by summing the percentage of Black Americans who moved from a different county, state, or country into the county of interest (county-level covariate) or who moved from a different state or country into the state of interest (state-level covariate). We took these metrics from the ACS table S0701.

##### Crime
We computed estimates of the number of violent crimes per person by taking the number of crimes reported divided by the population size for each year, and averaging the resulting proportions across the 5 years of data.

##### Segregation
We computed a dissimiliarity index as described by @massey1988dimensions. This metric is computed based on the racial dissimilarity of census tracts within a county, and reflects the proportion of a group within the county that would have to move in order for all census tracts to have group distributions that matched the overall distribution of the county. These computations were done using data from the ACS table B02001.

<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak

### 2018-05-18 written_docs/Riddle_Sinclair_PNAS/osf_supplement.Rmd
PNAS submitted

---
title: "OSF Supplementary Information"
author: "Travis Riddle & Stacey Sinclair"
date: "5/7/2018"
output: 
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
bibliography: ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache.lazy = FALSE, warning = FALSE, 
                      cache=TRUE, message=FALSE)
library(knitr)
library(dplyr)
library(ggplot2)
library(haven)
library(forcats)
library(tidyr)
library(kableExtra)
library(stringr)
```

```{r iat-data, echo=FALSE, results='hide', message=FALSE, cache=TRUE}

#Aggregate IAT data.
df <- rbind(read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2013.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2012.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2010.sav'))

df2 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2011.sav')
df3 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2009.sav')
df4 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2008.sav')
df5 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2007.sav')
df6 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2006.sav')
df7 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2005.sav')
df8 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2004.sav')
df9 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2002-2003.sav')
df10 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2014.sav')

"%ni%" <- Negate("%in%")

df %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  filter(raceomb==6) -> subdat

df2 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all,  
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df3 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df4 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df5 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df6 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df7 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df8 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df9 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(raceomb = ethnic) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

df10 %>%
  filter(CountyNo!='' &
           #!is.na(D_biep.White_Good_all) &
           #!is.na(tblack_0to10) &
           !is.na(age)) %>%
  dplyr::select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==6) -> subdat

subdat %>%
  mutate(explicit_bias=tblack_0to10) %>%
  #filter(!is.na(twhite_0to10)) %>%
  mutate(explicit_bias = twhite_0to10 - tblack_0to10) %>%
  mutate(age_bin = cut(age, breaks=c(14, 24, 34, 54, 75, 120))) %>%
  mutate(race = fct_recode(as.character(raceomb), 
                           'Black'='5', 'White'='6')) %>%
  filter(!is.na(age_bin)) %>%
  filter(race=='White') %>%
  filter(STATE %ni% 
           c('AA', 'AE', 'AP', 'AS', 'FM', 'GU', 'MH', 'MP', 'PR', 'VI')) %>%
  mutate(county_id = paste(STATE, CountyNo, sep='-')) -> individual_data

```

# Session details
Data analysis was done in R [@rcitation] version 3.3.2 running under OS X 10.11.6. Post-stratification was done with lme4, version 1.1.14 [@bates2015fitting]. Final model fitting was done on the university cluster running Springdale Linux, release 6.9 using rstanarm, version 2.17.2 [@rstanarm2016]. We used the implementation of the consensus monte carlo algorithm found in parallelMCMCcombine, version 1.0 [@miroshnikov2014parallel]. Figures were made with ggplot2, version 2.2.1 [@wickham2009ggplot2], with data manipulation done using dplyr version 0.7.2 [@wickham2017dplyr] and tidyr, version 0.7.1 [@wickham2017tidyr]. Session info for computations on the local machine is as follows:

```{r sessioninfo}
sessionInfo()
```

Session info for the computations run on the university cluster is as follows:

```
R version 3.4.3 (2017-11-30)
Platform: x86_64-redhat-linux-gnu (64-bit)
Running under: Springdale Linux 7.4 (Verona)

Matrix products: default
BLAS/LAPACK: /usr/lib64/R/lib/libRblas.so

locale:
 [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C
 [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8
 [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8
 [7] LC_PAPER=en_US.UTF-8       LC_NAME=C
 [9] LC_ADDRESS=C               LC_TELEPHONE=C
[11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
[1] rstanarm_2.15.3 Rcpp_0.12.13    ggplot2_2.2.1   forcats_0.2.0
[5] tidyr_0.7.2     dplyr_0.7.4

loaded via a namespace (and not attached):
 [1] lattice_0.20-35      zoo_1.8-0            gtools_3.5.0
 [4] assertthat_0.2.0     digest_0.6.12        mime_0.5
 [7] R6_2.2.2             plyr_1.8.4           stats4_3.4.3
[10] colourpicker_1.0     rlang_0.1.2          lazyeval_0.2.1
[13] minqa_1.2.4          miniUI_0.1.1         nloptr_1.0.4
[16] Matrix_1.2-12        DT_0.2               shinythemes_1.1.1
[19] splines_3.4.3        shinyjs_0.9.1        lme4_1.1-14
[22] stringr_1.2.0        htmlwidgets_0.9      loo_1.1.0
[25] igraph_1.1.2         munsell_0.4.3        shiny_1.0.5
[28] compiler_3.4.3       httpuv_1.3.5         rstan_2.16.2
[31] pkgconfig_2.0.1      base64enc_0.1-3      rstantools_1.3.0
[34] htmltools_0.3.6      tibble_1.3.4         gridExtra_2.3
[37] threejs_0.3.1        codetools_0.2-15     matrixStats_0.52.2
[40] MASS_7.3-47          grid_3.4.3           nlme_3.1-131
[43] xtable_1.8-2         gtable_0.2.0         magrittr_1.5
[46] StanHeaders_2.16.0-1 scales_0.5.0         stringi_1.1.5
[49] reshape2_1.4.2       bindrcpp_0.2         dygraphs_1.1.1.4
[52] xts_0.10-0           tools_3.4.3          glue_1.2.0
[55] shinystan_2.4.0      markdown_0.8         purrr_0.2.4
[58] crosstalk_1.0.0      rsconnect_0.8.5      parallel_3.4.3
[61] inline_0.3.14        colorspace_1.3-2     bayesplot_1.4.0
[64] bindr_0.1
```

# Preregistered analysis
```{r summary_counts}
#aggregate corporal punishment and preschool metrics
schools_dat <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')
county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv')

###############
schools_dat %>%
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
  gather(group, total_number, SCH_PSENR_HI_M:TOT_PSENR_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>%
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='NR_', 
                          biracial='TR_', white='WH_')) %>%
  filter(total_number>=0) %>%
  group_by(COMBOKEY, group) %>%
  mutate(total_number = sum(total_number)) %>%
  ungroup() %>%
  select(-prefix, -gender) %>%
  distinct() -> ps_enrollment

schools_dat %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>% 
  filter(SCH_GRADE_PS=='YES') %>%
  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:TOT_PSDISC_SINGOOS_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>% 
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='OS_', 
                          biracial='TR_', white='WH_')) %>%
  select(-prefix) -> ps_susp

schools_dat %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
         SCH_GRADE_PS, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>% 
  filter(SCH_GRADE_PS=='YES') %>%
  gather(group, number_2, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>%
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='OS_', 
                          biracial='TR_', white='WH_')) %>%
  select(COMBOKEY, group, gender, number_2) %>%
  right_join(ps_susp) %>%
  filter(number>=0 & number_2 >= 0) %>%
  mutate(number = number+number_2) -> ps_susp

schools_dat %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
         SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>% 
  gather(group, number, SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>% 
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='XP_', 
                          biracial='TR_', white='WH_')) %>%
  filter(number>=0) %>%
  select(-prefix) -> ps_exp

ps_susp %>% group_by(group) %>% summarise(n_students=sum(number)) -> susp_summ
ps_exp %>% group_by(group) %>% summarise(n_students=sum(number)) -> exp_summ
ps_enrollment %>% group_by(group) %>% summarise(n_students=sum(total_number)) -> enr_summ

schools_dat %>%
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_ENR_HI_M:TOT_ENR_F) %>%
  gather(group, total_number, SCH_ENR_HI_M:TOT_ENR_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>%
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='NR_', 
                          biracial='TR_', white='WH_')) %>%
  group_by(COMBOKEY, group) %>%
  mutate(total_number = sum(total_number)) %>%
  ungroup() %>%
  select(-prefix, -gender) %>%
  distinct() -> enrollment

schools_dat %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
         SCH_DISCWODIS_CORP_HI_M:TOT_DISCWODIS_CORP_F, 
         SCH_DISCWDIS_CORP_IDEA_HI_M:TOT_DISCWDIS_CORP_IDEA_F) %>% 
  gather(group, number, SCH_DISCWODIS_CORP_HI_M:TOT_DISCWDIS_CORP_IDEA_F) %>%
  mutate(g=group) %>%
  separate(group, into=c('group', 'gender'), -2) %>% 
  separate(group, into=c('prefix', 'group'), -4) %>%
  separate(prefix, into=c('prefix', 'disability'), -3) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='RP_', 
                          total='EA_', biracial='TR_', white='WH_')) %>%
  mutate(disability=fct_recode(disability, not_disabled='P_', disabled='ID',
                               not_disabled='CO', disabled='A_')) %>%
  select(-prefix) -> corp

corp %>%
  filter(group=='black'|group=='white') %>%
  group_by(COMBOKEY, LEA_STATE, LEA_NAME, LEAID,
           CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
  summarise(number = sum(number, na.rm=T)) %>%
  left_join(enrollment) %>%
  mutate(proportion = number/total_number) -> corp_stats

corp %>% 
  select(COMBOKEY, LEA_STATE, number) %>% 
  group_by(COMBOKEY, LEA_STATE) %>% 
  summarise(number=sum(number, na.rm=T)) -> corp_stats

schools_dat$south <- schools_dat$LEA_STATE %in% c('AL', 'AR', 'FL', 'GA', 'LA', 'MO', 'MS', 'OK', 'TN', 'TX')
corp_stats %>% filter(number>0) -> uses_corp
uses_corp$south <- uses_corp$LEA_STATE %in% c('AL', 'AR', 'FL', 'GA', 'LA', 'MO', 'MS', 'OK', 'TN', 'TX')

```
Here, we describe, in detail, the differences between the registered analysis plan and that presented in the main text, along with our reasons for making the alterations.

## Data exclusions
In our preanalysis plan, we specified our analyses to focus on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.  However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes.

First, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others.

Next, we discovered that the administration of corporal punishment was extremely irregular across counties, with almost all cases ocurring in the South. Specifically, Alabama, Arkansas, Florida, Georgia, Louisiana, Missouri, Mississippi, Oklahoma, Tennessee, and Texas accounted for `r round((table(uses_corp$south)[2]/dim(uses_corp)[1])*100, 1)`% of all educational institutions that administered corporal punishment at least once, despite containing only `r round((table(schools_dat$south)[2]/dim(schools_dat)[1])*100, 1)`% of the educational institutions in the full dataset. It is not clear whether this uneven distribution is due to legal prohibitions, cultural differences, or some combination of the two. Because this makes interpretation of the parameters uncertain, this analysis is also placed in the appendix.

We also found that the number of preschool students who are expelled or suspended is vanishingly small (`r exp_summ$n_students[8]` total expulsions and `r susp_summ$n_students[6]` total suspensions out of over 1.4 million enrolled preschool students), making reliably estimating any association across counties exceedingly unlikely. We additionally discovered that counts of one expulsion category (expulsion under zero-tolerance policies) overlapped with counts in other categories, and so excluded this category from the main text. We also decided to combine the remaining two expulsion categories to yield one overall count of the number of students expelled. We did this to remain consistent with previous research, and to obtain district-level rates of expulsion that were slightly higher, and therefore more amenable to exploring variation across counties.

When preparing the preregistration, we also had not known about the issues with juvenile justice facilities, or with the school districts with reporting errors, and so excluded these schools from analyses in the main text.

## Changes in measurement & estimation of bias
We preregistered our explicit bias as a simple feeling thermometer towards balcks (i.e. *how warm or cold do you feel towards Blacks? 0=very cold, 10=very warm*). However, the majority of past research (see for example @leitner2016blacks and @hehman2017disproportionate)  has used the difference in reported warmth towards whites and blacks, and so in the main text, we report models using this metric of explicit bias.

Our main text presents analyses with poststratified estimates of bias. Our registered analysis also described analyses with raw, county-based means. Unlike previous papers that used poststratification, we see substantial differences in results between the two estimation methods. Because post-stratification is known to yield better estimates, we suspect that the difference in results is largely driven by extreme county-level observations with little data to support these extreme estimates. Nevertheless, in the interest of completion, we present the model using raw county-level means as well.

We also registered an exploratory analysis using the responses of people who visited Project Implicit and identified themselves as teachers. Unfortunately, without population-level data on the demographic characteristics of teachers, we are not able to post-stratify these estimates. Additionally, there were relatively few teachers in the Project Implicit from which to derive estimates. Ultimately, we have little confidence that this analysis tells us anything about whether the biases of teachers specifically contribute to disparities in disciplinary actions, and so this analysis is also limited to the appendix.

## Additional covariates
We included four covariates that were not specified in the registered analysis. Specifically, the mobility, housing density, crime rate, and dissimilarity covariates. We also interacted all covariates with race. These modifications were done in order to cover more of the covariates that were included in previous papers [@leitner2016blacks].

# Registered analysis results
Here, we present the results of the preregistered analyses exactly. Though we described a number of changes above, it is not practical to run all possible combination of changes, as each set of models takes several days to just to estimate, and there are at least several dozen possible combinations of decisions to test, depending on how fine-grained one wishes to be with the decisions. Accordingly, we limit the presentation of results below to just the three models that we registered.

## County-level estimates
```{r raw-county-estimates, fig.cap='\\label{fig:raw-county-estimates}Difference in raw county means and post-stratifed estimates of implicit bias as a function of the number of observations in each county. The X axis is on the log scale.'}
individual_data %>%
  group_by(county_id) %>%
  summarise(n_obs=n()) %>%
  left_join(county_means) -> cm

cm %>%
  mutate(post_diff = as.numeric(bias) - as.numeric(weighted_bias)) %>%
  ggplot(aes(x=n_obs, y=post_diff)) + 
  geom_point() + 
  scale_x_log10(breaks=c(1, 10, 100, 1000, 10000)) +
  theme_classic() +
  ylab('Raw Bias - Weighted Bias difference') +
  xlab('Number of Observations')

```

Examining the models fit with simple county means, we find that the county-level estimates vary more than with the post-stratified estimates. This pattern is expected, as one desireable feature of post-stratification is that it regularizes extreme observations without a large amount of data to support them. This phenomena is illustrated in figure \@ref(fig:raw-county-estimates), which shows the distribution of differences between the raw county means and the post-stratified estimates of implicit bias as a function of the number of observations in each county. More specifically, while the mean of the county-level means does not meaningfully differ from the post-stratified estimates, the standard deviation is much larger ($M_{implicit}$ = `r round(mean(as.numeric(cm$bias), na.rm=T), 2)`, $SD_{implicit}$ = `r round(sd(as.numeric(cm$bias), na.rm=T), 2)`; $M_{explicit}$ = `r round(mean(as.numeric(cm$explicit), na.rm=T), 2)`, $SD_{explicit}$ = `r round(sd(as.numeric(cm$explicit), na.rm=T), 2)`)[^4]

[^4]: This explicit estimate is based on the simple warmth towards blacks question, and not the difference between warmth towards whites and warmth twoards blacks.

The county-level implicit bias estimates for the poststratified model are as reported in the main text. The estimates for explicit bias are slightly different, as they are not computed based on the difference in reporting warmth toward whites and warmth toward blacks. Specifically, the average amount of bias at the county-level is `r round(mean(as.numeric(cm$weighted_explicit)*-1, na.rm=T), 2)` (SD = `r round(sd(as.numeric(cm$weighted_explicit), na.rm=T), 2)`).

```{r teachers-summary}
df <- rbind(read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2013.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2012.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2010.sav'))

educators <- c('25-2000', '25-3000')

df %>%
  filter(CountyNo!='') %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         twhite_0to10, tblack_0to10, raceomb, age, occupation)  -> subdat

df2 %>%
  filter(CountyNo!='') %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
  rbind(subdat) -> subdat

df3 %>%
  filter(CountyNo!='') %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
  rbind(subdat) -> subdat

df4 %>%
  filter(CountyNo!='') %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
  rbind(subdat) -> subdat

df5 %>%
  filter(CountyNo!='') %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
  rbind(subdat) -> subdat

df6 %>%
  filter(CountyNo!='') %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
  rbind(subdat) -> subdat

df10 %>%
  filter(CountyNo!='') %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         twhite_0to10, tblack_0to10, raceomb, age, occupation) %>%
  rbind(subdat) -> subdat

subdat %>%
  filter(raceomb==6) %>%
  mutate(explicit_diff=twhite_0to10 - tblack_0to10,
         explicit = tblack_0to10-1) %>%
  filter(STATE %ni% 
           c('AA', 'AE', 'AP', 'AS', 'FM', 'GU', 'MH', 'MP', 'PR', 'VI')) %>%
  filter(occupation %in% educators) %>%
  mutate(county_id = paste(STATE, CountyNo, sep='-')) -> individual_data_teachers

teacher_estimates <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_teacher_means.csv')
```

A subset of years in the Project Implicit data also collected occupational information from respondents. As identified in our pre-analysis plan, we took advantage of the presence of primary and secondary educators in these data to test whether any associations between bias and race-based differences in the rates of disciplinary action were stronger among these respondents. Filtering for only white individuals who identified as primary, secondary, special education, and other teachers and instructors (occupation codes 25-2000 and 25-3000) reduced the dataset to `r dim(individual_data_teachers)[1]` respondents. In order to assure that our estimates were reasonably stable, we limited analysis to only counties that had at least 50 respondents. As such, our teacher analysis is limited to just `r dim(teacher_estimates)[1]` counties. Additionally, because we do not know of any state-level demographic estimates for teachers, we were unable to perform post-stratification for these data. Nonetheless, across these limited counties, the overall estimate of implicit and explicit bias are similar to those for the whole dataset ($M_{implicit}$ = `r round(mean(as.numeric(teacher_estimates$teacher_bias), na.rm=T), 2)`, $SD_{implicit}$ = `r round(sd(as.numeric(teacher_estimates$teacher_bias), na.rm=T), 2)`; $M_{explicit}$ = `r round(mean(as.numeric(teacher_estimates$teacher_explicit*-1), na.rm=T), 2)`, $SD_{explicit}$ = `r round(sd(as.numeric(teacher_estimates$teacher_explicit), na.rm=T), 2)`).

For all models, explicit and implicit bias were scored such that higher scores reflect a greater preference for whites/reduced preference for blacks.

## Model estimates

Tables \@ref(tab:mod-raw-table1) and \@ref(tab:mod-raw-table2) present results from the model fit to the raw, county-level means (we broke results from each set of estimates into two table to improve legibility). The table presents parameter estimates (operationalized as the mean of the posterior) and 95% uncertainty intervals for each of the population-level effects for each of the outcomes. Tables \@ref(tab:mod-post-table1) and \@ref(tab:mod-post-table2) present the results for post-stratified estimates, with tables \@ref(tab:mod-teach-table1) and \@ref(tab:mod-teach-table2) showing results for the teacher models.

```{r mod-raw}
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/raw_means/'
plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA, model=NA)

metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
model_raw <- data.frame()
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(14,4000,length(files)))

  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    model_raw <- rbind(model_raw, m$data)
    posterior_combo[,,j] <- t(dat[,c(1:14)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (p in 1:14){
    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
  }

  data.frame(coef=colnames(dat)[1:14],
             est=rowMeans(p_cons),
             lower=apply(p_cons,1,quantile, probs=.025),
             upper=apply(p_cons,1,quantile, probs=.975),
             p_neg = rowSums(p_cons<0)/4000,
             metric=rep(i, 14),
             model = rep('A', 14)) %>%
    rbind(plot.dat) %>%
    filter(!is.na(est)) -> plot.dat
}
plot.dat$metric <- fct_recode(plot.dat$metric,
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Corporal Punishment' = 'corporal',
                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
                               'Expulsion with Educational Services' = 'expulsion_w_ed',
                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
                               'Mechanical Restraint' = 'mechanical_restraint',
                               'Physical Restraint' = 'physical_restraint',
                               'Preschool Expulsion' = 'preschool_expulsion',
                               'Preschool Suspension' = 'preschool_susp',
                               'Seclusion' = 'seclusion')

plot.dat %>%
  mutate(est = round(est, 2)) %>%
  mutate(estimate=paste(est, ' [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
  mutate(estimate=kableExtra::cell_spec(estimate, 'latex', bold=ifelse(p_neg<.025|p_neg>.975, TRUE, FALSE))) %>%
  dplyr::select(metric, estimate, coef) %>%
  spread(metric, estimate) -> tabledata_raw

tabledata_raw$coef <- fct_recode(tabledata_raw$coef,
                            'Intercept' = '(Intercept)',
                            'total population' = 'total_pop',
                            'proportion black' = 'black_prop',
                            'proportion white' = 'white_prop',
                            'black-white ratio' = 'b.w.ratio',
                            'college grads' = 'col_grads',
                            'income' = 'med_income',
                            'poverty' = 'poverty_rate',
                            'unemployment' = 'unemp_rate',
                            'race: white' = 'groupwhite',
                            'implicit bias' = 'bias',
                            'explicit bias' = 'warmth',
                            'implicit bias*race: white' = 'groupwhite:bias',
                            'explicit bias*race: white' = 'groupwhite:warmth')
tabledata_raw <- tabledata_raw[c(1,4,14,2,11,5,9,10,12,3,13,6,7,8),]
names(tabledata_raw)[1] <- ''
```

```{r mod-raw-table1, results='asis'}
subtab <- tabledata_raw[,c(1, 2, 5, 7, 14, 3, 4)]
knitr::kable(subtab, 
             caption = '\\label{tab:model-raw-table1}Regression coefficient estimates for the population-level (i.e. fixed) effects, 
             along with 95\\% uncertainty intervals for each of the disciplinary metrics. Model uses raw 
             county-means as bias estimates.', 
           format = 'latex', booktabs=T, escape=F, row.names = F,
             col.names=linebreak(c('', 'Seclusion', 'Physical\nRestraint',
                                   'Mechanical\nRestraint', 
                                   'Corporal\nPunishment',
                                   'Preschool\nSuspension', 
                                   'Preschool\nExpulsion'))) %>%
  landscape() %>%
  kableExtra::kable_styling(latex_options=c('scale_down'))

```

```{r mod-raw-table2, results='asis'}
subtab <- tabledata_raw[,c(1, 6, 8, 9, 10, 11, 12, 13)]
knitr::kable(subtab, 
             caption = '\\label{tab:model-raw-table1}Regression coefficient estimates for the population-level (i.e. fixed) effects, 
             along with 95\\% uncertainty intervals for each of the disciplinary metrics. Model uses raw 
             county-means as bias estimates.', 
           format = 'latex', booktabs=T, escape=F, row.names = F,
             col.names=linebreak(c('', 'Out-of-School\nSuspension', 
                                   'Law Enf.\nReferral', 
                                   'In-School\nSuspension',
                                   'School-Related\nArrest', 
                                   'Expulsion w/o\n Ed. Services', 
                                   'Expulsion w/Ed.\nServices', 
                                   'Expulsion Under\nZero-Tolerance'))) %>%
  landscape() %>%
  kableExtra::kable_styling(latex_options=c('scale_down'))

```

```{r mod-post}
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/mrp_estimates/'
plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA, model=NA)

metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
model_post <- data.frame()
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(14,4000,length(files)))

  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    model_post <- rbind(model_post, m$data)
    posterior_combo[,,j] <- t(dat[,c(1:14)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (p in 1:14){
    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
  }

  data.frame(coef=colnames(dat)[1:14],
             est=rowMeans(p_cons),
             lower=apply(p_cons,1,quantile, probs=.025),
             upper=apply(p_cons,1,quantile, probs=.975),
             p_neg = rowSums(p_cons<0)/4000,
             metric=rep(i, 14),
             model = rep('A', 14)) %>%
    rbind(plot.dat) %>%
    filter(!is.na(est)) -> plot.dat
}
plot.dat$metric <- fct_recode(plot.dat$metric,
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Corporal Punishment' = 'corporal',
                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
                               'Expulsion with Educational Services' = 'expulsion_w_ed',
                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
                               'Mechanical Restraint' = 'mechanical_restraint',
                               'Physical Restraint' = 'physical_restraint',
                               'Preschool Expulsion' = 'preschool_expulsion',
                               'Preschool Suspension' = 'preschool_susp',
                               'Seclusion' = 'seclusion')

plot.dat %>%
  mutate(est = round(est, 2)) %>%
  mutate(estimate=paste(est, ' [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
  mutate(estimate=kableExtra::cell_spec(estimate, 'latex', bold=ifelse(p_neg<.025|p_neg>.975, TRUE, FALSE))) %>%
  dplyr::select(metric, estimate, coef) %>%
  spread(metric, estimate) -> tabledata_post

tabledata_post$coef <- fct_recode(tabledata_post$coef,
                            'Intercept' = '(Intercept)',
                            'total population' = 'total_pop',
                            'proportion black' = 'black_prop',
                            'proportion white' = 'white_prop',
                            'black-white ratio' = 'b.w.ratio',
                            'college grads' = 'col_grads',
                            'income' = 'med_income',
                            'poverty' = 'poverty_rate',
                            'unemployment' = 'unemp_rate',
                            'race: white' = 'groupwhite',
                            'implicit bias' = 'weighted_bias',
                            'explicit bias' = 'weighted_warmth',
                            'implicit bias*race: white' = 'groupwhite:weighted_bias',
                            'explicit bias*race: white' = 'groupwhite:weighted_warmth')

tabledata_post <- tabledata_post[c(1,3,14,2,10,4,8,9,11,12,13,5,6,7),]
names(tabledata_post)[1] <- ''
```

```{r mod-post-table1, results='asis'}
subtab <- tabledata_post[,c(1, 2, 5, 7, 14, 3, 4)]
knitr::kable(subtab, 
             caption = '\\label{tab:model-post-table1}Regression coefficient estimates for the 
             population-level (i.e. fixed) effects, along with 95\\% 
             uncertainty intervals for each of the disciplinary metrics. 
             Model uses bias estimates as derived from poststratification 
             scheme.', 
             format='latex', booktabs=T, escape=F, row.names = F,
             col.names=linebreak(c('', 'Seclusion', 'Physical\nRestraint',
                                   'Mechanical\nRestraint', 
                                   'Corporal\nPunishment',
                                   'Preschool\nSuspension', 
                                   'Preschool\nExpulsion'))) %>%
  kableExtra::kable_styling(latex_options=c('scale_down')) %>%
  kableExtra::landscape()

```

```{r mod-post-table2, results='asis'}
subtab <- tabledata_post[,c(1, 6, 8, 9, 10, 11, 12, 13)]
knitr::kable(subtab, 
             caption = '\\label{tab:model-post-table2}Regression coefficient estimates for the 
             population-level (i.e. fixed) effects, along with 95\\% 
             uncertainty intervals for each of the disciplinary metrics. 
             Model uses bias estimates as derived from poststratification 
             scheme.', 
             format='latex', booktabs=T, escape=F, row.names = F,
             col.names=linebreak(c('', 'Out-of-School\nSuspension', 
                                   'Law Enf.\nReferral', 
                                   'In-School\nSuspension',
                                   'School-Related\nArrest', 
                                   'Expulsion w/o\n Ed. Services', 
                                   'Expulsion w/Ed.\nServices', 
                                   'Expulsion Under\nZero-Tolerance'))) %>%
  kableExtra::kable_styling(latex_options=c('scale_down')) %>%
  kableExtra::landscape()

```

```{r mod-teach}
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/teacher_metrics/'
plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA, model=NA)

metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'inschool_susp', 'oos_susp')
model_teacher <- data.frame()
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(14,4000,length(files)))

  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    model_teacher <- rbind(model_teacher, m$data)
    posterior_combo[,,j] <- t(dat[,c(1:14)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (p in 1:14){
    m$stanfit@sim$samples[[1]][[p]] <- p_cons[p,1:1000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[2]][[p]] <- p_cons[p,1001:2000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[3]][[p]] <- p_cons[p,2001:3000]
  }
  for (p in 1:14){
    m$stanfit@sim$samples[[4]][[p]] <- p_cons[p,3001:4000]
  }

  data.frame(coef=colnames(dat)[1:14],
             est=rowMeans(p_cons),
             lower=apply(p_cons,1,quantile, probs=.025),
             upper=apply(p_cons,1,quantile, probs=.975),
             p_neg = rowSums(p_cons<0)/4000,
             metric=rep(i, 14),
             model = rep('A', 14)) %>%
    rbind(plot.dat) %>%
    filter(!is.na(est)) -> plot.dat
}
plot.dat$metric <- fct_recode(plot.dat$metric,
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Corporal Punishment' = 'corporal',
                               'Expulsion Under Zero-Tolerance' = 'expulsion_0_tolerance',
                               'Expulsion with Educational Services' = 'expulsion_w_ed',
                               'Expulsion without Educational Services' = 'expulsion_wo_ed',
                               'Mechanical Restraint' = 'mechanical_restraint',
                               'Physical Restraint' = 'physical_restraint',
                               'Preschool Expulsion' = 'preschool_expulsion',
                               'Preschool Suspension' = 'preschool_susp',
                               'Seclusion' = 'seclusion')

plot.dat %>%
  mutate(est = round(est, 2)) %>%
  mutate(estimate=paste(est, ' [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
  mutate(estimate=kableExtra::cell_spec(estimate, 'latex', bold=ifelse(p_neg<.025|p_neg>.975, TRUE, FALSE))) %>%
  dplyr::select(metric, estimate, coef) %>%
  spread(metric, estimate) -> tabledata_teachers

tabledata_teachers$coef <- fct_recode(tabledata_teachers$coef,
                            'Intercept' = '(Intercept)',
                            'total population' = 'total_pop',
                            'propotion black' = 'black_prop',
                            'proportion white' = 'white_prop',
                            'black-white ratio' = 'b.w.ratio',
                            'college grads' = 'col_grads',
                            'income' = 'med_income',
                            'poverty' = 'poverty_rate',
                            'unemployment' = 'unemp_rate',
                            'race: white' = 'groupwhite',
                            'implicit bias' = 'county_bias',
                            'explicit bias' = 'county_warmth',
                            'implicit bias*race: white' = 'groupwhite:county_bias',
                            'explicit bias*race: white' = 'groupwhite:county_warmth')

tabledata_teachers <- tabledata_teachers[c(1,3,14,2,12,4,10,11,13,5,6,7,8,9),]
names(tabledata_teachers)[1] <- ''
```

```{r mod-teach-table1, results='asis'}
subtab <- tabledata_teachers[,c(1, 2, 5, 7, 14, 3, 4)]
knitr::kable(subtab, 
             caption = '\\label{tab:model-teach-table1}Regression coefficient estimates for the 
             population-level (i.e. fixed) effects, along with 95\\% 
             uncertainty intervals for each of the disciplinary metrics. 
             Model uses only data from Project Implicit respondents who 
             identified as teachers.', 
             format='latex', booktabs=T, escape=F, row.names = F,
             col.names=linebreak(c('', 'Seclusion', 'Physical\nRestraint',
                                   'Mechanical\nRestraint', 
                                   'Corporal\nPunishment',
                                   'Preschool\nSuspension', 
                                   'Preschool\nExpulsion'))) %>%
  kableExtra::kable_styling(latex_options=c('scale_down')) %>%
  kableExtra::landscape()
```

```{r mod-teach-table2, results='asis'}
subtab <- tabledata_teachers[,c(1, 6, 8, 9, 10, 11, 12, 13)]
knitr::kable(subtab, 
             caption = '\\label{tab:model-teach-table2}Regression coefficient estimates for the 
             population-level (i.e. fixed) effects, along with 95\\% 
             uncertainty intervals for each of the disciplinary metrics. 
             Model uses only data from Project Implicit respondents who 
             identified as teachers.', 
             format='latex', booktabs=T, escape=F, row.names = F,
             col.names=linebreak(c('', 'Out-of-School\nSuspension', 
                                   'Law Enf.\nReferral', 
                                   'In-School\nSuspension',
                                   'School-Related\nArrest', 
                                   'Expulsion w/o\n Ed. Services', 
                                   'Expulsion w/Ed.\nServices', 
                                   'Expulsion Under\nZero-Tolerance'))) %>%
  kableExtra::kable_styling(latex_options=c('scale_down')) %>%
  kableExtra::landscape()
```

## References

### 2018-05-18 written_docs/Riddle_Sinclair_PNAS/pnas_supporting.Rmd
PNAS submitted

---
title: "PNAS Supporting Information"
author: "Travis Riddle & Stacey Sinclair"
date: "5/11/2018"
output: 
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
bibliography: ref.bib
---

The OSF page for this project (https://osf.io/pu79a/) hosts many additional details. These include interactive maps for all disciplinary outcomes, information for a preregistered version of these analyses, explanations for departures from these plans, code, and aggregated data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(dplyr)
library(knitr)
library(kableExtra)
library(forcats)
library(tidyr)
library(stringr)
library(ggplot2)
```

```{r, reg-coefs, results='asis'}
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/full_models/'
plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA)

# weighted bias w/exclusions & explicit difference
metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'in_school_arrest', 'law_enforcement', 'oos_susp')#1245
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  for (k in files){
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }
  data.frame(coef=colnames(dat)[1:30],
             est=rowMeans(p_cons),
             lower=apply(p_cons,1,quantile, probs=.025),
             upper=apply(p_cons,1,quantile, probs=.975),
             p_neg = rowSums(p_cons<0)/4000,
             metric=rep(i, 30)) %>%
    rbind(plot.dat) %>%
    filter(!is.na(est)) -> plot.dat
}

plot.dat$metric <- factor(plot.dat$metric, c('oos_susp', 'inschool_susp',
                                             'law_enforcement', 
                                             'expulsion_combined',
                                             'in_school_arrest'))

plot.dat$metric <- fct_recode(plot.dat$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat$coef <- fct_recode(plot.dat$coef,
                            'Intercept' = '(Intercept)',
                            'total population' = 'total_pop',
                            'proportion black' = 'black_prop',
                            'proportion white' = 'white_prop',
                            'black-white ratio' = 'b.w.ratio',
                            'college grads' = 'col_grads',
                            'income' = 'med_income',
                            'poverty' = 'poverty_rate',
                            'unemployment' = 'unemp_rate',
                            'crime' = 'crime_rate',
                            'housing density' = 'density',
                            'mobility' = 'mobility',
                            'dissimilarity' = 'dissim',
                            'total population*race: white' = 'groupwhite:total_pop',
                            'proportion black*race: white' = 'groupwhite:black_prop',
                            'proportion white*race: white' = 'groupwhite:white_prop',
                            'black-white ratio*race: white' = 'groupwhite:b.w.ratio',
                            'college grads*race: white' = 'groupwhite:col_grads',
                            'income*race: white' = 'groupwhite:med_income',
                            'poverty*race: white' = 'groupwhite:poverty_rate',
                            'unemployment*race: white' = 'groupwhite:unemp_rate',
                            'crime*race: white' = 'groupwhite:crime_rate',
                            'housing density*race: white' = 'groupwhite:density',
                            'mobility*race: white' = 'groupwhite:mobility',
                            'dissimilarity*race: white' = 'groupwhite:dissim',
                            'race: white' = 'groupwhite',
                            'implicit bias' = 'weighted_bias',
                            'explicit bias' = 'weighted_explicit_diff',
                            'implicit bias*race: white' = 'groupwhite:weighted_bias',
                            'explicit bias*race: white' = 'groupwhite:weighted_explicit_diff')

plot.dat %>%
  mutate(est = round(est, 2)) %>%
  mutate(estimate=paste(est, ' [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
  mutate(estimate=kableExtra::cell_spec(estimate, 'latex', bold=ifelse(p_neg<.025|p_neg>.975, TRUE, FALSE))) %>%
  dplyr::select(metric, estimate, coef) %>%
  spread(metric, estimate) -> tabledata

tabledata <- tabledata[c(1:7, 23:27, 30, 8, 28, 29, 9:19, 22, 20, 21),]
names(tabledata)[1] <- ''

knitr::kable(tabledata, 
             caption = '\\label{tab:reg-coefs}Regression coefficient estimates for the population-level (i.e. fixed) effects, 
             along with 95\\% uncertainty intervals for each of the disciplinary metrics. Bias estimates are from race data.', 
           format = 'latex', booktabs=T, escape=F, row.names = F) %>%
  landscape() 
```

```{r disc-probs, fig.cap="\\label{fig:disc-probs}Association between bias and the probability of receiving discipline for black and white students. Lines represent the mean of the posterior. Bands indicate 95% uncertainty intervals. Printed p_{pos} represents the posterior probability that the association is positive."}
plot.dat <- data.frame(group=NA, weighted_bias=NA, est=NA, lower=NA, 
                       upper=NA, bias=NA, metric=NA, p_neg_black=NA, p_neg_white=NA)
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    df <- as.matrix(m)
    posterior_combo[,,j] <- t(df[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }

  newdat <- expand.grid(group = rownames(table(m$data$group)),
                        total_pop=0,
                        unemp_rate=0,
                        med_income=0,
                        poverty_rate=0,
                        col_grads=0,
                        white_prop=0,
                        black_prop=0,
                        b.w.ratio=0,
                        mobility=0,
                        crime_rate=0,
                        density=0,
                        dissim=0,
                        weighted_bias = seq(-3,3,.25),
                        weighted_explicit_diff = 0,
                        number=0,
                        total_number=1)
  y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

  cbind(newdat, t(y_hat)) %>% 
    mutate(index=row_number()) %>%
    gather(sample, value, `1`:`4000`) -> posterior_distribution

  p_neg_blackimp <- sum(p_cons[3,]>0)/4000
  p_neg_blackexp <- sum(p_cons[4,]>0)/4000
  p_neg_whiteimp <- sum((p_cons[3,]+p_cons[17,])>0)/4000
  p_neg_whiteexp <- sum((p_cons[4,]+p_cons[18,])>0)/4000

  posterior_distribution %>%
    select(group, weighted_bias, sample, value, index) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Implicit',
           metric = i,
           p_neg_black=p_neg_blackimp,
           p_neg_white=p_neg_whiteimp) %>%
    rbind(plot.dat) -> plot.dat

  newdat <- expand.grid(group = rownames(table(m$data$group)),
                        total_pop=0,
                        unemp_rate=0,
                        med_income=0,
                        poverty_rate=0,
                        col_grads=0,
                        white_prop=0,
                        black_prop=0,
                        b.w.ratio=0,
                        mobility=0,
                        crime_rate=0,
                        density=0,
                        dissim=0,
                        weighted_bias = 0,
                        weighted_explicit_diff = seq(-3, 3, .25),
                        number=0,
                        total_number=1)
  y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

  cbind(newdat, t(y_hat)) %>% 
    mutate(index=row_number()) %>%
    gather(sample, value, `1`:`4000`) -> posterior_distribution

  posterior_distribution %>%
    select(group, weighted_explicit_diff, sample, value, index) %>%
    mutate(weighted_bias = weighted_explicit_diff) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Explicit',
           metric=i,
           p_neg_black=p_neg_blackexp,
           p_neg_white=p_neg_whiteexp) %>%
    rbind(plot.dat) -> plot.dat

}

plot.dat$metric <- factor(plot.dat$metric, c('in_school_arrest', 
                                         'expulsion_combined', 
                                         'law_enforcement',
                                         'inschool_susp', 
                                         'oos_susp'))

plot.dat$metric <- fct_recode(plot.dat$metric, 
                               'Out-of\nSchool\nSusp' = 'oos_susp',
                               'In-School\nSusp' = 'inschool_susp',
                               'School\nArrest' = 'in_school_arrest',
                               'Law Enf.\nReferral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')

plot.dat <- plot.dat[!is.na(plot.dat$weighted_bias),]

plot.dat %>%
  filter(weighted_bias==2.5) %>%
  mutate(text_y = est,
         text_x = 2.75) %>%
  select(group, text_y, text_x, metric, p_neg_black, p_neg_white, bias) %>%
  distinct() %>%
  gather(group1, p_neg, p_neg_black, p_neg_white) %>%
  mutate(group1 = str_sub(group1, 7, -1)) %>%
  filter(group==group1) %>%
  select(-group1) -> text_data

text_data$text_y[8] <- text_data$text_y[8]+.00025

text_data$text_comp <- '=='
text_data$text_comp[text_data$p_neg>.99] <- '>'
text_data$text_comp[text_data$p_neg<.01] <- '<'
text_data$p_neg[text_data$p_neg>.99] <- .99
text_data$p_neg[text_data$p_neg<.01] <- .01

plot.dat %>%
  filter(!is.na(weighted_bias)) %>%
  ggplot(aes(x=weighted_bias, y=est, group=group, linetype=group)) +
    geom_line(aes(color=group)) +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.15) +
    theme_classic() +
    geom_text(data=text_data, parse=T, size=2.5, 
              aes(x=2.8, y=text_y, vjust=0,
                  label=paste0('p[pos]', text_comp, round(p_neg,2)))) +
    ylab('Probability of \n disciplinary action') +
    xlab('Standardized Bias') + 
    facet_grid(metric~bias, scales='free') +
    xlim(-2.75, 3.25) +
    theme(axis.title.y = element_text(size=10),
        legend.title = element_blank(),
        legend.position='top',
        axis.text = element_text(size=10),
        axis.title.x = element_text(size=12, face='bold'),
        legend.text = element_text(size=10),
        strip.text.y=element_text(size=7)) -> p1

p1
```

```{r, reg-coefs-sex, results='asis'}
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/sex_bias/'
plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA)

# weighted bias w/exclusions & explicit difference
metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'in_school_arrest', 'law_enforcement', 'oos_susp')#1245
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  for (k in files){
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }
  data.frame(coef=colnames(dat)[1:30],
             est=rowMeans(p_cons),
             lower=apply(p_cons,1,quantile, probs=.025),
             upper=apply(p_cons,1,quantile, probs=.975),
             p_neg = rowSums(p_cons<0)/4000,
             metric=rep(i, 30)) %>%
    rbind(plot.dat) %>%
    filter(!is.na(est)) -> plot.dat
}

plot.dat$metric <- factor(plot.dat$metric, c('oos_susp', 'inschool_susp',
                                             'law_enforcement', 
                                             'expulsion_combined',
                                             'in_school_arrest'))

plot.dat$metric <- fct_recode(plot.dat$metric,
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat$coef <- fct_recode(plot.dat$coef,
                            'Intercept' = '(Intercept)',
                            'total population' = 'total_pop',
                            'proportion black' = 'black_prop',
                            'proportion white' = 'white_prop',
                            'black-white ratio' = 'b.w.ratio',
                            'college grads' = 'col_grads',
                            'income' = 'med_income',
                            'poverty' = 'poverty_rate',
                            'unemployment' = 'unemp_rate',
                            'crime' = 'crime_rate',
                            'housing density' = 'density',
                            'mobility' = 'mobility',
                            'dissimilarity' = 'dissim',
                            'total population*race: white' = 'groupwhite:total_pop',
                            'proportion black*race: white' = 'groupwhite:black_prop',
                            'proportion white*race: white' = 'groupwhite:white_prop',
                            'black-white ratio*race: white' = 'groupwhite:b.w.ratio',
                            'college grads*race: white' = 'groupwhite:col_grads',
                            'income*race: white' = 'groupwhite:med_income',
                            'poverty*race: white' = 'groupwhite:poverty_rate',
                            'unemployment*race: white' = 'groupwhite:unemp_rate',
                            'crime*race: white' = 'groupwhite:crime_rate',
                            'housing density*race: white' = 'groupwhite:density',
                            'mobility*race: white' = 'groupwhite:mobility',
                            'dissimilarity*race: white' = 'groupwhite:dissim',
                            'race: white' = 'groupwhite',
                            'implicit bias' = 'weighted_bias_sex',
                            'explicit bias' = 'weighted_explicit_sex',
                            'implicit bias*race: white' = 'groupwhite:weighted_bias_sex',
                            'explicit bias*race: white' = 'groupwhite:weighted_explicit_sex')
plot.dat %>%
  mutate(est = round(est, 2)) %>%
  mutate(estimate=paste(est, ' [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
  mutate(estimate=kableExtra::cell_spec(estimate, 'latex', bold=ifelse(p_neg<.025|p_neg>.975, TRUE, FALSE))) %>%
  dplyr::select(metric, estimate, coef) %>%
  spread(metric, estimate) -> tabledata

tabledata <- tabledata[c(1:7, 23:27, 30, 8, 28, 29, 9:19, 22, 20, 21),]
names(tabledata)[1] <- ''

knitr::kable(tabledata, 
             caption = '\\label{tab:reg-coefs-sex}Regression coefficient estimates for the population-level (i.e. fixed) effects, 
             along with 95\\% uncertainty intervals for each of the disciplinary metrics. Bias estimates are from sexuality data.', 
           format = 'latex', booktabs=T, escape=F, row.names = F) %>%
  landscape()

```

```{r cor-tab}
county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv')
county_means_sex <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means_sexuality.csv')
covs <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/covariates.csv')
county_means %>%
  mutate(race_implicit = weighted_bias,
         race_explicit = weighted_explicit_diff) %>%
  select(county_id, race_implicit, race_explicit) %>%
  left_join(county_means_sex) %>%
  mutate(sex_implicit = weighted_bias,
         sex_explicit = weighted_explicit) %>%
  select(county_id, race_implicit, race_explicit, sex_implicit, sex_explicit) %>%
  full_join(covs) %>%
  filter(county_id!='NA-NA') -> preds

preds %>% 
  select(race_implicit:sex_explicit, total_pop:poverty_rate, crime_rate:dissim) %>%
  cor(., use='pairwise.complete.obs') -> cormat

preds %>% 
  mutate(crime_rate = crime_rate*1000, 
         total_pop = total_pop/1000,
         white_pop = white_pop/1000,
         black_pop = black_pop/1000,
         med_income = med_income/1000) %>%
  select(race_implicit:sex_explicit, total_pop:poverty_rate, crime_rate:dissim) %>%
  summarise_all(funs(mean(., na.rm=T))) -> means

preds %>% mutate(crime_rate = crime_rate*1000,
                 total_pop = total_pop/1000,
                 white_pop = white_pop/1000,
                 black_pop = black_pop/1000,
                 med_income = med_income/1000) %>%
  select(race_implicit:sex_explicit, total_pop:poverty_rate, crime_rate:dissim) %>%
  summarise_all(funs(sd(., na.rm=T))) -> sdevs

upper<-round(cormat, 2)
upper[upper.tri(cormat, diag = T)]<-""
upper<-as.data.frame(upper)
upper %>%
  mutate_all(funs(as.character)) %>%
  rbind(round(means, 2)) %>%
  rbind(round(sdevs, 2)) -> cortab

vars <- c("1) Implicit Race", "2) Explicit Race", "3) Implicit Sex", 
          "4) Explicit Sex", "5) Population Size", "6) White Population",
          "7) Black Population", "8) College Grads", "9) Unemployment", 
          "10) Income", "11) Poverty", "12) Crime", "13) Housing Density", 
          "14) Mobility", "15) Proportion White", "16) Proportion Black", 
          "17) Black-White Ratio", "18) Segregation", "Mean", "SD")

cortab <- cbind(vars, cortab)

kable(cortab, 
      caption = '\\label{tab:cor-tab}Correlation matrix for all county-level predictor variables.',
      format = 'latex', booktabs=T, escape=F, row.names = F,
             col.names=c('', seq(1:18))) %>%
  landscape() %>%
  kable_styling(latex_options=c('scale_down')) %>%
  footnote(number = c("Crime rate is scaled to rate per 1000 people",
                      "Total population, black population, and white population are in 1000's of people",
                      "Income is in 1000's of dollars")) %>%
  group_rows("", 19, 20, latex_gap_space = ".5em")
```

### 2018-05-18 aggregate_data.R

@@ -13,7 +13,8 @@ library(stringr)
 #read in data
 df_school <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')
 county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv')
-df_haley <- read.csv('/Users/travis/Documents/gits/Data/Haley_countylinks/_Master Spreadsheet.csv')
+df_haley <- read.csv('/Users/travis/Documents/gits/Data/linking_files/_Master Spreadsheet.csv')
+county_means_sex <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means_sexuality.csv')
 
 # Get enrollment figures by race for all schools
 df_school %>%
@@ -116,7 +117,7 @@ df_school %>%
 #multiple for all schools but these, we need to remove these schools
 oos_susp$number[which(oos_susp$number<0)] <- NA
 oos_susp$number_2[which(oos_susp$number_2<0)] <- NA
-oos_susp$number <- oos_susp$number+oos_susp_number_2
+oos_susp$number <- oos_susp$number+oos_susp$number_2
 #the 16 (along with the others who don't report any) now have NA
 
 #summarise across gender
@@ -339,6 +340,16 @@ county_means %>%
          weighted_explicit_diff, COMBOKEY, group, number, total_number, metric, 
          LEAID) -> full_data
 
+county_means_sex %>%
+  select(-X) %>%
+  mutate(sex_bias = bias,
+         sex_explicit = explicit,
+         weighted_bias_sex = weighted_bias,
+         weighted_explicit_sex = weighted_explicit) %>%
+  select(county_id, sex_bias, sex_explicit, 
+         weighted_bias_sex, weighted_explicit_sex) %>%
+  right_join(full_data) -> full_data
+
 ################
 # Exclude schools #
 ################
@@ -485,7 +496,13 @@ covs_pop %>%
   left_join(covs_mob) %>%
   mutate(white_prop = white_pop/total_pop,
          black_prop = black_pop/total_pop) %>%
-  mutate(b.w.ratio = black_prop/white_prop) -> covs
+  mutate(b.w.ratio = black_prop/white_prop) %>%
+  mutate(county_id = paste(state_code, 
+                           stringr::str_sub(county_fips, -3, -1), sep='-')) %>%
[…]

### 2018-05-18 consensusMCMC_model_prep.R

@@ -13,7 +13,7 @@ dft %>%
   distinct() -> teacher_counties
 
 counties %>% 
-  mutate(grouping=sample(1:15, n(), replace=T)) -> county_assignments
+  mutate(grouping=sample(1:20, n(), replace=T)) -> county_assignments
 
 teacher_counties %>%
   mutate(grouping=sample(1:12, n(), replace=T)) -> teacher_county_assignments
@@ -205,12 +205,16 @@ string2 <- "') %>% select(county_id, bias:weighted_explicit_diff) %>% distinct()
 df %>% filter(grouping=="
 
 string3 <- ") %>% filter(exclude==FALSE) %>% filter(metric=='"
-string4 <- "') %>% select(county_id, total_pop:poverty_rate, crime_rate:b.w.ratio, COMBOKEY:metric) %>% left_join(scaled_county_level) -> mod.dat
+string4 <- "') %>% select(county_id, total_pop:poverty_rate, crime_rate:b.w.ratio, dissim, COMBOKEY:metric) %>% left_join(scaled_county_level) -> mod.dat
 
 m <- stan_glmer(cbind(number, total_number-number) ~ group + weighted_bias + 
 weighted_explicit_diff + group:weighted_bias + group:weighted_explicit_diff + total_pop + 
 unemp_rate + med_income + poverty_rate + col_grads + white_prop + 
-black_prop + b.w.ratio + mobility + crime_rate + density + 
+black_prop + b.w.ratio + mobility + crime_rate + density + dissim +
+total_pop:group + unemp_rate:group + med_income:group + 
+poverty_rate:group + col_grads:group + white_prop:group + 
+black_prop:group + b.w.ratio:group + mobility:group + 
+crime_rate:group + density:group + dissim:group +
 (group|county_id), data=mod.dat, prior = normal(0,5), 
 prior_intercept = normal(0,5), family=binomial, adapt_delta=.99)
 
@@ -218,9 +222,54 @@ save(m, file='/tigress/triddle/educational_disparities/mw_2013uclaexcl/"
 
 
 for(i in rownames(table(dfs$metric))){
-  for (j in 1:15){
+  for (j in 1:20){
     fs <- paste(string1, i, string2, j, string3, i, string4, i, "/m", j, ".rdata')", sep='')
     fileConn <- file(paste("/Users/travis/Documents/gits/educational_disparities/cluster/model_scripts/mw_2013uclaexcl/",i,"/m", j, ".R", sep=''))
[…]

### 2018-05-18 iat_analysis.R

@@ -8,32 +8,26 @@ library(stringr)
 library(lme4)
 
 "%ni%" <- Negate("%in%")
-custom_scale <- function(x){
-  out <- x-mean(x, na.rm=T)/sd(x, na.rm=T)
-  return(out)
-}
 ######load datasets
 #IAT
-df <- rbind(read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2013.sav'),
-            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2012.sav'),
-            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2010.sav'))
+df <- rbind(read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2013.sav'),
+            read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2012.sav'),
+            read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2010.sav'))
 
-df2 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2011.sav')
-df3 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2009.sav')
-df4 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2008.sav')
-df5 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2007.sav')
-df6 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2006.sav')
-df7 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2005.sav')
-df8 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2004.sav')
-df9 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2002-2003.sav')
-df10 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2014.sav')
+df2 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2011.sav')
+df3 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2009.sav')
+df4 <- read_sav('/Users/travis/Documents/gits/Data/iat_race/Race IAT.public.2008.sav')
[…]

### 2018-05-18 iat_analysis_sexuality.R

@@ -1,0 +1,287 @@
+library(haven)
+library(dplyr)
+library(ggplot2)
+library(forcats)
+library(rstanarm)
+library(tidyr)
+library(stringr)
+library(lme4)
+
+"%ni%" <- Negate("%in%")
+
+######load datasets
+#IAT
+df <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2004.sav')
+df2 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2005.sav')
+df3 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2006.sav')
+df4 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2007.sav')
+df5 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2008.sav')
+df6 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2009.sav')
+df7 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2010.sav')
+df8 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2011.sav')
+df9 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2012.sav')
+df10 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2013.sav')
+df11 <- read_sav('/Users/travis/Documents/gits/Data/iat_sexuality/Sexuality IAT.public.2014.sav')
+
+#County linking information
+df_haley <- read.csv('/Users/travis/Documents/gits/Data/Haley_countylinks/_Master Spreadsheet.csv', stringsAsFactors = F)
+df_haley[1791,6] <- 'Doña Ana County, New Mexico' #because of the "ñ"
+
+df_county_linking_info <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_linking_table.csv', stringsAsFactors = F)
+df_county_linking_info$county_name[1904] <- 'Doña Ana'
+df_states <- data.frame(state=c(state.name, 'District of Columbia'),
+                        state_abb=c(state.abb, 'DC'))
+
+#Covariates
[…]

### 2018-05-18 make_biasmaps.py

@@ -1,0 +1,86 @@
+import bokeh
+from bokeh.plotting import figure, save, show
+from bokeh.models import (
+    ColumnDataSource,
+    HoverTool,
+    ColorMapper,
+    LinearColorMapper
+)
+from bokeh.palettes import Viridis10 as palette
+import numpy as np
+import pandas as pd
+import sys
+
+def filter_lower48(counties):
+	counties = {
+		code: county for code, county in counties.items() if 
+			county["state"] in ["ca", "az", "nv", "or", "wa", "nm", "tx", "id", 
+								"co", "ut", "mt", "wy", "nd", "sd", "ok", "ne", 
+								"ks", "mo", "al", "la", "wi", "ms", "ar", "wi", 
+								"il", "fl", "in", "ia", "mn", "ky", "oh", "tn", 
+								"ga", "sc", "nc", "de", "dc", "md", "pa", "va", 
+								"wv", "nj",	"ct", "me", "ma", "nh", "ny", "ri", 
+								"vt", "mi"]
+	}
+	return(counties)
+
+def make_mapdat(counties, plotting_dat, metric):
+	map_dat = pd.DataFrame({'county_number':[str(county_id[1]) for county_id in counties],
+                       'state_name':[county['state'] for county in counties.values()],
+                       'x':[county["lons"] for county in counties.values()],
+                       'y':[county["lats"] for county in counties.values()],
+                       'county_name':[county['detailed name'] for county in counties.values()]})
+
+	map_dat['county_number'] = map_dat['county_number'].apply(lambda x: x.zfill(3))
+	map_dat["county_id"] = map_dat["state_name"] + '-' + map_dat["county_number"]
+	map_dat = plotting_dat.merge(map_dat, on='county_id')
+	map_dat = map_dat[~map_dat[metric].isnull()]
+	return(map_dat)
+
+
+def main(metric):
+	from bokeh.sampledata.us_counties import data as counties
+	counties = filter_lower48(counties)
+	df_comeans = pd.read_csv('/Users/travis/Documents/gits/educational_disparities/output/map_data.csv')
+	df_comeans['county_id'] = df_comeans.county_id.str.lower()
+
+	map_dat = make_mapdat(counties, df_comeans, metric=metric)
+
+	color_mapper = LinearColorMapper(palette=palette)
+	source = ColumnDataSource(data=dict(
[…]

### 2018-05-18 make_discmaps.py

@@ -1,0 +1,101 @@
+import bokeh
+from bokeh.plotting import figure, save, show
+from bokeh.models import (
+    ColumnDataSource,
+    HoverTool,
+)
+from bokeh.palettes import Viridis10 as palette
+from bokeh.transform import factor_cmap
+import numpy as np
+import pandas as pd
+import sys
+
+def filter_lower48(counties):
+	counties = {
+		code: county for code, county in counties.items() if 
+			county["state"] in ["ca", "az", "nv", "or", "wa", "nm", "tx", "id", 
+								"co", "ut", "mt", "wy", "nd", "sd", "ok", "ne", 
+								"ks", "mo", "al", "la", "wi", "ms", "ar", "wi", 
+								"il", "fl", "in", "ia", "mn", "ky", "oh", "tn", 
+								"ga", "sc", "nc", "de", "dc", "md", "pa", "va", 
+								"wv", "nj",	"ct", "me", "ma", "nh", "ny", "ri", 
+								"vt", "mi"]
+	}
+	return(counties)
+
+def make_mapdat(counties, plotting_dat, metric):
+	map_dat = pd.DataFrame({'county_number':[str(county_id[1]) for county_id in counties],
+                       'state_name':[county['state'] for county in counties.values()],
+                       'x':[county["lons"] for county in counties.values()],
+                       'y':[county["lats"] for county in counties.values()],
+                       'county_name':[county['detailed name'] for county in counties.values()]})
+
+	map_dat['county_number'] = map_dat['county_number'].apply(lambda x: x.zfill(3))
+	map_dat["county_id"] = map_dat["state_name"] + '-' + map_dat["county_number"]
+	map_dat = plotting_dat.merge(map_dat, on='county_id')
+	map_dat = map_dat[~map_dat[metric].isnull()]
+	map_dat[metric+'_cat'] = pd.cut(map_dat[metric], bins=[0, .99, 1, 2, 3, 4, 5, 6, 7, 8, np.inf], include_lowest=True)
+	return(map_dat)
+
+
+def main(metric):
+	from bokeh.sampledata.us_counties import data as counties
+	counties = filter_lower48(counties)
+	df_comeans = pd.read_csv('/Users/travis/Documents/gits/educational_disparities/output/map_data.csv')
+	df_comeans['county_id'] = df_comeans.county_id.str.lower()
+
[…]

### 2018-05-18 make_map_data.R

@@ -1,0 +1,56 @@
+library(dplyr)
+library(tidyr)
+
+county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv')
+county_means_sex <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means_sexuality.csv')
+names(county_means_sex)[c(5:8)] <- c('n_bias_obs_sex', 'n_explicit_obs_sex', 'weighted_bias_sex', 'weighted_explicit_sex')
+model.data <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/full_model_data.csv')
+
+model.data %>%
+  filter(exclude==FALSE) %>%
+  group_by(county_id, metric, group) %>%
+  summarise(total = sum(number),
+            total_students = sum(total_number)) %>%
+  mutate(prop=total/total_students) %>%
+  select(county_id, metric, group, prop) %>%
+  spread(group, prop) %>%
+  ungroup() %>%
+  mutate(relative_risk = case_when(black == 0 & white == 0 ~ 1, 
+                              TRUE ~ black/white)) %>%
+  select(-black, -white) %>%
+  spread(metric, relative_risk) -> relative_risks
+
+model.data %>%
+  filter(exclude==FALSE) %>%
+  group_by(county_id, metric, group) %>%
+  summarise(total = sum(number)) %>%
+  ungroup() %>%
+  mutate(gmet=paste(metric, group, sep='_')) %>%
+  select(county_id, gmet, total) %>%
+  spread(gmet, total) -> number_incidents
+
+model.data %>%
+  filter(exclude==FALSE) %>%
+  group_by(county_id, group) %>%
+  summarise(total = sum(total_number)) %>%
+  ungroup() %>%
+  spread(group, total) -> number_students
+
+county_means %>%
+  select(county_id, n_bias_obs:weighted_bias, weighted_explicit_diff) %>%
+  left_join(county_means_sex) %>%
+  select(-X, -bias, -explicit) %>%
+  left_join(relative_risks) %>%
+  left_join(number_incidents) %>%
+  left_join(number_students) %>%
+  mutate(implicit_obs = n_bias_obs,
+         explicit_obs = n_explicit_obs,
+         implicit = weighted_bias,
+         explicit = weighted_explicit_diff,
+         implicit_sex_obs = n_bias_obs_sex,
+         explicit_sex_obs = n_explicit_obs_sex,
[…]

### 2018-05-18 written_docs/Riddle_Sinclair_PNAS/Riddle_Sinclair_PNAS.tex
.gitignore is now working

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\templatetype{pnasresearcharticle}  % Choose template

\title{Race-based disparities in academic disciplinary actions are associated
with county-level rates of racial bias}

\author[a]{Travis Riddle}
\author[a,b]{Stacey Sinclair}

  \affil[a]{Princeton University, Department of Psychology, Princeton, NJ, 08544}
  \affil[b]{African American Studies}

% Please give the surname of the lead author for the running footer
\leadauthor{Riddle}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Black students in the United States are subject to disciplinary action
at rates much higher than their white counterparts. These disciplinary
actions put students at higher risk for negative life outcomes,
including involvement in the criminal justice system. Using government
data covering over 32 million students at nearly 93k schools, our
research demonstrates that the disciplinary gap between black and white
students across 5 different types of disciplinary actions is associated
with county-level rates of explicit racial bias, and suggests solving
racial disparities in education must involve combating explicit racial
bias.}

\authorcontributions{TR and SS designed the research and wrote the paper. TR analyzed the
data.}

\authordeclaration{Authors have no known conflicts of interest}

\correspondingauthor{\textsuperscript{} }

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  one |  two |  optional |  optional |  optional   }

\begin{abstract}
There are substantial gaps in academic achievement between black and
white students in the United States. Recently, increased attention has
focused on differences in the rates at which black and white students
are disciplined, finding that black students are more likely to be seen
as problematic and more likely to be punished than white students are
for the same offense. Although these disparities suggest that racial
biases are a contributor, no previous research has shown associations
with psychological measurements of bias and disciplinary outcomes. We
show that county-level estimates of explicit racial bias, as measured
using data from approximately 1.3 million visitors to the Project
Implicit website, are associated with racial disciplinary disparities
across approximately 93k schools in the United States, covering around
32 million white and black students. These associations do not extend to
explicit or implicit sexuality biases, showing the specificity of the
effect. These findings suggest that reducing racial disparities in
education may require efforts to reduce explicit racial bias.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{Please include your acknowledgments here, set in a single paragraph.
Please do not include any acknowledgments in the Supporting Information,
or anywhere else in the manuscript.}

In comparison to White Americans, Black Americans exhibit poorer
educational outcomes across a range of metrics. One outcome of
particular concern is the gap in disciplinary actions (1, 2). Research
using administrative datasets and longitudinal samples clearly show that
Black American students are far more likely to be suspended or expelled
(3, 4), and conditional on an office referral, are more likely to
receive stiffer punishments (5, 6). These disparities are particularly
concerning as they are associated with long-term life outcomes,
including prospects for employment (7) and involvement in the criminal
justice system (8).

As complex social phenomena, these racial differences in disciplinary
outcomes are multiply determined (2). There are clear structural and
socieconomic contributors such as the practice of segregating students
into achievement-based ``tracks'' of low and high performers (9) or
racial differences in socioeconomic status (10). However, racial bias is
also thought to contribute to the problem. For instance, in a controlled
study, (11) found that in comparison to white students, teachers were
more likely to view the same behavior from black students as being
indicative of a long term problem and deserving of suspension.
Similarly, using discipline data from an urban high school, (12) showed
that black students were especially likely to be referred to the office
for disciplinary action on the basis of defiant behavior - a relatively
subjective category of misbehavior in comparison to others they
examined, including truancy or fighting. Overall, there has been
consistent evidence that black students' behaviors are both perceived as
more problematic and are punished more harshly as compared to white
students'. However, to our knowledge, there has been no work assessing
whether racial bias is directly associated with disciplinary
disparities. Additionally, there has been no work assessing how racial
bias at the community level is associated with educational disparities.

Psychological measurements of racial bias typically occur through one of
two ways. Either individuals are asked to self-report their relative
attitudes toward different racial groups, or via methods designed to
assess automatic associations with people of different races. The latter
is thought to reflect the associations learned by early and frequent
exposure to environmental stimuli, whereas the former reflects top-down
modulation of those associations (13). Recently, researchers have begun
aggregating these measures up to geographical regions such as counties
or states, finding that regional-level measures of implicit and explicit
racial bias are associated with racial disparities in key social
outcomes, although the relative contributions are not consistent across
studies (14--16). For example, (14) found that Black Americans had
reduced access to health care and increased rates of death due to
circulatory disease in comparison to whites in counties with higher
levels of explicit racial bias against blacks. They found no such
associations for implicit racial bias. In contrast, (16), found that the
disproportionate use of lethal force by police on Black Americans was
associated with regional implicit biases, but were not associated with
explicit biases. As such, it is important to assess both types of bias
when seeking to understand the relationship between regional-level bias
and behavioral outcomes.

Regional levels of bias could be associated with the size of racial
student disciplinary disparities for a number of reasons. First, being
in an area characterized by higher racial bias likely means encountering
individuals who have negative feelings and beliefs about one's group.
The actions of such individuals within and/or outside of an educational
setting could contribute to disciplinary disparities. For example, if
teachers and administrators are biased, then they may be more likely to
make decisions that are unfavorable to black students, such as deciding
that a given misbehavior is worthy of official disciplinary action.
Biased administrators or local voters might also support school or
district policies thought to disproportionally punish students of color,
such as zero-tolerance policies or implementation of random drug sweeps
(17). Second, Black students who live in high bias areas may be more apt
to engage in behavior that warrants disciplinary action due to behavior
confirmation of the negative expectancies held by area peers and adults
(18--20). Third, the norms and structural factors (e.g., laws, policies)
that characterize regions higher in bias may constrain even those
individuals who are not biased themselves into engaging in or suborning
actions that negatively impact students of color (21, 22). Fourth,
biases assessed at the regional level might reflect affordances of the
local environment (e.g., confederate statues, biased media) that
undergird these biases and prime behaviors that contribute to
disciplinary disparities (23). Overall, these reasons, and the likely
possibility that they work in concert to inform behavior (24),
substantiate the possibility that there will be a relationship between
regional bias and disciplinary outcomes.

Most previous research has focused on out-of-school suspensions - likely
because they are the most frequently used and are regularly found to be
associated with negative outcomes (25, 26). However, other disciplinary
outcomes, though used less often, may be at least as damaging to
students as out-of-school suspensions (27). For instance school arrests
have been associated with increased likelihood of engaging in
anti-social behavior (28), and with increased risk of dropping out (8).
In addition, though alternative forms of discipline (e.g., in-school
suspension) are intended to insulate students from the negative
consequences of exclusionary discipline, the criteria by which students
are assigned the former kind of discipline often remain vulnerable to
bias (29). As such, examining the presence and basis of disparities in
the application of a wide range of disciplinary actions is warranted.

The present analyses combine regionally-coded implicit and explicit
racial bias measures from approximately 1.3 million respondents who
visited Project Implicit (30) with the most recent available data from
the Civil Rights Data Collection (CRDC) conducted by the US Department
of Education, a mandated census of disciplinary action in all US public
k-12 schools. The CRDC allowed well-powered examinations of five
different disciplinary metrics: in-school suspensions, out-of-school
suspensions, law enforcement referrals, school-related arrests, and
expulsions. We designed the analyses to determine whether, and if so the
extent to which regional estimates of pro-White/anti-Black implicit and
explicit bias are associated with black-white outcomes in disciplinary
gaps. We also use regional estimates of sexuality bias from Project
Implicit to determine whether demonstrated patterns are distinct to
racial bias, or an epiphenomenal associate of bias measures in general.

\subsection{Results}\label{results}

\subsubsection{Project Implicit
estimates}\label{project-implicit-estimates}

We first report the results of estimating the implicit and explicit
biases from Project Implicit data. When examining the unstandardized
county-level estimates adjusted with poststratification, we see that
there is a pro-white bias in both implicit (\(mean\) = 0.4, \(sd\) =
0.02) and explicit measures (\(mean\) = 0.83, \(sd\) = 0.17), where on
both scales 0 = no bias, and positive numbers indicate a pro-white bias.

\subsection{Disciplinary action
frequency}\label{disciplinary-action-frequency}

\begin{table}[!htbp] \centering 
  \caption{Percentage of students of each race receiving each type of disciplinary action} 
  \label{tab:disc-count} 
\begin{tabular}{@{\extracolsep{5pt}} ccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
metric & black & white \\ 
\hline \\[-1.8ex] 
school arrests & 0.29\% & 0.09\% \\ 
expulsions & 0.48\% & 0.22\% \\ 
law enforcement referral & 0.75\% & 0.34\% \\ 
in-school suspension & 11.11\% & 4.23\% \\ 
out-of-school suspension & 13.87\% & 3.63\% \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

Table \ref{tab:disc-count} shows the percentage of students of each race
who were reported having received each of the actions under
consideration. Additionally, figure \ref{fig:suspension-disp-map} shows
the relative risk ratio for out-of-school suspensions across counties.
Our statistical models (described below) show that the racial
differences seen here are extremely unlikely to be due to chance.

\begin{figure}
\centering
\includegraphics{Riddle_Sinclair_PNAS_files/figure-latex/suspension-disp-map-1.pdf}
\caption{Relative Risk Ratio for out-of-school suspensions for each
county in the continental United States. Relative risk ratio is computed
as the percentage of black students suspended to the percentage of white
students suspended, as reported in the CRDC. Higher values indicate more
black students suspended, relative to white students. Counties with the
value NA either have no schools, have zero black students enrolled, or
have zero white students enrolled. Interactive maps for all disciplinary
outcomes and bias estimates can be found at \url{https://osf.io/pu79a/}}
\end{figure}

\subsection{Associations across
counties}\label{associations-across-counties}

Figure \[fig:overall-associations\] shows the estimate of primary
interest for each of the models. The estimates displayed are the
coefficients for the interaction between race and each of the two bias
measurements. Given that Black Americans are the baseline group,
negative values for this coefficient indicate that as one moves into
counties with higher levels of bias, the gap between the probability of
a black student being disciplined and the probability of a white student
being disciplined grows. Table \[tab:reg-coefs\] displays the estimated
coefficients and 95\% uncertainty intervals for each of the fixed
effects for each model.

\begin{figure}
\centering
\includegraphics{Riddle_Sinclair_PNAS_files/figure-latex/overall-associations-1.pdf}
\caption{Association between each metric and county-level estimates of
explicit and implicit bias. Negative values indicate that the rate of
increase (or decrease) for blacks is faster (or slower) than for whites.
Point is the mean of the posterior and error bars represent 95\%
bayesian uncertainty intervals}
\end{figure}

Several patterns are apparent from this figure. First, county-level
explicit biases are reliably associated with disciplinary disparities,
such that counties with more bias are expected to have a larger racial
gap. The effect with the largest magnitude of these is school-related
arrests (\(M_{slope}\) = -0.16, {[}-0.28, -0.03{]}), followed by
expulsions (\(M_{slope}\) = -0.12, {[}-0.21, -0.02{]}), out-of-school
suspensions (\(M_{slope}\) = -0.11, {[}-0.14, -0.08{]}), in-school
suspensions (\(M_{slope}\) = -0.1, {[}-0.13, -0.06{]}), and law
enforcement referrals (\(M_{slope}\) = -0.09, {[}-0.17, -0.02{]}).

In contrast, the estimated effects for implicit bias are more ambiguous.
Expulsions (\(M_{slope}\) = 0.06, {[}-0.05, 0.17{]}), school-related
arrests (\(M_{slope}\) = 0.04, {[}-0.11, 0.17{]}), and law enforcement
referrals (\(M_{slope}\) = 0.01, {[}-0.07, 0.09{]}) are all estimated to
be close to zero, or with substantial uncertainty with respect to the
direction of the effect. Surprisingly, there is a reliable association
for in-school suspensions, with schools higher in implicit bias having
smaller in-school suspension gap (\(M_{slope}\) = 0.05, {[}0.01,
0.08{]}). Though not fully conclusive, the the association between
implicit biases and out-of-school suspensions tends to be in a similar,
counter-intuitive direction (\(M_{slope}\) = 0.03, {[}0, 0.07{]}).
However, it is important to point out that at no point do these
associations ``cross over'' such that white students are actually
suspended more frequently than black students. Even at a county with
implicit bias 3 standard deviations above the mean, for every white
student who receives an out-of-school suspension, about 2.43 {[}2.2,
2.66{]} black students are expected to be suspended.

\begin{table}[!htbp] \centering 
  \caption{Regression coefficient estimates for the population-level (i.e. fixed) effects, along with 95\% uncertainty intervals for each of the disciplinary metrics.} 
  \label{tab:reg-coefs} 
\begin{tabular}{@{\extracolsep{5pt}} cccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Out-of-School Suspension & In-School Suspension & Law Enf. Referral & Expulsion & School-Related Arrest \\ 
\hline \\[-1.8ex] 
Intercept & \textbackslash bf\{-2.31\}, [-2.34,-2.28] & \textbackslash bf\{-2.23\}, [-2.27,-2.2] & \textbackslash bf\{-5.77\}, [-5.87,-5.67] & \textbackslash bf\{-6.54\}, [-6.66,-6.42] & \textbackslash bf\{-7.97\}, [-8.18,-7.77] \\ 
black-white ratio & \textbackslash bf\{-0.15\}, [-0.21,-0.1] & \textbackslash bf\{-0.32\}, [-0.4,-0.24] & 0.13, [-0.04,0.29] & -0.09, [-0.29,0.12] & 0.16, [-0.18,0.48] \\ 
proportion black & \textbackslash bf\{0.22\}, [0.13,0.31] & \textbackslash bf\{0.16\}, [0.05,0.28] & \textbackslash bf\{-0.36\}, [-0.6,-0.12] & -0.23, [-0.52,0.07] & -0.37, [-0.79,0.07] \\ 
college grads & \textbackslash bf\{0.08\}, [0.04,0.12] & \textbackslash bf\{-0.06\}, [-0.1,-0.01] & \textbackslash bf\{0.15\}, [0.05,0.26] & -0.03, [-0.16,0.1] & \textbackslash bf\{0.22\}, [0.04,0.41] \\ 
crime & \textbackslash bf\{0.04\}, [0.01,0.07] & \textbackslash bf\{0.05\}, [0.01,0.09] & 0.04, [-0.05,0.12] & \textbackslash bf\{0.12\}, [0.03,0.22] & 0.02, [-0.13,0.16] \\ 
housing density & \textbackslash bf\{-0.05\}, [-0.07,-0.02] & -0.03, [-0.07,0] & 0.01, [-0.06,0.08] & -0.05, [-0.13,0.03] & \textbackslash bf\{-0.13\}, [-0.23,-0.03] \\ 
dissimilarity & \textbackslash bf\{0.2\}, [0.16,0.23] & \textbackslash bf\{0.09\}, [0.05,0.13] & \textbackslash bf\{0.19\}, [0.1,0.29] & \textbackslash bf\{0.3\}, [0.18,0.41] & \textbackslash bf\{0.48\}, [0.3,0.65] \\ 
income & \textbackslash bf\{-0.08\}, [-0.13,-0.03] & -0.07, [-0.13,0] & -0.05, [-0.2,0.09] & \textbackslash bf\{-0.29\}, [-0.46,-0.11] & 0.03, [-0.19,0.27] \\ 
mobility & -0.03, [-0.07,0.01] & 0, [-0.05,0.04] & 0.09, [-0.02,0.21] & 0.01, [-0.14,0.15] & 0.01, [-0.2,0.21] \\ 
poverty & \textbackslash bf\{-0.09\}, [-0.17,-0.01] & \textbackslash bf\{0.11\}, [0.01,0.21] & \textbackslash bf\{-0.58\}, [-0.81,-0.35] & \textbackslash bf\{-0.42\}, [-0.69,-0.14] & -0.35, [-0.73,0.04] \\ 
total population & 0, [-0.03,0.02] & \textbackslash bf\{-0.04\}, [-0.08,0] & -0.01, [-0.08,0.07] & 0, [-0.09,0.08] & \textbackslash bf\{0.12\}, [0.01,0.23] \\ 
unemployment & \textbackslash bf\{0.22\}, [0.18,0.26] & 0.02, [-0.03,0.07] & \textbackslash bf\{0.16\}, [0.03,0.28] & \textbackslash bf\{0.22\}, [0.08,0.35] & \textbackslash bf\{0.25\}, [0.03,0.46] \\ 
proportion white & -0.03, [-0.11,0.04] & -0.02, [-0.11,0.07] & \textbackslash bf\{-0.25\}, [-0.46,-0.05] & \textbackslash bf\{-0.27\}, [-0.51,-0.02] & -0.15, [-0.47,0.19] \\ 
race: white & \textbackslash bf\{-1.06\}, [-1.08,-1.03] & \textbackslash bf\{-0.91\}, [-0.93,-0.88] & \textbackslash bf\{-0.84\}, [-0.92,-0.76] & \textbackslash bf\{-0.89\}, [-0.99,-0.78] & \textbackslash bf\{-0.89\}, [-1.06,-0.7] \\ 
implicit bias & \textbackslash bf\{0.12\}, [0.07,0.17] & \textbackslash bf\{0.17\}, [0.1,0.23] & -0.05, [-0.18,0.08] & 0.06, [-0.1,0.22] & -0.17, [-0.4,0.04] \\ 
explicit bias & \textbackslash bf\{-0.05\}, [-0.1,-0.01] & \textbackslash bf\{0.12\}, [0.06,0.17] & -0.04, [-0.16,0.08] & 0.04, [-0.1,0.19] & \textbackslash bf\{0.36\}, [0.16,0.56] \\ 
black-white ratio\textasteriskcentered race: white & \textbackslash bf\{0.1\}, [0.05,0.15] & 0.05, [-0.01,0.11] & \textbackslash bf\{-0.26\}, [-0.47,-0.04] & 0.13, [-0.1,0.34] & -0.29, [-0.7,0.1] \\ 
proportion black\textasteriskcentered race: white & -0.02, [-0.09,0.05] & -0.01, [-0.08,0.07] & 0.2, [-0.01,0.39] & 0.04, [-0.2,0.28] & 0.27, [-0.1,0.62] \\ 
college grads\textasteriskcentered race: white & \textbackslash bf\{-0.12\}, [-0.15,-0.09] & \textbackslash bf\{-0.06\}, [-0.09,-0.02] & \textbackslash bf\{-0.15\}, [-0.22,-0.08] & \textbackslash bf\{-0.14\}, [-0.24,-0.04] & -0.08, [-0.21,0.04] \\ 
crime\textasteriskcentered race: white & -0.02, [-0.04,0] & 0, [-0.02,0.03] & 0.03, [-0.02,0.08] & -0.05, [-0.12,0.01] & 0.05, [-0.04,0.14] \\ 
housing density\textasteriskcentered race: white & \textbackslash bf\{-0.03\}, [-0.04,-0.01] & -0.01, [-0.03,0] & -0.02, [-0.05,0.02] & -0.03, [-0.08,0.02] & -0.06, [-0.21,0.08] \\ 
dissimilarity\textasteriskcentered race: white & \textbackslash bf\{-0.11\}, [-0.14,-0.08] & \textbackslash bf\{-0.05\}, [-0.07,-0.02] & \textbackslash bf\{-0.09\}, [-0.16,-0.02] & -0.04, [-0.13,0.05] & \textbackslash bf\{-0.22\}, [-0.34,-0.1] \\ 
income\textasteriskcentered race: white & \textbackslash bf\{-0.05\}, [-0.09,-0.01] & \textbackslash bf\{-0.09\}, [-0.12,-0.05] & -0.04, [-0.13,0.04] & 0.03, [-0.1,0.16] & -0.09, [-0.24,0.06] \\ 
mobility\textasteriskcentered race: white & 0.03, [-0.01,0.06] & 0, [-0.03,0.03] & -0.04, [-0.12,0.05] & -0.01, [-0.14,0.11] & 0.04, [-0.13,0.21] \\ 
poverty\textasteriskcentered race: white & 0.05, [-0.01,0.12] & 0, [-0.06,0.07] & -0.04, [-0.2,0.12] & 0.01, [-0.2,0.22] & 0, [-0.29,0.28] \\ 
total population\textasteriskcentered race: white & \textbackslash bf\{-0.03\}, [-0.05,-0.01] & -0.01, [-0.03,0.01] & 0, [-0.04,0.03] & -0.02, [-0.06,0.03] & -0.01, [-0.05,0.04] \\ 
unemployment\textasteriskcentered race: white & -0.02, [-0.05,0.01] & 0, [-0.04,0.03] & -0.02, [-0.11,0.06] & 0.03, [-0.07,0.14] & -0.03, [-0.18,0.11] \\ 
proportion white\textasteriskcentered race: white & 0.03, [-0.03,0.09] & 0.02, [-0.03,0.08] & -0.07, [-0.2,0.06] & 0.09, [-0.09,0.27] & -0.03, [-0.22,0.17] \\ 
implicit bias\textasteriskcentered race: white & 0.03, [0,0.07] & \textbackslash bf\{0.05\}, [0.01,0.08] & 0.01, [-0.08,0.09] & 0.06, [-0.05,0.18] & 0.04, [-0.1,0.18] \\ 
explicit bias\textasteriskcentered race: white & \textbackslash bf\{-0.11\}, [-0.14,-0.08] & \textbackslash bf\{-0.1\}, [-0.13,-0.06] & \textbackslash bf\{-0.09\}, [-0.17,-0.02] & \textbackslash bf\{-0.12\}, [-0.22,-0.02] & \textbackslash bf\{-0.16\}, [-0.28,-0.04] \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

\begin{figure}
\centering
\includegraphics{Riddle_Sinclair_PNAS_files/figure-latex/detail-risks-1.pdf}
\caption{Association between bias and relative risk ratio for black
students to white students. Line is the mean of the posterior. Bands
indicate 95\% uncertainty intervals. Points represent counties, whose
sizes are scaled to the number of schools in that county. Printed
p\_\{pos\} represents the posterior probability that the association is
positive. Note that the y axis is cut at 8 for legibility despite data
extending beyond. Additionally, a number of schools cannot be
represented because the ratio of discipline proportions (black/white) is
undefined, either because there are zero students of a given race
enrolled (e.g.~NA/white), or the proportion of white students
disciplined is zero (e.g.~black/0).}
\end{figure}

To better illustrate the nature of these relationships, figure
@ref(fig:detail-risks) shows the estimated relative risk ratios for each
type of bias measurement and each disciplinary metric. This quantity is
the ratio of the probability that a black student will receive a given
disciplinary action to the probability that a white students will
receive a given disciplinary action. Higher values indicate that black
students have a higher probability of being punished than white
students. Examining the risk ratio for out-of-school suspensions, in a
hypothetical county with 0 standardized bias on both implicit and
explicit, the model predicts that for every white student suspended, we
should expect 2.71 {[}2.65, 2.78{]} black students to be suspended. If
we move to a county one standard deviation above the mean of explicit
bias, the ratio of black to white students suspended increases to 3.02
{[}2.9, 3.14{]}, while the same movement for implicit bias decreases the
ratio to 2.61 {[}2.52, 2.71{]}.

\subsection{Additional analyses: Sexuality bias as
predictor}\label{additional-analyses-sexuality-bias-as-predictor}

We sought to test whether the relationships observed above were specific
to estimates of racial bias. Accordingly, we ran the same set of
analyses with sexuality bias replacing racial bias.

\subsubsection{Project Implicit
Estimates}\label{project-implicit-estimates-1}

The unstandardized county-level estimates of bias adjusted with
poststratification, evidence a pro-straight bias in both implicit
(\(mean\) = 0.38, \(sd\) = 0.05) and explicit measures (\(mean\) = 1.65,
\(sd\) = 0.49), where on both scales 0 = no bias, and positive numbers
indicate a pro-straight bias.

\subsubsection{Associations across
counties}\label{associations-across-counties-1}

Figure @ref(fig:overall-associations-sex) shows the estimate of primary
interest for each of the models. The estimates displayed are the
coefficients for the interaction between race and each of the two bias
measurements. Given that Black Americans are the baseline group,
negative values for this coefficient indicate that as one moves into
counties with higher levels of bias, the gap between the probability of
a black student being disciplined and the probability of a white student
being disciplined grows. Table @ref(tab:reg-coefs-sex) displays the
estimated coefficients and 95\% uncertainty intervals for each of the
fixed effects for each model.

\begin{figure}
\centering
\includegraphics{Riddle_Sinclair_PNAS_files/figure-latex/overall-associations-sex-1.pdf}
\caption{Association between each metric and county-level estimates of
explicit and implicit bias. Negative values indicate that the rate of
increase (or decrease) for blacks is faster (or slower) than for whites.
Point is the mean of the posterior and error bars represent 95\%
bayesian uncertainty intervals.}
\end{figure}

\begin{table}[!htbp] \centering 
  \caption{Regression coefficient estimates for the population-level (i.e. fixed) effects, along with 95\% uncertainty intervals for each of the disciplinary metrics.} 
  \label{tab:reg-coefs-sex} 
\begin{tabular}{@{\extracolsep{5pt}} cccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Out-of-School Suspension & In-School Suspension & Law Enf. Referral & Expulsion & School-Related Arrest \\ 
\hline \\[-1.8ex] 
Intercept & \textbackslash bf\{-2.29\}, [-2.32,-2.26] & \textbackslash bf\{-2.23\}, [-2.27,-2.19] & \textbackslash bf\{-5.72\}, [-5.83,-5.62] & \textbackslash bf\{-6.54\}, [-6.67,-6.41] & \textbackslash bf\{-7.98\}, [-8.2,-7.77] \\ 
black-white ratio & \textbackslash bf\{-0.19\}, [-0.25,-0.14] & \textbackslash bf\{-0.43\}, [-0.51,-0.36] & \textbackslash bf\{0.18\}, [0.03,0.35] & -0.11, [-0.31,0.08] & 0.11, [-0.21,0.44] \\ 
proportion black & \textbackslash bf\{0.33\}, [0.25,0.41] & \textbackslash bf\{0.44\}, [0.34,0.54] & \textbackslash bf\{-0.42\}, [-0.64,-0.2] & -0.16, [-0.42,0.1] & -0.29, [-0.68,0.1] \\ 
college grads & \textbackslash bf\{0.05\}, [0.01,0.09] & \textbackslash bf\{-0.05\}, [-0.1,0] & \textbackslash bf\{0.14\}, [0.02,0.25] & -0.01, [-0.15,0.12] & \textbackslash bf\{0.23\}, [0.04,0.42] \\ 
crime & 0.03, [0,0.06] & \textbackslash bf\{0.06\}, [0.02,0.1] & 0.03, [-0.05,0.12] & \textbackslash bf\{0.14\}, [0.04,0.24] & 0.03, [-0.11,0.18] \\ 
housing density & \textbackslash bf\{-0.04\}, [-0.07,-0.02] & -0.03, [-0.06,0.01] & 0, [-0.07,0.06] & -0.05, [-0.13,0.03] & \textbackslash bf\{-0.14\}, [-0.24,-0.04] \\ 
dissimilarity & \textbackslash bf\{0.17\}, [0.14,0.21] & \textbackslash bf\{0.09\}, [0.05,0.13] & \textbackslash bf\{0.17\}, [0.07,0.27] & \textbackslash bf\{0.29\}, [0.17,0.41] & \textbackslash bf\{0.48\}, [0.31,0.66] \\ 
income & \textbackslash bf\{-0.08\}, [-0.13,-0.03] & -0.03, [-0.09,0.04] & -0.07, [-0.2,0.07] & \textbackslash bf\{-0.27\}, [-0.43,-0.1] & 0.01, [-0.21,0.24] \\ 
mobility & \textbackslash bf\{-0.05\}, [-0.09,-0.02] & -0.03, [-0.07,0.02] & 0.09, [-0.02,0.2] & -0.01, [-0.16,0.14] & 0, [-0.21,0.2] \\ 
poverty & -0.02, [-0.1,0.06] & \textbackslash bf\{0.15\}, [0.05,0.25] & \textbackslash bf\{-0.46\}, [-0.69,-0.23] & \textbackslash bf\{-0.38\}, [-0.66,-0.09] & -0.3, [-0.68,0.08] \\ 
total population & -0.01, [-0.04,0.01] & -0.02, [-0.06,0.01] & -0.03, [-0.11,0.04] & 0, [-0.08,0.08] & \textbackslash bf\{0.11\}, [0.01,0.22] \\ 
unemployment & \textbackslash bf\{0.2\}, [0.16,0.24] & 0.04, [-0.01,0.09] & 0.09, [-0.03,0.2] & \textbackslash bf\{0.21\}, [0.06,0.35] & 0.15, [-0.06,0.36] \\ 
proportion white & 0.03, [-0.04,0.1] & \textbackslash bf\{0.09\}, [0,0.18] & \textbackslash bf\{-0.23\}, [-0.43,-0.04] & -0.23, [-0.47,0.02] & -0.15, [-0.49,0.18] \\ 
race: white & \textbackslash bf\{-1.06\}, [-1.09,-1.03] & \textbackslash bf\{-0.91\}, [-0.93,-0.88] & \textbackslash bf\{-0.86\}, [-0.94,-0.77] & \textbackslash bf\{-0.88\}, [-0.99,-0.76] & \textbackslash bf\{-0.89\}, [-1.08,-0.7] \\ 
implicit bias & \textbackslash bf\{0.05\}, [0,0.09] & \textbackslash bf\{0.13\}, [0.07,0.19] & 0.02, [-0.1,0.14] & \textbackslash bf\{0.2\}, [0.05,0.35] & \textbackslash bf\{0.4\}, [0.18,0.61] \\ 
explicit bias & \textbackslash bf\{-0.14\}, [-0.19,-0.1] & -0.03, [-0.08,0.03] & \textbackslash bf\{-0.26\}, [-0.38,-0.15] & -0.15, [-0.29,0] & \textbackslash bf\{-0.36\}, [-0.57,-0.14] \\ 
black-white ratio\textasteriskcentered race: white & \textbackslash bf\{0.13\}, [0.08,0.18] & 0.06, [0,0.12] & \textbackslash bf\{-0.22\}, [-0.43,-0.02] & 0.16, [-0.05,0.36] & -0.2, [-0.6,0.17] \\ 
proportion black\textasteriskcentered race: white & \textbackslash bf\{-0.09\}, [-0.16,-0.03] & -0.05, [-0.12,0.01] & 0.13, [-0.05,0.3] & -0.06, [-0.28,0.16] & 0.12, [-0.21,0.45] \\ 
college grads\textasteriskcentered race: white & \textbackslash bf\{-0.12\}, [-0.15,-0.09] & \textbackslash bf\{-0.06\}, [-0.09,-0.02] & \textbackslash bf\{-0.18\}, [-0.25,-0.1] & \textbackslash bf\{-0.13\}, [-0.24,-0.02] & -0.09, [-0.23,0.06] \\ 
crime\textasteriskcentered race: white & -0.02, [-0.04,0] & 0.01, [-0.02,0.03] & 0.02, [-0.03,0.07] & -0.04, [-0.1,0.03] & 0.05, [-0.04,0.14] \\ 
housing density\textasteriskcentered race: white & \textbackslash bf\{-0.02\}, [-0.04,0] & -0.01, [-0.03,0.01] & -0.01, [-0.04,0.02] & -0.03, [-0.08,0.02] & -0.06, [-0.22,0.09] \\ 
dissimilarity\textasteriskcentered race: white & \textbackslash bf\{-0.11\}, [-0.14,-0.08] & \textbackslash bf\{-0.05\}, [-0.08,-0.02] & \textbackslash bf\{-0.09\}, [-0.16,-0.02] & -0.04, [-0.13,0.05] & \textbackslash bf\{-0.2\}, [-0.33,-0.08] \\ 
income\textasteriskcentered race: white & \textbackslash bf\{-0.06\}, [-0.09,-0.02] & \textbackslash bf\{-0.09\}, [-0.13,-0.05] & -0.06, [-0.15,0.02] & 0.03, [-0.1,0.17] & -0.1, [-0.25,0.05] \\ 
mobility\textasteriskcentered race: white & 0.03, [0,0.07] & 0.01, [-0.03,0.04] & -0.04, [-0.13,0.04] & -0.01, [-0.13,0.12] & 0.05, [-0.12,0.22] \\ 
poverty\textasteriskcentered race: white & 0.05, [-0.02,0.11] & 0, [-0.07,0.06] & -0.09, [-0.25,0.07] & 0.02, [-0.21,0.24] & -0.05, [-0.33,0.23] \\ 
total population\textasteriskcentered race: white & \textbackslash bf\{-0.03\}, [-0.05,-0.01] & -0.01, [-0.03,0.01] & 0, [-0.04,0.03] & -0.02, [-0.06,0.03] & -0.01, [-0.05,0.04] \\ 
unemployment\textasteriskcentered race: white & 0, [-0.03,0.03] & 0.01, [-0.02,0.04] & -0.02, [-0.1,0.06] & 0.04, [-0.06,0.15] & -0.01, [-0.16,0.14] \\ 
proportion white\textasteriskcentered race: white & 0.01, [-0.04,0.07] & 0.01, [-0.05,0.06] & -0.1, [-0.23,0.02] & 0.06, [-0.13,0.24] & -0.07, [-0.29,0.14] \\ 
implicit bias\textasteriskcentered race: white & -0.01, [-0.05,0.02] & -0.01, [-0.04,0.02] & \textbackslash bf\{-0.11\}, [-0.19,-0.04] & 0.05, [-0.05,0.16] & -0.1, [-0.24,0.03] \\ 
explicit bias\textasteriskcentered race: white & 0, [-0.03,0.03] & -0.01, [-0.04,0.02] & 0.08, [0,0.15] & -0.03, [-0.14,0.07] & 0.08, [-0.07,0.23] \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

This figure illustrates that in general, county-level explicit and
implicit biases in favor of straight individuals are not consistently
associated with racial disciplinary disparities. The one exception is
for law enforcement referrals, where we see that counties with higher
levels of implicit bias are expected to have a larger racial gap
(\(M_{slope}\) = -0.11, {[}-0.19, -0.03{]}). For this particular
disciplinar action, explicit biases tend to show the opposite
association, with counties higher in explicit bias expected to have a
smaller racial gap, though the evidence is not as conclusive
(\(M_{slope}\) = 0.08, {[}0, 0.15{]}). All other estimates are located
near zero, or are inconclusive with respect to the direction of the
effect.

\section{Discussion}\label{discussion}

These analyses across five types of disciplinary actions are fully
consistent with county-level estimates of explicit racial bias being
associated with racial disciplinary disparities. Specifically, in
counties where the white population is estimated to have higher rates of
explicit biases that favor whites, the difference in suspensions,
expulsions, law enforcement referrals, and school-related arrests
between black and white students is expected to be greater than in those
counties where the white population has lower rates of explicit racial
biases. In a secondary set of analyses with sexuality bias, we found
evidence that the aforementioned pattern cannot be explained by a more
general tendency to exhibit explicit biases to stigmatized groups, as
opposed to racial biases specifically, since most of the associations
with respect to county-level explicit and implicit sexuality bias are
directionally inconclusive. Additionally, our analyses cover the vast
majority of school-aged students in the United States, and our models
include a large set of covariates, suggesting that the relationships
observed are not attributable to other indicators that can often
co-occur with racial disparities, such as socioeconomic status or
population demographics (10, 31).

Of course, this research is not without limitations and anomalies. The
relationship between implicit racial bias and suspension was in an
unexpected direction. That is, greater implicit bias is somewhat
associated with smaller racial disparities in suspensions. One
possibility is that counties higher in implicit bias are not as likely
to suspend black students, but rather opt to use other disciplinary
tools unmeasured in these data. Another possibility is that people in
counties characterized by higher implicit bias are somewhat willing and
able to correct their behavioral responses in order to conform to their
egalitarian explicit beliefs or egalitarian social norms {[}(32). Much
research at the individual level shows that people are able to
counteract the automatic impulses associated with implicit bias when
they have the motivation and cognitive bandwidth to do so (33).
Additionally, because of the correlational nature of the analyses, it is
impossible to definitively establish the causal relationship between
explicit bias and disciplinary disparities. The conclusion that explicit
biases predict disciplinary disparities is consonant with a great deal
of research on disciplinary disparities {[}(34). However, it is also
possible that living in a region in which black students are disciplined
to a greater extent than white students exacerbates and/or reinforces
the explicit racial biases of community members, or that the
relationship between explicit racial biases and disciplinary disparities
is bi-directional. Finally, although we used a poststratification scheme
to make our estimates more representative with respect to county age
distributions, we cannot account for other ways in which Project
Implicit data are not representative of the general population. Our
analyses trade off the ability to ask these more detailed question with
the strengths of statistical power and population coverage offered by
the large datasets we used here. A full understanding of the complex
issues shaping racial disciplinary disparities will require integrating
findings from carefully controlled experiments, targeted surveys,
administrative databases, and large-scale observational data.

Nevertheless, our work compliments other research indicating that racial
dynamics are an important source for the observed differences in
disciplinary rates between black and white students. For instance,
students, caregivers, and administrators perceive suspensions and the
disproportionate use of them as at least partially racially motivated
(35--37). Additionally, other work has shown that even after controlling
for a range of other factors, race remains associated with the
likelihood of receiving disciplinary actions (10, 29). Additionally,
experimental evidence shows that disciplinary decision-making for
teachers differs depending on the race of the student (11). The present
research adds to this work by showing for the first time associations
between disciplinary actions and measurements of racial bias.

Our work also compliments existing studies examining the degree to which
implicit and explicit racial biases are associated with racial
disparities in key areas, such as health and policing (14--16) by
extending this type of inquiry to educational outcomes. To properly
assess the meaning of these findings, it is imperative that future work
focus specifically on what it means to exist in a community that is
estimated to have high or low levels of implicit or explicit bias.

As we have already highlighted, students who are subject to the
disciplinary actions we examined here are at substantially higher risk
for negative life outcomes (8). In dispensing these disciplinary
decisions differentially across racial groups, educational agencies are
also differentially allocating life prospects. Although implementing
training and education focused on implicit bias is popular among
educational administrators (38), and has been recommended by the US
Department of Education (39), our analysis suggests that effectively
combating racial disparities in discipline requires confronting more
explicit forms of racial bias, perhaps, for example, through peer
influence or interventions designed to change social norms (40).

We offer the research presented here to prompt additional scrutiny with
respect to how and why educational agencies in the United States
differentially administer disciplinary actions, especially when those
actions are known to have dire consequences for student welfare.
Although the focus here is on actions taking place within educational
agencies, the mechanisms responsible for these disparities likely exist,
at least partly, in the larger community. Through understanding and
reducing these disciplinary disparities specifically and the biases that
exist in the community more broadly, there exists an avenue for
education to maximize its promise as the great equalizer it has the
potential to be.

\subsection{Materials \& Methods}\label{materials-methods}

\subsubsection{Analytic Approach}\label{analytic-approach}

Because Project Implicit is a nonrandom sample, we used multilevel
regression and post-stratification to obtain accurate geographical
population-based estimates of implicit and explicit bias. This procedure
corrects for biased sampling and regularizes extreme observations with
little data to support them (e.g.~a county with only a handful of
respondents with especially high or low scores) (41, 42). Following past
work (14), we identified age as one dimension along which IAT
respondents differed from the general population in ways that could bias
our conclusions (43). Our post-stratification weighting scheme is as
follows: We first grouped respondents into five age group categories
(15-24, 25-34, 35-54, 55-75, and 75+). We next fit multilevel models
estimating bias (implicit and explicit biases separately) as a function
of our state-level covariates (the ``fixed'' effects: all identical to
those included in the final model, described below), and allowed the
estimates to vary by age bin, county, and state (the ``random''
effects). Next, we determined the population of whites in each county in
these age groups using the American Community Survey's 5-year estimates
ending in 2014. Finally, we used our estimated models to predict the
expected response for each age bin, in each county. Our final
county-level estimates are the average of the values predicted for the 5
age bins, weighted by the population size of that bin in that county. As
a result of this procedure, we can be confident that our county-level
estimates should more closely approximate what our estimates would look
like if the Project Implicit data were truly representative along the
age dimension in all counties.

We analyzed these data using a series of bayesian multilevel logistic
regressions. We modeled the probability that a student would be expelled
as a function of a set of effects that are constant across observations
(i.e.~fixed effects: race (dummy coded), implicit bias, explicit bias,
an interaction term between race and implicit bias, an interaction term
between race and explicit bias, and all covariates described below, plus
each of the covariates interacting with race) and a set of effects that
vary across counties (i.e.~random effects: overall intercept \& race).
We fit separate models for each of the outcomes. All numerical
predictors were standardized at the appropriate level (county, state)
before model estimation (for post-stratification as well as for final
inference) to help with estimation efficiency and interpretability. We
also set priors for the intercept and coefficients in the bayesian model
to be weakly informative normal distributions centered on zero with a
standard deviation of five. This corresponds to a prior belief that all
parameters take values between -15 and 15 with a probability of more
than .99. Realistically, values outside this range are extremely
unlikely given that all variables were standardized prior to estimation.
Priors on all other parameters (e.g.~variance of the county-level
intercepts) were left to default settings from the software package used
to fit the models.

Because of the computational demands of fitting such a high-dimensional
model to such a large dataset (the full model for each metric would
consist of over 6k parameters to approximately 170k observations), we
used a consensus monte carlo algorithm to obtain approximate posterior
distributions for the parameters of interest (44). The approximate
posteriors derived from this algorithm have been shown to be nearly
indistinguishable from the true posterior, a result we verified using a
small subset of our own data. \#\#\# Data Sources \#\#\#\# Disciplinary
actions

To assess rates of disciplinary action, we used data from the Civil
Rights Data Collection (CRDC) conducted by the US Department of
Education. The dataset we used comes from the 2013-2014 academic year
and has data on ``all public local and educational agencies and schools,
including long-term secure juvenile justice facilities, charter schools,
alternative schools, and schools serving students with disabilities.''
In total, the CRDC data represents 95507 institutions enrolling
approximately 50 million students, of which approximately 25.2 million
are white and 7.8 million are Black\footnote{We note that we initially
  preregistered a number of analyses concerning this work. There are a
  number of differences between the analyses we registered and those
  presented in the main text. The registered analyses can be found, in
  full, in the project's OSF page (\url{https://osf.io/pu79a/}).}.
Previous work using CRDC data have identified a number of districts
whose data are in error, and have excluded juvenile justice facilities,
as these institutions constitute dramatically different educational
environments, where the meaning of disciplinary actions may be quite
different (45). We followed similar practices, excluding all juvenile
justice facilities. Additionally, we excluded data for a specific
disciplinary metric for any schools which reported disciplining more
students than it reported enrolling for \emph{any} race for that metric
(e.g.~a school reported expelling 5 white students when they reported
enrolling less than that number). We also excluded any school for all
disciplinary actions if they had an overreporting error for 3 or more
metrics. After these exclusions are applied, the final sample used for
modeling consists of 93493 institutions, enrolling 49.8 million students
of which 25.1 million are white and 7.7 million are black.\footnote{Researchers
  who compared individual schools' out-of-school suspension rates
  reported in the 2011-2012 CRDC with suspension rates of the same
  schools as reported on state websites where available found a number
  of mismatches (45). We did not have the resources to conduct similar
  comparisons for all outcomes in all 93493 institutions; however,
  informal inspection of the data led us to find one school district for
  which this was the case. Including versus excluding schools from this
  district do not change the results in any meaningful way. The analyses
  presented herein include this district. The results of corresponding
  analyses without this district can be requested from the first author.
  It should be noted that the United States Government Accountability
  Office's recent descriptive overview of the 2013-2014 CRDC also did
  not exclude cases for which there was a discrepancy between
  disciplinary rates reported in this dataset and any disciplinary data
  reported on state websites(46). As such, the analyses here are
  consistent with the only other published examination of the 2013-2014
  CRDC data that we are aware of.} From these data, we focus on the
number of students by race (black and white) who were subjected to each
of the disciplinary actions described below.

\subparagraph{Types of disciplinary
actions}\label{types-of-disciplinary-actions}

We report here on five different types of disciplinary actions:
in-school suspensions, out-of-school suspensions, law enforcement
referrals, school-related arrests, and expulsions.

The most extreme outcome, \emph{expulsion}, is defined as when a student
is prohibited from returning to the educational institution for the
remainder of the school year or longer. The institution may or may not
set up alternative educational services for the student. \emph{Law
enforcement referrals} are actions where a student is ``reported to any
law enforcement agency or official for an incident including a school
police unit, for an incident that occurs on school grounds, during
school-related events, or while taking school transportation, regardless
of whether official action is taken. Citations, tickets, and court
referrals are considered referrals to law enforcement.''
\emph{School-related arrests} refer to ``an arrest of a student for any
activity conducted on school grounds, during off-campus school
activities (including while taking school transportation)''
\emph{Out-of-school suspensions} are actions where the student ``is
temporarily removed from his or her regular school for at least half a
day (but less than the remainder of the school year)''. Finally,
\emph{in-school suspensions} are actions where the student is
``temporarily removed from his or her regular classroom(s) for at least
half a day\ldots{},but remains under the direct supervision of school
personnel. Direct supervision means school personnel are physically in
the same location as the student under their supervision.'' (47)

\paragraph{Racial Bias}\label{racial-bias}

We used measurements of implicit and explicit bias available from data
collected through Project Implicit (30). For a full description of the
implicit and explicit bias measures available in these data, refer to
(30) and (14). We note only that we used the IAT D-score as a measure of
implicit bias, and the difference between reported warmth towards whites
and warmth towards black (both measured from 0=very cold to 10=very
warm) as a measure of explicit bias. Additionally, we used only
respondents who had geographic information that would allow us to place
them in a United States county, identified as White, and visited the
site anytime after it went live in 2002 through the end of 2014. This
consisted of approximately 1.3 million total respondents from 3099
counties. 1.19 million respondents provided data for the IAT, and 1.15
million provided explicit bias ratings.

\paragraph{Sexuality Bias}\label{sexuality-bias}

We used measurements of implicit and explicit sexuality bias available
from data collected through Project Implicit (30). We opted for these
measures as robustness checks because there were enough observations to
closely mimic the racial bias analyses, and they have been used for
similar purposes in previous work (15) As in the previous study, we used
the IAT D-score as a measure of implicit bias. Our warmth score
represented the average of the difference between straight men and gay
men and the difference between straight women and lesbian women. We used
only respondents who had geographic information that would allow us to
place them in a United States county and visited the anytime between
when it went live in 2002 through the end of 2014. This consisted of
approximately 798K total respondents from 3044 counties. Of these
respondents, 720K respondents provided data for the IAT, and 761K
respondents provided explicit bias ratings.

\paragraph{Covariates}\label{covariates}

Each county-level variable used as a covariate in the final model and
the corresponding state-level variable used as a predictor in the
post-stratification scheme (described below) were taken from the same
source. Population size and proportions, socioeconomic indicators,
mobility, and segregation indices were all taken from the American
Community Survey (ACS) 5-year estimates for the time period ending in
2014. Urban-rural indicators were taken from the 2010 US Census, and
crime rates were taken from the FBI Uniform Crime Reporting program, as
made available through the National Archive of Criminal Justice Data for
each year from 2010-2014\footnote{These years are chosen to for the FBI
  data to yield estimates that are similar to the 5-year estimates from
  the ACS data.}. Each of these variables is described below.

\subparagraph{Population size and
proportions}\label{population-size-and-proportions}

We obtained the total population, the proportion of the population that
is white, the proportion of the population that is black, and the ratio
of black-to-white people in the population from ACS table B02001.

\subparagraph{Socioeconomic indicators}\label{socioeconomic-indicators}

We obtained estimates for the percentage of population with a Bachelor's
degree or higher, the percentage of the population aged 16 or over in
the labor force that is unemployed, the median household income, and the
percentage of families and people whose income in the last year was
below the poverty line from the ACS table DP03.

\subparagraph{Urban-rural indicator}\label{urban-rural-indicator}

We obtained estimates of housing density per square mile of land area
from Census table GCT-PH1.

\subparagraph{Mobility}\label{mobility}

We obtained estimates of population mobility by summing the percentage
of Black Americans who moved from a different county, state, or country
into the county of interest (county-level covariate) or who moved from a
different state or country into the state of interest (state-level
covariate). We took these metrics from the ACS table S0701.

\subparagraph{Crime}\label{crime}

We computed estimates of the number of violent crimes per person by
taking the number of crimes reported divided by the population size for
each year, and averaging the resulting proportions across the 5 years of
data.

\subparagraph{Segregation}\label{segregation}

We computed a dissimiliarity index as described by (48). This metric is
computed based on the racial dissimilarity of census tracts within a
county, and reflects the proportion of a group within the county that
would have to move in order for all census tracts to have group
distributions that matched the overall distribution of the county. These
computations were done using data from the ACS table B02001.

\showmatmethods
\showacknow
\pnasbreak

\hypertarget{refs}{}
\hypertarget{ref-kinsler2011understanding}{}
1. Kinsler J (2011) Understanding the black--white school discipline
gap. \emph{Economics of Education Review} 30(6):1370--1383.

\hypertarget{ref-okonofua2016vicious}{}
2. Okonofua JA, Walton GM, Eberhardt JL (2016) A vicious cycle: A
social--psychological account of extreme racial disparities in school
discipline. \emph{Perspectives on psychological science} 11(3):381--398.

\hypertarget{ref-aud2011america}{}
3. Aud S, KewalRamani A, Frohlich L (2011) America's youth: Transitions
to adulthood. nces 2012-026. \emph{National Center for Education
Statistics}.

\hypertarget{ref-yeager2017loss}{}
4. Yeager DS, Purdie-Vaughns V, Hooper SY, Cohen GL (2017) Loss of
institutional trust among racial and ethnic minority adolescents: A
consequence of procedural injustice and a cause of life-span outcomes.
\emph{Child development} 88(2):658--676.

\hypertarget{ref-gregory1995crime}{}
5. Gregory JF (1995) The crime of punishment: Racial and gender
disparities in the use of corporal punishment in us public schools.
\emph{Journal of Negro Education}:454--462.

\hypertarget{ref-skiba2011race}{}
6. Skiba RJ, et al. (2011) Race is not neutral: A national investigation
of african american and latino disproportionality in school discipline.
\emph{School Psychology Review} 40(1):85.

\hypertarget{ref-pager2009sequencing}{}
7. Pager D, Western B, Sugie N (2009) Sequencing disadvantage: Barriers
to employment facing young black and white men with criminal records.
\emph{The ANNALS of the American Academy of Political and Social
Science} 623(1):195--213.

\hypertarget{ref-hirschfield2009another}{}
8. Hirschfield P (2009) Another way out: The impact of juvenile arrests
on high school dropout. \emph{Sociology of Education} 82(4):368--393.

\hypertarget{ref-togut2011gestalt}{}
9. Togut TD (2011) The gestalt of the school-to-prison pipeline: The
duality of overrepresentation of minorities in special education and
racial disparity in school discipline eon minorities. \emph{American
University Journal of Gender, Social Policy \& the Law} 20:163.

\hypertarget{ref-skiba2002color}{}
10. Skiba RJ, Michael RS, Nardo AC, Peterson RL (2002) The color of
discipline: Sources of racial and gender disproportionality in school
punishment. \emph{The urban review} 34(4):317--342.

\hypertarget{ref-okonofua2015two}{}
11. Okonofua JA, Eberhardt JL (2015) Two strikes: Race and the
disciplining of young students. \emph{Psychological science}
26(5):617--624.

\hypertarget{ref-gregory2008discipline}{}
12. Gregory A, Weinstein RS (2008) The discipline gap and african
americans: Defiance or cooperation in the high school classroom.
\emph{Journal of School Psychology} 46(4):455--475.

\hypertarget{ref-rudman2004sources}{}
13. Rudman LA (2004) Sources of implicit attitudes. \emph{Current
Directions in Psychological Science} 13(2):79--82.

\hypertarget{ref-leitner2016blacks}{}
14. Leitner JB, Hehman E, Ayduk O, Mendoza-Denton R (2016) Blacks' death
rate due to circulatory diseases is positively related to whites'
explicit racial bias: A nationwide investigation using project implicit.
\emph{Psychological science} 27(10):1299--1311.

\hypertarget{ref-orchard2017county}{}
15. Orchard J, Price J (2017) County-level racial prejudice and the
black-white gap in infant health outcomes. \emph{Social Science \&
Medicine} 181:191--198.

\hypertarget{ref-hehman2017disproportionate}{}
16. Hehman E, Flake JK, Calanchini J (2017) Disproportionate use of
lethal force in policing is associated with regional racial biases of
residents. \emph{Social Psychological and Personality
Science}:1948550617711229.

\hypertarget{ref-verdugo2002race}{}
17. Verdugo RR (2002) Race-ethnicity, social class, and zero-tolerance
policies: The cultural and structural wars. \emph{Education and urban
society} 35(1):50--75.

\hypertarget{ref-cadinu2003stereotype}{}
18. Cadinu M, Maass A, Frigerio S, Impagliazzo L, Latinotti S (2003)
Stereotype threat: The effect of expectancy on performance.
\emph{European Journal of Social Psychology} 33(2):267--285.

\hypertarget{ref-darley1980expectancy}{}
19. Darley JM, Fazio RH (1980) Expectancy confirmation processes arising
in the social interaction sequence. \emph{American psychologist}
35(10):867.

\hypertarget{ref-major2005social}{}
20. Major B, O'brien LT (2005) The social psychology of stigma.
\emph{Annu Rev Psychol} 56:393--421.

\hypertarget{ref-jetten1996intergroup}{}
21. Jetten J, Spears R, Manstead AS (1996) Intergroup norms and
intergroup discrimination: Distinctive self-categorization and social
identity effects. \emph{Journal of personality and social psychology}
71(6):1222.

\hypertarget{ref-pettigrew1958personality}{}
22. Pettigrew TF (1958) Personality and sociocultural factors in
intergroup attitudes: A cross-national comparison. \emph{Journal of
conflict resolution} 2(1):29--42.

\hypertarget{ref-payne2017bias}{}
23. Payne BK, Vuletich HA, Lundberg KB (2017) The bias of crowds: How
implicit bias bridges personal and systemic prejudice.
\emph{Psychological Inquiry} 28(4):233--248.

\hypertarget{ref-markus2010cultures}{}
24. Markus HR, Kitayama S (2010) Cultures and selves: A cycle of mutual
constitution. \emph{Perspectives on psychological science}
5(4):420--430.

\hypertarget{ref-cuellar2015school}{}
25. Cuellar AE, Markowitz S (2015) School suspension and the
school-to-prison pipeline. \emph{International Review of Law and
Economics} 43:98--106.

\hypertarget{ref-matjasko2011effective}{}
26. Matjasko JL (2011) How effective are severe disciplinary policies?
School policies and offending from adolescence into young adulthood.
\emph{Journal of school psychology} 49(5):555--572.

\hypertarget{ref-rausch2005academic}{}
27. Rausch MK, Skiba RJ, Simmons A (2005) The academic cost of
discipline: The relationship between suspension/expulsion and school
achievement. \emph{Annual Meeting of the American Educational Research
Association, Montreal, Canada}.

\hypertarget{ref-hemphill2006effect}{}
28. Hemphill SA, Toumbourou JW, Herrenkohl TI, McMorris BJ, Catalano RF
(2006) The effect of school suspensions and arrests on subsequent
adolescent antisocial behavior in australia and the united states.
\emph{Journal of Adolescent Health} 39(5):736--744.

\hypertarget{ref-anyon2014persistent}{}
29. Anyon Y, et al. (2014) The persistent effect of race and the promise
of alternatives to suspension in school discipline outcomes.
\emph{Children and Youth Services Review} 44:379--386.

\hypertarget{ref-xu2014psychology}{}
30. Xu K, Nosek B, Greenwald A (2014) Psychology data from the race
implicit association test on the project implicit demo website.
\emph{Journal of Open Psychology Data} 2(1).

\hypertarget{ref-williams1999race}{}
31. Williams DR (1999) Race, socioeconomic status, and health the added
effects of racism and discrimination. \emph{Annals of the New York
Academy of Sciences} 896(1):173--188.

\hypertarget{ref-pryor2004dual}{}
32. Pryor JB, Reeder GD, Yeadon C, Hesson-McInnis M (2004) A
dual-process model of reactions to perceived stigma. \emph{Journal of
personality and social psychology} 87(4):436.

\hypertarget{ref-devine2002regulation}{}
33. Devine PG, Plant EA, Amodio DM, Harmon-Jones E, Vance SL (2002) The
regulation of explicit and implicit race bias: The role of motivations
to respond without prejudice. \emph{Journal of personality and social
psychology} 82(5):835.

\hypertarget{ref-mcintosh2014education}{}
34. McIntosh K, Girvan EJ, Horner RH, Smolkowski K (2014) Education not
incarceration: A conceptual model for reducing racial and ethnic
disproportionality in school discipline. \emph{Journal of Applied
Research on Children: Informing Policy for Children at Risk} 5(2):4.

\hypertarget{ref-haight2014ecological}{}
35. Haight W, Gibson PA, Kayama M, Marshall JM, Wilson R (2014) An
ecological-systems inquiry into racial disproportionalities in
out-of-school suspensions from youth, caregiver and educator
perspectives. \emph{Children and Youth Services Review} 46:128--138.

\hypertarget{ref-gibson2013caregivers}{}
36. Gibson PA, Haight W (2013) Caregivers' moral narratives of their
african american children's out-of-school suspensions: Implications for
effective family--school collaborations. \emph{Social work}
58(3):263--272.

\hypertarget{ref-ruck2002racial}{}
37. Ruck MD, Wortley S (2002) Racial and ethnic minority high school
students' perceptions of school disciplinary practices: A look at some
canadian findings. \emph{Journal of Youth and Adolescence}
31(3):185--195.

\hypertarget{ref-staats2016understanding}{}
38. Staats C (2016) Understanding implicit bias: What educators should
know. \emph{American Educator} 39(4):29.

\hypertarget{ref-us2014dear}{}
39. US Department of Education \& US Department of Justice (2014)
\emph{Dear colleague letter on the nondiscriminatory administration of
school discipline} (US Department of Education, Office for Civil Rights;
US Department of Justice Washington, DC) Available at:
\url{https://www2.ed.gov/about/offices/list/ocr/letters/colleague-201401-title-vi.pdf}.

\hypertarget{ref-paluck2009prejudice}{}
40. Paluck EL, Green DP (2009) Prejudice reduction: What works? A review
and assessment of research and practice. \emph{Annual review of
psychology} 60:339--367.

\hypertarget{ref-park2004bayesian}{}
41. Park DK, Gelman A, Bafumi J (2004) Bayesian multilevel estimation
with poststratification: State-level estimates from national polls.
\emph{Political Analysis} 12(4):375--385.

\hypertarget{ref-gelman1997poststratification}{}
42. Gelman A, Little TC (1997) Poststratification into many categories
using hierarchical logistic regression. \emph{Survey Methodology}
23(2):127--35.

\hypertarget{ref-gonsalkorale2009aging}{}
43. Gonsalkorale K, Sherman JW, Klauer KC (2009) Aging and prejudice:
Diminished regulation of automatic race bias among older adults.
\emph{Journal of Experimental Social Psychology} 45(2):410--414.

\hypertarget{ref-scott2016bayes}{}
44. Scott SL, et al. (2016) Bayes and big data: The consensus monte
carlo algorithm. \emph{International Journal of Management Science and
Engineering Management} 11(2):78--88.

\hypertarget{ref-losen2015we}{}
45. Losen DJ, et al. (2015) Are we closing the school discipline gap?

\hypertarget{ref-usgao2018discipline}{}
46. United States Government Accountability Office (2018)
\emph{Discipline disparities for black students, boys, and students with
disabilities} Available at:
\url{https://www.gao.gov/assets/700/690828.pdf}.

\hypertarget{ref-crdc2013survey}{}
47. CRDC Survey (2014) 2013-2014 civil rights data collection survey.
Available at:
\href{/url\%7Bhttp://www2.ed.gov/about/offices/list/ocr/docs/crdc-2013-14-p1-p4.doc\%7D}{\textbackslash{}url\{http://www2.ed.gov/about/offices/list/ocr/docs/crdc-2013-14-p1-p4.doc\}}.

\hypertarget{ref-massey1988dimensions}{}
48. Massey DS, Denton NA (1988) The dimensions of residential
segregation. \emph{Social forces} 67(2):281--315.

% Bibliography
% \bibliography{pnas-sample}

\end{document}

### 2018-05-18 written_docs/Riddle_Sinclair_PNAS/osf_supplement.tex
.gitignore is now working

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={OSF Supplementary Information},
            pdfauthor={Travis Riddle \& Stacey Sinclair},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{OSF Supplementary Information}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Travis Riddle \& Stacey Sinclair}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{5/7/2018}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\begin{document}
\maketitle

\section{Session details}\label{session-details}

Data analysis was done in R (R Core Team 2016) version 3.3.2 running
under OS X 10.11.6. Post-stratification was done with lme4, version
1.1.14 (Bates et al. 2015). Final model fitting was done on the
university cluster running Springdale Linux, release 6.9 using rstanarm,
version 2.17.2 (Stan Development Team 2016). We used the implementation
of the consensus monte carlo algorithm found in parallelMCMCcombine,
version 1.0 (Miroshnikov and Conlon 2014). Figures were made with
ggplot2, version 2.2.1 (Wickham 2009), with data manipulation done using
dplyr version 0.7.2 (Wickham et al. 2017) and tidyr, version 0.7.1
(Wickham and Henry 2017). Session info for computations on the local
machine is as follows:

\begin{verbatim}
## R version 3.3.2 (2016-10-31)
## Platform: x86_64-apple-darwin13.4.0 (64-bit)
## Running under: OS X El Capitan 10.11.6
## 
## locale:
## [1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
## 
## attached base packages:
## [1] stats     graphics  grDevices utils     datasets  methods   base     
## 
## other attached packages:
## [1] bindrcpp_0.2     stringr_1.2.0    kableExtra_0.8.0 tidyr_0.7.1     
## [5] forcats_0.3.0    haven_1.0.0      ggplot2_2.2.1    dplyr_0.7.4     
## [9] knitr_1.20      
## 
## loaded via a namespace (and not attached):
##  [1] Rcpp_0.12.14      plyr_1.8.4        bindr_0.1        
##  [4] tools_3.3.2       digest_0.6.13     viridisLite_0.3.0
##  [7] evaluate_0.10     tibble_1.3.4      gtable_0.2.0     
## [10] pkgconfig_2.0.1   rlang_0.2.0.9000  rstudioapi_0.6   
## [13] yaml_2.1.14       httr_1.3.1        xml2_1.1.1       
## [16] rprojroot_1.2     grid_3.3.2        glue_1.1.1       
## [19] R6_2.2.2          rmarkdown_1.9     bookdown_0.4     
## [22] purrr_0.2.4       readr_1.0.0       magrittr_1.5     
## [25] codetools_0.2-15  backports_1.0.5   scales_0.5.0.9000
## [28] htmltools_0.3.6   assertthat_0.2.0  rvest_0.3.2      
## [31] colorspace_1.3-2  stringi_1.1.2     lazyeval_0.2.1   
## [34] munsell_0.4.3
\end{verbatim}

Session info for the computations run on the university cluster is as
follows:

\begin{verbatim}
R version 3.4.3 (2017-11-30)
Platform: x86_64-redhat-linux-gnu (64-bit)
Running under: Springdale Linux 7.4 (Verona)

Matrix products: default
BLAS/LAPACK: /usr/lib64/R/lib/libRblas.so

locale:
 [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C
 [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8
 [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8
 [7] LC_PAPER=en_US.UTF-8       LC_NAME=C
 [9] LC_ADDRESS=C               LC_TELEPHONE=C
[11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
[1] rstanarm_2.15.3 Rcpp_0.12.13    ggplot2_2.2.1   forcats_0.2.0
[5] tidyr_0.7.2     dplyr_0.7.4

loaded via a namespace (and not attached):
 [1] lattice_0.20-35      zoo_1.8-0            gtools_3.5.0
 [4] assertthat_0.2.0     digest_0.6.12        mime_0.5
 [7] R6_2.2.2             plyr_1.8.4           stats4_3.4.3
[10] colourpicker_1.0     rlang_0.1.2          lazyeval_0.2.1
[13] minqa_1.2.4          miniUI_0.1.1         nloptr_1.0.4
[16] Matrix_1.2-12        DT_0.2               shinythemes_1.1.1
[19] splines_3.4.3        shinyjs_0.9.1        lme4_1.1-14
[22] stringr_1.2.0        htmlwidgets_0.9      loo_1.1.0
[25] igraph_1.1.2         munsell_0.4.3        shiny_1.0.5
[28] compiler_3.4.3       httpuv_1.3.5         rstan_2.16.2
[31] pkgconfig_2.0.1      base64enc_0.1-3      rstantools_1.3.0
[34] htmltools_0.3.6      tibble_1.3.4         gridExtra_2.3
[37] threejs_0.3.1        codetools_0.2-15     matrixStats_0.52.2
[40] MASS_7.3-47          grid_3.4.3           nlme_3.1-131
[43] xtable_1.8-2         gtable_0.2.0         magrittr_1.5
[46] StanHeaders_2.16.0-1 scales_0.5.0         stringi_1.1.5
[49] reshape2_1.4.2       bindrcpp_0.2         dygraphs_1.1.1.4
[52] xts_0.10-0           tools_3.4.3          glue_1.2.0
[55] shinystan_2.4.0      markdown_0.8         purrr_0.2.4
[58] crosstalk_1.0.0      rsconnect_0.8.5      parallel_3.4.3
[61] inline_0.3.14        colorspace_1.3-2     bayesplot_1.4.0
[64] bindr_0.1
\end{verbatim}

\section{Preregistered analysis}\label{preregistered-analysis}

Here, we describe, in detail, the differences between the registered
analysis plan and that presented in the main text, along with our
reasons for making the alterations.

\subsection{Data exclusions}\label{data-exclusions}

In our preanalysis plan, we specified our analyses to focus on 13
actions - corporal punishment, in-school suspension, out-of-school
suspension, expulsion with educational services, expulsion without
educational services, expulsion under zero-tolerance policies, referral
to law enforcement, school-related arrests, mechanical restraint,
physical restraint, seclusion, preschool suspension, and preschool
expulsion. However, upon further study, we discovered reasons we thought
justified excluding a number of these outcomes.

First, seclusion, physical restraint, and mechanical restraint are not
disciplinary actions, but are rather used as means to restrain students
who are at risk of harming themselves or others.

Next, we discovered that the administration of corporal punishment was
extremely irregular across counties, with almost all cases ocurring in
the South. Specifically, Alabama, Arkansas, Florida, Georgia, Louisiana,
Missouri, Mississippi, Oklahoma, Tennessee, and Texas accounted for 98\%
of all educational institutions that administered corporal punishment at
least once, despite containing only 26.9\% of the educational
institutions in the full dataset. It is not clear whether this uneven
distribution is due to legal prohibitions, cultural differences, or some
combination of the two. Because this makes interpretation of the
parameters uncertain, this analysis is also placed in the appendix.

We also found that the number of preschool students who are expelled or
suspended is vanishingly small (131 total expulsions and 6751 total
suspensions out of over 1.4 million enrolled preschool students), making
reliably estimating any association across counties exceedingly
unlikely. We additionally discovered that counts of one expulsion
category (expulsion under zero-tolerance policies) overlapped with
counts in other categories, and so excluded this category from the main
text. We also decided to combine the remaining two expulsion categories
to yield one overall count of the number of students expelled. We did
this to remain consistent with previous research, and to obtain
district-level rates of expulsion that were slightly higher, and
therefore more amenable to exploring variation across counties.

When preparing the preregistration, we also had not known about the
issues with juvenile justice facilities, or with the school districts
with reporting errors, and so excluded these schools from analyses in
the main text.

\subsection{Changes in measurement \& estimation of
bias}\label{changes-in-measurement-estimation-of-bias}

We preregistered our explicit bias as a simple feeling thermometer
towards balcks (i.e. \emph{how warm or cold do you feel towards Blacks?
0=very cold, 10=very warm}). However, the majority of past research (see
for example Leitner et al. (2016) and Hehman, Flake, and Calanchini
(2017)) has used the difference in reported warmth towards whites and
blacks, and so in the main text, we report models using this metric of
explicit bias.

Our main text presents analyses with poststratified estimates of bias.
Our registered analysis also described analyses with raw, county-based
means. Unlike previous papers that used poststratification, we see
substantial differences in results between the two estimation methods.
Because post-stratification is known to yield better estimates, we
suspect that the difference in results is largely driven by extreme
county-level observations with little data to support these extreme
estimates. Nevertheless, in the interest of completion, we present the
model using raw county-level means as well.

We also registered an exploratory analysis using the responses of people
who visited Project Implicit and identified themselves as teachers.
Unfortunately, without population-level data on the demographic
characteristics of teachers, we are not able to post-stratify these
estimates. Additionally, there were relatively few teachers in the
Project Implicit from which to derive estimates. Ultimately, we have
little confidence that this analysis tells us anything about whether the
biases of teachers specifically contribute to disparities in
disciplinary actions, and so this analysis is also limited to the
appendix.

\subsection{Additional covariates}\label{additional-covariates}

We included four covariates that were not specified in the registered
analysis. Specifically, the mobility, housing density, crime rate, and
dissimilarity covariates. We also interacted all covariates with race.
These modifications were done in order to cover more of the covariates
that were included in previous papers (Leitner et al. 2016).

\section{Registered analysis results}\label{registered-analysis-results}

Here, we present the results of the preregistered analyses exactly.
Though we described a number of changes above, it is not practical to
run all possible combination of changes, as each set of models takes
several days to just to estimate, and there are at least several dozen
possible combinations of decisions to test, depending on how
fine-grained one wishes to be with the decisions. Accordingly, we limit
the presentation of results below to just the three models that we
registered.

\subsection{County-level estimates}\label{county-level-estimates}

\begin{figure}
\centering
\includegraphics{osf_supplement_files/figure-latex/raw-county-estimates-1.pdf}
\caption{\label{fig:raw-county-estimates}\label{fig:raw-county-estimates}Difference
in raw county means and post-stratifed estimates of implicit bias as a
function of the number of observations in each county. The X axis is on
the log scale.}
\end{figure}

Examining the models fit with simple county means, we find that the
county-level estimates vary more than with the post-stratified
estimates. This pattern is expected, as one desireable feature of
post-stratification is that it regularizes extreme observations without
a large amount of data to support them. This phenomena is illustrated in
figure \ref{fig:raw-county-estimates}, which shows the distribution of
differences between the raw county means and the post-stratified
estimates of implicit bias as a function of the number of observations
in each county. More specifically, while the mean of the county-level
means does not meaningfully differ from the post-stratified estimates,
the standard deviation is much larger (\(M_{implicit}\) = 0.4,
\(SD_{implicit}\) = 0.11; \(M_{explicit}\) = 6.24, \(SD_{explicit}\) =
0.59)\footnote{This explicit estimate is based on the simple warmth
  towards blacks question, and not the difference between warmth towards
  whites and warmth twoards blacks.}

The county-level implicit bias estimates for the poststratified model
are as reported in the main text. The estimates for explicit bias are
slightly different, as they are not computed based on the difference in
reporting warmth toward whites and warmth toward blacks. Specifically,
the average amount of bias at the county-level is 6.25 (SD = 0.14).

A subset of years in the Project Implicit data also collected
occupational information from respondents. As identified in our
pre-analysis plan, we took advantage of the presence of primary and
secondary educators in these data to test whether any associations
between bias and race-based differences in the rates of disciplinary
action were stronger among these respondents. Filtering for only white
individuals who identified as primary, secondary, special education, and
other teachers and instructors (occupation codes 25-2000 and 25-3000)
reduced the dataset to 76959 respondents. In order to assure that our
estimates were reasonably stable, we limited analysis to only counties
that had at least 50 respondents. As such, our teacher analysis is
limited to just 327 counties. Additionally, because we do not know of
any state-level demographic estimates for teachers, we were unable to
perform post-stratification for these data. Nonetheless, across these
limited counties, the overall estimate of implicit and explicit bias are
similar to those for the whole dataset (\(M_{implicit}\) = 0.37,
\(SD_{implicit}\) = 0.06; \(M_{explicit}\) = 6.68, \(SD_{explicit}\) =
0.27).

For all models, explicit and implicit bias were scored such that higher
scores reflect a greater preference for whites/reduced preference for
blacks.

\subsection{Model estimates}\label{model-estimates}

Tables \ref{tab:mod-raw-table1} and \ref{tab:mod-raw-table2} present
results from the model fit to the raw, county-level means (we broke
results from each set of estimates into two table to improve
legibility). The table presents parameter estimates (operationalized as
the mean of the posterior) and 95\% uncertainty intervals for each of
the population-level effects for each of the outcomes. Tables
\ref{tab:mod-post-table1} and \ref{tab:mod-post-table2} present the
results for post-stratified estimates, with tables
\ref{tab:mod-teach-table1} and \ref{tab:mod-teach-table2} showing
results for the teacher models.

\begin{landscape}\begin{table}

\caption{\label{tab:mod-raw-table1}\label{tab:model-raw-table1}Regression coefficient estimates for the population-level (i.e. fixed) effects, 
             along with 95\% uncertainty intervals for each of the disciplinary metrics. Model uses raw 
             county-means as bias estimates.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{lllllll}
\toprule
 & Seclusion & \makecell[l]{Physical\\Restraint} & \makecell[l]{Mechanical\\Restraint} & \makecell[l]{Corporal\\Punishment} & \makecell[l]{Preschool\\Suspension} & \makecell[l]{Preschool\\Expulsion}\\
\midrule
Intercept & \textbf{-11.19 [-11.64,-10.76]} & \textbf{-8.78 [-9,-8.58]} & \textbf{-14.1 [-14.98,-13.31]} & \textbf{-3.21 [-3.31,-3.13]} & \textbf{-8.84 [-9.16,-8.56]} & \textbf{-16.07 [-17.98,-14.37]}\\
proportion black & \textbf{-1.51 [-2.19,-0.83]} & \textbf{-0.67 [-1.04,-0.31]} & 0.8 [-0.13,1.75] & 0.11 [-0.15,0.37] & \textbf{1.02 [0.6,1.46]} & 0.99 [-0.51,2.52]\\
proportion white & 0.26 [-0.2,0.73] & 0.19 [-0.08,0.45] & 0.26 [-0.5,1.04] & \textbf{-0.3 [-0.5,-0.08]} & \textbf{-0.44 [-0.8,-0.07]} & -0.34 [-1.54,0.95]\\
black-white ratio & \textbf{0.81 [0.15,1.44]} & 0.26 [-0.08,0.6] & -0.59 [-1.47,0.26] & \textbf{-0.34 [-0.52,-0.15]} & \textbf{-0.78 [-1.14,-0.44]} & -0.34 [-1.61,0.81]\\
total population & \textbf{0.16 [0,0.31]} & 0.02 [-0.07,0.12] & \textbf{0.41 [0.21,0.64]} & \textbf{-0.48 [-0.67,-0.3]} & 0.05 [-0.06,0.17] & \textbf{0.49 [0.24,0.77]}\\
\addlinespace
college grads & \textbf{0.25 [0.01,0.49]} & \textbf{0.3 [0.17,0.43]} & -0.04 [-0.47,0.36] & \textbf{-0.15 [-0.24,-0.05]} & \textbf{-0.21 [-0.4,-0.02]} & -0.24 [-0.96,0.45]\\
income & \textbf{-0.32 [-0.63,-0.02]} & -0.18 [-0.37,0] & 0 [-0.53,0.53] & \textbf{-0.33 [-0.48,-0.18]} & \textbf{-0.3 [-0.57,-0.03]} & -0.35 [-1.33,0.56]\\
poverty & \textbf{-0.64 [-1.05,-0.22]} & -0.14 [-0.37,0.09] & 0.3 [-0.34,0.95] & 0.01 [-0.15,0.17] & \textbf{-0.32 [-0.65,0]} & \textbf{-1.21 [-2.46,-0.03]}\\
unemployment & \textbf{0.4 [0.15,0.68]} & \textbf{0.23 [0.08,0.38]} & -0.21 [-0.65,0.23] & \textbf{-0.14 [-0.23,-0.05]} & \textbf{-0.29 [-0.48,-0.09]} & -0.3 [-1.12,0.53]\\
implicit bias & -0.02 [-0.38,0.36] & 0 [-0.23,0.22] & -0.07 [-0.59,0.46] & -0.01 [-0.1,0.08] & -0.16 [-0.4,0.07] & 0.04 [-0.96,1.09]\\
\addlinespace
explicit bias & 0.33 [-0.04,0.69] & 0.07 [-0.13,0.28] & -0.14 [-0.65,0.37] & -0.06 [-0.15,0.03] & -0.17 [-0.41,0.06] & -0.05 [-1.11,0.99]\\
race: white & \textbf{-1.04 [-1.39,-0.67]} & \textbf{-1.01 [-1.19,-0.82]} & \textbf{-1.72 [-2.36,-1.05]} & \textbf{-0.68 [-0.73,-0.62]} & \textbf{-1.65 [-1.95,-1.35]} & \textbf{-1.91 [-3.45,-0.41]}\\
implicit bias*race: white & 0.12 [-0.24,0.49] & 0.08 [-0.15,0.3] & 0.06 [-0.45,0.57] & -0.04 [-0.1,0.03] & 0.2 [-0.06,0.46] & -0.03 [-1.28,1.19]\\
explicit bias*race: white & -0.22 [-0.57,0.14] & -0.01 [-0.21,0.2] & -0.09 [-0.61,0.42] & 0 [-0.06,0.06] & 0.08 [-0.17,0.35] & -0.4 [-1.72,0.88]\\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\begin{landscape}\begin{table}

\caption{\label{tab:mod-raw-table2}\label{tab:model-raw-table1}Regression coefficient estimates for the population-level (i.e. fixed) effects, 
             along with 95\% uncertainty intervals for each of the disciplinary metrics. Model uses raw 
             county-means as bias estimates.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{llllllll}
\toprule
 & \makecell[l]{Out-of-School\\Suspension} & \makecell[l]{Law Enf.\\Referral} & \makecell[l]{In-School\\Suspension} & \makecell[l]{School-Related\\Arrest} & \makecell[l]{Expulsion w/o\\ Ed. Services} & \makecell[l]{Expulsion w/Ed.\\Services} & \makecell[l]{Expulsion Under\\Zero-Tolerance}\\
\midrule
Intercept & \textbf{-2.68 [-2.71,-2.65]} & \textbf{-6.12 [-6.2,-6.03]} & \textbf{-2.63 [-2.67,-2.59]} & \textbf{-8.2 [-8.37,-8.04]} & \textbf{-8.22 [-8.39,-8.06]} & \textbf{-8.11 [-8.28,-7.95]} & \textbf{-9.74 [-10,-9.49]}\\
proportion black & \textbf{0.34 [0.27,0.42]} & \textbf{-0.45 [-0.64,-0.26]} & \textbf{0.61 [0.51,0.71]} & -0.19 [-0.5,0.13] & \textbf{0.32 [0.03,0.62]} & -0.29 [-0.6,0.02] & 0.02 [-0.36,0.42]\\
proportion white & 0.02 [-0.03,0.08] & \textbf{-0.34 [-0.48,-0.2]} & \textbf{0.13 [0.05,0.21]} & -0.06 [-0.3,0.19] & 0.23 [-0.03,0.48] & \textbf{-0.25 [-0.47,-0.02]} & 0.21 [-0.09,0.52]\\
black-white ratio & \textbf{-0.21 [-0.27,-0.15]} & 0.06 [-0.1,0.23] & \textbf{-0.53 [-0.62,-0.45]} & 0.09 [-0.17,0.35] & -0.13 [-0.35,0.09] & -0.22 [-0.53,0.08] & -0.23 [-0.61,0.14]\\
total population & 0.01 [-0.02,0.03] & 0.01 [-0.06,0.07] & -0.03 [-0.07,0] & 0.09 [0,0.19] & 0.01 [-0.08,0.1] & 0.06 [-0.03,0.16] & 0.09 [-0.02,0.2]\\
\addlinespace
college grads & \textbf{-0.04 [-0.07,-0.01]} & 0.03 [-0.06,0.11] & \textbf{-0.12 [-0.16,-0.07]} & \textbf{0.15 [0.01,0.29]} & \textbf{-0.2 [-0.34,-0.06]} & -0.06 [-0.19,0.07] & -0.08 [-0.26,0.09]\\
income & \textbf{-0.09 [-0.14,-0.05]} & -0.05 [-0.16,0.06] & \textbf{-0.08 [-0.14,-0.02]} & 0 [-0.18,0.17] & \textbf{-0.31 [-0.49,-0.13]} & -0.09 [-0.26,0.09] & -0.08 [-0.3,0.14]\\
poverty & 0.02 [-0.03,0.08] & \textbf{-0.4 [-0.53,-0.27]} & \textbf{0.2 [0.12,0.26]} & \textbf{-0.22 [-0.44,0]} & \textbf{-0.35 [-0.57,-0.13]} & -0.15 [-0.38,0.06] & 0.2 [-0.07,0.46]\\
unemployment & \textbf{0.22 [0.18,0.25]} & \textbf{0.19 [0.1,0.27]} & 0.03 [-0.01,0.07] & \textbf{0.25 [0.1,0.4]} & \textbf{0.45 [0.31,0.59]} & \textbf{0.27 [0.13,0.41]} & \textbf{0.31 [0.14,0.48]}\\
implicit bias & 0.01 [-0.03,0.05] & -0.06 [-0.17,0.04] & \textbf{0.08 [0.03,0.12]} & -0.09 [-0.26,0.1] & 0.04 [-0.12,0.21] & -0.1 [-0.28,0.08] & -0.03 [-0.27,0.21]\\
\addlinespace
explicit bias & 0.03 [-0.01,0.06] & 0.09 [-0.01,0.19] & 0 [-0.04,0.04] & -0.02 [-0.19,0.15] & -0.01 [-0.17,0.14] & 0.02 [-0.14,0.19] & \textbf{0.3 [0.08,0.52]}\\
race: white & \textbf{-1.06 [-1.09,-1.03]} & \textbf{-0.83 [-0.9,-0.77]} & \textbf{-0.85 [-0.87,-0.82]} & \textbf{-1.07 [-1.18,-0.94]} & \textbf{-0.88 [-1.02,-0.74]} & \textbf{-0.85 [-0.98,-0.72]} & \textbf{-0.49 [-0.71,-0.27]}\\
implicit bias*race: white & -0.01 [-0.04,0.03] & -0.03 [-0.13,0.06] & \textbf{-0.04 [-0.07,-0.01]} & 0.01 [-0.16,0.16] & -0.09 [-0.24,0.05] & 0.11 [-0.04,0.26] & -0.19 [-0.41,0.04]\\
explicit bias*race: white & 0 [-0.03,0.03] & \textbf{-0.1 [-0.18,-0.01]} & -0.02 [-0.05,0.01] & -0.02 [-0.17,0.12] & 0 [-0.14,0.14] & -0.02 [-0.17,0.12] & -0.14 [-0.34,0.07]\\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\begin{landscape}\begin{table}

\caption{\label{tab:mod-post-table1}\label{tab:model-post-table1}Regression coefficient estimates for the 
             population-level (i.e. fixed) effects, along with 95\% 
             uncertainty intervals for each of the disciplinary metrics. 
             Model uses bias estimates as derived from poststratification 
             scheme.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{lllllll}
\toprule
 & Seclusion & \makecell[l]{Physical\\Restraint} & \makecell[l]{Mechanical\\Restraint} & \makecell[l]{Corporal\\Punishment} & \makecell[l]{Preschool\\Suspension} & \makecell[l]{Preschool\\Expulsion}\\
\midrule
Intercept & \textbf{-11.11 [-11.55,-10.7]} & \textbf{-8.73 [-8.94,-8.53]} & \textbf{-13.96 [-14.8,-13.17]} & \textbf{-3.24 [-3.33,-3.15]} & \textbf{-9 [-9.35,-8.69]} & \textbf{-14.17 [-15.97,-12.57]}\\
proportion black & \textbf{-0.79 [-1.45,-0.11]} & -0.4 [-0.8,0.02] & 0.86 [-0.19,1.9] & 0.13 [-0.17,0.41] & \textbf{0.78 [0.32,1.25]} & 0.58 [-0.89,2.15]\\
proportion white & \textbf{0.49 [0.02,0.99]} & \textbf{0.28 [0,0.56]} & 0.18 [-0.6,0.99] & \textbf{-0.25 [-0.47,-0.02]} & \textbf{-0.56 [-0.91,-0.19]} & -0.7 [-1.84,0.56]\\
black-white ratio & 0.52 [-0.08,1.06] & 0.15 [-0.26,0.52] & -0.84 [-1.95,0.14] & \textbf{-0.33 [-0.51,-0.14]} & \textbf{-0.76 [-1.13,-0.4]} & -0.37 [-1.88,0.84]\\
total population & \textbf{0.14 [0.01,0.28]} & 0.03 [-0.06,0.11] & \textbf{0.32 [0.14,0.52]} & \textbf{-0.44 [-0.62,-0.26]} & 0.03 [-0.06,0.13] & 0.22 [0,0.45]\\
\addlinespace
college grads & 0.2 [-0.04,0.45] & \textbf{0.27 [0.14,0.41]} & -0.02 [-0.43,0.39] & \textbf{-0.12 [-0.22,-0.02]} & -0.12 [-0.32,0.07] & -0.2 [-0.92,0.48]\\
income & -0.27 [-0.56,0.04] & -0.17 [-0.34,0.01] & 0.03 [-0.48,0.55] & \textbf{-0.38 [-0.53,-0.22]} & \textbf{-0.38 [-0.64,-0.12]} & -0.15 [-1.03,0.71]\\
poverty & \textbf{-0.55 [-0.96,-0.14]} & -0.14 [-0.38,0.1] & 0.3 [-0.37,0.94] & 0 [-0.17,0.16] & \textbf{-0.44 [-0.76,-0.12]} & -1.02 [-2.21,0.12]\\
unemployment & \textbf{0.44 [0.18,0.71]} & \textbf{0.26 [0.11,0.42]} & -0.17 [-0.6,0.27] & \textbf{-0.12 [-0.21,-0.02]} & \textbf{-0.21 [-0.41,-0.02]} & -0.21 [-1,0.56]\\
implicit bias & \textbf{-0.42 [-0.71,-0.14]} & \textbf{-0.33 [-0.51,-0.16]} & -0.07 [-0.51,0.41] & 0.01 [-0.11,0.13] & 0.14 [-0.09,0.39] & 0.01 [-1,1.02]\\
\addlinespace
explicit bias & 0.11 [-0.15,0.37] & -0.04 [-0.19,0.12] & -0.06 [-0.45,0.35] & \textbf{-0.14 [-0.24,-0.05]} & \textbf{-0.21 [-0.41,-0.02]} & -0.51 [-1.39,0.4]\\
race: white & \textbf{-1.06 [-1.4,-0.69]} & \textbf{-1.02 [-1.2,-0.82]} & \textbf{-1.77 [-2.44,-1.13]} & \textbf{-0.64 [-0.7,-0.57]} & \textbf{-1.6 [-1.92,-1.26]} & \textbf{-2.09 [-3.6,-0.61]}\\
implicit bias*race: white & -0.1 [-0.32,0.12] & \textbf{0.18 [0.05,0.31]} & -0.02 [-0.39,0.36] & \textbf{-0.11 [-0.18,-0.05]} & -0.06 [-0.28,0.16] & -0.72 [-1.93,0.46]\\
explicit bias*race: white & -0.15 [-0.37,0.07] & 0.03 [-0.11,0.16] & -0.32 [-0.65,0.03] & -0.02 [-0.08,0.05] & -0.13 [-0.32,0.07] & -0.48 [-1.64,0.63]\\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\begin{landscape}\begin{table}

\caption{\label{tab:mod-post-table2}\label{tab:model-post-table2}Regression coefficient estimates for the 
             population-level (i.e. fixed) effects, along with 95\% 
             uncertainty intervals for each of the disciplinary metrics. 
             Model uses bias estimates as derived from poststratification 
             scheme.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{llllllll}
\toprule
 & \makecell[l]{Out-of-School\\Suspension} & \makecell[l]{Law Enf.\\Referral} & \makecell[l]{In-School\\Suspension} & \makecell[l]{School-Related\\Arrest} & \makecell[l]{Expulsion w/o\\ Ed. Services} & \makecell[l]{Expulsion w/Ed.\\Services} & \makecell[l]{Expulsion Under\\Zero-Tolerance}\\
\midrule
Intercept & \textbf{-2.69 [-2.72,-2.66]} & \textbf{-6.12 [-6.21,-6.03]} & \textbf{-2.63 [-2.67,-2.59]} & \textbf{-8.19 [-8.36,-8.03]} & \textbf{-8.2 [-8.37,-8.03]} & \textbf{-8.09 [-8.26,-7.92]} & \textbf{-9.66 [-9.92,-9.4]}\\
proportion black & \textbf{0.28 [0.21,0.36]} & \textbf{-0.28 [-0.48,-0.08]} & \textbf{0.31 [0.2,0.41]} & -0.26 [-0.59,0.08] & \textbf{0.47 [0.14,0.79]} & \textbf{-0.37 [-0.72,-0.03]} & 0.17 [-0.25,0.59]\\
proportion white & 0 [-0.06,0.06] & \textbf{-0.29 [-0.43,-0.15]} & 0.07 [0,0.15] & -0.1 [-0.34,0.16] & \textbf{0.28 [0.01,0.56]} & \textbf{-0.26 [-0.5,-0.03]} & 0.27 [-0.03,0.58]\\
black-white ratio & \textbf{-0.19 [-0.24,-0.13]} & -0.01 [-0.17,0.15] & \textbf{-0.36 [-0.44,-0.28]} & 0.08 [-0.17,0.33] & -0.19 [-0.42,0.04] & -0.17 [-0.48,0.11] & -0.24 [-0.62,0.1]\\
total population & 0 [-0.03,0.02] & 0.02 [-0.04,0.08] & \textbf{-0.04 [-0.07,0]} & 0.09 [0,0.19] & 0 [-0.08,0.09] & 0.06 [-0.03,0.15] & 0.09 [-0.01,0.2]\\
\addlinespace
college grads & \textbf{-0.04 [-0.07,-0.01]} & -0.01 [-0.09,0.07] & \textbf{-0.08 [-0.12,-0.03]} & \textbf{0.17 [0.03,0.31]} & \textbf{-0.23 [-0.37,-0.09]} & -0.05 [-0.19,0.08] & -0.14 [-0.31,0.04]\\
income & \textbf{-0.09 [-0.14,-0.05]} & -0.03 [-0.14,0.08] & \textbf{-0.11 [-0.17,-0.05]} & -0.01 [-0.18,0.17] & \textbf{-0.29 [-0.47,-0.11]} & -0.11 [-0.28,0.07] & -0.06 [-0.28,0.17]\\
poverty & 0.02 [-0.03,0.07] & \textbf{-0.38 [-0.51,-0.25]} & \textbf{0.16 [0.09,0.23]} & -0.2 [-0.43,0.02] & \textbf{-0.3 [-0.52,-0.09]} & -0.17 [-0.39,0.04] & 0.23 [-0.04,0.5]\\
unemployment & \textbf{0.21 [0.18,0.25]} & \textbf{0.18 [0.09,0.26]} & \textbf{0.06 [0.01,0.1]} & \textbf{0.26 [0.11,0.41]} & \textbf{0.41 [0.28,0.55]} & \textbf{0.27 [0.13,0.42]} & \textbf{0.26 [0.09,0.44]}\\
implicit bias & \textbf{0.08 [0.04,0.13]} & -0.08 [-0.19,0.01] & \textbf{0.13 [0.08,0.19]} & 0 [-0.16,0.16] & -0.03 [-0.19,0.13] & -0.03 [-0.2,0.13] & -0.15 [-0.36,0.06]\\
\addlinespace
explicit bias & \textbf{0.06 [0.02,0.09]} & \textbf{0.12 [0.03,0.21]} & \textbf{-0.16 [-0.2,-0.11]} & -0.08 [-0.23,0.06] & \textbf{0.14 [0,0.28]} & -0.06 [-0.2,0.09] & \textbf{0.25 [0.06,0.43]}\\
race: white & \textbf{-1.05 [-1.08,-1.02]} & \textbf{-0.83 [-0.89,-0.76]} & \textbf{-0.84 [-0.86,-0.81]} & \textbf{-1.05 [-1.18,-0.92]} & \textbf{-0.9 [-1.04,-0.76]} & \textbf{-0.87 [-1,-0.73]} & \textbf{-0.53 [-0.75,-0.3]}\\
implicit bias*race: white & \textbf{-0.04 [-0.07,-0.01]} & -0.04 [-0.1,0.02] & \textbf{-0.04 [-0.07,-0.02]} & 0 [-0.09,0.09] & -0.02 [-0.12,0.09] & 0.09 [-0.01,0.18] & 0.07 [-0.08,0.22]\\
explicit bias*race: white & 0 [-0.03,0.03] & -0.04 [-0.1,0.01] & 0 [-0.03,0.02] & -0.02 [-0.1,0.07] & 0.02 [-0.08,0.12] & 0.04 [-0.05,0.12] & 0.02 [-0.11,0.16]\\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\begin{landscape}\begin{table}

\caption{\label{tab:mod-teach-table1}\label{tab:model-teach-table1}Regression coefficient estimates for the 
             population-level (i.e. fixed) effects, along with 95\% 
             uncertainty intervals for each of the disciplinary metrics. 
             Model uses only data from Project Implicit respondents who 
             identified as teachers.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{lllllll}
\toprule
 & Seclusion & \makecell[l]{Physical\\Restraint} & \makecell[l]{Mechanical\\Restraint} & \makecell[l]{Corporal\\Punishment} & \makecell[l]{Preschool\\Suspension} & \makecell[l]{Preschool\\Expulsion}\\
\midrule
Intercept & \textbf{-10.12 [-10.61,-9.66]} & \textbf{-8.22 [-8.47,-8]} & \textbf{-12.87 [-13.84,-12.01]} & \textbf{-6.25 [-7.29,-5.37]} & \textbf{-8.6 [-9.05,-8.21]} & \textbf{-13.68 [-15.63,-12.13]}\\
propotion black & 0.51 [-0.83,1.84] & 0.06 [-0.57,0.69] & \textbf{1.98 [0.26,3.78]} & 0.02 [-2.7,3.11] & \textbf{1.2 [0.3,2.16]} & \textbf{2.88 [0.91,5.07]}\\
proportion white & \textbf{1.5 [0.46,2.55]} & \textbf{0.74 [0.23,1.26]} & 0.79 [-0.68,2.32] & -1.73 [-4.31,0.84] & 0.25 [-0.64,1.14] & 1.39 [-0.39,3.46]\\
black-white ratio & 0.18 [-1.36,1.73] & -0.01 [-0.65,0.59] & \textbf{-1.82 [-3.63,-0.1]} & -0.85 [-2.88,0.99] & -0.51 [-1.17,0.16] & -1.76 [-3.86,0.1]\\
total population & \textbf{0.35 [0,0.72]} & 0.02 [-0.19,0.24] & \textbf{0.66 [0.12,1.24]} & -0.97 [-2.25,0.07] & 0.16 [-0.15,0.48] & 0.43 [-0.07,0.95]\\
\addlinespace
college grads & 0.14 [-0.36,0.64] & \textbf{0.33 [0.09,0.58]} & -0.5 [-1.27,0.23] & -0.11 [-1.01,0.78] & -0.1 [-0.52,0.3] & -0.3 [-1.64,1.06]\\
income & -0.4 [-1.18,0.4] & \textbf{-0.42 [-0.84,-0.01]} & -0.13 [-1.4,1.1] & -0.38 [-1.72,0.95] & -0.21 [-0.89,0.47] & 1.4 [-0.26,3.05]\\
poverty & -0.33 [-1.32,0.62] & -0.01 [-0.51,0.46] & 0.41 [-0.96,1.77] & -0.16 [-2.01,1.48] & 0.31 [-0.41,1.04] & 0.39 [-1.63,2.29]\\
unemployment & -0.17 [-0.7,0.36] & -0.07 [-0.33,0.19] & -0.8 [-1.63,0.02] & -0.77 [-1.76,0.11] & \textbf{-0.55 [-0.97,-0.14]} & -0.04 [-1.05,1.04]\\
implicit bias & -0.38 [-0.97,0.17] & \textbf{-0.4 [-0.73,-0.06]} & -0.65 [-1.63,0.32] & 0.09 [-1.2,1.47] & 0.15 [-0.33,0.62] & -1.04 [-2.74,0.57]\\
\addlinespace
explicit bias & 0.12 [-0.42,0.67] & -0.2 [-0.51,0.11] & 0.81 [-0.09,1.78] & 0.51 [-0.71,1.75] & 0.05 [-0.45,0.55] & -0.07 [-1.39,1.23]\\
race: white & \textbf{-1.34 [-1.72,-0.96]} & \textbf{-1.13 [-1.32,-0.93]} & \textbf{-1.38 [-2.14,-0.58]} & -0.53 [-1.16,0.12] & \textbf{-1.63 [-2.05,-1.23]} & \textbf{-3.01 [-5.12,-1.03]}\\
implicit bias*race: white & 0.01 [-0.41,0.42] & 0.21 [-0.04,0.45] & 0.18 [-0.71,1.04] & -0.01 [-0.69,0.72] & -0.27 [-0.7,0.16] & 0.43 [-2.07,2.96]\\
explicit bias*race: white & 0.11 [-0.29,0.5] & 0.11 [-0.12,0.34] & -0.68 [-1.5,0.08] & -0.21 [-0.98,0.51] & -0.12 [-0.56,0.31] & -2.05 [-4.42,0.3]\\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\begin{landscape}\begin{table}

\caption{\label{tab:mod-teach-table2}\label{tab:model-teach-table2}Regression coefficient estimates for the 
             population-level (i.e. fixed) effects, along with 95\% 
             uncertainty intervals for each of the disciplinary metrics. 
             Model uses only data from Project Implicit respondents who 
             identified as teachers.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{llllllll}
\toprule
 & \makecell[l]{Out-of-School\\Suspension} & \makecell[l]{Law Enf.\\Referral} & \makecell[l]{In-School\\Suspension} & \makecell[l]{School-Related\\Arrest} & \makecell[l]{Expulsion w/o\\ Ed. Services} & \makecell[l]{Expulsion w/Ed.\\Services} & \makecell[l]{Expulsion Under\\Zero-Tolerance}\\
\midrule
Intercept & \textbf{-2.34 [-2.4,-2.29]} & \textbf{-5.57 [-5.7,-5.43]} & \textbf{-2.7 [-2.77,-2.62]} & \textbf{-7.31 [-7.55,-7.06]} & \textbf{-8.41 [-8.73,-8.1]} & \textbf{-7.61 [-7.87,-7.36]} & \textbf{-9.38 [-9.74,-9.05]}\\
propotion black & 0.12 [-0.01,0.27] & \textbf{-0.58 [-0.96,-0.2]} & \textbf{0.49 [0.28,0.69]} & -0.25 [-0.96,0.48] & 0.22 [-0.56,0.98] & -0.4 [-1.04,0.23] & -0.6 [-1.51,0.31]\\
proportion white & 0.07 [-0.05,0.18] & \textbf{-0.29 [-0.56,-0.02]} & \textbf{0.23 [0.05,0.4]} & 0.31 [-0.21,0.83] & \textbf{0.81 [0.12,1.52]} & 0.2 [-0.29,0.69] & 0.17 [-0.49,0.82]\\
black-white ratio & -0.08 [-0.19,0.03] & 0.07 [-0.31,0.45] & \textbf{-0.32 [-0.49,-0.16]} & 0.27 [-0.48,0.98] & 0.04 [-0.64,0.73] & -0.09 [-0.73,0.52] & -0.01 [-1.03,1]\\
total population & -0.05 [-0.11,0.01] & -0.02 [-0.16,0.12] & 0 [-0.09,0.09] & 0.09 [-0.15,0.33] & 0.12 [-0.17,0.4] & 0.06 [-0.21,0.33] & 0.07 [-0.19,0.34]\\
\addlinespace
college grads & \textbf{-0.15 [-0.22,-0.08]} & -0.03 [-0.19,0.13] & \textbf{-0.26 [-0.36,-0.16]} & 0 [-0.28,0.29] & -0.16 [-0.53,0.2] & -0.22 [-0.53,0.07] & -0.25 [-0.62,0.11]\\
income & \textbf{-0.23 [-0.33,-0.12]} & \textbf{-0.37 [-0.61,-0.13]} & \textbf{-0.2 [-0.36,-0.05]} & -0.05 [-0.51,0.4] & -0.53 [-1.13,0.06] & 0.11 [-0.33,0.57] & -0.11 [-0.66,0.42]\\
poverty & -0.06 [-0.18,0.05] & \textbf{-0.26 [-0.54,0]} & 0.05 [-0.1,0.21] & 0.08 [-0.46,0.59] & -0.12 [-0.78,0.54] & 0.28 [-0.18,0.73] & 0.4 [-0.2,0.97]\\
unemployment & \textbf{0.11 [0.04,0.19]} & 0.04 [-0.12,0.21] & -0.09 [-0.19,0.02] & 0.12 [-0.16,0.42] & 0.36 [-0.02,0.73] & 0.11 [-0.2,0.43] & 0.02 [-0.34,0.39]\\
implicit bias & -0.02 [-0.11,0.06] & 0.08 [-0.11,0.26] & -0.02 [-0.13,0.09] & -0.07 [-0.42,0.28] & -0.2 [-0.62,0.22] & \textbf{-0.56 [-0.93,-0.19]} & -0.15 [-0.59,0.29]\\
\addlinespace
explicit bias & 0.04 [-0.03,0.12] & 0.09 [-0.1,0.27] & 0 [-0.11,0.11] & -0.05 [-0.39,0.3] & 0.08 [-0.33,0.51] & -0.05 [-0.43,0.3] & 0.07 [-0.37,0.53]\\
race: white & \textbf{-1.46 [-1.5,-1.42]} & \textbf{-1.01 [-1.09,-0.93]} & \textbf{-1.14 [-1.18,-1.1]} & \textbf{-1.1 [-1.24,-0.94]} & \textbf{-1.07 [-1.26,-0.87]} & \textbf{-1.06 [-1.21,-0.91]} & \textbf{-0.72 [-0.96,-0.47]}\\
implicit bias*race: white & 0 [-0.06,0.05] & -0.07 [-0.18,0.03] & 0 [-0.05,0.05] & 0.04 [-0.13,0.2] & -0.16 [-0.36,0.04] & -0.02 [-0.19,0.15] & -0.1 [-0.39,0.18]\\
explicit bias*race: white & -0.01 [-0.06,0.05] & 0.03 [-0.07,0.14] & -0.01 [-0.06,0.05] & 0 [-0.17,0.17] & \textbf{-0.25 [-0.47,-0.04]} & -0.04 [-0.22,0.13] & -0.07 [-0.34,0.19]\\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-bates2015fitting}{}
Bates, Douglas, Martin Mächler, Ben Bolker, and Steve Walker. 2015.
``Fitting Linear Mixed-Effects Models Using lme4.'' \emph{Journal of
Statistical Software} 67 (1): 1--48.
doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}.

\hypertarget{ref-hehman2017disproportionate}{}
Hehman, Eric, Jessica K Flake, and Jimmy Calanchini. 2017.
``Disproportionate Use of Lethal Force in Policing Is Associated with
Regional Racial Biases of Residents.'' \emph{Social Psychological and
Personality Science}. SAGE Publications Sage CA: Los Angeles, CA,
1948550617711229.

\hypertarget{ref-leitner2016blacks}{}
Leitner, Jordan B, Eric Hehman, Ozlem Ayduk, and Rodolfo Mendoza-Denton.
2016. ``Blacks' Death Rate Due to Circulatory Diseases Is Positively
Related to Whites' Explicit Racial Bias: A Nationwide Investigation
Using Project Implicit.'' \emph{Psychological Science} 27 (10). SAGE
Publications Sage CA: Los Angeles, CA: 1299--1311.

\hypertarget{ref-miroshnikov2014parallel}{}
Miroshnikov, Alexey, and Erin Conlon. 2014. \emph{ParallelMCMCcombine:
Methods for Combining Independent Subset Markov Chain Monte Carlo (Mcmc)
Posterior Samples to Estimate a Posterior Density Given the Full Data
Set}. \url{https://CRAN.R-project.org/package=parallelMCMCcombine}.

\hypertarget{ref-rcitation}{}
R Core Team. 2016. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\hypertarget{ref-rstanarm2016}{}
Stan Development Team. 2016. ``Rstanarm: Bayesian Applied Regression
Modeling via Stan.'' \url{http://mc-stan.org/}.

\hypertarget{ref-wickham2009ggplot2}{}
Wickham, Hadley. 2009. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York. \url{http://ggplot2.org}.

\hypertarget{ref-wickham2017tidyr}{}
Wickham, Hadley, and Lionel Henry. 2017. \emph{Tidyr: Easily Tidy Data
with 'Spread()' and 'Gather()' Functions}.
\url{https://CRAN.R-project.org/package=tidyr}.

\hypertarget{ref-wickham2017dplyr}{}
Wickham, Hadley, Romain Francois, Lionel Henry, and Kirill Müller. 2017.
\emph{Dplyr: A Grammar of Data Manipulation}.
\url{https://CRAN.R-project.org/package=dplyr}.

\end{document}

### 2018-05-18 written_docs/draft.tex
.gitignore is now working

@@
 
+In comparison to White Americans, African Americans exhibit poorer
+educational outcomes across a range of metrics. One outcome of
+particular concern is the gap in disciplinary actions (Kinsler, 2011;
+Okonofua, Walton, \& Eberhardt, 2016). Research using administrative
+datasets and longitudinal samples clearly show that African American
+students are far more likely to be suspended or expelled (Aud,
+KewalRamani, \& Frohlich, 2011; Yeager, Purdie-Vaughns, Hooper, \&
+Cohen, 2017), and conditional on an office referral, are more likely to
+receive stiffer punishments (J. F. Gregory, 1995). Concerns about these
+disparities are exacerbated by their associations with long-term life
+outcomes, including prospects for employment (Pager, Western, \& Sugie,
+2009), mental and physical health (Pascoe \& Smart Richman, 2009), and
+involvement in the criminal justice system (Hirschfield, 2009). In this
+paper, we aim to establish that these disciplinary gaps are related to
+environmental levels of bias in predictable ways.
+
+It is perhaps not surprising that levels of bias would be associated
+with racial disciplinary gaps. This straightforward statement, however,
+is contextualized in a number of ways. First, the operationalization of
+bias should play a role in the association. As we describe below, levels
+of implicit and explicit biases should be differentially associated with
+disciplinary actions. This differential association should be
+determined, in part, by the type of disciplinary action under
+examination. For a given student committing a specific offense, the
+educational institution may opt for one of several disciplinary actions.
+These actions vary in severity, institutional costs, and the degree of
+formal administrative involvement.
+
+To illustrate this problem in the context of previous research, we now
+describe research on the associations between implicit bias and explicit
+bias and decision making. This sets the stage for understanding what
+types of decisions administrators and educators make when electing a
+specific disciplinary action. We then describe the types of disciplinary
+actions that are taken and the characteristics of these actions in terms
+of severity, institutional costs, and formal administrative involvement.
+Finally, in crossing this information, we concretely specify the set of
+associations expected when examining bias and disciplinary actions in
+American schools.
+
+\subsection{Bias and decision-making}\label{bias-and-decision-making}
+
+Previous research on biased attitudes and behavior has shown that
+although implicit and explicit attitudes are typically correlated
+(Hofmann, Gawronski, Gschwendner, Le, \& Schmitt, 2005; McConnell \&
+Leibold, 2001), they tend to behaviorally manifest slightly differently.
+As a default, a naiive expectation is that explicit attitudes should
+guide behavior, since if people tell us they like something, we should
+expect that to generally correspond to their behavior. However, under
+some circumstances, implicit biases are more strongly associated with
+behavioral outcomes. B. A. Nosek, Hawkins, and Frazier (2011) suggest
+that, in comparison to explicit attitudes, implicit attitudes are more
+strongly associated with behavior when the individual lacks motivated to
+explicitly respond differently than their implicit attitude would
+suggest, when they do not have the opportunity or ability to modify
+their initial response inclination, or when they do not have awareness
+of the implicit response, the behavior, or the link between them. In
+contrast, the more motivated, more able, or more aware the person is
+over their behavioral response, the more explicit biases should play a
+role. One key situational factor that determines the response is
+situational uncertainty. In uncertain situations, the respondent is not
+likely to have strong motivations to change their initial response,
+otherwise the situation would not be uncertain. Research supports this
+idea, as when the situation is characterized by greater uncertainty,
+implicit biases are known to be stronger determinants of behavior
+(Friese, Hofmann, \& Schmitt, 2009; Galdi, Arcuri, \& Gawronski, 2008;
+Gawronski, Geschke, \& Banse, 2003). In short, when taking (or not
+taking) an action can be attributed to multiple reasons due to
+situational ambiguity, then implicit biases should result in higher
+rates of discriminatory behavior.
+
+Accordingly, this research suggests that disciplinary actions that are
+implemented quickly, have relatively little administrative oversight,
+and require little in the way of insitutional resources should be
+characterized by strong associations with implicit bias. On the other
+hand, any disciplinary action that requires higher levels of oversight,
+resources, or lengthy time investments to complete should be more
+strongly associated with explicit bias.
+
+\subsection{Types of disciplinary
+actions}\label{types-of-disciplinary-actions}
+
+We report here on five different types of disciplinary actions:
+in-school suspensions, out-of-school suspensions, law enforcement
+referrals, school-related arrests, and expulsions. As described above,
+these actions vary in severity, required resources, and administrative
+oversight.
+
+The most extreme outcome, expulsion, is defined as when a student is
+prohibited from ever returning to the educational institution. The
+institution may or may not set up alternative educational services for
+the student. Because of the the severe nature of this punishment, it
+also requires considerable administrative oversight, and is often
+conferred for behaviors that are relatively unambiguous (e.g.~bringing a
+gun to school). We believe similar cases can be made for law enforcement
+referrals and school-related arrests insofar as these actions tend to be
+more severe, and require resources, or occasionally the involvement of
+outside agencies.
+
+Somewhat less severe than expulsions, but one of the most frequently
+utilized disciplinary actions, out-of-school suspensions occupy
+something of a middle ground. There is a substantial body of research
+documenting the negative associations between this particular
+disciplinary action and life outcome (Cuellar \& Markowitz, 2015;
+Matjasko, 2011), as well as their disproportionate use on stigmatized
+students (Haight, Gibson, Kayama, Marshall, \& Wilson, 2014; Kinsler,
+2011). Out-of-school suspensions are administered for a wide range of
+behaviors. Some of these behaviors are apt to be determined with a fair
+degree of subjectivity (e.g.~defiance) (A. Gregory \& Weinstein, 2008),
+while others may be more unambiguous (e.g.~fighting, drug use). In
+comparison to expulsions and arrests, this type of action demands less
+administrative oversight.
+
+In-school suspensions, wherein the student is removed from the classroom
+but remains under the direct supervision of school personnel, are also
+administered for a wide range of behaviors. However, in-school
+suspensions have been highlighted as a promising alternative to more
+exclusionary types of disciplinary actions (e.g.~out-of-school
+suspensions, expulsions) partly because students remain under the
+supervision of school personnel, which keeps them in a structured,
+educational setting (Anyon et al., 2014). Like out-of-school
+suspensions, this type of action requires relatively little
+administrative oversight. But unlike out-of-school suspension, removing
+a student from a classroom to a \emph{different} educational setting
+requires additional space and personnel, and so this represents an
+inherent investment in the student.
+
+\subsection{Concrete predictions of
+association}\label{concrete-predictions-of-association}
+
+Given these disciplinary outcomes and what is known about implicit and
+explicit attitudes, what should we expect in examining associations
+between them? First, it's clear that we should expect African American
+students to receive higher levels of disciplinary actions. Second, we
+should expect the disparity in disciplinary actions to generally be
+positively associated with explicit bias. Finally, we should expect
+implicit bias to only show associations in instances where there is
+little administrative overhead, and where the behaviors to-be-punished
+are often ambiguous. By this logic, out-of-school suspensions are the
+only metric that meet these criteria.
+
+\subsection{The present study}\label{the-present-study}
+
+The present study uses a national dataset of disciplinary outcomes to
+evaluate the degree to which implicit and explicit bias are associated
+with each of these outcomes. After an initial, registered analytic
+approach, we subsequently made several adjustments to strengthen our
+analyses and streamlined them with respect to previous research. In
+general, the data support the conclusion that implicit biases are
+specifically associated with out-of-school suspensions, while explicit
+biases tend to be associated with disciplinary actions of all types.
+
 \section{Methods}\label{methods}
@@
 
-We used three distinct data sources for the work described here. The
-academic disciplinary data is part of the Civil Rights Data Collection
-(CRDC) through the US Department of Education. The dataset we used comes
-from the 2013-2014 academic year and has data on \enquote{all public
-local and educational agencies and schools, including long-term secure
-juvenile justice facilities, charter schools, alternative schools, and
-schools serving students with disabilities.} In total, the CRDC data
-represents 95507 institutions enrolling approximately 50 million
-students, of which approximately 25 million are white and 7.8 million
-are Black\footnote{We note that there are a number of differences
-  between the analyses we registered and those presented in the main
-  text. Our general conclusions are largely the same for both sets of
-  analyses. We opted to report the modified analyses for reasons of
-  clarity and to remain congruent with previous research on the same
-  topics. The registered analyses can be found, in full, in the
-  appendix.}. Previous work using these data have identified a number of
-districts whose data are in error, and have excluded juvenile justice
-facilities, as these institutions constitute dramatically different
-educational environments, where the meaning of disciplinary actions may
-be quite different (Losen et al., 2015). After these exclusions are
-applied, the final sample used for modeling consists of 90002
-institutions, enrolling 32 million black or white students, of which 25
-million are white and 7 million are black.
+\subsubsection{Disciplinary actions}\label{disciplinary-actions}
 
-We obtained county-level demographic information for use as covariates
-and state-level demographic information for use in post-stratification
-from the US Census Bureau. For both county-level and state-level
-demographics, we use 5-year estimates for the period ending in 2014 from
-the American Community Survey, which surveys around 295k households per
-month.
+To assess rates of disciplinary action, we used data from the Civil
+Rights Data Collection (CRDC) conducted by the US Department of
+Education. The dataset we used comes from the 2013-2014 academic year
+and has data on \enquote{all public local and educational agencies and
+schools, including long-term secure juvenile justice facilities, charter
+schools, alternative schools, and schools serving students with
+disabilities.} In total, the CRDC data represents 95507 institutions
+enrolling approximately 50 million students, of which approximately 25
+million are white and 7.8 million are Black\footnote{We note that there
+  are a number of differences between the analyses we registered and
+  those presented in the main text. Our general conclusions are largely
+  the same for both sets of analyses. We opted to report the modified
+  analyses for reasons of clarity and to remain congruent with previous
+  research on the same topics. The registered analyses can be found, in
+  full, in the appendix.}. Previous work using these data have
+identified a number of districts whose data are in error, and have
+excluded juvenile justice facilities, as these institutions constitute
+dramatically different educational environments, where the meaning of
+disciplinary actions may be quite different (Losen et al., 2015). After
+these exclusions are applied, the final sample used for modeling
+consists of 90002 institutions, enrolling 32 million black or white
+students, of which 24.70 million are white and 7.30 million are black.
+From these data, we focus on the number of students by race (black and
+white) who received one of several types of disciplinary action. We
+report here rates of out-of-school suspension, in-school suspension,
+school-related arrests, law enforcement referrals, and total number of
+expulsions of any type.
+
+\subsubsection{Racial Bias}\label{racial-bias}
 
 We used measurements of implicit and explicit bias available from data
-collected through Project Implicit (Xu, Nosek, \& Greenwald, 2014). From
-this data, we used only respondents who had geographic information that
-would allow us to place them in a United States county, identified as
-White, and visited the site before 2015. This consisted of approximately
-1.1 million total respondents from 3091 counties.
+collected through Project Implicit (Xu, Nosek, \& Greenwald, 2014). For
+a full description of the implicit and explicit bias measures available
+in these data, refer to Xu et al. (2014) and Leitner, Hehman, Ayduk, and
+Mendoza-Denton (2016). We note only that we used the IAT D-score as a
+measure of implicit bias, and the difference between reported warmth
+towards whites and warmth towards black (both measured from 0=very cold,
+10=very warm) as a measure of explicit bias. Additionally, we used only
+respondents who had geographic information that would allow us to place
+them in a United States county, identified as White, and visited the
+site before 2015. This consisted of approximately 1.1 million total
+respondents from 3091 counties.
+
+\subsubsection{Covariates}\label{covariates}
+
+Each county-level variable used as a covariates in the final model and
+the corresponding state-level variable used as a predictor in the
+post-stratification scheme were taken from the from the same source.
+Population size and proportions, socioeconomic indicators, and mobility
+were all taken from the American Community Survey (ACS) 5-year estimates
+for the time period ending in 2014. Urban-rural indicators were taken
+from the 2010 US Census, and crime rates were taken from the FBI Uniform
+Crime Reporting program, as made available through the National Archive
+of Criminal Justice Data for each year from 2010-2014. Each of these
+variables is described below.
+
+\paragraph{Population size and
+proportions}\label{population-size-and-proportions}
+
+We obtained the total population, the proportion of the population that
+is white, the proportion of the population that is black, and the ratio
+of black-to-white people in the population from ACS table B02001.
+
+\paragraph{Socioeconomic indicators}\label{socioeconomic-indicators}
+
+We obtained estimates for the percentage of population with a Bachelor's
+degree or higher, the percentage of the population aged 16 or over in
+the labor force that is unemployed, the median household income, and the
+percentage of families and people whose income in the last year was
+below the poverty line from the ACS table DP03.
+
+\paragraph{Urban-rural indicator}\label{urban-rural-indicator}
+
+We obtained estimates of housing density per square mile of land area
+from Census table GCT-PH1.
+
+\paragraph{Mobility}\label{mobility}
+
+We obtained estimates of population mobility by summing the percentage
+of African Americans who moved from a different county, state, or
+country into the county of interest (county-level covariate) or who
+moved from a different state or country into the state of interest
+(state-level covariate). We took these metrics from the ACS table S0701.
+
+\paragraph{Crime}\label{crime}
+
+We computed estimates of the number of violent crimes per person by
+taking the number of crimes reported divided by the population size for
+each year, and averaging the resulting proportions across the 5 years of
+data.
+
+\section{Results}\label{results}
+
+\subsection{Data analysis}\label{data-analysis}
+
+Because Project Implicit is a nonrandom sample, we used multilevel
+regression and post-stratification to obtain accurate geographical
+population-based estimates of implicit and explicit bias. Multilevel
+regression and post-stratification corrects for biased sampling and
+regularizes extreme observations with little data to support them
+(e.g.~a county with only a handful of respondents with especially high
+or low scores) (Gelman \& Little, 1997; Park, Gelman, \& Bafumi, 2004).
+Following past work (Leitner et al., 2016), we identified age as one
+dimension along which IAT respondents differed from the general
+population in ways that could bias our conclusions (Gonsalkorale,
+Sherman, \& Klauer, 2009). Our post-stratification weighting scheme is
+as follows: We first grouped respondents into five age group categories
+(15-24, 25-34, 35-54, 55-75, and 75+). We next fit multilevel models
+estimating bias (implicit and explicit biases seperately) as a function
+of our state-level covariates (the \enquote{fixed} effects: proportion
+of population that is white; proportion of the population that is black,
+the ratio of black to white people in the population, the percentage of
+college graduates, the unemployment rate, the median household income,
+the poverty rate, housing density, mobility rate, \& crime rate; each of
+these is described in more detail below), and allowed the estimates to
+vary by age bin, county, and state (the \enquote{random} effects). Next,
+we determined the population of whites in each county in these age
+groups using the American Community Survey's 5-year estimates ending in
+2014. Finally, we used our estimated models to predict the expected
+response for each age bin, in each county. Our final county-level
+estimates are the average of the values predicted for the 5 age bins,
+weighted by the population size of that bin in that county. As a result
+of this procedure, we can be confident that our county-level estimates
+should more closely approximate what our estimates would look like if
+the Project Implicit data were truly representative along the age
+dimension in all counties.
+
+We analyzed these data using a series of bayesian multilevel logistic
+regressions. We modeled the probability that a student would be expelled
+as a function of a set of effects that are constant across observations
+(i.e.~fixed effects: race (dummy coded), implicit bias, explicit bias,
+an interaction term between race and implicit bias, an interaction term
+between race and explicit bias, and all covariates described above) and
+a set of effects that vary across counties (i.e.~random effects: overall
+intercept \& race). We fit separate models for each of the outcomes. All
+numerical predictors were standardized at the appropriate level (county,
+state) before model estimation (for post-stratification as well as for
+final inference) to help with estimation efficiency and
+interpretability. We also set priors for the intercept and coefficients
+in the bayesian model to be weakly informative normal distributions
+centered on zero with a standard deviation of five. This corresponds to
+a prior belief of all parameters taking values between -15 and 15.
+Values outside this range are extremely unlikely given that all
+variables were standardized prior to estimation. All other parameters
+were left to default values.
+
+Because of the computational demands of fitting such a high-dimensional
+model to such a large dataset (the full model for each metric would
+consist of over 6k parameters to approximately 170k observations), we
+used a consensus monte carlo algorithm to obtain approximate posterior
+distributions for the parameters of interest (Scott et al., 2016). The
+approximate posteriors derived from this algorithm have been shown to be
+nearly indistinguishable from the true posterior, a result we verified
+using a small subset of our own data.
+
+\subsection{Project Implicit
+Estimates}\label{project-implicit-estimates}
+
+We first report the results of estimating the implicit and explicit
+biases in from Project Implicit data. When examining the unstandardized
+county-level estimates adjusted with poststratification, we see that
+there is a pro-white bias in both implicit (\(mean\) = 0.40, \(sd\) =
+0.02) and explicit measures (\(mean\) = 0.82, \(sd\) = 0.17, where on
+both scales 0 = no bias, and positive numbers indicate a pro-white
+bias).
+
+\subsection{Disciplinary action
+frequency}\label{disciplinary-action-frequency}
+
+\begin{table}[tbp]
+\begin{center}
+\begin{threeparttable}
+\caption{\label{tab:disc-count}Percentage of students of each race receiving each type of disciplinary action}
+\begin{tabular}{lll}
+\toprule
+metric & \multicolumn{1}{c}{black} & \multicolumn{1}{c}{white}\\
+\midrule
+expulsions & 0.5\% & 0.23\%\\
+school arrests & 0.3\% & 0.09\%\\
+in-school suspension & 11.47\% & 4.29\%\\
+law enforcement referral & 0.76\% & 0.33\%\\
+out-of-school suspension & 26.19\% & 7.2\%\\
+\bottomrule
+\end{tabular}
+\end{threeparttable}
+\end{center}
+\end{table}
+
+Table \ref{tab:disc-count} shows the percentage of students of each race
+who were reported having received each of the actions under
+consideration. Our statistical models (described below) show that these
+differences are extremely unlikely to be due to chance.
+
+\subsection{Associations across
+counties}\label{associations-across-counties}
+
+Figure \ref{fig:overall-associations} shows the estimate of primary
+interest for each of the models. The estimates displayed are the
+coefficients for the interaction between race and each of the two bias
+measurements. Given that African Americans are the baseline group,
+negative values for this coefficient indicate that as one moves into
+counties with higher levels of bias, the gap between the probability of
+a black student being disciplined and the probability of a white student
+being disciplined grows. Table \ref{tab:reg-coefs} displays the
+estimated coefficients and 95\% uncertainty intervals for each of the
+fixed effects for each model.
+
+\begin{figure}
+\centering
+\includegraphics{draft_files/figure-latex/overall-associations-1.pdf}
+\caption{\label{fig:overall-associations}Association between each metric and
+county-level estimates of explicit and implicit bias. Negative values
+indicate that the rate of increase (or decrease) for blacks is faster
+(or slower) than for whites. Point is the mean of the posterior and
+error bars represent 95\% bayesian uncertainty intervals.}
+\end{figure}
+
+Several patterns are apparent from this figure. First, in all cases but
+school-related arrests, county-level explicit biases are reliably
+associated with disciplinary disparities, such that counties with more
+bias are expected to have a larger racial gap. The effect with the
+largest magnitude of these is out-of-school suspensions (\(M_{slope}\) =
+-0.12, {[}-0.18, -0.06{]}), followed by expulsions (\(M_{slope}\) =
+-0.11, {[}-0.21, -0.01{]}), law enforcement referrals (\(M_{slope}\) =
+-0.10, {[}-0.16, -0.03{]}), and in-school suspensions (\(M_{slope}\) =
+-0.08, {[}-0.12, -0.05{]}). The model indicates that the evidence for
+school-related arrests is inconclusive (\(M_{slope}\) = -0.03, {[}-0.13,
+0.08{]}).
+
+Second, the effect with the largest magnitude of all is that between
+implicit biases and out-of-school suspension (\(M_{slope}\) = -0.22,
+{[}-0.28, -0.16{]}). All other estimated parameters are inconclusive.
+
+\begin{table}[tbp]
+\begin{center}
+\begin{threeparttable}
+\caption{\label{tab:reg-coefs}Regression coefficient estimates for the population-level (i.e. fixed) effects, along with 95\% uncertainty intervals for each of the disciplinary metrics.}
+\small{
+\begin{tabular}{llllll}
+\toprule
+ & \multicolumn{1}{c}{Out-of-School Suspension} & \multicolumn{1}{c}{Law Enf.
+Referral} & \multicolumn{1}{c}{In-School Suspension} & \multicolumn{1}{c}{School-Related Arrest} & \multicolumn{1}{c}{Expulsion}\\
+\midrule
+Intercept & -2.02, [-2.07,-1.97] & -5.75, [-5.83,-5.67] & -2.3, [-2.33,-2.26] & -7.82, [-7.98,-7.66] & -6.5, [-6.6,-6.4]\\
+total population & 0.01, [-0.02,0.03] & 0.03, [-0.03,0.09] & -0.04, [-0.07,-0.01] & 0.13, [0.04,0.22] & 0.03, [-0.04,0.11]\\
+propotion black & 0.4, [0.31,0.49] & -0.25, [-0.46,-0.03] & 0.19, [0.09,0.3] & -0.14, [-0.5,0.24] & -0.13, [-0.38,0.12]\\
+proportion white & 0.03, [-0.03,0.09] & -0.2, [-0.36,-0.04] & 0, [-0.08,0.08] & 0.02, [-0.26,0.3] & -0.2, [-0.39,-0.01]\\
+black-white ratio & -0.26, [-0.32,-0.19] & 0.07, [-0.1,0.24] & -0.34, [-0.42,-0.27] & -0.09, [-0.38,0.18] & -0.18, [-0.39,0.01]\\
+college grads & -0.03, [-0.06,0] & 0.03, [-0.05,0.11] & -0.09, [-0.13,-0.05] & 0.17, [0.04,0.31] & -0.12, [-0.22,-0.02]\\
+income & -0.13, [-0.18,-0.08] & -0.08, [-0.2,0.03] & -0.13, [-0.18,-0.07] & 0.03, [-0.15,0.21] & -0.28, [-0.42,-0.14]\\
+poverty & -0.06, [-0.13,0.01] & -0.63, [-0.81,-0.46] & 0.1, [0.01,0.18] & -0.29, [-0.59,0] & -0.4, [-0.61,-0.19]\\
+unemployment & 0.23, [0.19,0.27] & 0.2, [0.1,0.29] & 0.03, [-0.01,0.07] & 0.3, [0.15,0.46] & 0.29, [0.19,0.39]\\
+crime & 0.04, [0.01,0.07] & 0.07, [0,0.14] & 0.06, [0.03,0.09] & 0.11, [-0.01,0.22] & 0.09, [0.01,0.17]\\
+housing density & -0.01, [-0.03,0.02] & -0.07, [-0.14,0.01] & 0, [-0.04,0.03] & -0.06, [-0.15,0.04] & 0, [-0.07,0.08]\\
+mobility & 0.02, [-0.01,0.05] & 0.07, [0,0.14] & 0.01, [-0.02,0.04] & 0.05, [-0.07,0.17] & 0.01, [-0.07,0.1]\\
+race: white & -0.65, [-0.69,-0.61] & -0.87, [-0.93,-0.81] & -0.85, [-0.87,-0.82] & -1.1, [-1.22,-0.99] & -0.93, [-1.01,-0.85]\\
+implicit bias & 0.36, [0.29,0.43] & -0.1, [-0.23,0.02] & 0.21, [0.15,0.28] & -0.25, [-0.45,-0.05] & 0.03, [-0.12,0.18]\\
+explicit bias & -0.05, [-0.11,0.02] & -0.01, [-0.13,0.1] & 0.1, [0.04,0.16] & 0.37, [0.19,0.55] & 0.04, [-0.09,0.18]\\
+implicit bias*race: white & -0.22, [-0.28,-0.16] & 0.05, [-0.02,0.12] & 0.01, [-0.02,0.05] & -0.01, [-0.13,0.09] & 0.07, [-0.03,0.17]\\
+explicit bias*race: white & -0.12, [-0.18,-0.06] & -0.1, [-0.16,-0.03] & -0.08, [-0.12,-0.05] & -0.03, [-0.13,0.08] & -0.11, [-0.2,-0.02]\\
+\bottomrule
+\end{tabular}
+}
+\end{threeparttable}
+\end{center}
+\end{table}
+
+The findings for other disciplinary metrics suggest a differential
+effect of explicit and implicit bias. For example, there is a reliable
+association between in-school suspensions and explicit bias. The
+difference in the slope of the association between explicit bias and the
+log of the odds for in-school suspensions between white and black
+students is estimated to be -0.12, with 95\% of the posterior
+distribution between -0.18 and -0.06 and a proportion \textgreater{}.99
+of the posterior distribution consistent with a negative effect. Similar
+effects can be seen for explicit bias and law enforcement referrals
+(-0.10, {[}-0.16, -0.03{]}, \(p_{neg}\) \textgreater{}.99), and explicit
+bias and expulsions (-0.11, {[}-0.21, -0.01{]}, \(p_{neg}\) = .99). In
+contrast, zero is included as a plausible value for the associations
+between implicit bias and in-school suspensions (0.01, {[}-0.02,
+0.05{]}, \(p_{neg}\) = .27), law enforcement referrals (0.05, {[}-0.02,
+0.12{]}, \(p_{neg}\) = .08), and expulsions (0.07, {[}-0.03, 0.17{]},
+\(p_{neg}\) = .07).
+
+\begin{figure}
+\centering
+\includegraphics{draft_files/figure-latex/detail-risks-1.pdf}
+\caption{\label{fig:detail-risks}Association between bias and relative risk
+ratio for black students to white students. Line is the mean of the
+posterior. Bands indicate 95\% uncertainty intervals. Points represent
+counties, whose sizes are scaled to the number of schools in that
+county. Note that the y axis is cut at 8 for legibility despite data
+extending beyond.}
+\end{figure}
+
+To better illustrate the nature of these relationships, figure
+\ref{fig:detail-risks} shows the estimated relative risk ratios for each
+type of bias measurement and each disciplinary metric. Examining the
+risk ratio for out-of-school suspensions, in a hypothetical county with
+0 standardized bias on both implicit and explicit, the model predicts
+that for every white student expelled, we should expect 1.81 {[}1.74,
+1.88{]} black students to be expelled. If we move to a county one
+standard deviation above the mean of explicit bias, the ratio of black
+to white students expelled increases to 2.17 {[}2.05, 2.30{]}, while the
+same movement for implicit bias increase the ratio to 2.03 {[}1.90,
+2.17{]}.`
+
+\begin{figure}
+\centering
+\includegraphics{draft_files/figure-latex/detail-rates-1.pdf}
+\caption{\label{fig:detail-rates}Association between bias and probability of
+receiving disciplinary action for black and white students. Lines are
+the means of the posterior. Bands indicate 95\% uncertainty intervals.}
+\end{figure}
+
+Figure \ref{fig:detail-rates} unpacks these risk ratios by detailing the
+probabilities of receiving disciplinary action as a function of race,
+bias, and type of action. In a hypotheticl county with 0 standardized
+bias on both implicit and explicit, the model predicts that 11.70\%
+{[}11.20, 12.20{]} of black students will receive an out-of-school
+suspension . The corresponding rate for white students is just over half
+that for black students, with about 6.50\% {[}6.30, 6.60{]} expected to
+be suspended. Holding all other variables constant, moving to a
+hypothetical county one standard deviation above the mean of explicit
+bias has the effect of slightly decreasing the estimated percentage of
+black students expected to be suspended to 11.20\% {[}10.40, 12.10{]},
+while the percentage of white students expected to be suspended would
+decrease at a faster rate to 5.50\% {[}5.30, 5.80{]}. The same movement
+for implicit bias would dramatically increase the expected suspensions
+for black students (16\% {[}14.90, 17.10), and increase the expected
+suspensions for white students a much smaller amount to (7.40\% {[}7,
+7.70).
+
+\section{Discussion}\label{discussion}
+
+\newpage
+
+\section{References}\label{references}
+
+\newpage
+
+\section{Appendix}\label{appendix}
+
+\subsection{Session details}\label{session-details}
+
+Data analysis was done in R (R Core Team, 2016) version 3.3.2 running
+under OS X 10.11.6. Post-stratification was done with lme4, version
+1.1.14 (Bates, Mächler, Bolker, \& Walker, 2015). Final model fitting
+was done on the university cluster running Springdale Linux, release 6.9
+using rstanarm, version 2.17.2 (Stan Development Team, 2016). We used
+the implementation of the consensus monte carlo algorithm found in
+parallelMCMCcombine, version 1.0 (Miroshnikov \& Conlon, 2014). Figures
+were made with ggplot2, version 2.2.1 (Wickham, 2009), with data
+manipulation done using dplyr version 0.7.2 (Wickham, Francois, Henry,
+\& Müller, 2017) and tidyr, version 0.7.1 (Wickham \& Henry, 2017). A
+full report of session information can be found on the OSF page
+(\ldots{}.)
+
+\subsection{Preregistered analysis}\label{preregistered-analysis}
+
+Here, we describe, in detail, the differences between the registered
+analysis plan and that presented in the main text, along with our
+reasons for making the alterations.
+
+\subsubsection{Data exclusions}\label{data-exclusions}
+
+In our preanalysis plan, we specified our analyses to focus on 13
+actions - corporal punishment, in-school suspension, out-of-school
+suspension, expulsion with educational services, expulsion without
+educational services, expulsion under zero-tolerance policies, referral
+to law enforcement, school-related arrests, mechanical restraint,
+physical restraint, seclusion, preschool suspension, and preschool
+expulsion. However, upon further study, we discovered reasons we thought
+justified excluding a number of these outcomes.
+
+First, seclusion, physical restraint, and mechanical restraint are not
+disciplinary actions, but are rather used as means to restrain students
+who are at risk of harming themselves or others.
+
+Next, we discovered that the administration of corporal punishment was
+extremely irregular across counties, with almost all cases ocurring in
+the South. Specifically, Alabama, Arkansas, Florida, Georgia, Louisiana,
+Missouri, Mississippi, Oklahoma, Tennessee, and Texas accounted for 98\%
+of all educational institutions that administered corporal punishment at
+least once, despite containing only 26.90\% of the educational
+institutions in the full dataset. It is not clear whether this uneven
+distribution is due to legal prohibitions, cultural differences, or some
+combination of the two. Because this makes interpretation of the
+parameters uncertain, this analysis is also placed in the appendix.
+
+We also found that the number of preschool students who are expelled or
+suspended is vanishingly small (131 total expulsions and 6751 total
+suspensions out of over 1.4 million enrolled preschool students), making
+reliably estimating any association across counties exceedingly
+unlikely. We additionally discovered that counts of one expulsion
+category (expulsion under zero-tolerance policies) overlapped with
+counts in other categories, and so excluded this category from the main
+text. We also decided to combine the remaining two expulsion categories
+to yield one overall count of the number of students expelled. We did
+this to remain consistent with previous research, and to obtain
+district-level rates of expulsion that were slightly higher, and
+therefore more amenable to exploring variation across counties.
+
+When preparing the preregistration, we also had not known about the
+issues with juvenile justice facilities, or with the school districts
+with reporting errors, and so excluded these schools from analyses in
+the main text.
+
+\subsubsection{Changes in measurement \& estimation of
+bias}\label{changes-in-measurement-estimation-of-bias}
+
+We preregistered our explicit bias as a simple feeling thermometer
+towards balcks (i.e. \emph{how warm or cold do you feel towards Blacks?
+0=very cold, 10=very warm}). However, the majority of past research (see
+for example Leitner et al. (2016) and Hehman, Flake, and Calanchini
+(2017)) has used the difference in reported warmth towards whites and
+blacks, and so in the main text, we report models using this metric of
+explicit bias.
+
+Our main text presents analyses with poststratified estimates of bias.
+Our registered analysis also described analyses with raw, county-based
+means. Unlike previous papers that used poststratification, we see
+substantial differences in results between the two estimation methods.
+Because post-stratification is known to yield better estimates, we
+suspect that the difference in results is largely driven by extreme
+county-level observations with little data to support these extreme
+estimates. Nevertheless, in the interest of completion, we present the
+model using raw county-level means as well.
+
+We also registered an exploratory analysis using the responses of people
+who visited Project Implicit and identified themselves as teachers.
+Unfortunately, without population-level data on the demographic
+characteristics of teachers, we are not able to post-stratify these
+estimates. Additionally, there were relatively few teachers in the
+Project Implicit from which to derive estimates. Ultimately, we have
+little confidence that this analysis tells us anything about whether the
+biases of teachers specifically contribute to disparities in
+disciplinary actions, and so this analysis is also limited to the
+appendix.
+
+\subsubsection{Additional covariates}\label{additional-covariates}
+
+We included three covariates that were not specified in the registered
+analysis. Specifically, the mobility, housing density, and crime rate
+covariates. These covariates were included in order to cover more of the
+covariates that were included in previous papers (Leitner et al., 2016).
+
+\subsection{Registered analysis
+results}\label{registered-analysis-results}
+
+Here, we present the results of the preregistered analyses exactly.
+Though we described a number of changes above, it is not practical to
+run all possible combination of changes, as each set of models takes
+four to five days to just to estimate, and there are at least several
+dozen possible combinations of decisions to test, depending on how
+fine-grained one wishes to be with the decisions. Accordingly, we limit
+the presentation of results below to just the three models that we
+registered.
+
+\subsubsection{County-level estimates}\label{county-level-estimates}
+
+\begin{figure}
+\centering
+\includegraphics{draft_files/figure-latex/raw-county-estimates-1.pdf}
+\caption{\label{fig:raw-county-estimates}Difference in raw county means and
+post-stratifed estimates of implicit bias as a function of the number of
+observations in each county. The X axis is on the log scale.}
+\end{figure}
+
+Examining the models fit with simple county means, we find that the
+county-level estimates vary more than with the post-stratified
+estimates. This pattern is expected, as one desireable feature of
+post-stratification is that it regularizes extreme observations without
+a large amount of data to support them. This phenomena is illustrated in
+figure \ref{fig:raw-county-estimates}, which shows the distribution of
+differences between the raw county means and the post-stratified
+estimates of implicit bias as a function of the number of observations
+in each county. More specifically, while the mean of the county-level
+means does not meaningfully differ from the post-stratified estimates,
+the standard deviation is much larger (\(M_{implicit}\) = 0.40,
+\(SD_{implicit}\) = 0.11; \(M_{explicit}\) = 6.25, \(SD_{explicit}\) =
+0.60)\footnote{This explicit estimate is based on the simple warmth
+  towards blacks question, and not the difference between warmth towards
+  whites and warmth twoards blacks.}
+
+The county-level implicit bias estimates for the poststratified model
+are as reported in the main text. The estimates for explicit bias are
+slightly different, as they are not computed based on the difference in
+reporting warmth toward whites and warmth toward blacks. Specifically,
+the average amount of bias at the county-level is 6.26 (SD = 0.13).
 
@@
 limited to just 287 counties. Additionally, because we do not know of
-any state-level demographic estimates for teachers, we are unable to
-perform post-stratification for these data.
+any state-level demographic estimates for teachers, we were unable to
+perform post-stratification for these data. Nonetheless, across these
+limited counties, the overall estimate of implicit and explicit bias are
+similar to those for the whole dataset (\(M_{implicit}\) = 0.38,
+\(SD_{implicit}\) = 0.06; \(M_{explicit}\) = -6.70, \(SD_{explicit}\) =
+0.27)
 
-\subsection{Measures}\label{measures}
+\subsubsection{Model estimates}\label{model-estimates}
 
-The primary outcome in this analysis is a count of the number of
-students by race (black and white) who received one of several types of
-disciplinary action. We report here rates of out-of-school suspension,
-in-school suspension, school-related arrests, law enforcement referrals,
-and total number of expulsions of any type.
+Table \ref{tab:mod-raw} presents results from the model fit to the raw,
+county-level means. The table presents parameter estimates
+(operationalized as the mean of the posterior) and 95\% uncertainty
+intervals for each of the population-level effects for each of the
+outcomes. Table \ref{tab:mod-post} and table \ref{tab:mod-teach} present
+the corresponding results for the post-stratified and teacher models,
+respectively.
 
-For both the post-stratification procedure (described below) and the
-actual statistical models used for inference, we used the same set of
-covariates. These covariates were at the state-level for
-post-stratification and were at the county-level for the statistical
-models used for final estimation and inference. Specifically, our
-population based covariates were the total population count, the
-proportion of the population that is black, the proportion of the
-population that is white and the ratio of black-to-white persons. We
-also used socioeconomic covariates. We used the percent of individuals
-aged 16 or over who were in the labor force but unemployed, the median
-household income, and the percentage of all families whose income is
-below the poverty line.
+\begin{lltable}
+\small{
+\begin{longtable}{llllllllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:mod-raw}Regression coefficient estimates for the population-level (i.e.
+fixed) effects, along with 95\% uncertainty intervals for each of the disciplinary metrics.
+Model uses raw, county-level means as estimates of implicit and explicit bias.}\\
+\toprule
+ & \multicolumn{1}{c}{Seclusion} & \multicolumn{1}{c}{Preschool Suspension} & \multicolumn{1}{c}{Preschool Expulsion} & \multicolumn{1}{c}{Physical Restraint} & \multicolumn{1}{c}{Out-of-School Suspension} & \multicolumn{1}{c}{Mechanical Restraint} & \multicolumn{1}{c}{Law Enf.
+Referral} & \multicolumn{1}{c}{In-School Suspension} & \multicolumn{1}{c}{School-Related Arrest} & \multicolumn{1}{c}{Expulsion without Educational Services} & \multicolumn{1}{c}{Expulsion with Educational Services} & \multicolumn{1}{c}{Expulsion Under Zero-Tolerance} & \multicolumn{1}{c}{Corporal Punishment}\\
+\midrule
+Intercept & -11.19, [-11.64,-10.77] & -8.84, [-9.15,-8.56] & -16.07, [-17.96,-14.35] & -8.78, [-8.99,-8.58] & -2.68, [-2.71,-2.65] & -14.1, [-14.96,-13.31] & -6.12, [-6.21,-6.03] & -2.63, [-2.67,-2.59] & -8.2, [-8.37,-8.04] & -8.22, [-8.39,-8.07] & -8.11, [-8.28,-7.94] & -9.74, [-10.01,-9.49] & -3.21, [-3.31,-3.13]\\
+total population & 0.16, [0.01,0.31] & 0.05, [-0.07,0.17] & 0.49, [0.24,0.76] & 0.02, [-0.07,0.12] & 0.01, [-0.02,0.03] & 0.41, [0.2,0.65] & 0.01, [-0.06,0.07] & -0.03, [-0.07,0.01] & 0.09, [0,0.19] & 0.01, [-0.08,0.1] & 0.06, [-0.03,0.15] & 0.09, [-0.01,0.2] & -0.48, [-0.67,-0.3]\\
+propotion black & -1.51, [-2.18,-0.85] & 1.02, [0.59,1.46] & 0.99, [-0.47,2.53] & -0.67, [-1.04,-0.3] & 0.34, [0.27,0.42] & 0.8, [-0.15,1.75] & -0.45, [-0.64,-0.25] & 0.61, [0.51,0.7] & -0.19, [-0.5,0.12] & 0.32, [0.04,0.61] & -0.29, [-0.62,0.02] & 0.02, [-0.36,0.42] & 0.11, [-0.14,0.37]\\
+proportion white & 0.26, [-0.23,0.74] & -0.44, [-0.8,-0.07] & -0.34, [-1.53,0.91] & 0.19, [-0.08,0.45] & 0.02, [-0.03,0.08] & 0.26, [-0.48,1.06] & -0.34, [-0.48,-0.2] & 0.13, [0.05,0.2] & -0.06, [-0.31,0.19] & 0.23, [-0.03,0.49] & -0.25, [-0.47,-0.02] & 0.21, [-0.08,0.51] & -0.3, [-0.51,-0.08]\\
+black-white ratio & 0.81, [0.17,1.45] & -0.78, [-1.14,-0.44] & -0.34, [-1.63,0.82] & 0.26, [-0.08,0.6] & -0.21, [-0.27,-0.15] & -0.59, [-1.5,0.27] & 0.06, [-0.1,0.22] & -0.53, [-0.62,-0.45] & 0.09, [-0.18,0.36] & -0.13, [-0.36,0.09] & -0.22, [-0.53,0.07] & -0.23, [-0.62,0.13] & -0.34, [-0.52,-0.15]\\
+college grads & 0.25, [0.01,0.48] & -0.21, [-0.4,-0.01] & -0.24, [-0.98,0.45] & 0.3, [0.17,0.43] & -0.04, [-0.07,-0.01] & -0.04, [-0.46,0.37] & 0.03, [-0.06,0.11] & -0.12, [-0.16,-0.07] & 0.15, [0.02,0.29] & -0.2, [-0.34,-0.06] & -0.06, [-0.2,0.07] & -0.08, [-0.25,0.09] & -0.15, [-0.25,-0.05]\\
+income & -0.32, [-0.64,-0.01] & -0.3, [-0.57,-0.03] & -0.35, [-1.27,0.55] & -0.18, [-0.36,0] & -0.09, [-0.13,-0.05] & 0, [-0.5,0.53] & -0.05, [-0.16,0.06] & -0.08, [-0.14,-0.02] & 0, [-0.18,0.18] & -0.31, [-0.49,-0.14] & -0.09, [-0.26,0.09] & -0.08, [-0.3,0.14] & -0.33, [-0.47,-0.18]\\
+poverty & -0.64, [-1.05,-0.22] & -0.32, [-0.65,-0.01] & -1.21, [-2.46,0.02] & -0.14, [-0.38,0.09] & 0.02, [-0.03,0.08] & 0.3, [-0.32,0.95] & -0.4, [-0.54,-0.27] & 0.2, [0.13,0.27] & -0.22, [-0.44,0] & -0.35, [-0.56,-0.14] & -0.15, [-0.37,0.06] & 0.2, [-0.07,0.46] & 0.01, [-0.15,0.17]\\
+unemployment & 0.4, [0.14,0.68] & -0.29, [-0.49,-0.1] & -0.3, [-1.08,0.48] & 0.23, [0.08,0.38] & 0.22, [0.18,0.25] & -0.21, [-0.64,0.24] & 0.19, [0.1,0.27] & 0.03, [-0.01,0.08] & 0.25, [0.1,0.39] & 0.45, [0.31,0.59] & 0.27, [0.13,0.41] & 0.31, [0.13,0.49] & -0.14, [-0.23,-0.04]\\
+race: white & -1.04, [-1.39,-0.66] & -1.65, [-1.94,-1.36] & -1.91, [-3.47,-0.42] & -1.01, [-1.19,-0.82] & -1.06, [-1.09,-1.03] & -1.72, [-2.36,-1.09] & -0.83, [-0.9,-0.77] & -0.85, [-0.87,-0.82] & -1.07, [-1.19,-0.94] & -0.88, [-1.02,-0.74] & -0.85, [-0.99,-0.72] & -0.49, [-0.71,-0.26] & -0.68, [-0.73,-0.62]\\
+implicit bias & -0.02, [-0.37,0.37] & -0.16, [-0.41,0.08] & 0.04, [-0.92,1.04] & 0, [-0.23,0.22] & 0.01, [-0.03,0.05] & -0.07, [-0.6,0.48] & -0.06, [-0.16,0.04] & 0.08, [0.03,0.12] & -0.09, [-0.27,0.1] & 0.04, [-0.12,0.21] & -0.1, [-0.29,0.08] & -0.03, [-0.26,0.2] & -0.01, [-0.11,0.08]\\
+explicit bias & 0.33, [-0.04,0.69] & -0.17, [-0.41,0.06] & -0.05, [-1.12,0.94] & 0.07, [-0.13,0.28] & 0.03, [-0.01,0.07] & -0.14, [-0.65,0.36] & 0.09, [-0.01,0.2] & 0, [-0.05,0.04] & -0.02, [-0.18,0.15] & -0.01, [-0.17,0.15] & 0.02, [-0.15,0.2] & 0.3, [0.07,0.52] & -0.06, [-0.15,0.02]\\
+implicit bias*race: white & 0.12, [-0.24,0.46] & 0.2, [-0.06,0.47] & -0.03, [-1.25,1.16] & 0.08, [-0.15,0.3] & -0.01, [-0.04,0.03] & 0.06, [-0.43,0.57] & -0.03, [-0.12,0.06] & -0.04, [-0.07,-0.01] & 0.01, [-0.16,0.17] & -0.09, [-0.24,0.05] & 0.11, [-0.04,0.25] & -0.19, [-0.42,0.04] & -0.04, [-0.1,0.04]\\
+explicit bias*race: white & -0.22, [-0.57,0.14] & 0.08, [-0.18,0.34] & -0.4, [-1.66,0.91] & -0.01, [-0.21,0.2] & 0, [-0.03,0.03] & -0.09, [-0.59,0.43] & -0.1, [-0.19,-0.01] & -0.02, [-0.05,0.01] & -0.02, [-0.17,0.12] & 0, [-0.14,0.14] & -0.02, [-0.17,0.12] & -0.14, [-0.35,0.07] & 0, [-0.06,0.06]\\
+\bottomrule
+\end{longtable}
+}
+\end{lltable}
 
-Finally, for the implicit and explicit bias measures, we relied on two
-primary variables from the Project Implicit data. Implicit bias was
-assessed via an Implicit Association Test (IAT). This test uses a
-speeded dual-categorization task in which individuals must quickly
-categorize black and white faces and \enquote{good} and \enquote{bad}
-words with key presses. The difference in how quickly and accurately
-participants are able to pair white faces with \enquote{good} words and
-black faces with \enquote{bad} words in comparison to the inverse is
-thought to reflect implicit associations between the two races and
-positive and negative affective reactions. This association is indicated
-in the IAT D-score, which we used as a measure of implicit bias. Our
-measure of explicit bias is the difference between reported warmth
-towards whites (i.e. \emph{how warm or cold do you feel towards Whites?
-0=very cold, 10=very warm}) and reported warmth toward blacks.
+\begin{lltable}
+\small{
+\begin{longtable}{llllllllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:mod-post}Regression coefficient estimates for the population-level (i.e.
+fixed) effects, along with 95\% uncertainty intervals for each of the disciplinary metrics.
+Model uses poststratified estimates of implicit and explicit bias.}\\
+\toprule
+ & \multicolumn{1}{c}{Seclusion} & \multicolumn{1}{c}{Preschool Suspension} & \multicolumn{1}{c}{Preschool Expulsion} & \multicolumn{1}{c}{Physical Restraint} & \multicolumn{1}{c}{Out-of-School Suspension} & \multicolumn{1}{c}{Mechanical Restraint} & \multicolumn{1}{c}{Law Enf.
+Referral} & \multicolumn{1}{c}{In-School Suspension} & \multicolumn{1}{c}{School-Related Arrest} & \multicolumn{1}{c}{Expulsion without Educational Services} & \multicolumn{1}{c}{Expulsion with Educational Services} & \multicolumn{1}{c}{Expulsion Under Zero-Tolerance} & \multicolumn{1}{c}{Corporal Punishment}\\
+\midrule
+Intercept & -11.11, [-11.56,-10.7] & -9, [-9.34,-8.7] & -14.17, [-15.91,-12.61] & -8.73, [-8.94,-8.52] & -2.69, [-2.72,-2.66] & -13.96, [-14.8,-13.19] & -6.12, [-6.21,-6.04] & -2.63, [-2.67,-2.59] & -8.19, [-8.36,-8.03] & -8.2, [-8.36,-8.03] & -8.09, [-8.26,-7.92] & -9.66, [-9.92,-9.41] & -3.24, [-3.34,-3.15]\\
+unemployment & 0.44, [0.18,0.71] & -0.21, [-0.4,-0.01] & -0.21, [-0.98,0.57] & 0.26, [0.11,0.41] & 0.21, [0.18,0.25] & -0.17, [-0.62,0.29] & 0.18, [0.09,0.26] & 0.06, [0.01,0.1] & 0.26, [0.11,0.41] & 0.41, [0.27,0.55] & 0.27, [0.13,0.41] & 0.26, [0.09,0.44] & -0.12, [-0.21,-0.02]\\
+college grads & 0.2, [-0.04,0.45] & -0.12, [-0.32,0.08] & -0.2, [-0.92,0.49] & 0.27, [0.14,0.41] & -0.04, [-0.07,-0.01] & -0.02, [-0.44,0.4] & -0.01, [-0.1,0.07] & -0.08, [-0.12,-0.03] & 0.17, [0.04,0.31] & -0.23, [-0.37,-0.09] & -0.05, [-0.19,0.08] & -0.14, [-0.31,0.03] & -0.12, [-0.22,-0.02]\\
+proportion white & 0.49, [0.02,0.98] & -0.56, [-0.92,-0.18] & -0.7, [-1.84,0.51] & 0.28, [0.01,0.56] & 0, [-0.06,0.06] & 0.18, [-0.58,1.02] & -0.29, [-0.44,-0.14] & 0.07, [-0.01,0.15] & -0.1, [-0.35,0.16] & 0.28, [0.01,0.54] & -0.26, [-0.5,-0.03] & 0.27, [-0.03,0.57] & -0.25, [-0.48,-0.02]\\
+black-white ratio & 0.52, [-0.08,1.05] & -0.76, [-1.14,-0.41] & -0.37, [-1.86,0.9] & 0.15, [-0.25,0.54] & -0.19, [-0.24,-0.13] & -0.84, [-1.92,0.18] & -0.01, [-0.18,0.15] & -0.36, [-0.44,-0.28] & 0.08, [-0.17,0.33] & -0.19, [-0.42,0.04] & -0.17, [-0.47,0.11] & -0.24, [-0.62,0.11] & -0.33, [-0.52,-0.14]\\
+race: white & -1.06, [-1.41,-0.67] & -1.6, [-1.92,-1.27] & -2.09, [-3.64,-0.56] & -1.02, [-1.2,-0.83] & -1.05, [-1.08,-1.02] & -1.77, [-2.43,-1.13] & -0.83, [-0.89,-0.76] & -0.84, [-0.86,-0.81] & -1.05, [-1.18,-0.92] & -0.9, [-1.04,-0.76] & -0.87, [-1,-0.73] & -0.53, [-0.75,-0.29] & -0.64, [-0.7,-0.58]\\
+poverty & -0.55, [-0.97,-0.14] & -0.44, [-0.76,-0.11] & -1.02, [-2.23,0.14] & -0.14, [-0.37,0.1] & 0.02, [-0.03,0.07] & 0.3, [-0.36,0.93] & -0.38, [-0.52,-0.24] & 0.16, [0.1,0.23] & -0.2, [-0.43,0.02] & -0.3, [-0.52,-0.09] & -0.17, [-0.39,0.05] & 0.23, [-0.04,0.49] & 0, [-0.16,0.16]\\
+total population & 0.14, [0,0.28] & 0.03, [-0.06,0.13] & 0.22, [0.01,0.45] & 0.03, [-0.05,0.11] & 0, [-0.03,0.02] & 0.32, [0.14,0.52] & 0.02, [-0.04,0.08] & -0.04, [-0.07,0] & 0.09, [0,0.18] & 0, [-0.08,0.09] & 0.06, [-0.03,0.16] & 0.09, [-0.02,0.2] & -0.44, [-0.62,-0.25]\\
+implicit bias & -0.42, [-0.7,-0.14] & 0.14, [-0.09,0.39] & 0.01, [-0.97,1.07] & -0.33, [-0.5,-0.16] & 0.08, [0.04,0.13] & -0.07, [-0.53,0.43] & -0.08, [-0.18,0.02] & 0.13, [0.08,0.19] & 0, [-0.16,0.16] & -0.03, [-0.19,0.12] & -0.03, [-0.2,0.13] & -0.15, [-0.35,0.06] & 0.01, [-0.11,0.14]\\
+implicit bias*race: white & -0.1, [-0.33,0.11] & -0.06, [-0.28,0.15] & -0.72, [-1.97,0.5] & 0.18, [0.05,0.31] & -0.04, [-0.06,-0.01] & -0.02, [-0.39,0.35] & -0.04, [-0.1,0.02] & -0.04, [-0.07,-0.01] & 0, [-0.1,0.09] & -0.02, [-0.12,0.09] & 0.09, [-0.01,0.18] & 0.07, [-0.08,0.22] & -0.11, [-0.18,-0.05]\\
+propotion black & -0.79, [-1.46,-0.11] & 0.78, [0.33,1.25] & 0.58, [-0.87,2.16] & -0.4, [-0.81,0.02] & 0.28, [0.21,0.36] & 0.86, [-0.2,1.92] & -0.28, [-0.48,-0.07] & 0.31, [0.2,0.41] & -0.26, [-0.58,0.07] & 0.47, [0.15,0.8] & -0.37, [-0.69,-0.03] & 0.17, [-0.24,0.59] & 0.13, [-0.17,0.42]\\
+explicit bias & 0.11, [-0.15,0.37] & -0.21, [-0.41,-0.02] & -0.51, [-1.39,0.42] & -0.04, [-0.2,0.13] & 0.06, [0.02,0.09] & -0.06, [-0.48,0.36] & 0.12, [0.03,0.21] & -0.16, [-0.2,-0.11] & -0.08, [-0.23,0.06] & 0.14, [0,0.28] & -0.06, [-0.2,0.08] & 0.25, [0.07,0.43] & -0.14, [-0.25,-0.04]\\
+explicit bias*race: white & -0.15, [-0.36,0.07] & -0.13, [-0.32,0.06] & -0.48, [-1.66,0.62] & 0.03, [-0.11,0.17] & 0, [-0.02,0.03] & -0.32, [-0.66,0.03] & -0.04, [-0.1,0.01] & 0, [-0.03,0.02] & -0.02, [-0.11,0.07] & 0.02, [-0.08,0.11] & 0.04, [-0.05,0.12] & 0.02, [-0.12,0.16] & -0.02, [-0.08,0.04]\\
+income & -0.27, [-0.58,0.03] & -0.38, [-0.65,-0.11] & -0.15, [-1.06,0.75] & -0.17, [-0.35,0.01] & -0.09, [-0.14,-0.05] & 0.03, [-0.49,0.54] & -0.03, [-0.14,0.08] & -0.11, [-0.16,-0.05] & -0.01, [-0.18,0.17] & -0.29, [-0.47,-0.11] & -0.11, [-0.28,0.07] & -0.06, [-0.28,0.16] & -0.38, [-0.53,-0.22]\\
+\bottomrule
+\end{longtable}
+}
+\end{lltable}
 
-\subsection{Data analysis}\label{data-analysis}
-
-Data analysis proceded in two steps. We first estimated county-level
-implict and explicit bias using multilevel regression and
-post-stratification. Post-stratification is a valuable procedure in
-obtaining accurate geographical population-based estimates because it
-allows a non-representative sample (e.g Project Implicit) to more
-closely resemble the true population, and it regularizes extreme
-observations with little data to support them (e.g.~a county with only a
-handful of respondents with especially high or low scores) (Gelman \&
-Little, 1997; Park, Gelman, \& Bafumi, 2004). Following past work
-(Leitner, Hehman, Ayduk, \& Mendoza-Denton, 2016), we identified age as
-one dimension along which IAT respondents differed from the general
-population in ways that could bias our conclusions (Gonsalkorale,
-Sherman, \& Klauer, 2009). Our post-stratification weighting scheme is
-as follows: We first grouped respondents into five age group categories
-(15-24, 25-34, 35-54, 55-75, and 75+). We next fit multilevel models
-estimating bias (implicit and explicit biases seperately) as a function
-of our state-level covariates (the \enquote{fixed} effects), and allowed
-the estimates to vary by age bin, county, and state (the
-\enquote{random} effects). Next, we determined the population of whites
-in each county in these age groups using the American Community Survey's
-5-year estimates ending in 2014. Finally, we used our estimated models
-to predict the expected response for each age bin, in each county. Our
-final county-level estimates are the average of the values predicted for
-the 5 age bins, weighted by the population size of that bin in that
-county. As a result of this procedure, we can be confident that our
-county-level estimates should more closely approximate what our
-estimates would look like if the Project Implicit data were truly
-representative along the age dimension in all counties.
-
-After obtaining these estimates, we use them as predictors in bayesian
-multilevel logistic regressions. Formally, the likelihood for a given
-observation is written as a binomial function:
-
-\[
-{n\choose y}\pi^y(1-\pi)^{n-y}
-\]
-
-Where \(y\) is the observed count of incidents (e.g.~number of black or
-white students suspended), \(n\) is the number of at-risk students
-(e.g.~total number of black or white students), and
-\(\pi = g^{-1}(\eta)\) is the probability of the incident ocurring. For
-this analysis, the linear predictor takes the form of a multilevel model
-with a set of effects that vary over county:
-
-\[
-\eta = \alpha + X\beta + \gamma_{county}
-\]
-
-Where \(\alpha\) is an intercept that is constant across observations,
-\(\beta\) represents a set of effects that are also constant across
-observations (i.e.~fixed effects), and \(\gamma_{county}\) represents
-intercepts and effects of ethnicity that vary across the counties
-(i.e.~random efffects).
-
-In addition to the covariates described above, we also include effects
-of race, implicit bias, explicit bias, and the two-way interactions
-between implicit bias and race and explicit bias and race. We fit
-separate models for each of the outcomes.
-
-Because of the computational demands of fitting such a high-dimensional
-model to such a large dataset (the full model for each metric would
-consist of over 6k parameters to approximately 170k observations), we
-used a consensus monte carlo algorithm to obtain approximate posterior
-distributions for the parameters of interest (Scott et al., 2016). The
-approximate posteriors derived from this algorithm have been shown to be
-nearly indistinguishable from the true posterior, a result we verified
-using a small subset of our own data.
-
-All numerical predictor variables were standardized at the appropriate
-level (county, state) before model estimation to help with estimation
-efficiency and interpretability. We set priors for the intercept and
-coefficients in the bayesian models to be weakly informative normal
-distributions centered on zero with a standard deviation of five. All
-other parameters were left to default values. Data analysis was done in
-R (R Core Team, 2016) version 3.3.2 running under OS X 10.11.6.
-Post-stratification was done with lme4, version 1.1.14 (Bates, Mächler,
-Bolker, \& Walker, 2015). Final model fitting was done on the university
-cluster running Springdale Linux, release 6.9 using rstanarm, version
-2.17.2 (Stan Development Team, 2016). We used the implementation of the
-consensus monte carlo algorithm found in parallelMCMCcombine, version
-1.0 (Miroshnikov \& Conlon, 2014). Figures were made with ggplot2,
-version 2.2.1 (Wickham, 2009), with data manipulation done using dplyr
-version 0.7.2 (Wickham, Francois, Henry, \& Müller, 2017) and tidyr,
-version 0.7.1 (Wickham \& Henry, 2017). A full report of session
-information can be found on the OSF page (\ldots{}.)
-
-\section{Results}\label{results}
-
-\subsection{Project Implicit
-Estimates}\label{project-implicit-estimates}
-
-We first report the results of estimating the implicit and explicit
-biases in from Project Implicit data. Overall, the individuals in
-project implicit show a pro-white bias in both implicit (\emph{mean} =
-0.40, \emph{sd} = 0.41), and explicit measures (\emph{mean} = 0.88
-\emph{sd} = 1.83). When aggregated at the county level and adjusted with
-poststratification, the summary statistics across counties are similar
-in terms of their location, but as expected, the variability is much
-diminished (\(mean_{implicit}\) = 0.40, \(sd_{implicit}\) = 0.02;
-\(mean_{explicit}\) = 0.79, \(sd_{explicit}\) = 0.15, where on both
-scales 0 = no bias, and positive numbers indicate a pro-white bias).
-
-\subsection{Disciplinary action
-frequency}\label{disciplinary-action-frequency}
-
-\begin{table}[tbp]
-\begin{center}
-\begin{threeparttable}
-\caption{\label{tab:disc-count}Count of students by race receiving each type of disciplinary action}
-\begin{tabular}{lll}
+\begin{lltable}
+\small{
+\begin{longtable}{llllllllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:mod-teach}Regression coefficient estimates for the population-level (i.e.
+fixed) effects, along with 95\% uncertainty intervals for each of the disciplinary metrics.
+Model uses only data from Project Implicit respondents who identified as teachers.}\\
 \toprule
-group & \multicolumn{1}{c}{metric} & \multicolumn{1}{c}{students}\\
+ & \multicolumn{1}{c}{Seclusion} & \multicolumn{1}{c}{Preschool Suspension} & \multicolumn{1}{c}{Preschool Expulsion} & \multicolumn{1}{c}{Physical Restraint} & \multicolumn{1}{c}{Out-of-School Suspension} & \multicolumn{1}{c}{Mechanical Restraint} & \multicolumn{1}{c}{Law Enf.
+Referral} & \multicolumn{1}{c}{In-School Suspension} & \multicolumn{1}{c}{School-Related Arrest} & \multicolumn{1}{c}{Expulsion without Educational Services} & \multicolumn{1}{c}{Expulsion with Educational Services} & \multicolumn{1}{c}{Expulsion Under Zero-Tolerance} & \multicolumn{1}{c}{Corporal Punishment}\\
 \midrule
-black & expulsions & 36,755.00\\
-black & school arrests & 21,456.00\\
-black & in-school suspension & 829,706.00\\
-black & law enforcement referral & 53,306.00\\
-black & out-of-school suspension & 1,850,492.00\\
-white & expulsions & 55,832.00\\
-white & school arrests & 21,420.00\\
-white & in-school suspension & 1,054,172.00\\
-white & law enforcement referral & 81,565.00\\
-white & out-of-school suspension & 1,768,326.00\\
+Intercept & -10.12, [-10.63,-9.66] & -8.6, [-9.04,-8.21] & -13.68, [-15.54,-12.13] & -8.22, [-8.46,-8] & -2.34, [-2.4,-2.29] & -12.87, [-13.88,-11.99] & -5.57, [-5.7,-5.44] & -2.7, [-2.77,-2.62] & -7.31, [-7.55,-7.07] & -8.41, [-8.72,-8.09] & -7.61, [-7.86,-7.36] & -9.38, [-9.72,-9.06] & -6.25, [-7.29,-5.37]\\
+poverty & -0.33, [-1.31,0.63] & 0.31, [-0.45,1.04] & 0.39, [-1.6,2.26] & -0.01, [-0.51,0.47] & -0.06, [-0.17,0.04] & 0.41, [-0.99,1.8] & -0.26, [-0.53,0] & 0.05, [-0.1,0.21] & 0.08, [-0.43,0.57] & -0.12, [-0.82,0.56] & 0.28, [-0.19,0.73] & 0.4, [-0.17,0.95] & -0.16, [-2.01,1.48]\\
+college grads & 0.14, [-0.36,0.66] & -0.1, [-0.5,0.31] & -0.3, [-1.69,0.99] & 0.33, [0.08,0.58] & -0.15, [-0.22,-0.08] & -0.5, [-1.27,0.27] & -0.03, [-0.2,0.13] & -0.26, [-0.36,-0.15] & 0, [-0.29,0.27] & -0.16, [-0.52,0.2] & -0.22, [-0.52,0.07] & -0.25, [-0.63,0.11] & -0.11, [-1.01,0.78]\\
+proportion white & 1.5, [0.46,2.52] & 0.25, [-0.66,1.16] & 1.39, [-0.37,3.33] & 0.74, [0.24,1.25] & 0.07, [-0.04,0.19] & 0.79, [-0.7,2.31] & -0.29, [-0.56,-0.01] & 0.23, [0.06,0.4] & 0.31, [-0.21,0.85] & 0.81, [0.14,1.5] & 0.2, [-0.29,0.7] & 0.17, [-0.48,0.83] & -1.73, [-4.31,0.84]\\
+black-white ratio & 0.18, [-1.4,1.69] & -0.51, [-1.2,0.16] & -1.76, [-3.74,0.07] & -0.01, [-0.65,0.63] & -0.08, [-0.18,0.03] & -1.82, [-3.66,-0.06] & 0.07, [-0.3,0.44] & -0.32, [-0.49,-0.16] & 0.27, [-0.5,0.99] & 0.04, [-0.62,0.73] & -0.09, [-0.74,0.54] & -0.01, [-1.09,1.04] & -0.85, [-2.88,0.99]\\
+implicit bias & -0.38, [-0.97,0.21] & 0.15, [-0.32,0.63] & -1.04, [-2.7,0.62] & -0.4, [-0.72,-0.07] & -0.02, [-0.11,0.06] & -0.65, [-1.64,0.3] & 0.08, [-0.12,0.27] & -0.02, [-0.13,0.09] & -0.07, [-0.43,0.29] & -0.2, [-0.62,0.23] & -0.56, [-0.94,-0.18] & -0.15, [-0.59,0.28] & 0.09, [-1.2,1.47]\\
+explicit bias*race: white & 0.11, [-0.27,0.48] & -0.12, [-0.57,0.31] & -2.05, [-4.45,0.26] & 0.11, [-0.12,0.34] & -0.01, [-0.06,0.05] & -0.68, [-1.51,0.1] & 0.03, [-0.07,0.14] & -0.01, [-0.06,0.05] & 0, [-0.18,0.16] & -0.25, [-0.46,-0.04] & -0.04, [-0.22,0.13] & -0.07, [-0.34,0.21] & -0.21, [-0.98,0.51]\\
+income & -0.4, [-1.23,0.39] & -0.21, [-0.91,0.49] & 1.4, [-0.21,3] & -0.42, [-0.84,-0.01] & -0.23, [-0.33,-0.13] & -0.13, [-1.45,1.13] & -0.37, [-0.6,-0.13] & -0.2, [-0.36,-0.05] & -0.05, [-0.49,0.4] & -0.53, [-1.15,0.05] & 0.11, [-0.34,0.57] & -0.11, [-0.69,0.41] & -0.38, [-1.72,0.95]\\
+total population & 0.35, [0.01,0.72] & 0.16, [-0.16,0.47] & 0.43, [-0.06,0.94] & 0.02, [-0.19,0.23] & -0.05, [-0.12,0.01] & 0.66, [0.11,1.25] & -0.02, [-0.16,0.12] & 0, [-0.1,0.09] & 0.09, [-0.14,0.33] & 0.12, [-0.16,0.4] & 0.06, [-0.22,0.33] & 0.07, [-0.19,0.34] & -0.97, [-2.25,0.07]\\
+explicit bias & 0.12, [-0.43,0.66] & 0.05, [-0.45,0.56] & -0.07, [-1.32,1.22] & -0.2, [-0.51,0.13] & 0.04, [-0.03,0.12] & 0.81, [-0.11,1.8] & 0.09, [-0.1,0.28] & 0, [-0.11,0.11] & -0.05, [-0.4,0.31] & 0.08, [-0.34,0.5] & -0.05, [-0.43,0.32] & 0.07, [-0.37,0.52] & 0.51, [-0.71,1.75]\\
+propotion black & 0.51, [-0.79,1.84] & 1.2, [0.27,2.13] & 2.88, [0.86,4.95] & 0.06, [-0.55,0.68] & 0.12, [-0.01,0.26] & 1.98, [0.19,3.79] & -0.58, [-0.97,-0.21] & 0.49, [0.29,0.69] & -0.25, [-1,0.51] & 0.22, [-0.55,0.99] & -0.4, [-1.04,0.26] & -0.6, [-1.54,0.31] & 0.02, [-2.7,3.11]\\
+unemployment & -0.17, [-0.69,0.37] & -0.55, [-0.98,-0.15] & -0.04, [-1.06,1.05] & -0.07, [-0.33,0.2] & 0.11, [0.04,0.18] & -0.8, [-1.65,0.06] & 0.04, [-0.12,0.21] & -0.09, [-0.19,0.02] & 0.12, [-0.17,0.41] & 0.36, [-0.02,0.73] & 0.11, [-0.21,0.43] & 0.02, [-0.36,0.39] & -0.77, [-1.76,0.11]\\
+race: white & -1.34, [-1.71,-0.97] & -1.63, [-2.07,-1.24] & -3.01, [-5.02,-1.01] & -1.13, [-1.31,-0.93] & -1.46, [-1.5,-1.42] & -1.38, [-2.13,-0.55] & -1.01, [-1.09,-0.93] & -1.14, [-1.18,-1.1] & -1.1, [-1.24,-0.95] & -1.07, [-1.25,-0.87] & -1.06, [-1.21,-0.92] & -0.72, [-0.96,-0.46] & -0.53, [-1.16,0.12]\\
+implicit bias*race: white & 0.01, [-0.4,0.41] & -0.27, [-0.69,0.14] & 0.43, [-2.21,2.98] & 0.21, [-0.03,0.45] & 0, [-0.06,0.05] & 0.18, [-0.67,1.05] & -0.07, [-0.18,0.03] & 0, [-0.06,0.05] & 0.04, [-0.13,0.2] & -0.16, [-0.36,0.04] & -0.02, [-0.2,0.17] & -0.1, [-0.38,0.18] & -0.01, [-0.69,0.72]\\
 \bottomrule
-\end{tabular}
-\end{threeparttable}
-\end{center}
-\end{table}
-
-Table \ref{tab:disc-count} shows the number of students of each race who
-were reported having received each of the actions under consideration.
-The counts range from a low of just 21420 white students arrested to a
-high of 1850492 black students receiving an out-of-school suspension.
-Considering the vast differences in the overall number of black and
-white students, this simple count already illustrates that black
-students are disciplined at rates far higher than their white
-counterparts.
-
-\subsection{Associations across
-county}\label{associations-across-county}
-
-Figure \ref{fig:overall-associations} shows the estimate of primary
-interest for each of the models. The estimates displayed are the
-coefficients for the interaction between race and each of the two bias
-measurements. Given that African Americans are the baseline group,
-negative values for this coefficient indicate that as the bias in a
-county increases, the gap between the probability of a black student
-being disciplined and the probability of a white student being
-disciplined grows.
-
-\begin{figure}
-\centering
-\includegraphics{draft_files/figure-latex/overall-associations-1.pdf}
-\caption{\label{fig:overall-associations}Association between each metric and
-county-level estimates of explicit and implicit bias. Negative values
-indicate that the rate of increase (or decrease) for blacks is faster
-(or slower) than for whites. Point is the mean of the posterior and
-error bars represent 95\% bayesian uncertainty intervals.}
-\end{figure}
-
-Several patterns are apparent from this figure. First, with the
-exceptions of implicit bias and expulsions and law enforcement
-referrals, all estimates are directionally consistent with higher levels
-of bias leading to larger differences between groups. Second, these
-effects are especially consistent for the two types of suspensions. The
-largest effect estimated is between implicit bias and out-of-school
-suspensions. The difference in the slope of the association between
-implicit bias and the log of the odds for out-of-school suspensions
-between white and black students is estimated to be -0.25, with 95\% of
-the posterior distribution between -0.31 and -0.20 and a proportion
-\textgreater{}.99 of the posterior distribution consistent with a
-negative effect.
-
-Although not nearly as large of a difference, we have similar certainty
-with respect to the association between out-of-school suspensions and
-explicit bias. The difference in the slope of the association between
-explicit bias and the log of the odds for out-of-school suspensions
-between white and black students is estimated to be -0.08, with 95\% of
-the posterior distribution between -0.14 and -0.03 and a proportion
-\textgreater{}.99 of the posterior distribution consistent with a
-negative effect.
-
-The estimated associations for in-school suspensions are smaller still,
-but are generally consistent with effects of the same direction. For
-implicit bias, the relevant parameter is estimated at -0.02 {[}-0.06,
-0.01{]} \(p_{neg}\) = 0.93, and for explicit bias, the effect is
-slightly larger, and a positive effect is essentially not credible,
-given the data and model -0.05 {[}-0.08, -0.02{]} \(p_{neg}\)
-\textgreater{}.99.
-
-Other outcomes are estimated with less precision, or with patterns that
-are inconsistent between implicit and explicit bias. For instance,
-examining the associations for expulsions, the effect of explicit biases
-are generally in the expected direction (\emph{est} = -0.07, {[}-0.15,
-0.02{]}, \(p_{neg}\) = .94), but the effect for implicit bias is
-estimated to be close to zero, with a enough uncertainty (\emph{est} =
-0.04, {[}-0.05, 0.12{]}, \(p_{neg}\) = .19) to make it difficult to
-claim an effect of one direction or the other. The model indicates
-similar uncertainty with respect to school-related arrests and both
-estimates of bias (\(est_{explicit}\) = -0.02, {[}-0.11, 0.07{]},
-\(p_{neg}\) = .67; \(est_{implicit}\) = -0.02, {[}-0.12, 0.08{]},
-\(p_{neg}\) = .66).
-
-\begin{figure}
-\centering
-\includegraphics{draft_files/figure-latex/detail-fig-exp-1.pdf}
-\caption{\label{fig:detail-fig-exp}Association between bias and expulsions.
-Top: Association between bias and the estimated probability of
-expulsion. Line is the mean of the posterior. Bands indicate 95\%
-uncertainty intervals; Bottom: Association between bias and the relative
-risk ratio for black students to white students. Points represent
-counties, whose size are scaled to the number of schools in that
-county.}
-\end{figure}
-
-To better illustrate the nature of these relationships, figure
-\ref{fig:detail-fig-exp} shows the estimated probabilities of expulsion
-for black and white students as a function of bias, along with the
-relative risk ratio for black students. The relative risk ratio is the
-ratio of the probability that a black student will be expelled to the
-probability that a white student will get expelled. Values over 1
-reflect higher levels of punishment for black students. As previously
-indicated in table \ref{tab:disc-count}, the top part of this figure
-makes plain the higher probability of expulsion for black children. In a
-county at the mean of the distribution of bias, approximately 0.15\% of
-black students are expected to be expelled {[}0.14, 0.17{]}. The
-corresponding rate for white students is much lower, with about 0.06\%
-expected to be expelled {[}0.06, 0.07{]}. Moving to a county one
-standard deviation above the mean of explicit bias has the effect of
-increasing the estimated percentage of black students expected to be
-expelled to 0.16\% {[}0.13, 0.18{]}, while the percentage of white
-students expected to be expelled would decline to 0.06\% {[}0.05,
-0.07{]}. The same movement for implicit bias would slightly increase the
-expected expulsions for black students ( 0.15\% {[}0.13, 0.18), and
-increase the expected expulsions for white students a very small amount
-more to (0.06\% {[}0.06, 0.07). In real terms, in a county at the mean
-of the distributions of bias, for every white student expelled, we
-should expect 2.45 {[}2.25, 2.45{]} black students to be expelled. If we
-move to a county one standard deviation above the mean of explicit bias,
-the ratio of black to white students expelled increases to 2.63 {[}2.33,
-2.63{]}, while the same movement for implicit bias slightly decreases
-the ratio to 2.36 {[}2.13, 2.36{]}.
-
-\begin{figure}
-\centering
-\includegraphics{draft_files/figure-latex/detail-fig-susp-1.pdf}
-\caption{\label{fig:detail-fig-susp}Association between bias and
-out-of-school suspensions Top: Association between bias and the
-estimated probability of suspension Line is the mean of the posterior.
-Bands indicate 95\% uncertainty intervals; Bottom: Association between
-bias and the relative risk ratio for black students to white students.
-Points represent counties, whose size are scaled to the number of
-schools in that county.}
-\end{figure}
-
-Figure \ref{fig:detail-fig-susp} shows similar patterns, but for
-out-of-school suspensions. In a county at the mean of the distribution
-of bias, approximately 11.50\% of black students are expected to be
-suspended {[}11, 12{]}. The corresponding rate for white students is
-just over half that for black students, with about 6.40\% expected to be
-suspended {[}6.20, 6.50{]}. Moving to a county one standard deviation
-above the mean of explicit bias has the effect of slightly decreasing
-the estimated percentage of black students expected to be suspended to
-11.10\% {[}10.30, 11.80{]}, while the percentage of white students
-expected to be suspended would decrease at a faster rate to 5.60\%
-{[}5.40, 5.90{]}. The same movement for implicit bias would dramatically
-increase the expected suspensions for black students (15.40\% {[}14.40,
-16.40), and increase the expected suspensions for white students a much
-smaller amount to (6.90\% {[}6.60, 7.30). In real terms, in a county at
-the mean of the distributions of bias, for every white student expelled,
-we should expect 1.80 {[}1.73, 1.80{]} black students to be expelled. If
-we move to a county one standard deviation above the mean of explicit
-bias, the ratio of black to white students expelled increases to 1.96
-{[}1.84, 1.96{]}, while the same movement for implicit bias slightly
-decreases the ratio to 2.23 {[}2.11, 2.23{]}.`
-
-\section{Discussion}\label{discussion}
-
-\newpage
-
-\section{References}\label{references}
-
-\newpage
-
-\section{Appendix}\label{appendix}
-
-\subsection{Preregistered analysis}\label{preregistered-analysis}
-
-In our preanalysis plan, we specified our analyses to focus on 13
-actions - corporal punishment, in-school suspension, out-of-school
-suspension, expulsion with educational services, expulsion without
-educational services, expulsion under zero-tolerance policies, referral
-to law enforcement, school-related arrests, mechanical restraint,
-physical restraint, seclusion, preschool suspension, and preschool
-expulsion. However, upon further study, we discovered reasons we thought
-justified excluding a number of these outcomes. In particular,
-seclusion, physical restraint, and mechanical restraint are not
-disciplinary actions, but are rather used as means to restrain students
-who are at risk of harming themselves or others. Additionally, the
-number of preschool students who are expelled or suspended is
-vanishingly small (131 total expulsions and 6751 total suspensions out
-of over 1.4 million enrolled preschool students), making reliably
-estimating any association across counties exceedingly unlikely. We
-additionally discovered that counts of one expulsion category (expulsion
-under zero-tolerance policies) overlapped with counts in other
-categories, and so excluded this category. To remain consistent with
-previous studies, we opted to combine the remaining two expulsion
-categories to yield one overall count of the number of students
-expelled.
-
-We also preregistered our explicit bias as a simple feeling thermometer
-towards balcks (i.e. \emph{how warm or cold do you feel towards Blacks?
-0=very cold, 10=very warm}). However, past research (Hehman, Flake, \&
-Calanchini, 2017; Leitner et al., 2016) has used the difference in
-reported warmth towards whites and blacks, and so in the main text, we
-report models using this metric of explicit bias. Additionally, we
-preregistered analyses with poststratified estimates (as presented in
-the main text) along with raw, county-based means. Finally, we had not
-known about the issues with juvenile justice facilities, or with the
-school districts with reporting errors. Here, we present the results of
-the preregistered analyses exactly.
-
-\section{Simple county means}\label{simple-county-means}
-
-\section{Post stratified estimates}\label{post-stratified-estimates}
-
-\section{Teacher analyses}\label{teacher-analyses}
+\end{longtable}
+}
+\end{lltable}
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-anyon2014persistent}{}
+Anyon, Y., Jenson, J. M., Altschul, I., Farrar, J., McQueen, J., Greer,
+E., \ldots{} Simmons, J. (2014). The persistent effect of race and the
+promise of alternatives to suspension in school discipline outcomes.
+\emph{Children and Youth Services Review}, \emph{44}, 379--386.
+
+\hypertarget{ref-aud2011america}{}
+Aud, S., KewalRamani, A., \& Frohlich, L. (2011). America's youth:
+Transitions to adulthood. nces 2012-026. \emph{National Center for
+Education Statistics}.
+
 \hypertarget{ref-bates2015fitting}{}
@@
 doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}
+
+\hypertarget{ref-cuellar2015school}{}
+Cuellar, A. E., \& Markowitz, S. (2015). School suspension and the
+school-to-prison pipeline. \emph{International Review of Law and
+Economics}, \emph{43}, 98--106.
+
+\hypertarget{ref-friese2009and}{}
+Friese, M., Hofmann, W., \& Schmitt, M. (2009). When and why do implicit
+measures predict behaviour? Empirical evidence for the moderating role
+of opportunity, motivation, and process reliance. \emph{European Review
+of Social Psychology}, \emph{19}(1), 285--338.
+
+\hypertarget{ref-galdi2008automatic}{}
+Galdi, S., Arcuri, L., \& Gawronski, B. (2008). Automatic mental
+associations predict future choices of undecided decision-makers.
+\emph{Science}, \emph{321}(5892), 1100--1102.
+
+\hypertarget{ref-gawronski2003implicit}{}
+Gawronski, B., Geschke, D., \& Banse, R. (2003). Implicit bias in
+impression formation: Associations influence the construal of
+individuating information. \emph{European Journal of Social Psychology},
+\emph{33}(5), 573--589.
 
@@
 
+\hypertarget{ref-gregory2008discipline}{}
+Gregory, A., \& Weinstein, R. S. (2008). The discipline gap and african
+americans: Defiance or cooperation in the high school classroom.
+\emph{Journal of School Psychology}, \emph{46}(4), 455--475.
+
+\hypertarget{ref-gregory1995crime}{}
+Gregory, J. F. (1995). The crime of punishment: Racial and gender
+disparities in the use of corporal punishment in us public schools.
+\emph{Journal of Negro Education}, 454--462.
+
+\hypertarget{ref-haight2014ecological}{}
+Haight, W., Gibson, P. A., Kayama, M., Marshall, J. M., \& Wilson, R.
+(2014). An ecological-systems inquiry into racial disproportionalities
+in out-of-school suspensions from youth, caregiver and educator
+perspectives. \emph{Children and Youth Services Review}, \emph{46},
+128--138.
+
 \hypertarget{ref-hehman2017disproportionate}{}
@@
 1948550617711229.
+
+\hypertarget{ref-hirschfield2009another}{}
+Hirschfield, P. (2009). Another way out: The impact of juvenile arrests
+on high school dropout. \emph{Sociology of Education}, \emph{82}(4),
+368--393.
+
+\hypertarget{ref-hofmann2005meta}{}
+Hofmann, W., Gawronski, B., Gschwendner, T., Le, H., \& Schmitt, M.
+(2005). A meta-analysis on the correlation between the implicit
+association test and explicit self-report measures. \emph{Personality
+and Social Psychology Bulletin}, \emph{31}(10), 1369--1385.
+
+\hypertarget{ref-kinsler2011understanding}{}
+Kinsler, J. (2011). Understanding the black--white school discipline
+gap. \emph{Economics of Education Review}, \emph{30}(6), 1370--1383.
 
@@
 
+\hypertarget{ref-matjasko2011effective}{}
+Matjasko, J. L. (2011). How effective are severe disciplinary policies?
+School policies and offending from adolescence into young adulthood.
+\emph{Journal of School Psychology}, \emph{49}(5), 555--572.
+
+\hypertarget{ref-mcconnell2001relations}{}
+McConnell, A. R., \& Leibold, J. M. (2001). Relations among the implicit
+association test, discriminatory behavior, and explicit measures of
+racial attitudes. \emph{Journal of Experimental Social Psychology},
+\emph{37}(5), 435--442.
+
 \hypertarget{ref-miroshnikov2014parallel}{}
@@
 
+\hypertarget{ref-nosek2011implicit}{}
+Nosek, B. A., Hawkins, C. B., \& Frazier, R. S. (2011). Implicit social
+cognition: From measures to mechanisms. \emph{Trends in Cognitive
+Sciences}, \emph{15}(4), 152--159.
+
+\hypertarget{ref-okonofua2016vicious}{}
+Okonofua, J. A., Walton, G. M., \& Eberhardt, J. L. (2016). A vicious
+cycle: A social--psychological account of extreme racial disparities in
+school discipline. \emph{Perspectives on Psychological Science},
+\emph{11}(3), 381--398.
+
+\hypertarget{ref-pager2009sequencing}{}
+Pager, D., Western, B., \& Sugie, N. (2009). Sequencing disadvantage:
+Barriers to employment facing young black and white men with criminal
+records. \emph{The ANNALS of the American Academy of Political and
+Social Science}, \emph{623}(1), 195--213.
+
 \hypertarget{ref-park2004bayesian}{}
@@
 polls. \emph{Political Analysis}, \emph{12}(4), 375--385.
+
+\hypertarget{ref-pascoe2009perceived}{}
+Pascoe, E. A., \& Smart Richman, L. (2009). Perceived discrimination and
+health: A meta-analytic review. \emph{Psychological Bulletin},
+\emph{135}(4), 531.
 
@@
 
+\hypertarget{ref-yeager2017loss}{}
+Yeager, D. S., Purdie-Vaughns, V., Hooper, S. Y., \& Cohen, G. L.
+(2017). Loss of institutional trust among racial and ethnic minority
+adolescents: A consequence of procedural injustice and a cause of
+life-span outcomes. \emph{Child Development}, \emph{88}(2), 658--676.
+
 \end{document}
